---
title: "The Silicon Sample Tournament: A multi-team benchmark of AI approaches for predicting behavioral-experiment effects"

abstract: |
    This is an amendment to our preregistration *"Predicting the effects of behavioral interventions on humans with LLMs"*. Writing the initial preregistration, we conceived of the project as a single-team simulation project. With this amendment, we turn the project into a **multi-team prediction tournament**: we invite independent research teams to predict the results of a novel behavioral megastudy (16 message interventions to increase trust in climate scientists, 13 preregistered outcomes, a representative U.S. sample of ~18,000) using any AI-based approach they choose—behavioral clones, group-level simulation, or direct effect forecasting. Submissions enter at one of **three tiers** defined by the granularity of the prediction (individual-level synthetic data, cell-level statistics, or effect-level estimates with prediction intervals), and every approach is scored on the preregistered metric ladder against the same human–human ceiling. The initial preregistration remains valid: it specifies in detail how we measure the (dis)similarity between the human sample and simulated samples, and all analytic procedures laid out there are kept in place. At the time of writing this amendment, human data are partially collected (~10,000 of the ~18,000 responses); none of these data have been shared, published, or made accessible to any prospective team, and no team will be granted access before the prediction lock. Before any human data are released, every team registers its approach using a standardized method-registration form (an extended version of the GUIDE-LLM reporting checklist) and deposits its predictions on Zenodo—publicly, or partly escrowed under a disclosure policy that keeps every entry locked and verifiable even where it cannot be fully public.

date: "2026-06-11"

format:
  html:
    code-fold: true
    toc: true
    toc-depth: 4
  # PDF disabled for now to speed up iteration — re-enable for the final render.
  # pdf:
  #   toc: true
  #   toc-depth: 5
  #   include-before-body:
  #     text: |
  #       \clearpage
  #   colorlinks: true
  #   tbl-colwidths: true
  #   header-includes:
  #     - \usepackage{float}
  #     - \floatplacement{figure}{H}
  #     - \floatplacement{table}{H}
  #     - \usepackage{listings}
  #     - \lstset{breaklines=true, basicstyle=\small\ttfamily}

execute:
  echo: true
  warning: false
  message: false
  fig-width: 8
  fig-height: 5
  fig-dpi: 300
---

```{r}
#| label: setup
#| output: false

# packages
library(tidyverse)
library(sandwich)
library(lmtest)
library(marginaleffects)
library(broom)
library(MetBrewer)
library(patchwork)
library(gt)
library(tinytable)
library(here)

# custom functions (reused unchanged from the original preregistration)
source(here("R/functions/statistics.R"))
source(here("R/functions/plots.R"))
source(here("R/functions/tables.R"))
source(here("R/functions/reporting.R"))

# simulated (random) placeholder data
human_data     <- readRDS(here("data/simulation_preregistration/human_data_preregistration.rds"))
llm_data_teams <- readRDS(here("data/simulation_preregistration/llm_data_preregistration_teams.rds"))

# Same preregistered random split of the human sample as in the original
# preregistration. Human 1 is the reference half all submissions are scored
# against; Human 2 is the human-human ceiling.
set.seed(42)
split_ids    <- sample(unique(human_data$id), size = floor(length(unique(human_data$id)) / 2))
human_data_1 <- human_data |> filter(id %in% split_ids)
human_data_2 <- human_data |> filter(!id %in% split_ids)
```


::: {.callout-warning}
**This is an ongoing draft, to share with collaborators. The amendment is not yet registered and will change still.**
:::

::: {.callout-important}
**This is an amendment.** It builds on, and does not replace, the original preregistration ([Preregistration](preregistration.qmd)). All analytical procedures are kept in place. This document adds (i) the tournament design with three submission tiers, (ii) a standardized method-registration framework and disclosure policy, (iii) per-tier submission specifications, and (iv) a new cross-team comparison.

The tournament's public-facing design—motivation, rules, the three tiers, the full method-registration form, the disclosure policy, scoring, and the timeline—is described in detail in the [call for participation](index.html). To avoid duplication, this amendment summarizes that design only as far as the analysis needs it, and concentrates on the **preregistered analysis plan**. Where the two documents overlap, the call is the canonical description of the procedure and this amendment is the canonical description of the analysis.

All datasets used in this amendment are simulated random noise with the correct structure.
:::

## Motivation

Whether large language models can stand in for human participants is one of the most consequential open questions in behavioral science, and the evidence so far is mixed and hard to interpret: most existing evaluations target experiments that were already published—and plausibly in the models' training data—and almost none are preregistered, so reported successes are entangled with researcher degrees of freedom in model and prompt choice [@cumminsThreatAnalyticFlexibility2025b]. The original preregistration addresses this with a single approach—a group simulation originally planned by the core team, built from demographically representative U.S. profiles with a particular model and prompt—but a single team's result is only one point in a large design space.

This amendment therefore turns the project into a **multi-team prediction tournament**, the *Silicon Sample Tournament*: independent teams predict the same novel megastudy (N ≈ 18,000) with any AI-based approach, and every approach is scored on the same preregistered metric ladder against the same human–human ceiling. The design removes the two problems of existing evaluations by construction—the target is genuinely novel (none of the interventions has ever been tested in this form), and every prediction is deposited and time-stamped before any human data are released. The full rationale, rules, and incentives for participating teams are laid out in the [call for participation](index.html); here we keep only what is needed to motivate and interpret the analysis.

The benchmark scores the 16 non-interactive text interventions of the megastudy^[The original experiment contains 20 conditions; the tournament scores only the 16 non-interactive text interventions. The four left out are the three interactive LLM-chatbot arms and the "Value similarity" quiz, all of which depend on a live back-and-forth that would substantially complicate the simulation task.], across self-reported attitudes (e.g., support for funding of climate science) and behavioral measures (e.g., donations to a scientific organization, a newsletter sign-up).

@tbl-interventions gives an overview of the 16 interventions, grouped by the trust dimension they target. All are short text messages (an article, an interview excerpt, a corrected statistic, a portrait) shown once before the outcome battery; together with the shared control condition they make up the 17 conditions every submission covers.

```{r}
#| label: tbl-interventions
#| tbl-cap: "The 16 text-based interventions scored in the tournament, grouped by the broad angle each takes on trust in climate scientists. The original megastudy also fields three interactive LLM-chatbot arms and a 'Value similarity' quiz; these four are interactive and excluded from the simulation tournament, leaving these 16 plus a shared control condition (17 conditions in total)."

interventions <- read_csv(here("data/interventions.csv"), show_col_types = FALSE)

tag_order <- c(
  "Collaboration and peer-review",
  "Scientific methods and results",
  "Applications and impact",
  "Others' endorsement",
  "Values",
  "Other"
)

# Same table layout as the original preregistration: category headers inserted
# as bold full-width rows, with a running intervention number.
intervention_table_data <- interventions |>
  mutate(across(c(title, summary, tag), str_trim)) |>
  filter(tag != "LLM-chatbot", title != "Value similarity") |>
  mutate(tag = factor(tag, levels = tag_order)) |>
  arrange(tag, title) |>
  mutate(number = as.character(row_number())) |>
  group_by(tag) |>
  group_modify(~ bind_rows(
    tibble(title = as.character(.y$tag), summary = ""),
    .x
  )) |>
  ungroup() |>
  mutate(` ` = ifelse(title %in% levels(tag), "", number)) |>
  select(` `, title, summary) |>
  rename(`Intervention Title` = title, Summary = summary)

# identify header rows
header_rows <- which(intervention_table_data$` ` == "")

intervention_table_data |>
  tt(width = c(0.05, 0.25, 0.70)) |>
  format_tt(escape = TRUE) |>
  style_tt(align = "l", alignv = "t") |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = header_rows, bold = TRUE, j = 2, colspan = 2) |>
  style_tt(fontsize = 0.85)
```

The first goal of the tournament is *not* to crown a single champion team. Per the original preregistration, we evaluate approaches on a ladder of metrics of increasing strictness—from directional agreement to formal equivalence—plus distributional fidelity and stereotyping diagnostics. We expect different approaches to win on different rungs; the point is to map the design space, not to declare one winner. Second, and perhaps more importantly, we aim to provide a **catalogue**: a structured, comparable description of how research teams currently approach AI prediction of behavioral experiments—open- vs. closed-weight models, persona sources, administration protocols, fine-tuning, in-context learning—and which of these choices appear to matter.

## The human target sample

Predictions are scored against the human data from the megastudy itself. Participants are recruited from a national, non-probability opt-in panel of U.S. residents (CloudResearch), all aged 18 or older, who pass a series of attention and bot-detection checks before treatment assignment, so that the analyzed sample is relatively attentive to the intervention materials.

The benchmark targets **N ≈ 18,000 complete responses**: 1,000 per intervention across the 16 text-based interventions, plus 2,000 in the shared control condition. Recruitment uses **census-based cross quotas** on gender × age and gender × race/ethnicity to approximate the U.S. adult population; the targets below match the proportions of the parent megastudy, with counts simply rescaled to N ≈ 18,000. Participants selecting "Other" as their gender are not quota-constrained.

The targets below are drawn from the 2024 U.S. Census Bureau Population Estimates Program. *Total* is the share within each category; *Male* and *Female* give the within-category gender split.

```{r}
#| label: tbl-quotas
#| tbl-cap: "Sampling quota targets (N ≈ 18,000), from the 2024 U.S. Census Bureau Population Estimates Program. Proportions match the parent megastudy, with counts rescaled to N ≈ 18,000."

# Same layout as the interventions table: category headers inserted as bold
# full-width rows above their levels (tinytable, for a consistent look).
quotas <- read_csv(here("data/quotas_18000.csv"), show_col_types = FALSE)

quota_table_data <- quotas |>
  group_by(Variable) |>
  group_modify(~ bind_rows(
    tibble(Category = as.character(.y$Variable), Total = "", Male = "", Female = ""),
    .x
  )) |>
  ungroup() |>
  select(Category, Total, Male, Female) |>
  rename(` ` = Category)

header_rows_q <- which(quota_table_data$Total == "")

quota_table_data |>
  tt(width = c(0.40, 0.20, 0.20, 0.20)) |>
  format_tt(escape = TRUE) |>
  style_tt(align = "l", alignv = "t") |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = header_rows_q, bold = TRUE) |>
  style_tt(fontsize = 0.85)
```

Gender, age, and race are the only mandatory survey items (needed to fill the quotas) and so enter every analysis model as covariates with no missing data. The six demographic moderators used for the subgroup and calibration analyses—gender, age, race, education, income, and partisan identity—are documented in the [original preregistration](preregistration.qmd) and codebook; full sampling and exclusion detail is in the [human study preregistration](https://doi.org/10.5281/zenodo.20160212).

## Tournament design

Participation is by invitation. Each team builds an AI-based prediction pipeline—any models, personas, prompts, fine-tuning, retrieval, or prior literature—from the shipped materials (intervention texts, the full survey instrument, the codebook, the registration template, and the validator), but receives **no participant pool** (each team builds its own profiles) and **never any human outcome data**. Before the prediction lock the team registers its approach and deposits its predictions on Zenodo, hashed and time-stamped; once human data collection completes, every submission runs through the preregistered analysis pipeline unchanged and is reported on a public leaderboard alongside the human–human ceiling, and in a joint manuscript with all participating teams as co-authors. Three ground rules define a valid entry and matter for the analysis: **blinding is absolute** (no team accesses, solicits, or is shown any human outcome data—including pilots—before the lock; at registration time only ~10,000 of the planned ~18,000 responses had been collected and none released; a signed attestation is registration item G.3), pipelines are **AI-based and automated** (no human in the loop at prediction time), and **everything is registered though not necessarily public** (the disclosure policy below). The full procedure, ground rules, timeline, and incentives are in the [call for participation](index.html).

### Three ways to compete

We deliberately welcome very different approaches. Submissions enter at one of **three tiers**, defined by the granularity of the prediction. The common currency across tiers is the set of preregistered estimands—the same quantities the human analysis produces—so every approach is scored on the rungs it can reach. Teams choose the highest tier their approach supports; lower tiers are simply eligible for fewer analyses.

- **Tier 1 · Full — individual-level synthetic data.** A respondent-per-row dataset of behavioral clones: each synthetic participant completes the experiment in one condition. Flows through the preregistered analysis pipeline completely unchanged. Eligible for everything: ATE recovery, calibration regression, response distributions, subgroup heterogeneity, demographic baselines, and stereotyping diagnostics. *Grain: 1 row = 1 synthetic respondent.*
- **Tier 2 · Cell-level — group-level predictions.** A predicted mean and a **95% prediction interval** on that mean per cell—per condition × outcome, and per condition × moderator-level × outcome. For approaches that simulate or reason about groups rather than individuals. Eligible for ATEs, calibration, subgroup heterogeneity, and demographic baselines; not eligible for distribution-shape metrics (OVL, KS) or predictability regressions. *Grain: 1 row = 1 cell (condition × [moderator level ×] outcome).*
- **Tier 3 · Effect-level — direct effect predictions.** One predicted ATE per intervention × outcome, in original outcome units, with a **95% prediction interval** expressing the team's epistemic uncertainty. For zero-shot forecasting, in-context learning over the literature, meta-analytic priors, or hybrid pipelines. Eligible for directional agreement, Spearman ρ, Pearson r, RMSE, and calibration regression; inferential agreement and TOST are computed from the submitted intervals. *Grain: 1 row = 1 intervention × outcome estimate.*

### Analysis eligibility by tier

| Preregistered analysis | Tier 1 | Tier 2 | Tier 3 |
|---|:-:|:-:|:-:|
| ATE recovery — six-metric ladder (Section 1) | ✓ | ✓ | ✓ * |
| Calibration regression α / β (Section 1) | ✓ | ✓ | ✓ |
| Response distributions — variance ratio, OVL, KS (Section 1) | ✓ | — | — |
| Subgroup heterogeneity — condition × moderator (Section 2) | ✓ | ✓ | — |
| Within-subgroup distributions (Section 2) | ✓ | — | — |
| Demographic baseline calibration (Section 3) | ✓ | ✓ | — |
| Demographic predictability / stereotyping (Section 3) | ✓ | — | — |

\* For Tiers 2 and 3, inferential agreement and equivalence (TOST) are evaluated against the submitted 95% prediction intervals rather than model-based standard errors. For Tier 1, standard errors fall out of the fitted models, so no separate uncertainty input is required.

## Method registration and disclosure

Before the prediction lock, every team documents its pipeline on a structured **method-registration form**—built on the GUIDE-LLM reporting checklist [@feuerriegelReportingChecklistLarge2026] and extended where a reporting checklist alone does not guarantee, in principle, full replicability of a prediction pipeline—and deposits it alongside the predictions. The full item-by-item form and the **disclosure policy** (which lets proprietary teams seal items under escrow while keeping the lock publicly verifiable) are published in the [call for participation](index.html#registration-form); a fillable template ships in the [tournament repository](https://github.com/janpfander/silicon-sample-submission). Two aspects of the registration feed the analysis directly:

- **Approach characteristics** recorded on the form—open- vs. closed-weight model, persona source, administration protocol, fine-tuning, in-context learning, and so on—are the features joined to the leaderboard for the exploratory design-choice analysis (see *Cross-team comparison*).
- **Disclosure class** is a per-entry badge the policy assigns from the entry's most restricted item: **Class A · Open** (all items public), **Class B · Escrowed** (some items sealed but available to the core team and auditors under confidentiality, so compliance stays verifiable), and **Class C · Sealed** (one or more items withheld even from escrow). All three classes receive full leaderboard standing carrying their badge; only publicly disclosed features enter the design-choice analysis, and Class C entries are excluded from that analysis only—not from the leaderboard.

## Submission specification

Each tier defines the exact prediction file(s) a submission provides; the schemas below are the contract the scoring code consumes (`build_pairs_t1()` / `build_pairs_t2()` / `build_pairs_t3()`, see *Cross-team comparison*). The deposit logistics—Zenodo record structure, the `metadata.json` file, file naming, the self-check validator, and the public/escrowed split—are specified on the [materials page](materials.html) and in the [tournament repository](https://github.com/janpfander/silicon-sample-submission). Two requirements bear directly on the analysis: full **coverage** is all 16 interventions plus control across all 13 outcomes (declared subsets are permitted and scored pairwise, with coverage reported on the leaderboard), and **condition and column labels must match the published codebook exactly** (submissions are validated before the lock; cells that fail to parse are dropped).

### Tier 1 — individual-level dataset

One row per synthetic respondent, matching the structure of the cleaned analysis data (`cleaned_llm.rds` schema). Each team supplies its own synthetic respondents: `profile_id` is a team-assigned identifier, unique within the submission, and assignment of respondents to conditions (between-subjects, covering all 17 conditions) is the team's responsibility, registered under C.4/C.6.

| Column(s) | Type | Specification |
|---|---|------|
| `profile_id` | string | A unique respondent identifier you assign (unique within the submission). |
| `condition` | factor | Control or one of the 16 intervention labels, exactly as in the codebook. |
| `gender`, `age_band`, `race`, `education`, `income`, `party` | factor | The six moderators, levels per codebook (from your own profiles). |
| `trust_multidimensional` + 12 trust items | numeric | Primary composite 0–100; the 12 underlying items per codebook scale. |
| `trust_post`, `distrust_post`, `funding_perceptions`, `policy_role_mean`, `inst_trust_mean` | numeric | Secondary outcomes, original scales per codebook. |
| `belief_post`, `concern_mean`, `policy_general`, `policy_specific_mean`, `behavior_mean` | numeric | Tertiary outcomes, original scales per codebook. |
| `donation_ams` | numeric | Donation to AMS in $, per codebook. |
| `newsletter_signup` | 0 / 1 | Binary behavioral outcome. |

Because a Tier 1 submission has the same shape as the cleaned analysis data, it flows through the original analysis pipeline unchanged and is eligible for **all** analyses. The model functions in `R/functions/statistics.R` are agnostic to the specific condition labels and to the data source, so the same code path applies to the human data and to any team's submission.

### Tier 2 — cell-level file(s)

Two long-format files. `*_cells_main.csv`: one row per condition × outcome. `*_cells_moderator.csv`: one row per condition × moderator × moderator-level × outcome.

| Column | Type | Specification |
|---|---|------|
| `condition` | factor | Per codebook. |
| `moderator`, `moderator_level` | factor | Moderator file only; per codebook levels. |
| `outcome` | string | Outcome variable name per codebook. |
| `mean` | numeric | Predicted cell mean on the original scale (proportion for `newsletter_signup`). |
| `pi_lower`, `pi_upper` | numeric | 95% prediction interval bounds on the cell mean — the team's epistemic uncertainty about the *human* cell mean, not a within-simulation SD. The ATE interval is derived by combining the treatment- and control-cell intervals; used for inferential-agreement and equivalence scoring. |

### Tier 3 — effect-level file

One row per intervention × outcome (16 × 13 = 208 rows for full coverage). Estimates are ATEs versus control on the original outcome scale; the binary outcome is on the probability scale.

| Column | Type | Specification |
|---|---|------|
| `condition` | factor | Intervention label per codebook. |
| `outcome` | string | Outcome variable name per codebook. |
| `ate` | numeric | Predicted ATE, original units. |
| `pi_lower`, `pi_upper` | numeric | 95% prediction interval bounds — the team's epistemic uncertainty about the *human* ATE, not a within-simulation standard error. Used for inferential-agreement and equivalence scoring. |

## Analysis plan

### Per-team analysis (reused)

Scoring follows the preregistered analysis plan without modification. The human sample is split in half by a preregistered random seed: every submission is scored against **Human 1**, and the agreement between Human 1 and Human 2 defines the **human–human ceiling**—the upper bound any predictive method could reach—reported in the same row format as every team. We do not restate the original plan here; see the [original preregistration](preregistration.qmd) for the full specification of:

- **Section 1 — Average Treatment Effects:** ATE recovery across all outcomes (`run_main_treatment_model()` / `run_main_treatment_model_binary()`, compared with `compare_estimates()` and `run_calibration()`), and control-condition response distributions (`compare_distributions()`).
- **Section 2 — Subgroup Effects:** condition × moderator heterogeneity across the six moderators (`run_moderator_model()`), and within-subgroup distributions.
- **Section 3 — Demographic Baseline Calibration:** group-mean calibration (`compare_demographic_baselines()`) and demographic predictability / stereotyping (`compare_demographic_predictability()`).

Which of these analyses applies to a given submission follows from its tier (see *Analysis eligibility by tier*). One fixed **reference row** appears in every results table: the human–human ceiling (Human 1 vs. Human 2). Every submission is scored under identical rules and appears on the leaderboard like any other; no entry receives a special reference row beyond the ceiling.

### Tier-specific estimand recovery

The comparison-metric ladder requires, per intervention × outcome, an estimate, a standard error, and an inferential category. How these are obtained depends on the tier:

- **Tier 1:** the submission runs through `run_main_treatment_model()` exactly like the human data; estimates, HC2 robust standard errors, and BH-adjusted p-values fall out of the fitted models.
- **Tier 2:** the ATE for each intervention × outcome is the submitted cell mean minus the submitted control mean. Because only cell summaries are available, the preregistered OLS cannot be refit on a Tier 2 submission; instead we compute the ATE's standard error from the submitted cell prediction intervals. Each cell interval implies a standard error on that cell mean, $\mathrm{SE}_{cell} = (\mathrm{pi}_{upper} - \mathrm{pi}_{lower}) / (2 \times z_{0.975})$, and the treatment and control cells combine in the usual way for a between-subjects contrast, $\mathrm{SE} = \sqrt{\mathrm{SE}_t^2 + \mathrm{SE}_c^2}$. The test statistic $z = \mathrm{ATE} / \mathrm{SE}$ is referred to the standard normal distribution for a two-sided p-value, $p = 2\,\Phi(-|z|)$. These p-values are BH-adjusted within outcome, mirroring the Tier 1 models. Subgroup ATEs and demographic baselines are computed analogously from the moderator cell file.
- **Tier 3:** the submitted `ate` is the estimate. Inferential agreement uses the submitted 95% prediction interval: an effect counts as significantly positive (negative) if the interval lies entirely above (below) zero—the direct analog of a two-sided test at α = .05, so the significance criterion matches the one applied to Tiers 1–2. For the TOST equivalence test, the interval implies a standard error, $\mathrm{SE} = (\mathrm{pi}_{upper} - \mathrm{pi}_{lower}) / (2 \times z_{0.975})$.

### Cross-team comparison (new)

This is the new analytical contribution of the amendment. We assemble a **leaderboard**: each submission is scored with the same metric ladder used throughout the original preregistration—in increasing order of strictness: directional agreement, Spearman ρ, Pearson r, inferential agreement, RMSE, and TOST equivalence—so that approaches can be compared directly and against the human–human ceiling. Knowing on which rung an approach breaks is more informative than any single pass/fail verdict. On top of the ladder: the calibration regression (human ATEs on predicted ATEs) separates additive bias (α) from proportionality (β); distributional metrics (variance ratio, overlap coefficient, KS distance) test whether Tier 1 submissions reproduce response *shapes*, not just means; and the stereotyping diagnostics ask whether demographic variables predict synthetic responses with the same magnitude as in humans—or whether simulated respondents over-rely on demographic cues.

**Ranking principle: one score per metric, pooled across outcomes.** Each approach is scored on every rung of the ladder, with the intervention × outcome estimates **pooled across all outcomes** into a single value per metric. We deliberately do *not* collapse the rungs into one composite "champion" score: an approach can recover the *direction* of effects well while mis-scaling them, or nail the primary outcome while failing on the periphery, and a single number would hide exactly the structure the tournament exists to reveal. So the leaderboard is read **per metric**—each column (or panel) is its own ranking, and the interesting result is *which* approaches lead on *which* rung. Pooling across outcomes (rather than averaging per-outcome scores) keeps every intervention × outcome comparison as one data point and matches the pooled-row logic of the original preregistration; the one metric that cannot be pooled this way is RMSE (not comparable across outcome scales), which is therefore reported only in the per-outcome breakdowns. Per-outcome, distributional, subgroup, and demographic-calibration results follow for the tiers eligible for them, and every entry carries its disclosure-class badge (open / escrowed / sealed) so readers see at a glance how much of it is auditable.

**Presentation.** The headline view is a **forest-plot leaderboard** (one panel per metric): each approach is a point, ordered by performance, with the **human–human ceiling pinned at the top** and a dotted line dropped from its value so every approach is read against the ceiling; a grey dotted reference line marks the null where one is meaningful (zero for the correlations, chance agreement for directional agreement) and is omitted where it is not. Each point carries an **uncertainty interval** — a cluster bootstrap over interventions, introduced by this amendment (the original preregistration reports point estimates only) and detailed in the next subsection. The same forest-plot template is reused for the subgroup-heterogeneity and distributional metrics, so Sections 1–3 share one visual language. The precise numbers accompany every figure as a companion table.

#### Uncertainty intervals on the leaderboard

**Why the leaderboard carries intervals at all.** Each point on the forest plot is a single pooled score — one number per approach per metric. A bare point invites over-reading: if one approach scores Pearson *r* = 0.60 and another 0.55, is the first genuinely better, or is the gap an artifact of the particular interventions that happened to be tested? The interval answers exactly that question. It is the guard against ranking approaches on differences that would not survive a different, equally reasonable test.

**What the interval is, mechanically.** The pooled metric is computed once on the full data to give the point estimate. We then form a *cluster bootstrap over interventions*: the 16 interventions are resampled with replacement (some drawn twice, some dropped), the pooled metric is recomputed on each resampled set, and the 2.5th–97.5th percentiles of those recomputed scores give the 95% interval (`cluster_boot()` in `R/functions/statistics.R`). Clustering is *by intervention* — each intervention's full block of outcome rows is drawn or dropped together — because the outcome rows belonging to one intervention are not independent. Crucially, **no submitted prediction is ever altered**: each approach's predicted effect for each intervention stays fixed at the value the team submitted. What is resampled is *which interventions enter the score*, not how well anyone did on them. The interval therefore measures the spread of the recomputed *score*, not the spread of the raw effects: averaging or correlating cancels much of the effect-to-effect spread, which is why the bar is typically far narrower than the range of the underlying estimates.

**What the interval means, and the one assumption it rests on.** The interval treats the 16 interventions as a *test set* standing in for the broader space of interventions the benchmark cares about, exactly as a held-out test set stands in for future cases in predictive modelling. Read this way, the bar answers a generalisation question: *would this approach's score, and its rank, survive on a different set of interventions of the same kind?* This is the appropriate question for a benchmark, whose purpose is to identify general predictive skill rather than performance on sixteen specific stimuli. The assumption is therefore about the *interventions*, not the predictions: that the tested set is a sample from a population of possible interventions. Resampling does **not** assume that a team that predicted a given intervention well "could have" gotten it wrong — the predictions are fixed; it assumes only that a replication might have tested a different set of interventions. (If one instead regarded these exact sixteen as the entire universe of interest, there would be no sampling and hence no uncertainty to display — the scores would simply be exact. We do not adopt that reading: the tournament exists to compare *approaches*, not to characterise these particular stimuli.)

**Interval width is informative, not a penalty on accuracy.** Because the metrics are bounded agreement scores with a ceiling at perfect agreement, an approach that tracks the human effects closely sits on the diagonal and stays there under resampling — dropping or duplicating interventions barely moves a score already pinned near the ceiling — so accurate approaches carry *narrow* intervals. An approach whose predictions do not track the human effects has a score floating in the interior, free to swing with the resampled set, and so carries a *wide* interval. Interval width is thus entangled with accuracy, but in the helpful direction: a wide bar is the leaderboard signalling that a score is flimsy and may reflect test-set luck rather than skill — precisely the warning one wants attached to a weak entry. (Spread in the size of the true effects does not by itself widen the bar; the correlation and agreement metrics are scale-free, so what matters is whether predictions hug the line, not how far apart the interventions are.)

**Relation to the per-effect confidence and prediction intervals.** The bootstrap interval is a *different layer* of uncertainty from the confidence intervals (Tier 1) and prediction intervals (Tiers 2–3) attached to individual treatment effects, and the two should not be conflated. A per-effect CI or PI is *within-intervention*: it says how precisely a single effect is known — finite-respondent sampling noise for a Tier 1 model fit, or the team's own stated uncertainty for a Tier 2/3 submission. The bootstrap interval is *across-intervention*: it says how much the pooled score depends on which interventions were tested. The two are **nested, not competing**. The per-effect intervals enter the *value* of each row: for the correlation and directional rungs only the point estimates are used, so the per-effect intervals do not affect those scores at all and the bootstrap is their sole source of an error bar; for the inferential-agreement and TOST rungs the per-effect intervals set each row's significance or equivalence label, which the bootstrap then pools and resamples. In all cases the bootstrap holds each intervention's estimate fixed at its point value: it propagates the *which-interventions* uncertainty into the bars but **not** the within-effect sampling noise.

**A noted alternative.** Folding both layers into a single bar would require a *two-level* bootstrap — in each replicate, resample the interventions *and* perturb each drawn effect by its standard error (drawing the effect from a normal centred on the estimate with that SE). We adopt the single-level (point-estimate) cluster bootstrap as the default because, with only sixteen interventions, the *which-interventions* term dominates the uncertainty in any ranking, while the human study is powered such that each effect's standard error is comparatively small relative to the treatment-to-treatment spread. The per-effect CIs and PIs remain reported in the per-outcome breakdowns as the finer, within-effect layer.

**Scope across panels.** The ATE leaderboard scores interventions, so its bootstrap resamples interventions and is available to **every tier** — the resampling needs only one predicted effect per intervention, which all tiers supply (including effect-only Tier 3). The subgroup and distributional leaderboards are restricted to Tier 1, but for a different reason: their *metrics* (interaction models, response-shape comparisons) require respondent-level data, not the bootstrap. The distributional metrics are computed in the control condition only and have no intervention contrast, so that panel necessarily resamples outcomes rather than interventions; with a handful of outcomes those intervals are correspondingly coarser and are read with that caveat.

Registered approach characteristics (open- vs. closed-weight, persona source, administration protocol, fine-tuned vs. not, …) are joined to the leaderboard for a descriptive, hypothesis-generating analysis of which design choices accompany better predictions—preregistered as exploratory, and restricted to publicly disclosed features (Class C entries are excluded from this analysis, not from the leaderboard).

::: {.callout-note}
**What this amendment changes relative to the original analysis.** The estimands, models, and comparison metrics are taken from the original preregistration unchanged. This amendment adds three presentation-and-aggregation elements on top, none of which alters the original pipeline: (i) it names *pooling across outcomes, scored per metric* as the headline ranking principle (the original computes the same pooled metrics but does not frame them as a ranking); (ii) it attaches a **cluster-bootstrap uncertainty interval** to each pooled metric (the original reports point estimates only), via the new `cluster_boot()` / `pooled_metrics()` / `signed_metrics()` helpers in `R/functions/statistics.R`; and (iii) it introduces the **forest-plot leaderboard** (`plot_leaderboard_forest()` in `R/functions/plots.R`) as the shared display for ATE, subgroup, and distributional metrics. All additions are new functions; the original preregistration's functions and outputs are untouched.
:::

::: {.callout-note}
**Exploratory by design.** With a modest number of teams, the relationship between approach characteristics and predictive performance is reported **descriptively** (rankings, group comparisons, visual patterns). We do not preregister inferential tests of feature effects; the cross-team feature analysis is hypothesis-generating, intended to surface which design choices appear to matter for future, larger benchmarks.
:::

#### Scoring functions

The leaderboard reuses the original preregistration's comparison functions and adds four helpers, all in `R/functions/`. As in the original, each is shown in full below and explained before it.

`pooled_metrics()` computes the ATE-recovery ladder on intervention × outcome pairs **pooled across all outcomes**: directional agreement, Spearman ρ, Pearson r, inferential agreement, and TOST equivalence. It mirrors the original `compare_estimates()` but operates on the pooled set of pairs rather than a single outcome; pooled RMSE is deliberately omitted, as it is not comparable across outcome scales. It expects pre-joined pairs with the reference (Human 1) columns `estimate_h`, `se_h`, `p_adj_h` and the submission columns `estimate_l`, `se_l`, `infer_l`.

```{r, results='asis'}
#| label: fn-pooled-metrics
print_a_function_from_file("pooled_metrics")
```

`signed_metrics()` is the reduced ladder for the subgroup leaderboard, where only the signed-effect rungs are pooled (directional agreement, Spearman ρ, Pearson r). This matches the original preregistration's subgroup section, which reports these three on estimate-only pairs (no standard error or inferential category).

```{r, results='asis'}
#| label: fn-signed-metrics
print_a_function_from_file("signed_metrics")
```

`cluster_boot()` attaches the uncertainty interval described above. It resamples the levels of a clustering variable (interventions for the ATE and subgroup panels, outcomes for the distributional panel) with replacement, recomputes the supplied metric function on each replicate, and returns the point estimate together with percentile confidence bounds in long form (one row per metric, with `value`, `lo`, `hi`). It is generic: the metric function is passed in, so the same bootstrap serves every panel.

```{r, results='asis'}
#| label: fn-cluster-boot
print_a_function_from_file("cluster_boot")
```

`plot_leaderboard_forest()` renders the shared forest-plot display: one panel per metric, approaches ranked by a chosen metric with the human–human ceiling pinned on top and a dotted line dropped from its value, optional null/ideal reference lines, and the cluster-bootstrap intervals drawn as horizontal bars.

```{r, results='asis'}
#| label: fn-plot-leaderboard-forest
print_a_function_from_file("plot_leaderboard_forest", file = here::here("R/functions/plots.R"))
```

The remainder of this section demonstrates the leaderboard code on placeholder data: five mock "teams", each a stratified bootstrap resample (within condition) of the single individual-level placeholder dataset, stacked with a `team` label. To exercise all three tier code paths, three teams enter at Tier 1 (individual-level), one is collapsed into a Tier 2 cell-level file, and one into a Tier 3 effect-level file. All of this placeholder data—including the resampled teams and the collapsed Tier 2/3 files—is built in the simulation script (`R/simulate_data_preregistration.R`), not here; this document only reads it. The code will be applied to real submissions without modification.

```{r}
#| label: variables

outcomes_primary   <- "trust_multidimensional"
outcomes_secondary <- c("trust_post", "distrust_post",
                         "funding_perceptions", "policy_role_mean", "inst_trust_mean")
outcomes_tertiary  <- c("belief_post", "concern_mean", "policy_general",
                        "policy_specific_mean", "behavior_mean")
outcomes_continuous <- c(outcomes_primary, outcomes_secondary, outcomes_tertiary)

# Illustrative outcome subset and categorical moderators — same definitions as
# the original preregistration; used for the subgroup-effects leaderboard.
outcomes_illustrative <- c("trust_multidimensional", "donation_ams",
                           "funding_perceptions", "policy_general")
moderators_cat <- c("gender", "age_band", "race", "education", "income", "party")

outcome_label_map <- c(
  trust_multidimensional = "Multidimensional trust (primary)",
  trust_post             = "Overall trust",
  distrust_post          = "Distrust",
  funding_perceptions    = "Funding perceptions",
  policy_role_mean       = "Policy role",
  inst_trust_mean        = "Institutional trust",
  belief_post            = "Climate belief",
  concern_mean           = "Climate concern",
  policy_general         = "General climate policy",
  policy_specific_mean   = "Specific climate policy",
  behavior_mean          = "Pro-climate behavior"
)
```

```{r}
#| label: leaderboard-helpers

# Extract ATE pairs (estimate, SE, adjusted p) — same helper as the original
# preregistration's rq1-all chunk.
prep_ate <- function(df) {
  df |>
    select(condition, estimate) |>
    mutate(
      se    = if ("std.error"        %in% names(df)) df$std.error        else NA_real_,
      p_adj = if ("p.value_adjusted" %in% names(df)) df$p.value_adjusted else NA_real_
    )
}

# Inferential category from estimate sign + adjusted p-value (Tiers 1-2)
classify_sig <- function(estimate, p_adj) {
  case_when(
    !is.na(p_adj) & p_adj < .05 & estimate > 0 ~ "pos",
    !is.na(p_adj) & p_adj < .05 & estimate < 0 ~ "neg",
    TRUE                                       ~ "ns"
  )
}

# Reference (Human 1) ATE models, fit once and reused for every comparison.
human_models <- map(outcomes_continuous,
                    ~ run_main_treatment_model(human_data_1, outcome = .x)) |>
  set_names(outcomes_continuous)

human_side <- map_dfr(outcomes_continuous, function(out) {
  prep_ate(human_models[[out]]) |>
    rename(estimate_h = estimate, se_h = se, p_adj_h = p_adj) |>
    mutate(outcome = out)
})

control_lab <- levels(human_data$condition)[1]
z95 <- qnorm(0.975)

# Tier 1: pool ATE pairs from an individual-level dataset by refitting the
# preregistered treatment models on the submission.
build_pairs_t1 <- function(pred_data) {
  map_dfr(outcomes_continuous, function(out) {
    prep_ate(run_main_treatment_model(pred_data, outcome = out)) |>
      mutate(outcome = out)
  }) |>
    mutate(infer_l = classify_sig(estimate, p_adj)) |>
    rename(estimate_l = estimate, se_l = se) |>
    select(-p_adj) |>
    inner_join(human_side, by = c("condition", "outcome"))
}

# Tier 2: ATEs and SEs from submitted cell means and their 95% prediction
# intervals (mean, pi_lower, pi_upper per condition x outcome). Each cell
# interval implies an SE on the cell mean; the treatment and control cells
# combine for the between-subjects contrast. Normal-approximation p-values,
# BH-adjusted within outcome to mirror the Tier 1 models.
build_pairs_t2 <- function(cells_main) {
  cells_main |>
    mutate(
      condition = as.character(condition),
      se_cell   = (pi_upper - pi_lower) / (2 * z95)
    ) |>
    group_by(outcome) |>
    group_modify(function(cells, key) {
      ctrl <- cells |> filter(condition == control_lab)
      cells |>
        filter(condition != control_lab) |>
        mutate(
          estimate_l = mean - ctrl$mean,
          se_l       = sqrt(se_cell^2 + ctrl$se_cell^2),
          p_adj      = p.adjust(2 * pnorm(-abs(estimate_l / se_l)), method = "BH"),
          infer_l    = classify_sig(estimate_l, p_adj)
        ) |>
        select(condition, estimate_l, se_l, infer_l)
    }) |>
    ungroup() |>
    inner_join(human_side, by = c("condition", "outcome"))
}

# Tier 3: submitted ATEs taken as-is; inference from the submitted 95%
# prediction interval (significant iff the interval excludes zero, the direct
# analog of a two-sided test at alpha = .05); implied SE from the interval
# width for the TOST computation.
build_pairs_t3 <- function(effects) {
  effects |>
    mutate(
      condition  = as.character(condition),
      estimate_l = ate,
      se_l       = (pi_upper - pi_lower) / (2 * z95),
      infer_l    = case_when(
        pi_lower > 0 ~ "pos",
        pi_upper < 0 ~ "neg",
        TRUE         ~ "ns"
      )
    ) |>
    inner_join(human_side, by = c("condition", "outcome")) |>
    select(condition, outcome, estimate_l, se_l, infer_l,
           estimate_h, se_h, p_adj_h)
}

# `pooled_metrics()`, `signed_metrics()`, and `cluster_boot()` are defined in
# R/functions/statistics.R and shown under "Scoring functions" above.

# One leaderboard row: pooled ladder + calibration slope and R^2 on the
# primary outcome. Works identically for all tiers since the calibration
# regression only needs (condition, estimate) pairs.
leaderboard_row <- function(pairs, submission, tier, disclosure) {
  calib <- run_calibration(
    human_models[[outcomes_primary]],
    pairs |> filter(outcome == outcomes_primary) |>
      select(condition, estimate = estimate_l)
  )
  pooled_metrics(pairs) |>
    mutate(submission = submission, tier = tier, disclosure = disclosure,
           beta = calib$beta, calib_r2 = calib$r_squared)
}
```

```{r}
#| label: mock-submissions

# Mock Tier 2 (cell-level) and Tier 3 (effect-level) submissions, generated
# outside this script alongside the other placeholder data (see
# R/simulate_data_preregistration.R) so that no mock-data simulation lives in
# the preregistration itself. Tier 2 collapses one placeholder team to the
# cell-level schema (*_cells_main.csv); Tier 3 gives ATEs vs. control with a
# widened 95% prediction interval standing in for a team's epistemic uncertainty.
cells_t2   <- readRDS(here("data/simulation_preregistration/mock_submission_tier2_cells.rds"))
effects_t3 <- readRDS(here("data/simulation_preregistration/mock_submission_tier3_effects.rds"))
```

```{r}
#| label: leaderboard-build

# Three mock Tier 1 teams (mock disclosure classes shown to demonstrate the
# badge column), one Tier 2 team, one Tier 3 team, plus the fixed human-human
# ceiling (Human 1 vs. Human 2). Each submission's ATE pairs are built once
# here and reused for both the leaderboard table and the bootstrap forest plot.
pairs_list <- list(
  "Team 1"               = build_pairs_t1(llm_data_teams |> filter(team == "team_1")),
  "Team 2"               = build_pairs_t1(llm_data_teams |> filter(team == "team_2")),
  "Team 3"               = build_pairs_t1(llm_data_teams |> filter(team == "team_3")),
  "Team 4"               = build_pairs_t2(cells_t2),
  "Team 5"               = build_pairs_t3(effects_t3),
  "Human–human ceiling"  = build_pairs_t1(human_data_2)
)

submission_meta <- tribble(
  ~submission,            ~tier, ~disclosure,
  "Team 1",               "1",   "A",
  "Team 2",               "1",   "B",
  "Team 3",               "1",   "C",
  "Team 4",               "2",   "A",
  "Team 5",               "3",   "A",
  "Human–human ceiling",  "—",   "—"
)

leaderboard <- imap_dfr(pairs_list, function(pr, sub) {
  m <- submission_meta |> filter(submission == sub)
  leaderboard_row(pr, sub, tier = m$tier, disclosure = m$disclosure)
}) |>
  arrange(desc(submission == "Human–human ceiling"), desc(pearson_r))
```

```{r}
#| label: tbl-leaderboard
#| tbl-cap: "Cross-team leaderboard (placeholder data). Pooled ATE-recovery metrics across all continuous outcomes, plus calibration slope and R² on the primary outcome, scored against the Human 1 reference half. The Human–human ceiling (Human 1 vs. Human 2) is the upper bound for any approach. Metrics run least-to-most strict, left to right; Class is the disclosure-class badge (A open / B escrowed / C sealed)."

leaderboard |>
  select(submission, tier, disclosure, directional_pct, spearman_rho, pearson_r,
         inferential_pct, tost_pct, beta, calib_r2) |>
  gt() |>
  fmt_number(columns = c(spearman_rho, pearson_r, beta, calib_r2), decimals = 2) |>
  fmt_number(columns = c(directional_pct, inferential_pct, tost_pct), decimals = 0) |>
  cols_label(
    submission      = "Submission",
    tier            = "Tier",
    disclosure      = "Class",
    directional_pct = "Directional %",
    spearman_rho    = "Spearman ρ",
    pearson_r       = "Pearson r",
    inferential_pct = "Inferential %",
    tost_pct        = "TOST %",
    beta            = "Calib. β",
    calib_r2        = "Calib. R²"
  ) |>
  tab_style(
    style     = cell_fill(color = "#eef3f7"),
    locations = cells_body(rows = submission == "Human–human ceiling")
  )
```

```{r}
#| label: ate-ladder-labels

# Display labels / panel order for the metric ladder, and the null reference
# line per metric (50 = chance for directional agreement, 0 for the
# correlations; omitted for the percentage-agreement metrics where no clean
# null exists).
ladder_labels <- c(
  directional_pct = "Directional %",
  spearman_rho    = "Spearman ρ",
  pearson_r       = "Pearson r",
  inferential_pct = "Inferential %",
  tost_pct        = "TOST %"
)
ladder_nulls <- c("Directional %" = 50, "Spearman ρ" = 0, "Pearson r" = 0)

relabel_metrics <- function(df, labels) {
  df |> mutate(metric = factor(labels[metric], levels = labels))
}
```

```{r}
#| label: fig-leaderboard
#| fig-width: 11
#| fig-height: 3.6
#| fig-cap: "Forest-plot leaderboard for ATE recovery (placeholder data): one panel per evaluation metric, least-to-most strict, each approach scored on ATE pairs pooled across all continuous outcomes against Human 1. Points are ranked by Pearson r; the human–human ceiling (Human 1 vs. Human 2) is pinned on top with a dotted reference line dropped from its value. Grey dotted lines mark the null where meaningful (chance = 50% for directional agreement; 0 for the correlations). Bars are 95% cluster-bootstrap intervals over the 16 interventions. On placeholder data every submission is random noise, so all approaches sit near the null and overlap heavily."

ate_forest <- imap_dfr(pairs_list, function(pr, sub) {
  cluster_boot(pr, pooled_metrics, cluster = "condition", n_boot = 500) |>
    mutate(submission = sub)
}) |>
  relabel_metrics(ladder_labels)

plot_leaderboard_forest(ate_forest, sort_metric = "Pearson r",
                        null_lines = ladder_nulls)
```

#### Subgroup effects

The same template scores **subgroup recovery**: whether an approach reproduces *who* an intervention moves, not just its average effect. For each contribution we fit the preregistered condition × moderator interaction model (`run_moderator_model()`) across the illustrative outcomes and the six demographic moderators, then pool the interaction estimates against Human 1. Because interaction estimates require respondent-level data, this view covers Tier 1 (and Tier 2 from the moderator cell file); Tier 3 effect-only submissions are not eligible. As in the original preregistration's subgroup section, only the signed-effect rungs are pooled here (directional agreement, Spearman ρ, Pearson r); equivalence and inferential agreement across mixed moderator-level scales are left to the per-outcome breakdowns.

```{r}
#| label: subgroup-forest-data

# Human-1 interaction estimates, computed once and reused for every submission.
human_mod_side <- map_dfr(outcomes_illustrative, function(out) {
  map_dfr(moderators_cat, function(mod) {
    run_moderator_model(human_data_1, outcome = out, moderator = mod)$interaction_effects |>
      select(condition, moderator_level, estimate_h = estimate) |>
      mutate(outcome = out, moderator = mod)
  })
})

build_subgroup_pairs <- function(pred_data) {
  pred_side <- map_dfr(outcomes_illustrative, function(out) {
    map_dfr(moderators_cat, function(mod) {
      run_moderator_model(pred_data, outcome = out, moderator = mod)$interaction_effects |>
        select(condition, moderator_level, estimate_l = estimate) |>
        mutate(outcome = out, moderator = mod)
    })
  })
  inner_join(human_mod_side, pred_side,
             by = c("condition", "moderator_level", "outcome", "moderator"))
}

# Tier-1 submissions plus the ceiling are eligible for the subgroup view.
subgroup_inputs <- list(
  "Team 1"              = llm_data_teams |> filter(team == "team_1"),
  "Team 2"              = llm_data_teams |> filter(team == "team_2"),
  "Team 3"              = llm_data_teams |> filter(team == "team_3"),
  "Human–human ceiling" = human_data_2
)

subgroup_forest <- imap_dfr(subgroup_inputs, function(dat, sub) {
  cluster_boot(build_subgroup_pairs(dat), signed_metrics,
               cluster = "condition", n_boot = 500) |>
    mutate(submission = sub)
}) |>
  relabel_metrics(ladder_labels[c("directional_pct", "spearman_rho", "pearson_r")])
```

```{r}
#| label: fig-subgroup-leaderboard
#| fig-width: 8
#| fig-height: 3.2
#| fig-cap: "Forest-plot leaderboard for subgroup recovery (placeholder data, Tier 1): agreement between each approach and Human 1 on condition × moderator interaction estimates, pooled across the illustrative outcomes and the six demographic moderators. Layout as in the ATE figure — ceiling pinned on top, null lines where meaningful, 95% cluster-bootstrap intervals over interventions."

plot_leaderboard_forest(subgroup_forest, sort_metric = "Pearson r",
                        null_lines = ladder_nulls)
```

#### Response distributions

Finally, the template ranks how well an approach reproduces the **shape** of the human response distribution, not only its mean. Using the preregistered `compare_distributions()` in the shared control condition, we score the overlap coefficient (OVL, 1 = identical densities), the Kolmogorov–Smirnov distance (KS, 0 = indistinguishable), and the variance ratio (clone/human, 1 = matched spread), each pooled across the continuous outcomes. These shape metrics require respondent-level data, so the distributional view is Tier 1 only. Here the reference is not a null at zero but the **ideal value** of each metric (OVL = 1, KS = 0, variance ratio = 1), drawn as the grey dotted line; the human–human ceiling shows how close two independent human halves come to that ideal.

```{r}
#| label: distribution-forest-data

dist_per_outcome <- function(pred_data) {
  map_dfr(outcomes_continuous, function(out) {
    compare_distributions(human_data_1, pred_data, out, control_lab) |>
      mutate(outcome = out)
  })
}
dist_metrics <- function(df) {
  df |> summarise(across(c(ovl, ks_d, variance_ratio), ~ mean(.x, na.rm = TRUE)))
}

dist_labels <- c(ovl = "Overlap (OVL)", ks_d = "KS distance",
                 variance_ratio = "Variance ratio")
dist_ideals <- c("Overlap (OVL)" = 1, "KS distance" = 0, "Variance ratio" = 1)

distribution_forest <- imap_dfr(subgroup_inputs, function(dat, sub) {
  cluster_boot(dist_per_outcome(dat), dist_metrics,
               cluster = "outcome", n_boot = 500) |>
    mutate(submission = sub)
}) |>
  relabel_metrics(dist_labels)
```

```{r}
#| label: fig-distribution-leaderboard
#| fig-width: 8
#| fig-height: 3.2
#| fig-cap: "Forest-plot leaderboard for response-distribution match (placeholder data, Tier 1): overlap (OVL), KS distance, and variance ratio between each approach and Human 1 in the control condition, pooled across continuous outcomes. Grey dotted lines mark the ideal value of each metric (OVL = 1, KS = 0, variance ratio = 1) rather than a null; the ceiling is the human–human comparison. Bars are 95% cluster-bootstrap intervals over outcomes."

plot_leaderboard_forest(distribution_forest, sort_metric = "Overlap (OVL)",
                        null_lines = dist_ideals)
```

When real submissions arrive, these leaderboards are recomputed by replacing the placeholder datasets with the submitted prediction files routed through their tier's builder (`build_pairs_t1()` / `build_pairs_t2()` / `build_pairs_t3()`); the subgroup and distribution figures are produced for the tiers eligible for them, alongside the demographic-baseline and stereotyping diagnostics of Section 3. The approach-characteristics table coded from the registration forms is then joined to the leaderboard for the exploratory analysis of which design choices accompany better predictions (publicly disclosed features only; Class C entries excluded from this analysis, not from the leaderboard).

