---
title: "Predicting the effects of behavioral interventions on humans with LLMs"

abstract: |
    We test the potential of predicting human behavior in survey experiments with behavioral clones—large language model (LLM) generated artificial agents mimicking human participants. To evaluate this potential, we predict the outcomes of a large-scale field experiment (N = ~22,000) to be conducted in the United States. The experiment will test 20 text-based interventions designed to increase trust in climate scientists, assessing a broad range of secondary, outcomes including self-reported attitudes (e.g., support for funding of climate science) and behavioral measures (e.g., donations to a scientific organization). Before the human study is fielded, we will simulate it with behavioral clones using demographically representative U.S. participant profiles. We will lock the simulated dataset and analysis code. Once human data are available, we will compare clone-based predictions to human results using a set of preregistered metrics spanning (i) inferential agreement and estimate convergence on treatment effects, (ii) outcome distributions, and (iii) subgroup heterogeneity. Our study will provide a large-scale, credible benchmark test of LLMs' potential---and limitations---as tools for simulating human subjects in social science research. All analyses in this preregistration are demonstrated on simulated placeholder data with identical structure to the real datasets; no human outcome data have been observed at the time of registration.

# bibliography: ../references.bib
#csl: https://www.zotero.org/styles/apa
date: "2026-05-04"

format:
  html:
    code-fold: true
    toc: true
    toc-depth: 4
  pdf:
    toc: true
    toc-depth: 5
    include-before-body:
      text: |
        \clearpage
    colorlinks: true
    tbl-colwidths: true
    header-includes:
      - \usepackage{float}
      - \floatplacement{figure}{H}
      - \floatplacement{table}{H}

execute:
  echo: true
  warning: false
  message: false
  fig-width: 8
  fig-height: 5
  fig-dpi: 300
---

```{r}
#| label: setup
#| output: false

# packages
library(tidyverse)
library(sandwich)
library(lmtest)
library(marginaleffects)
library(broom)
library(MetBrewer)
library(patchwork)
library(gt)
library(here)

# custom functions
source(here("R/functions/statistics.R"))
source(here("R/functions/plots.R"))
source(here("R/functions/tables.R"))
source(here("R/functions/reporting.R"))

# simulated (random) data
human_data <- readRDS(here("data/simulation_preregistration/human_data_preregistration.rds"))
llm_data   <- readRDS(here("data/simulation_preregistration/llm_data_preregistration.rds"))

# actual data later
# human_data  <- readRDS("data/cleaned_llm.rds")
# llm_data <- readRDS("data/cleaned_llm.rds")
```

::: {.callout-important}
**Placeholder data.** Both datasets used throughout are simulated random noise with the correct structure. The analysis code will be applied to real human and LLM clone data without modification once both are available.
:::

## Research Questions

The main outcome of the megastudy is trust in climate scientists. However, there are several secondary and tertiary outcomes. For each of these outcomes we have the following research questions:

**RQ1 — ATE recovery.** How well do LLM clones predict the average treatment effects (ATEs) across 20 interventions?

**RQ2 — Distribution shape recovery.** How well do LLM clones mirror human response distributions (within experimental conditions)?

**RQ3 — Subgroup heterogeneity recovery.** How well do LLM clones recover differences in intervention effects across different demographic groups?

**RQ4 — Individual-level demographic calibration.** Are LLM clone responses calibrated to the demographic profile of each simulated participant? We examine (i) whether clones reproduce the correct baseline response level for each demographic group, (ii) whether demographic variables predict clone responses with the same magnitude as in humans (stereotyping test), and (iii) whether within-subgroup response distributions match between humans and clones.

**Secondary — Stacked-model discrepancy test.** As a complement to the TOST-based equivalence tests in RQ1, we pool human and clone data and estimate a condition × source interaction, providing a direct inferential test of whether each clone ATE significantly departs from its human counterpart.

## Measures

```{r}
#| label: variables

outcomes_primary <- "trust_multidimensional"

outcomes_secondary <- c(
  "trust_post", "distrust_post", 
  #"donation_ams",
  "funding_perceptions", "policy_role_mean", "inst_trust_mean"
)

outcomes_binary <- "newsletter_signup"

outcomes_tertiary <- c(
  "belief_post", "concern_mean", "policy_general",
  "policy_specific_mean", "behavior_mean"
)

outcomes_continuous <- c(outcomes_primary, outcomes_secondary, outcomes_tertiary)

trust_dimensions <- c(
  "trust_competence", "trust_integrity", "trust_benevolence", "trust_openness"
)

trust_items <- c(
  paste0("trust_competence_",  1:3),
  paste0("trust_integrity_",   1:3),
  paste0("trust_benevolence_", 1:3),
  paste0("trust_openness_",    1:3)
)

moderators_cat <- c("gender", "age_band", "race", "education", "income", "party")

outcome_label_map <- c(
  trust_multidimensional = "Multidimensional trust (primary)",
  trust_post             = "Overall trust",
  distrust_post          = "Distrust",
  donation_ams           = "Donation to AMS ($)",
  newsletter_signup      = "Newsletter sign-up",
  funding_perceptions    = "Funding perceptions",
  policy_role_mean       = "Policy role",
  inst_trust_mean        = "Institutional trust",
  belief_post            = "Climate belief",
  concern_mean           = "Climate concern",
  policy_general         = "General climate policy",
  policy_specific_mean   = "Specific climate policy",
  behavior_mean          = "Pro-climate behavior"
)
```

## Comparison Functions

Two helper functions operationalize the comparison metrics. They are applied symmetrically to human and LLM clone model outputs throughout the document.

### Comparison metrics

`compare_estimates()` is used throughout — for ATEs (RQ1), moderator interaction estimates (RQ3), and secondary analyses. It takes two tidy model result tibbles and a `join_by` argument specifying the key columns, and computes six metrics:

- **Spearman ρ** — rank correlation of point estimates. Answers: do clones rank the 20 interventions in the same order as humans? A value of 1 means perfect rank agreement; 0 means no correspondence.
- **Pearson r** — linear correlation of point estimates. Sensitive to outliers but captures linear proportionality.
- **RMSE** — root mean squared error. Captures average magnitude error in the same units as the outcome.
- **Directional agreement** — % of estimates with the same sign. Answers: do clones at least get the direction of each effect right?
- **Inferential agreement** — % of estimates with the same BH-adjusted conclusion (significantly positive, significantly negative, or non-significant at α = .05). Answers: would a researcher using clones reach the same inferential conclusion as one using human data?
- **TOST (%)** — % of individual estimate pairs that pass an equivalence test with preregistered bound Δ_k = 0.5 × |estimate_human_k|. Equivalence is declared (two one-sided tests, α = .05) when the clone estimate falls within ±50% of the human estimate. Note: when human estimates are near zero, Δ_k shrinks accordingly, making equivalence harder to achieve.

```{r}
#| label: fn-compare-estimates

compare_estimates <- function(h_result, l_result,
                              join_by = "condition",
                              tost_delta_pct = 0.5) {

  # Standardize columns; SE derived from std.error or from 95% CI width if absent
  prep <- function(df, suffix) {
    df |>
      select(all_of(join_by), estimate) |>
      mutate(
        se    = if ("std.error" %in% names(df)) df$std.error
                else if (all(c("conf.low", "conf.high") %in% names(df)))
                  (df$conf.high - df$conf.low) / (2 * 1.96)
                else NA_real_,
        p_adj = if ("p.value_adjusted" %in% names(df)) df$p.value_adjusted
                else NA_real_
      ) |>
      rename_with(~ paste0(.x, "_", suffix), c(estimate, se, p_adj))
  }

  joined <- inner_join(prep(h_result, "h"), prep(l_result, "l"), by = join_by)

  # Three-level inferential category based on BH-adjusted p and direction
  infer_cat <- function(est, p_adj) {
    case_when(
      p_adj < 0.05 & est > 0 ~ "sig_positive",
      p_adj < 0.05 & est < 0 ~ "sig_negative",
      TRUE                   ~ "not_significant"
    )
  }

  # TOST: equivalence declared when |ATE_h - ATE_l| < delta = tost_delta_pct × |ATE_h|
  # p_tost is the maximum of the two one-sided p-values; equivalent if p_tost < 0.05
  joined |>
    mutate(
      delta      = tost_delta_pct * abs(estimate_h),
      diff       = estimate_h - estimate_l,
      se_diff    = sqrt(se_h^2 + se_l^2),
      p_lower    = pnorm((diff + delta) / se_diff, lower.tail = FALSE),
      p_upper    = pnorm((diff - delta) / se_diff),
      p_tost     = pmax(p_lower, p_upper),
      equivalent = p_tost < 0.05,
      same_infer = infer_cat(estimate_h, p_adj_h) == infer_cat(estimate_l, p_adj_l)
    ) |>
    summarise(
      spearman_rho    = cor(estimate_h, estimate_l, method = "spearman",
                            use = "pairwise.complete.obs"),
      pearson_r       = cor(estimate_h, estimate_l, use = "pairwise.complete.obs"),
      rmse            = sqrt(mean((estimate_h - estimate_l)^2, na.rm = TRUE)),
      directional_pct = mean(sign(estimate_h) == sign(estimate_l), na.rm = TRUE) * 100,
      inferential_pct = if (!anyNA(p_adj_h) && !anyNA(p_adj_l))
        mean(same_infer, na.rm = TRUE) * 100 else NA_real_,
      tost_pct        = mean(equivalent, na.rm = TRUE) * 100
    )
}
```

### Distribution comparison metrics

`compare_distributions()` compares response distributions for a single condition and outcome using three metrics:

- **OVL (overlapping coefficient)** — proportion of area where the two kernel density estimates overlap; 0 = no overlap, 1 = identical distributions; primary metric for RQ2
- **KS D-statistic** — maximum absolute difference between the two empirical CDFs; 0 = identical
- **Variance ratio** — LLM clone variance divided by human variance; 1 = equal spread

```{r}
#| label: fn-compare-distributions

compute_ovl <- function(x, y, n_grid = 512) {
  lo <- min(c(x, y)); hi <- max(c(x, y))
  if (lo == hi) return(NA_real_)
  # Evaluate both KDEs on the same grid, then integrate the overlap area
  d_h <- density(x, from = lo, to = hi, n = n_grid)
  d_l <- density(y, from = lo, to = hi, n = n_grid)
  sum(pmin(d_h$y, d_l$y)) * (hi - lo) / n_grid
}

compare_distributions <- function(human_data, llm_data, outcome, condition_val) {
  x <- human_data |> filter(condition == condition_val) |> pull(all_of(outcome)) |> na.omit()
  y <- llm_data   |> filter(condition == condition_val) |> pull(all_of(outcome)) |> na.omit()
  tibble(
    condition      = as.character(condition_val),
    ovl            = compute_ovl(x, y),
    ks_d           = suppressWarnings(ks.test(x, y)$statistic),
    variance_ratio = var(y) / var(x)  # > 1: clones more variable; < 1: less variable
  )
}
```

### Calibration regression

`run_calibration()` regresses human ATEs on LLM clone ATEs: `ATE_human = α + β × ATE_clone`. Think of it as asking: if you know the clone's ATE for an intervention, how well can you predict the human ATE? The intercept α captures systematic bias — positive α means clones systematically underestimate, negative means they overestimate. The slope β captures proportionality: β = 1 is perfect calibration (clone effects scale 1:1 with human effects), β > 1 means clone effects are too small, β < 1 means they are too large. We report R² and a formal F-test of the hypothesis β = 1.

```{r}
#| label: fn-calibration

run_calibration <- function(h_result, l_result) {
  joined <- inner_join(
    h_result |> select(condition, estimate_h = estimate),
    l_result |> select(condition, estimate_l = estimate),
    by = "condition"
  )

  # ATE_h = α + β × ATE_l; perfect calibration: α = 0 and β = 1
  fit   <- lm(estimate_h ~ estimate_l, data = joined)
  coefs <- broom::tidy(fit, conf.int = TRUE)

  tibble(
    alpha     = coefs$estimate[1],
    alpha_lo  = coefs$conf.low[1],
    alpha_hi  = coefs$conf.high[1],
    beta      = coefs$estimate[2],
    beta_lo   = coefs$conf.low[2],
    beta_hi   = coefs$conf.high[2],
    r_squared = broom::glance(fit)$r.squared,
    # F-test H0: β = 1 (clone ATEs are proportionally calibrated to human ATEs)
    p_beta_1  = car::linearHypothesis(fit, "estimate_l = 1")$`Pr(>F)`[2]
  )
}
```

## RQ1: Average Treatment Effect Recovery

For each intervention, the average treatment effect (ATE) is the difference in mean outcome between participants assigned to that intervention and those in the control condition. We estimate ATEs by running `run_main_treatment_model()` — OLS with HC2 heteroskedasticity-robust standard errors and BH-adjusted p-values — on human and LLM clone data separately for each preregistered outcome. For `newsletter_signup` (binary), we use `run_main_treatment_model_binary()`, which returns marginal effects on the probability scale.

Good prediction means the clones rank the interventions in the same order as humans (high Spearman ρ), and that individual ATE pairs fall within a preregistered tolerance band (high TOST %). We first report metrics pooled across all outcomes — `r (length(outcomes_continuous) + length(outcomes_binary)) * 20` estimate pairs in total — to maximize statistical power, then break down by outcome and zoom in on the primary outcome. All models are fit first.

```{r}
#| label: rq1-all

# Run and store all models (avoids re-fitting in later chunks)
ate_model_results <- map(outcomes_continuous, function(out) {
  list(
    h = run_main_treatment_model(human_data, outcome = out),
    l = run_main_treatment_model(llm_data,   outcome = out)
  )
}) |> set_names(outcomes_continuous)

# Binary outcome: marginal effects on probability scale
h_bin <- run_main_treatment_model_binary(human_data, "newsletter_signup")
l_bin <- run_main_treatment_model_binary(llm_data,   "newsletter_signup")

# References to primary outcome models for the primary subsection below
h_primary <- ate_model_results[["trust_multidimensional"]]$h
l_primary <- ate_model_results[["trust_multidimensional"]]$l

# Extract ATE pairs with SE and adjusted p-value from a model result tibble
prep_ate <- function(df) {
  df |>
    select(condition, estimate) |>
    mutate(
      se    = if ("std.error"        %in% names(df)) df$std.error        else NA_real_,
      p_adj = if ("p.value_adjusted" %in% names(df)) df$p.value_adjusted else NA_real_
    )
}

# Pool ATE pairs across all outcomes (continuous + binary)
pooled_ates <- bind_rows(
  map_dfr(outcomes_continuous, function(out) {
    inner_join(
      prep_ate(ate_model_results[[out]]$h) |>
        rename(estimate_h = estimate, se_h = se, p_adj_h = p_adj),
      prep_ate(ate_model_results[[out]]$l) |>
        rename(estimate_l = estimate, se_l = se, p_adj_l = p_adj),
      by = "condition"
    ) |> mutate(outcome = out)
  }),
  inner_join(
    prep_ate(h_bin$marginal_effects) |>
      rename(estimate_h = estimate, se_h = se, p_adj_h = p_adj),
    prep_ate(l_bin$marginal_effects) |>
      rename(estimate_l = estimate, se_l = se, p_adj_l = p_adj),
    by = "condition"
  ) |> mutate(outcome = "newsletter_signup")
) |>
  mutate(
    tier = case_when(
      outcome == outcomes_primary                          ~ "Primary",
      outcome %in% c(outcomes_secondary, outcomes_binary) ~ "Secondary",
      TRUE                                                ~ "Tertiary"
    ),
    outcome_label = outcome_label_map[outcome]
  )

# Per-outcome summary metrics
ate_metrics <- map_dfr(outcomes_continuous, function(out) {
  compare_estimates(ate_model_results[[out]]$h, ate_model_results[[out]]$l) |>
    mutate(outcome = out)
})

calib_metrics <- map_dfr(outcomes_continuous, function(out) {
  run_calibration(ate_model_results[[out]]$h, ate_model_results[[out]]$l) |>
    mutate(outcome = out)
})

ate_metrics <- bind_rows(
  ate_metrics,
  compare_estimates(h_bin$marginal_effects, l_bin$marginal_effects) |>
    mutate(outcome = "newsletter_signup")
) |>
  mutate(
    tier = case_when(
      outcome == outcomes_primary                          ~ "Primary",
      outcome %in% c(outcomes_secondary, outcomes_binary) ~ "Secondary",
      TRUE                                                ~ "Tertiary"
    ),
    outcome_label = outcome_label_map[outcome]
  )

calib_metrics <- bind_rows(
  calib_metrics,
  run_calibration(h_bin$marginal_effects, l_bin$marginal_effects) |>
    mutate(outcome = "newsletter_signup")
) |>
  mutate(
    tier = case_when(
      outcome == outcomes_primary                          ~ "Primary",
      outcome %in% c(outcomes_secondary, outcomes_binary) ~ "Secondary",
      TRUE                                                ~ "Tertiary"
    ),
    outcome_label = outcome_label_map[outcome]
  )

# Pooled summary row — prepended to ate_metrics so it appears at the top of tbl-rq1
pooled_row <- pooled_ates |>
  mutate(
    delta      = 0.5 * abs(estimate_h),
    diff       = estimate_h - estimate_l,
    se_diff    = sqrt(se_h^2 + se_l^2),
    p_lower    = pnorm((diff + delta) / se_diff, lower.tail = FALSE),
    p_upper    = pnorm((diff - delta) / se_diff),
    p_tost     = pmax(p_lower, p_upper),
    equivalent = p_tost < 0.05,
    same_sign  = sign(estimate_h) == sign(estimate_l),
    infer_h    = case_when(
      !is.na(p_adj_h) & p_adj_h < 0.05 & estimate_h > 0 ~ "sig_positive",
      !is.na(p_adj_h) & p_adj_h < 0.05 & estimate_h < 0 ~ "sig_negative",
      TRUE ~ "not_significant"
    ),
    infer_l    = case_when(
      !is.na(p_adj_l) & p_adj_l < 0.05 & estimate_l > 0 ~ "sig_positive",
      !is.na(p_adj_l) & p_adj_l < 0.05 & estimate_l < 0 ~ "sig_negative",
      TRUE ~ "not_significant"
    )
  ) |>
  summarise(
    spearman_rho    = cor(estimate_h, estimate_l, method = "spearman",
                          use = "pairwise.complete.obs"),
    pearson_r       = cor(estimate_h, estimate_l, use = "pairwise.complete.obs"),
    rmse            = NA_real_,
    directional_pct = mean(same_sign,  na.rm = TRUE) * 100,
    inferential_pct = mean(infer_h == infer_l, na.rm = TRUE) * 100,
    tost_pct        = mean(equivalent, na.rm = TRUE) * 100
  ) |>
  mutate(outcome = "pooled", tier = "All outcomes", outcome_label = "All outcomes")

ate_metrics <- bind_rows(pooled_row, ate_metrics)
```

### All outcomes

The scatter below shows all `r nrow(pooled_ates)` estimate pairs — one per intervention × outcome combination — colored by outcome tier. Points above the diagonal indicate clone overestimation; points below indicate underestimation. Because estimates from different outcomes are on different scales, the spread also reflects scale variation across outcomes.

The comparison table below includes a pooled summary row at the top, followed by per-outcome rows grouped by tier. RMSE is left blank for the pooled row because it is not comparable across outcomes on different scales; all other metrics are scale-invariant. The calibration regression (second table) and faceted scatter (third figure) are per-outcome only.

```{r}
#| label: fig-rq1-pooled
#| fig-cap: "Human vs. LLM clone ATEs pooled across all outcomes. Each point is one intervention × outcome pair (N = 240). The dashed line is perfect agreement; dotted lines mark zero. Color indicates outcome tier."

tier_colors <- c("Primary" = "#C0392B", "Secondary" = "#2980B9", "Tertiary" = "#7F8C8D")

pooled_ates |>
  ggplot(aes(x = estimate_h, y = estimate_l, color = tier)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey60", linewidth = 0.5) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey80") +
  geom_vline(xintercept = 0, linetype = "dotted", color = "grey80") +
  geom_point(size = 2, alpha = 0.65) +
  scale_color_manual(values = tier_colors, name = "Outcome tier") +
  labs(x = "Human ATE", y = "LLM clone ATE") +
  plot_theme +
  theme(legend.position = "top")
```

```{r}
#| label: tbl-rq1
#| tbl-cap: "RQ1 comparison metrics. The 'All outcomes' row pools all estimate pairs across outcomes; RMSE (—) is omitted there because outcomes differ in scale. Per-outcome rows each have N = 20 intervention pairs. TOST: % of pairs achieving equivalence (Δ = 50% of human ATE)."

ate_metrics |>
  arrange(match(tier, c("All outcomes", "Primary", "Secondary", "Tertiary")), outcome_label) |>
  select(Tier = tier, Outcome = outcome_label,
         `Spearman ρ` = spearman_rho, `Pearson r` = pearson_r, RMSE = rmse,
         `Directional (%)` = directional_pct, `Inferential (%)` = inferential_pct,
         `TOST (%)` = tost_pct) |>
  gt(groupname_col = "Tier") |>
  fmt_number(columns = c(`Spearman ρ`, `Pearson r`, RMSE), decimals = 3) |>
  sub_missing(columns = "RMSE", missing_text = "—") |>
  fmt_number(columns = c(`Directional (%)`, `Inferential (%)`, `TOST (%)`), decimals = 1) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_row_groups()
  )
```

```{r}
#| label: tbl-rq1-calibration
#| tbl-cap: "Calibration regression by outcome. α: intercept (systematic bias, with 95% CI). β: slope (proportionality; 1 = perfect, with 95% CI). R²: variance explained. p(β=1): F-test of perfect calibration."

calib_metrics |>
  arrange(match(tier, c("Primary", "Secondary", "Tertiary")), outcome_label) |>
  mutate(
    beta_ci  = paste0("[", round(beta_lo,  2), ", ", round(beta_hi,  2), "]"),
    alpha_ci = paste0("[", round(alpha_lo, 2), ", ", round(alpha_hi, 2), "]")
  ) |>
  select(Tier = tier, Outcome = outcome_label,
         alpha, alpha_ci, beta, beta_ci, r_squared, p_beta_1) |>
  gt(groupname_col = "Tier") |>
  cols_label(alpha = "α", alpha_ci = "95% CI (α)",
             beta = "β", beta_ci = "95% CI (β)",
             r_squared = "R²", p_beta_1 = "p (β=1)") |>
  fmt_number(columns = c(alpha, beta, r_squared), decimals = 3) |>
  fmt_number(columns = p_beta_1, decimals = 3) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_row_groups()
  )
```

```{r}
#| label: fig-rq1-all
#| fig-cap: "Human vs. LLM clone ATEs for all continuous outcomes. Each point is one of the 20 interventions; axes are free-scaled across panels. Blue line: OLS regression through the panel's points. Faint dashed line: perfect calibration (slope = 1)."
#| fig-height: 12

map_dfr(outcomes_continuous, function(out) {
  inner_join(
    ate_model_results[[out]]$h |> select(condition, estimate_h = estimate),
    ate_model_results[[out]]$l |> select(condition, estimate_l = estimate),
    by = "condition"
  ) |> mutate(outcome = outcome_label_map[out])
}) |>
  ggplot(aes(x = estimate_h, y = estimate_l)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey60", linewidth = 0.4, alpha = 0.35) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey80") +
  geom_vline(xintercept = 0, linetype = "dotted", color = "grey80") +
  geom_smooth(method = "lm", se = FALSE, color = "#2980B9", linewidth = 0.6) +
  geom_point(size = 1.5, alpha = 0.7, color = "grey20") +
  facet_wrap(~ outcome, scales = "free", ncol = 3) +
  labs(x = "Human ATE", y = "LLM clone ATE") +
  plot_theme
```

### Primary outcome (focused)

The primary outcome, multidimensional trust in climate scientists, is the pre-specified main endpoint — a composite averaging four subdimensions (competence, integrity, benevolence, openness). We show its ATE scatter separately to allow easy reading of individual intervention labels. The calibration regression fits `ATE_human = α + β × ATE_clone`: α = 0 means no systematic bias; β = 1 means clone effects are proportionally accurate. We test β = 1 via a formal F-test.

```{r}
#| label: fig-rq1-primary
#| fig-cap: "Human vs. LLM clone ATEs for the primary outcome (multidimensional trust). Each point is one of the 20 interventions. Blue line: OLS regression through the points. Faint dashed line: perfect calibration (slope = 1). Dotted lines mark zero."

inner_join(
  h_primary |> select(condition, estimate_h = estimate),
  l_primary |> select(condition, estimate_l = estimate),
  by = "condition"
) |>
  ggplot(aes(x = estimate_h, y = estimate_l)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey60", linewidth = 0.5, alpha = 0.35) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey80") +
  geom_vline(xintercept = 0, linetype = "dotted", color = "grey80") +
  geom_smooth(method = "lm", se = FALSE, color = "#2980B9", linewidth = 0.7) +
  geom_point(size = 2.5, alpha = 0.8, color = "grey20") +
  geom_text(aes(label = str_remove(condition, "intervention_")),
            size = 2.5, vjust = -0.8, color = "grey40") +
  labs(x = "Human ATE (OLS estimate)", y = "LLM clone ATE (OLS estimate)") +
  plot_theme
```

```{r}
#| label: rq1-primary-calibration
#| tbl-cap: "Calibration regression for the primary outcome. α: intercept (systematic bias). β: slope (proportionality; 1 = perfect). p(β=1): F-test of perfect calibration."

run_calibration(h_primary, l_primary)
```

## RQ2: Response Distribution Shape Recovery

Even if clones match the average treatment effect for each intervention, they may diverge from humans in how responses are distributed: two groups with the same mean can differ in spread, skewness, or tail shape. RQ2 asks whether clones reproduce the full response distribution within each condition, not just its mean.

We compare distributions using three metrics. The **overlapping coefficient (OVL)** is the proportion of area shared between the two kernel density estimates: it equals 1 when distributions are identical and 0 when they share no support. Intuitively, imagine overlaying the two smoothed histograms — OVL is the fraction of the total area covered by both. The **Kolmogorov–Smirnov D-statistic** is the largest vertical gap between the two empirical cumulative distribution functions at any response value: 0 means the CDFs are indistinguishable everywhere, and larger values indicate a point on the scale where the distributions diverge substantially. The **variance ratio** (clone variance / human variance) captures relative spread: a ratio of 1 is ideal, > 1 means clones are more dispersed than humans, and < 1 means clones give more uniform responses.

We first examine distributions within each of the 21 conditions for the primary outcome in detail, then summarise the three metrics averaged across conditions for all outcomes.

```{r}
#| label: rq2-compute

conditions_all <- levels(human_data$condition)

# All continuous outcomes: per-condition metrics
dist_metrics_all <- map_dfr(outcomes_continuous, function(out) {
  map_dfr(conditions_all, function(cond) {
    compare_distributions(human_data, llm_data, out, cond)
  }) |> mutate(outcome = out)
})

# Primary outcome per-condition metrics (subset for density table and plot)
dist_metrics <- dist_metrics_all |> filter(outcome == "trust_multidimensional")

# Summary: average metrics across conditions, one row per outcome
dist_summary <- dist_metrics_all |>
  group_by(outcome) |>
  summarise(
    mean_ovl       = mean(ovl,            na.rm = TRUE),
    mean_ks_d      = mean(ks_d,           na.rm = TRUE),
    mean_var_ratio = mean(variance_ratio, na.rm = TRUE)
  ) |>
  mutate(
    tier = case_when(
      outcome == outcomes_primary     ~ "Primary",
      outcome %in% outcomes_secondary ~ "Secondary",
      TRUE                            ~ "Tertiary"
    ),
    outcome_label = outcome_label_map[outcome]
  )
```

```{r}
#| label: fig-rq2-densities
#| fig-cap: "Response distributions for multidimensional trust by experimental condition. Human participants shown in dark grey, LLM clones in orange."
#| fig-height: 14

source_colors <- c("Human" = "grey30", "LLM clone" = met.brewer("Juarez", n = 10)[1])

bind_rows(
  human_data |> select(condition, value = trust_multidimensional) |>
    mutate(source = "Human"),
  llm_data   |> select(condition, value = trust_multidimensional) |>
    mutate(source = "LLM clone")
) |>
  filter(!is.na(value)) |>
  ggplot(aes(x = value, fill = source, color = source)) +
  geom_density(alpha = 0.35, linewidth = 0.4) +
  scale_fill_manual(values  = source_colors, name = NULL) +
  scale_color_manual(values = source_colors, name = NULL) +
  facet_wrap(~ condition, ncol = 4) +
  labs(x = "Multidimensional trust (0–100)", y = "Density") +
  plot_theme +
  theme(legend.position = "top")
```

```{r}
#| label: tbl-rq2
#| tbl-cap: "RQ2 distribution comparison metrics by condition. OVL: overlapping coefficient (1 = identical). KS D: Kolmogorov–Smirnov D-statistic (0 = identical CDFs). Variance ratio: LLM clone / human variance."

dist_metrics |>
  mutate(condition = factor(condition, levels = conditions_all)) |>
  arrange(condition) |>
  mutate(condition = as.character(condition)) |>
  gt() |>
  cols_label(condition = "Condition", ovl = "OVL",
             ks_d = "KS D", variance_ratio = "Variance ratio") |>
  fmt_number(columns = c(ovl, ks_d, variance_ratio), decimals = 3)
```

Primary outcome summary — mean OVL across all conditions: `r round(mean(dist_metrics$ovl, na.rm = TRUE), 3)`. Mean KS D: `r round(mean(dist_metrics$ks_d, na.rm = TRUE), 3)`. Mean variance ratio: `r round(mean(dist_metrics$variance_ratio, na.rm = TRUE), 3)`.

### All outcomes

The table below summarises distribution similarity averaged across all 21 conditions for each continuous outcome. OVL close to 1 indicates near-identical distributions; KS D close to 0 indicates CDFs that track each other everywhere; variance ratio close to 1 indicates matched spread. Newsletter signup is excluded because kernel density estimates are not meaningful for a binary outcome.

```{r}
#| label: tbl-rq2-summary
#| tbl-cap: "RQ2 distribution metrics averaged across all 21 conditions, by outcome. OVL: mean overlapping coefficient (1 = identical). KS D: mean Kolmogorov–Smirnov D-statistic (0 = identical CDFs). Variance ratio: mean clone / human variance (1 = equal spread)."

dist_summary |>
  arrange(match(tier, c("Primary", "Secondary", "Tertiary")), outcome_label) |>
  select(Tier = tier, Outcome = outcome_label,
         `Mean OVL` = mean_ovl, `Mean KS D` = mean_ks_d,
         `Mean variance ratio` = mean_var_ratio) |>
  gt(groupname_col = "Tier") |>
  fmt_number(columns = c(`Mean OVL`, `Mean KS D`, `Mean variance ratio`), decimals = 3) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_row_groups()
  )
```

## RQ3: Subgroup Heterogeneity Recovery

Interventions may work differently for different demographic groups — for example, a message about scientific consensus might shift trust more strongly among Republicans than Democrats, or more among younger than older respondents. RQ3 asks whether clones reproduce these subgroup differences, or whether they produce a one-size-fits-all response that averages across demographic groups.

For each of six categorical moderators (gender, age band, race, education, income, and partisan identity), we estimate condition × moderator interactions using `run_moderator_model()` separately on human and clone data. Each interaction term captures how much more (or less) effective a given intervention is for one demographic group compared to the reference group. We then compare the full set of interaction estimates between humans and clones using `compare_estimates()` (with `join_by = c("condition", "moderator_level")`). A clone that predicts subgroup heterogeneity well will produce interaction estimates that correlate strongly with human estimates (high Spearman ρ) and agree on sign (high directional %).

```{r}
#| label: rq3-models

mod_results_h <- map(set_names(moderators_cat), function(mod) {
  run_moderator_model(human_data, outcome = "trust_multidimensional", moderator = mod)
})

mod_results_l <- map(set_names(moderators_cat), function(mod) {
  run_moderator_model(llm_data, outcome = "trust_multidimensional", moderator = mod)
})
```

```{r}
#| label: rq3-metrics

rq3_metrics <- map_dfr(moderators_cat, function(mod) {
  compare_estimates(
    mod_results_h[[mod]]$interaction_effects,
    mod_results_l[[mod]]$interaction_effects,
    join_by = c("condition", "moderator_level")
  ) |> mutate(
    moderator   = mod,
    n_estimates = nrow(mod_results_h[[mod]]$interaction_effects)
  )
})
```

```{r}
#| label: tbl-rq3
#| tbl-cap: "RQ3 interaction estimate comparison metrics by moderator. Spearman ρ: rank correlation across all condition × moderator level interaction estimates. Directional agreement: % with the same sign. N: number of interaction estimates compared."

rq3_metrics |>
  select(Moderator = moderator, `Spearman ρ` = spearman_rho, `Pearson r` = pearson_r,
         RMSE = rmse, `Directional (%)` = directional_pct,
         `Inferential (%)` = inferential_pct, `TOST (%)` = tost_pct,
         `N estimates` = n_estimates) |>
  gt() |>
  fmt_number(columns = c(`Spearman ρ`, `Pearson r`, RMSE), decimals = 3) |>
  fmt_number(columns = c(`Directional (%)`, `Inferential (%)`, `TOST (%)`), decimals = 1)
```

```{r}
#| label: fig-rq3
#| fig-cap: "Human vs. LLM clone condition × moderator interaction estimates for the primary outcome, by moderator. Each point is one condition × moderator level combination."
#| fig-height: 10

map_dfr(moderators_cat, function(mod) {
  inner_join(
    mod_results_h[[mod]]$interaction_effects |>
      select(condition, moderator_level, estimate_h = estimate),
    mod_results_l[[mod]]$interaction_effects |>
      select(condition, moderator_level, estimate_l = estimate),
    by = c("condition", "moderator_level")
  ) |> mutate(moderator = mod)
}) |>
  ggplot(aes(x = estimate_h, y = estimate_l)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey60", linewidth = 0.4) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey80") +
  geom_vline(xintercept = 0, linetype = "dotted", color = "grey80") +
  geom_point(size = 1.5, alpha = 0.65, color = "grey20") +
  facet_wrap(~ moderator, scales = "free", ncol = 2) +
  labs(x = "Human interaction estimate", y = "LLM clone interaction estimate") +
  plot_theme
```

## RQ4: Individual-level Demographic Calibration

RQ1–3 all operate on aggregated estimates — average treatment effects, interaction effects, and marginal distributions. RQ4 asks a different question: are individual clone responses calibrated to the demographic profile they were assigned? A clone given a profile of, say, a young Black woman with low income should give responses that match what real young Black women with low income give — not just what the average American gives.

All analyses use the primary outcome (`trust_multidimensional`) and the control condition unless stated otherwise. Three sub-questions are examined in turn.

### RQ4a: Demographic baseline calibration

For each moderator, we compute mean outcomes within each demographic group in the control condition separately for humans and clones, then correlate and compare those cell means. High correlation and low RMSE indicate that clones start from the right baseline response level for each demographic type — before any intervention is applied.

```{r}
#| label: rq4a-compute

# Cell-mean comparison in the control condition, one moderator at a time
baseline_cal <- compare_demographic_baselines(
  human_data, llm_data,
  outcome    = "trust_multidimensional",
  moderators = moderators_cat
)
```

```{r}
#| label: tbl-rq4a
#| tbl-cap: "Demographic baseline calibration. For each moderator, cell means in the control condition are compared between humans and clones. r: Pearson correlation of cell means. RMSE: root mean squared error. N cells: number of demographic groups."

baseline_cal |>
  select(Moderator = moderator, r, RMSE = rmse, `N cells` = n_cells) |>
  gt() |>
  fmt_number(columns = c(r, RMSE), decimals = 3)
```

### RQ4b: Demographic predictability — stereotyping test

We regress the primary outcome on all six demographic moderators plus condition fixed effects separately for humans and clones. If clones over-rely on demographic stereotypes — mapping demographic variables to outcomes too mechanically — they will show a much higher R² than humans: demographics explain more of the variance in clone responses than in human responses. We also compare individual regression coefficients to identify which demographic predictors are over- or underweighted by the clones.

```{r}
#| label: rq4b-compute

pred_test <- compare_demographic_predictability(
  human_data, llm_data,
  outcome    = "trust_multidimensional",
  predictors = moderators_cat
)
```

```{r}
#| label: tbl-rq4b-rsq
#| tbl-cap: "R² from regressing the primary outcome on demographics + condition FE, separately for humans and clones. Clone R² >> human R² signals demographic stereotyping."

pred_test$r_squared |>
  mutate(source = if_else(source == "human", "Human", "LLM clone")) |>
  gt() |>
  cols_label(source = "Source", r_squared = "R²") |>
  fmt_number(columns = r_squared, decimals = 3)
```

```{r}
#| label: fig-rq4b-coefs
#| fig-cap: "Demographic regression coefficients for the primary outcome in humans (x-axis) vs. LLM clones (y-axis). Each point is one dummy-coded level of a moderator. The dashed line marks perfect agreement."

pred_test$coefficients |>
  ggplot(aes(x = est_h, y = est_l)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey60", linewidth = 0.5) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey80") +
  geom_vline(xintercept = 0, linetype = "dotted", color = "grey80") +
  geom_point(size = 2, alpha = 0.8, color = "grey20") +
  geom_text(aes(label = term), size = 2.2, vjust = -0.8, color = "grey40") +
  labs(x = "Human coefficient", y = "LLM clone coefficient") +
  plot_theme
```

### RQ4c: Within-subgroup distribution shapes

A clone dataset could reproduce the overall response distribution while having miscalibrated within-group spreads — matching the average but not what specific demographic groups look like. We extend RQ2 by comparing response distributions within each demographic subgroup in the control condition, using the same OVL, KS D, and variance ratio metrics. Minimum group size of 10 is required for stable kernel density estimation.

```{r}
#| label: rq4c-compute

control_val <- levels(human_data$condition)[1]

# For each moderator × level, compare human and clone distributions in control
subgroup_dist <- map_dfr(moderators_cat, function(mod) {
  groups <- levels(human_data[[mod]])
  map_dfr(groups, function(grp) {
    x <- human_data |>
      filter(condition == control_val, .data[[mod]] == grp) |>
      pull(trust_multidimensional) |> na.omit()
    y <- llm_data |>
      filter(condition == control_val, .data[[mod]] == grp) |>
      pull(trust_multidimensional) |> na.omit()
    # Require minimum n for stable KDE
    if (length(x) < 10 | length(y) < 10) return(NULL)
    tibble(moderator = mod, group = grp,
           ovl = compute_ovl(x, y),
           ks_d = suppressWarnings(ks.test(x, y)$statistic),
           variance_ratio = var(y) / var(x))
  })
})
```

```{r}
#| label: tbl-rq4c
#| tbl-cap: "Within-subgroup distribution comparison in the control condition. OVL: overlapping coefficient (1 = identical). KS D: Kolmogorov–Smirnov statistic. Variance ratio: clone / human."

subgroup_dist |>
  arrange(moderator, group) |>
  gt(groupname_col = "moderator") |>
  cols_label(group = "Group", ovl = "OVL", ks_d = "KS D",
             variance_ratio = "Variance ratio") |>
  fmt_number(columns = c(ovl, ks_d, variance_ratio), decimals = 3) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_row_groups())
```


