---
title: "The Silicon Sample Benchmark: A multi-team preregistration for predicting behavioral-experiment effects"

abstract: |
    This is the preregistration for the **Silicon Sample Benchmark**, a multi-team prediction benchmark. We invite independent research teams to predict the results of a novel behavioral megastudy (16 message interventions to increase trust in climate scientists, 13 preregistered outcomes, a representative U.S. sample of ~18,000) with any AI-based approach they choose, from behavioral clones to group-level simulation to direct effect forecasting. Submissions enter at one of three tiers defined by how fine-grained the prediction is (individual-level synthetic data, cell-level statistics, or effect-level estimates), and every approach is scored with the same preregistered metrics against the same human replication reference. At the time of writing, human data collection is complete (~18,000 responses). None of these data have been shared, published, or made accessible to any prospective team, and no team will be granted access before the prediction lock. Before any human data are released, every team registers its approach using a standardized method-registration form (an extended version of the GUIDE-LLM reporting checklist) and deposits its predictions on Zenodo, either publicly or partly escrowed under a disclosure policy that keeps every entry locked and verifiable even where it cannot be fully public.

date: "2026-07-20"

format:
  html:
    code-fold: true
    toc: true
    toc-depth: 4
  # PDF disabled for now to speed up iteration — re-enable for the final render.
  # pdf:
  #   toc: true
  #   toc-depth: 5
  #   include-before-body:
  #     text: |
  #       \clearpage
  #   colorlinks: true
  #   tbl-colwidths: true
  #   header-includes:
  #     - \usepackage{float}
  #     - \floatplacement{figure}{H}
  #     - \floatplacement{table}{H}
  #     - \usepackage{listings}
  #     - \lstset{breaklines=true, basicstyle=\small\ttfamily}

execute:
  echo: true
  warning: false
  message: false
  fig-width: 8
  fig-height: 5
  fig-dpi: 300
---

```{r}
#| label: setup
#| output: false

# packages
library(tidyverse)
library(sandwich)
library(lmtest)
library(marginaleffects)
library(broom)
library(MetBrewer)
library(gghalves)
library(patchwork)
library(gt)
library(tinytable)
library(here)

# custom functions (reused unchanged from the original preregistration)
source(here("R/functions/statistics.R"))
source(here("R/functions/plots.R"))
source(here("R/functions/tables.R"))
source(here("R/functions/reporting.R"))

# simulated (random) placeholder data
human_data     <- readRDS(here("data/simulation_preregistration/human_data_preregistration.rds"))
llm_data_teams <- readRDS(here("data/simulation_preregistration/llm_data_preregistration_teams.rds"))

# Restrict the human data to the scored conditions (the 16 non-interactive
# interventions + control) BEFORE any model is fit. The real cleaned human
# data will additionally contain the three chatbot arms and the value-
# similarity arm, which no submission predicts; fitting the human models on
# those arms would put the human side's BH adjustment over a different test
# family than the submissions'. A no-op on the placeholder data (which has
# scored conditions only), load-bearing when the real loader is wired in.
scored_conditions <- levels(droplevels(factor(llm_data_teams$condition)))
human_data <- human_data |>
  filter(condition %in% scored_conditions) |>
  mutate(condition = droplevels(condition))

# Same preregistered random split of the human sample as in the original
# preregistration. Human 1 is the reference half all submissions are scored
# against; Human 2 is the human replication reference.
set.seed(42)
split_ids    <- sample(unique(human_data$id), size = floor(length(unique(human_data$id)) / 2))
human_data_1 <- human_data |> filter(id %in% split_ids)
human_data_2 <- human_data |> filter(!id %in% split_ids)
```


::: {.callout-warning}
**This is an ongoing draft, to share with collaborators. This preregistration is not yet registered and might change still. We will register right after the deadline for the intent to participate (August 1st). This allows us to be exact about the number of participating teams.**
:::

::: {.callout-important}
**This is the canonical preregistration for the benchmark.** It is self-contained. The full analysis plan is restated here (see *Analysis plan*), so a participating team does not need to open the project's earlier single-team preregistration ([v1](preregistration.qmd)). The section *About this preregistration* explains why this second preregistration exists, what it keeps from v1, and every way it changes.

The benchmark's public-facing design (motivation, rules, the three tiers, the full method-registration form, the disclosure policy, scoring, and the timeline) is also described in the [call for participation](index.html). To avoid duplication, this document summarizes that design only as far as the analysis needs it, and concentrates on the analysis plan. Where the two overlap, the call is the canonical description of the procedure and this document is the canonical description of the analysis.

All datasets used for demonstration here are simulated random noise with the correct structure.
:::

## Motivation

Whether large language models can stand in for human participants is a consequential open question in behavioral science. The evidence so far is mixed, and has methodological issues: Most existing evaluations target experiments that were already published, and plausibly in the models' training data. Almost none are preregistered, so reported successes are entangled with researcher degrees of freedom in model and prompt choice [@cumminsThreatAnalyticFlexibility2025b].

We therefore aim to provide a preregistered, multi-team prediction benchmark, the *Silicon Sample Benchmark*. Independent teams predict the same novel megastudy (N ≈ 18,000) with any AI-based approach, and every approach is scored with the same preregistered metrics. The design addresses both problems. The target is novel, since none of the interventions has been tested in this form, and every prediction is deposited and time-stamped before any human data are released. The full rationale, rules, and incentives for participating teams are laid out in the [call for participation](index.html).

The benchmark scores the 16 non-interactive text interventions of the megastudy^[The original experiment contains 20 intervention conditions plus a shared control. The benchmark scores only the 16 non-interactive text interventions. The four left out are the three interactive LLM-chatbot arms and the "Value similarity" quiz, all of which depend on a live back-and-forth that would substantially complicate the simulation task.], across self-reported attitudes (e.g., trust in climate scientists, support for funding of climate science) and behavioral measures (a donation to a scientific organization, a newsletter sign-up).

@tbl-interventions gives an overview of the 16 interventions, grouped by their content. All are short text messages (e.g., an article, an interview excerpt, corrected statistics, a portrait) shown once before the outcome battery. Together with the shared control condition they make up the 17 conditions every submission needs to cover.

```{r}
#| label: tbl-interventions
#| tbl-cap: "The 16 text-based interventions scored in the benchmark, grouped by the broad angle each takes on trust in climate scientists. The **Code name** column gives the internal condition identifier used in the raw survey files (`survey.qsf` / `survey.json`), for teams that parse those directly. Each intervention has exactly one code name; where intervention-author teams were combined, the code name is a semicolon-joined list of animal pairs, and the semicolons are part of the name, so match on the full string and never split it (see `survey/condition_codenames.csv` in the benchmark repository). The original megastudy also fields three interactive LLM-chatbot arms and a 'Value similarity' quiz; these four are interactive and excluded from the simulation benchmark, leaving these 16 plus a shared control condition (17 conditions in total)."

interventions <- read_csv(here("data/interventions.csv"), show_col_types = FALSE)

tag_order <- c(
  "Collaboration and peer-review",
  "Scientific methods and results",
  "Applications and impact",
  "Others' endorsement",
  "Values",
  "Other"
)

# Same table layout as v1: category headers inserted
# as bold full-width rows, with a running intervention number.
intervention_table_data <- interventions |>
  mutate(across(c(title, summary, tag, code_name), str_trim)) |>
  filter(tag != "LLM-chatbot", title != "Value similarity") |>
  mutate(tag = factor(tag, levels = tag_order)) |>
  arrange(tag, title) |>
  mutate(number = as.character(row_number())) |>
  # internal condition code names exactly as in the raw survey files
  # (survey.qsf / survey.json); semicolon-joined names are a
  # single identifier — semicolons are part of the name, never split.
  mutate(codename = code_name) |>
  group_by(tag) |>
  group_modify(~ bind_rows(
    tibble(title = as.character(.y$tag), codename = "", summary = ""),
    .x
  )) |>
  ungroup() |>
  mutate(` ` = ifelse(title %in% levels(tag), "", number)) |>
  select(` `, title, codename, summary) |>
  rename(`Intervention Title` = title, `Code name` = codename, Summary = summary)

# identify header rows
header_rows <- which(intervention_table_data$` ` == "")

intervention_table_data |>
  tt(width = c(0.04, 0.19, 0.17, 0.60)) |>
  format_tt(escape = TRUE) |>
  style_tt(align = "l", alignv = "t") |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = header_rows, bold = TRUE, j = 2, colspan = 3) |>
  style_tt(j = 3, monospace = TRUE) |>
  style_tt(fontsize = 0.85)
```

@tbl-outcomes summarizes the 13 preregistered outcomes every submission predicts: the primary trust composite, ten further attitude measures (all 0–100 sliders), and two behavioral measures. Multi-item scales enter the analysis as the mean of their items. The exact item wordings and response scales are in the shipped codebook.

```{r}
#| label: tbl-outcomes
#| tbl-cap: "The 13 preregistered outcomes. Variable names, item counts, example items, and response scales follow the shipped codebook (`codebook.csv` in the benchmark submission template repository), which lists every underlying item in full. Multi-item scales are analyzed as the mean of their items."

tribble(
  ~Outcome, ~Variable, ~Items, ~`Example item`, ~Scale,
  "Multidimensional trust (primary)", "trust_multidimensional", "12",
    "How incompetent or competent are most climate scientists? (four subscales: competence, integrity, benevolence, openness; three items each)",
    "0–100",
  "Single-item trust", "trust_post", "1",
    "How much do you trust climate scientists?", "0–100",
  "Distrust", "distrust_post", "1",
    "How much do you distrust climate scientists?", "0–100",
  "Funding perceptions", "funding_perceptions", "1",
    "Do you think the federal government is spending too much, too little or about the right amount of money on climate change research? (reverse-coded: higher = supports more funding)",
    "0–100",
  "Policy role", "policy_role_mean", "4",
    "Climate scientists should work closely with policy makers to integrate scientific results into policy-making.", "0–100",
  "Institutional trust", "inst_trust_mean", "5",
    "How much do you trust the following institutions? — Environmental Protection Agency (EPA)", "0–100",
  "Climate belief", "belief_post", "1",
    "How accurate do you think this statement is? \"Human activities are causing climate change.\"", "0–100",
  "Climate concern", "concern_mean", "3",
    "How concerned are you about climate change?", "0–100",
  "General climate policy", "policy_general", "1",
    "How much do you oppose or support: \"The U.S. government should do more to reduce global warming\"", "0–100",
  "Specific climate policy", "policy_specific_mean", "7",
    "How much do you support or oppose: Raising taxes on fossil fuels (e.g., gas, oil, coal)", "0–100",
  "Pro-climate behavior", "behavior_mean", "6",
    "How likely are you to engage in the following in the next 12 months? — Eat less meat", "0–100",
  "Donation to AMS", "donation_ams", "1",
    "Of the $10 bonus, how much would you like to donate to the American Meteorological Society (AMS)?", "$0–10",
  "Newsletter signup", "newsletter_signup", "1",
    "Subscription to the \"Talking Climate\" newsletter, offered on an earlier survey page (recorded yes/no).", "0 / 1"
) |>
  # zero-width break after each underscore so long variable names wrap
  mutate(Variable = str_replace_all(Variable, "_", "_\u200b")) |>
  tt(width = c(0.20, 0.18, 0.06, 0.46, 0.10)) |>
  format_tt(escape = TRUE) |>
  style_tt(align = "l", alignv = "t") |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(j = 2, monospace = TRUE) |>
  style_tt(fontsize = 0.85)
```

The goal of the benchmark is *not* to crown a single champion team. Instead, we aim to map the current design space of synthetic sample approaches and to evaluate, as a field, how well they predict human behavior. This benchmark study will provide a catalogue---a structured, comparable description of how research teams currently approach AI prediction of behavioral experiments (open- vs. closed-weight models, persona sources, administration protocols, fine-tuning, in-context learning). We will evaluate approaches on a set of metrics, from directional agreement to calibration, plus distributional fidelity and stereotyping diagnostics (all specified in full under *Analysis plan*, below). We expect different approaches to win on different metrics.

## About this preregistration

**Why a second preregistration.** The project began as a single-team simulation study, preregistered as *"Predicting the effects of behavioral interventions on humans with LLMs"* (the original preregistration, **v1**). That plan asked how well one approach, an individual-level simulation built by members of the core team from demographically representative U.S. profiles with one particular model and prompt, recovers a novel megastudy. But a single team's result is only one point in a large space of ways to build a synthetic sample. We therefore decided to turn the project into a **multi-team benchmark** in which many independent teams predict the same novel experiment with any AI-based approach, and every approach is scored identically. This new preregistration plan supersedes the original **v1** preregistration.

**What it keeps from v1.** The human target study and the estimands (average treatment effects per intervention × outcome, subgroup interactions, control-condition response distributions, and demographic baselines), the three-section analysis structure, the least-to-most-strict ordering of the comparison metrics, and the preregistered split-half **human replication reference** are all carried over from v1. All are restated below (see *Analysis plan*) to keep this new preregistration self-contained. The model functions are reused unchanged.

**What it changes and adds.** The changes fall into three groups. Group A redesigns the study, group B changes what it reports, and group C adds scoring tools from the recent silicon-sample literature. Each change is summarized here, with a pointer to the section that specifies it in full.

Because this preregistration is written for teams who are blinded to the human outcome data, revising the analysis plan does not carry the usual risk of data-contingent researcher degrees of freedom.

**A. From one approach to a field.** These change the *design*, leaving v1's analysis pipeline untouched.

1. **Single team → multi-team benchmark.** The project now evaluates a *field* of independent submissions rather than one core-team approach. It allows for submissions at different levels (Tier 1 individual-level synthetic data, Tier 2 cell-level statistics, Tier 3 effect-level estimates). See *Benchmark design*.
2. **Method registration and disclosure.** Every team documents its pipeline on a standardized method-registration form (an extended GUIDE-LLM checklist) and deposits its predictions under a disclosure policy that keeps even proprietary entries verifiable. See *Method registration and disclosure*.
3. **Per-tier submission specifications.** Each tier has an exact file schema, and each tier's file is reduced to the same set of effect estimates before scoring, so predictions are comparable across tiers. See *Submission specification* and *Tier-specific estimand recovery*.

**B. Additional benchmark design choices.** Benchmark specific results are mere add-ons to the original v1 pipeline. No estimand or model inherited from v1 is redefined.

4. **The headline result is the field, not a single scorecard.** The primary result is no longer one approach's report card but the distribution of predictive quality across the whole field of approaches. Every approach is scored against a split-half **human replication reference** (how well one half of the human sample predicts the other). Teams may enter more than once, so the field summary counts one primary entry per team. Every entry must cover the full study (all 16 interventions × 13 outcomes), so all scores are measured on the same basis. See *Cross-team comparison*.
5. **One score per metric, pooled across outcomes.** A single data point is one predicted effect placed next to the matching human effect, for example how much a particular message shifted trust in climate scientists. We deliberately do *not* collapse the metrics into one overall "champion" score. An approach can get the direction of effects right while mis-scaling their size, or do well on the main outcome and badly on the rest, and a single number would hide exactly the structure the benchmark exists to reveal. Each metric is therefore read on its own. See *Cross-team comparison* (*Scoring principle*). That said, in our main figures, we will report correlations (and adjusted correlations).
6. **An uncertainty interval on every score.** The benchmark tests 16 interventions, but these are only a sample of all possible kinds of messages one could have tested. A different study could have used 16 other messages. An approach's score is therefore an estimate, and like any estimate it would move somewhat under a different draw of interventions. With bootstrap sampling, we attach an interval to every score showing how much it might move based on message selection. For the same reason, we never call one approach better than another from the ranking alone. See *Uncertainty intervals on the leaderboard*.
7. **Point predictions only.** v1's comparison metrics included inferential agreement (does the prediction reach the same significance verdict as the human data?) and an equivalence test, both of which need an uncertainty around each predicted effect. In the benchmark, teams submit point predictions without uncertainty intervals, and these two metrics are not part of the cross-team scoring. Submitted intervals would have created their own problems: a team could make every prediction "not significant" simply by submitting wide intervals, which in a world of mostly small effects scores well without skill, and closing that loophole would have required extra interval-scoring metrics and cross-tier standardizations of no substantive interest. The remaining metrics (direction, rank and linear correlation, absolute error, calibration) capture predictive quality without any submitted uncertainty. See *Comparison functions* and *Cross-team comparison*.

**C. New evaluation add-ons from the silicon-sample literature.** The remaining additions extend the originally proposed evaluation metrics from v1. The new metrics are motivated by how recent studies have evaluated synthetic samples. They supplement or improve the inherited metrics from v1.

8. **A noise-corrected correlation.** The human effects an approach is scored against are estimated from a (here, half-split) sample which is not perfectly representing the population.
The resulting sampling noise can drag down the correlations between human and synthetic sample effects. Because the sampling noise is unrelated to predictive skill, we try to account for it. Following @ashokkumarLargeLanguageModels2026a, we report a noise-corrected correlation beside the raw one (the raw value stays the ranked metric). See *Comparison functions*.
9. **A floor for the metrics without a natural zero.** RMSE has no natural zero to read it against. Adding a naive "predict no effect for everything" approach as a floor reference makes such scores more interpretable. Each approach can then be interpreted by its distance to that floor, and to the human reference [@ashokkumarLargeLanguageModels2026a; @cumminsThreatAnalyticFlexibility2025b]. See *Comparison functions*.
10. **Subgroup distributions.** Averaging outcome distributions over demographic groups can hide an approach that serves most groups well but consistently misses one. We report the gap between the worst- and best-served group in terms of distribution similarity [@parkLLMAgentsGrounded2026]. Because too-similar synthetic responses (under-dispersion) are the problem the literature flags most often, we make under-dispersion the headline distribution diagnostic. See *Section 3*, and *Section 1* (response distributions).
11. **Human sample split robustness.** The human sample is divided in half at random to build the reference. That split is fixed before any data are seen, so it is not a researcher degree of freedom. But the particular split we drew is still arbitrary, and another random split would serve just as well. To show the results do not hinge on the one we happened to draw, we recompute them across many alternative splits and report how stable they are [@cumminsThreatAnalyticFlexibility2025b]. See *Cross-team comparison*.
12. **One unit for all outcomes.** The outcome measures use different scales, with 0–100 attitude sliders, a $0–10 donation, and a 0–1 signup probability. To make effects on these scales comparable, we convert them to **percentage points of its outcome's scale range** before any cross-outcome scoring, following the practice of @ashokkumarLargeLanguageModels2026a. The binary newsletter outcome is added to the subgroup analysis on the same footing, via the linear probability model. See *Cross-team comparison* (*Scoring principle*) and *Section 2*.
13. **A pooled calibration regression.** v1 runs the calibration regression per outcome, on 16 points each. The benchmark instead pools all condition × outcome pairs into one regression (16 × 13 = 208 points, in percentage points of scale range), the same way @ashokkumarLargeLanguageModels2026a estimate their calibration slope across all effects of an archive. See *Comparison functions* (*Calibration regression*).
14. **A fourth distribution metric.** Beside v1's variance ratio, OVL, and KS, the response-distribution analyses add the Wasserstein-1 distance (W1), a bandwidth-free companion to the KDE-based OVL. See *Comparison functions* (*Distribution comparison metrics*).

## The human target sample

Predictions are scored against the human data from the megastudy itself. Participants are recruited from a non-probability opt-in panel, quota-matched (age, gender and race) to the population of U.S. residents, all aged 18 or older, who pass a series of attention and bot-detection checks before treatment assignment.

The benchmark targets **N ≈ 18,000 complete responses**, with 1,000 per intervention across the 16 text-based interventions plus 2,000 in the shared control condition. Recruitment uses **census-based cross quotas** on gender × age and gender × race/ethnicity to approximate the U.S. adult population. The targets below match the proportions of the parent megastudy, with counts rescaled to a total N ≈ 18,000. Participants selecting "Other" as their gender are not quota-constrained.

The targets below are drawn from the 2024 U.S. Census Bureau Population Estimates Program. *Total* is the share within each category. *Male* and *Female* give the within-category gender split. The table describes the key demographics of the human sample. Participating teams are not required to use this information for their own synthetic sample.

```{r}
#| label: tbl-quotas
#| tbl-cap: "Sampling quota targets (N ≈ 18,000), from the 2024 U.S. Census Bureau Population Estimates Program. Proportions match the parent megastudy, with counts rescaled to N ≈ 18,000, while the target of the megastudy is N≈ 22,000."

# Same layout as the interventions table: category headers inserted as bold
# full-width rows above their levels (tinytable, for a consistent look).
quotas <- read_csv(here("data/quotas_18000.csv"), show_col_types = FALSE)

quota_table_data <- quotas |>
  group_by(Variable) |>
  group_modify(~ bind_rows(
    tibble(Category = as.character(.y$Variable), Total = "", Male = "", Female = ""),
    .x
  )) |>
  ungroup() |>
  select(Category, Total, Male, Female) |>
  rename(` ` = Category)

header_rows_q <- which(quota_table_data$Total == "")

quota_table_data |>
  tt(width = c(0.40, 0.20, 0.20, 0.20)) |>
  format_tt(escape = TRUE) |>
  style_tt(align = "l", alignv = "t") |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = header_rows_q, bold = TRUE) |>
  style_tt(fontsize = 0.85)
```

Full sampling and exclusion detail is in the [human study preregistration](https://doi.org/10.5281/zenodo.20160212). The human reference estimates follow that preregistration's analysis plan throughout; in particular, its preregistered checks for differential attrition may lead the human models to apply inverse-probability weights, in which case submissions are scored against the weighted estimates.

## Benchmark design

Participation is by invitation. Each team builds an AI-based prediction pipeline from the shipped materials, including intervention texts, the full survey instrument, the codebook, and the registration template.

Teams can use any models, personas, prompts, fine-tuning, retrieval, or prior literature they want. Before the prediction lock deadline, each team registers its approach by depositing its synthetic sample/predictions and materials on Zenodo, hashed and time-stamped. After the prediction lock deadline, every submission runs through the preregistered analysis pipeline unchanged and is reported on a public leaderboard alongside the human replication reference, and in a joint manuscript with all participating teams as co-authors. There are three ground rules for an entry to be valid: **No data access.** By design, no participating team has access to the human data. **Pipelines are AI-based and automated.** Predictions may be made based on previously published human data, but no new human data specific to this project may be collected. **Everything is registered** (see the disclosure policy below). The full procedure, ground rules, timeline, and incentives are in the [call for participation](index.html).

**Blinding and data access.** The collected human data (all ~18,000 responses; collection complete at the time of writing) have been accessed only by the research lead, Jan Pfänder, for data-quality monitoring and for the preliminary interim results presented at the talks disclosed below. As a result, members or affiliates of the core team who have had no access to any human outcome data may enter the benchmark as regular teams under the same rules as everyone else (same registration form, same blinding attestation, same deadlines). 

For full disclosure, Jan Pfänder presented some very preliminary results (human megastudy sample at the time N ≈ 3000) from an initial simulation attempt (following preregistration **v1**) in a five-minute talk at a closed academic workshop at the Max Planck Institute for Human Development in Berlin (May 2026). This is the only presentation that has targeted an audience of potential collaborators on the benchmark study. It focused on the prediction ability of the simulation approach and revealed nothing about how humans behaved in the megastudy. Two other talks were explicitly about the human megastudy results. Jan Pfänder presented preliminary results from the human megastudy at an invited talk at the Max Planck Institute for Human Development in Berlin (May 2026, N at the time ≈ 3000 of 22000), and at a small conference at the University of Hohenheim (June 2026, N at the time ≈ 13000 of 22000). No one in the audience of these talks will be invited to participate in the benchmark study. In addition, since we ask for extensive disclosure of materials, a "cheating" prompt (e.g., "make sure intervention XX does best") would surface in the reporting framework that teams are required to fill out.

### Three ways to compete

Submissions enter at one of **three tiers**, defined by the granularity of the prediction. Teams choose the highest tier their approach supports. Lower tiers are simply eligible for fewer analyses.

- **Tier 1 · Individual-level synthetic data.** Eligible for everything, meaning ATE recovery, calibration regression, response distributions, subgroup heterogeneity, demographic baselines, and stereotyping diagnostics. *Grain: 1 row = 1 synthetic respondent.*
- **Tier 2 · Group-level treatment effect predictions.** Eligible for ATEs, calibration, subgroup heterogeneity, and demographic baselines. Not eligible for the distribution-shape metrics (variance ratio, OVL, KS, W1) or the predictability regressions. *Grain: 1 row = 1 cell (condition × [moderator level ×] outcome).*
- **Tier 3 · Average treatment effect predictions.** Eligible for ATE recovery and the calibration regression. *Grain: 1 row = 1 intervention × outcome estimate.*

### Analysis eligibility by tier

| Preregistered analysis | Tier 1 | Tier 2 | Tier 3 |
|---|:-:|:-:|:-:|
| ATE recovery (Section 1) | ✓ | ✓ | ✓ |
| Calibration regression $\alpha$ / $\beta$ (Section 1) | ✓ | ✓ | ✓ |
| Response distributions, variance ratio, OVL, KS, W1 (Section 1) | ✓ | — | — |
| Subgroup heterogeneity, condition × moderator (Section 2) | ✓ | ✓ | — |
| Within-subgroup distributions (Section 2) | ✓ | — | — |
| Demographic baseline calibration (Section 3) | ✓ | ✓ | — |
| Demographic predictability / stereotyping (Section 3) | ✓ | — | — |

## Method registration and disclosure

Before the prediction lock, every team documents its pipeline on a structured **method-registration form** and deposits it alongside the predictions. The form builds on the GUIDE-LLM reporting checklist [@feuerriegelReportingChecklistLarge2026]. The form and the **disclosure policy** (which lets proprietary teams seal items under escrow while keeping the lock publicly verifiable) are published in the [call for participation](index.html#registration-form). We provide a template in the [benchmark submission template repository](https://github.com/janpfander/silicon-sample-submission). 

## Submission specification

The deposit logistics (Zenodo record structure, the `metadata.json` file, file naming, the self-check validator, and the public/escrowed split) are specified in the [call for participation](index.html) and the [benchmark submission template repository](https://github.com/janpfander/silicon-sample-submission). This section only specifies aspects that matter for the analysis. Column labels and factor levels must match the published codebook (`codebook.csv`) exactly; condition labels must match the intervention titles in `survey/condition_codenames.csv` and the example prediction files exactly, with the control condition written as lowercase `control`. Every submission must cover all 16 interventions and all 13 outcomes (see *Cross-team comparison*).

### Tier 1 — individual-level dataset

One row per synthetic respondent, with the columns listed below (an example file ships in the submission template repository). Each team supplies its own synthetic respondents. `profile_id` is a team-assigned identifier, unique within the submission, and assignment of respondents to conditions (between-subjects, covering all 17 conditions) is the team's responsibility, registered under form items D.1 and D.3.

| Column(s) | Type | Specification |
|---|---|------|
| `profile_id` | string | A unique respondent identifier you assign (unique within the submission). |
| `condition` | factor | `control` (lowercase) or one of the 16 intervention titles, exactly as in `survey/condition_codenames.csv`. |
| `gender`, `age_band`, `race`, `education`, `income`, `party` | factor | The six moderators, levels per codebook (from your own profiles). |
| `trust_multidimensional` + 12 trust items | numeric | Primary composite 0–100. The 12 underlying items per codebook scale. |
| `trust_post`, `distrust_post`, `funding_perceptions`, `policy_role_mean`, `inst_trust_mean` | numeric | Secondary outcomes, original scales per codebook. |
| `belief_post`, `concern_mean`, `policy_general`, `policy_specific_mean`, `behavior_mean` | numeric | Tertiary outcomes, original scales per codebook. |
| `donation_ams` | numeric | Donation to AMS in $, per codebook. |
| `newsletter_signup` | 0 / 1 | Binary behavioral outcome. |

Because a Tier 1 submission has the same shape as the human participant data, the same analysis code applies. 

**Precision requirement.** An approach's estimated effects are only as precise as the synthetic sample behind them. The fewer synthetic respondents, the noisier each estimated effect, and a very small sample would score poorly for a reason that has nothing to do with the method. A Tier 1 submission must therefore contain at least as many synthetic respondents as the human half it is scored against, meaning **500 per intervention and 1,000 in the control condition**. At that size the submission's effects carry about the same sampling noise as the human reference. We encourage teams to go well beyond this minimum. A larger synthetic sample gives more precise effect estimates that reflect the method itself rather than the luck of a particular draw, and synthetic respondents are relatively cheap to generate.

### Tier 2 — cell-level file(s)

Two long-format files. `*_cells_main.csv` has one row per condition × outcome, and `*_cells_moderator.csv` one row per condition × moderator × moderator-level × outcome. Overall ATEs are derived from the main file (treatment minus control cell mean) and subgroup estimates from the moderator file, so no separate effect-level file is needed.

| Column | Type | Specification |
|---|---|------|
| `condition` | factor | Per `survey/condition_codenames.csv` (lowercase `control`). |
| `moderator`, `moderator_level` | factor | Moderator file only. Levels per codebook. |
| `outcome` | string | Outcome variable name per codebook. |
| `mean` | numeric | Predicted cell mean on the original scale (proportion for `newsletter_signup`). Cells in the moderator file may be `NA` where an approach cannot produce a group mean; they are excluded pairwise and the validator warns. The main file must be complete. |

### Tier 3 — effect-level file

One row per intervention × outcome (16 × 13 = 208 rows). Estimates are ATEs versus control on the original outcome scale. The binary outcome is on the probability scale.

| Column | Type | Specification |
|---|---|------|
| `condition` | factor | Intervention title per `survey/condition_codenames.csv`. |
| `outcome` | string | Outcome variable name per codebook. |
| `ate` | numeric | Predicted ATE, original units. |

## Analysis plan

::: {.callout-note}
**Demo outcome subsets.** Some demonstration chunks below run on a reduced outcome set — a four-outcome subset (`outcomes_illustrative` in the code) for the expensive subgroup models, and the primary outcome for some single-outcome displays. These subsets exist only to keep this document fast to render. They are not part of the registered design: with real submissions, every analysis runs on all outcomes.
:::

### Overview and the human replication reference

Each submission is scored against the same human sample. The human sample is split into two halves by a preregistered random seed (seed 42, set in the setup chunk above). **Human 1** is the reference half every submission is scored against. **Human 2** predicts Human 1 like any submission would, and its scores define the **human replication reference**, which shows how well a fresh human sample of this size reproduces the reference half. This human reference is evaluated in the same way as a Tier-1 synthetic submission. A **floor baseline** ("no effect") anchors the other end for the metrics without a natural null (see *Cross-team comparison*). Because the split seed is an arbitrary choice, we will provide robustness checks across many alternative splits, so that results do not hinge on the seed we happened to draw (*Robustness of the scoring pipeline*, below). 

Which of the analyses below applies to a given submission follows from its tier (see *Analysis eligibility by tier*). Tier 1 is eligible for all of them, Tiers 2–3 for the estimate-based subset.

### Estimation functions

Every effect size compared below is produced by the same model functions, applied in the same way to the human reference and to each submission. They produce the average treatment effects per intervention for Section 1 and the condition × moderator interactions for Section 2. These functions are unchanged from v1, and their full statistical rationale is in the [human study preregistration](https://doi.org/10.5281/zenodo.20160212). We reproduce them here so this document stands alone.

`run_main_treatment_model()` estimates each intervention's ATE against the shared control by OLS with HC2 heteroskedasticity-robust standard errors and BH-adjusted p-values.

```{r, results='asis'}
#| label: fn-run-main-treatment-model
print_a_function_from_file("run_main_treatment_model")
```

`run_main_treatment_model_binary()` is the counterpart for the binary newsletter-signup outcome. It fits a logistic regression and returns marginal effects on the probability scale.

```{r, results='asis'}
#| label: fn-run-main-treatment-model-binary
print_a_function_from_file("run_main_treatment_model_binary")
```

`run_moderator_model()` estimates the condition × moderator interactions used for the subgroup analysis (Section 2). For the binary newsletter-signup outcome the benchmark's subgroup scoring runs this same OLS function as a linear probability model: in the saturated condition × moderator specification, the OLS interaction coefficient equals the difference-in-differences of the cell signup proportions, in probability units, which convert to percentage points like every other outcome. A logistic model gives the same number: both saturated models fit the cell proportions exactly, so converting the logit back to the probability scale (as `run_main_treatment_model_binary()` does via marginal effects for the ATEs) and taking the second difference reproduces the identical difference-in-differences — we verified this numerically. The linear probability model computes this quantity directly, with valid HC2 standard errors, and it stays defined when a small demographic cell contains no signups at all, a case in which the logit coefficients diverge.

```{r, results='asis'}
#| label: fn-run-moderator-model
print_a_function_from_file("run_moderator_model")
```

### Comparison functions

The comparison metrics are implemented by a small set of helper functions, applied symmetrically to the human reference and to each submission, and shown here in full.

#### Comparison metrics

The estimate-based metrics compare the point estimates of a submission with those of the human reference, for ATEs (Section 1) and moderator interaction estimates (Section 2). They are reported in increasing order of strictness and implemented in `pooled_metrics()` (shown under *Scoring functions*, below).

- **Directional agreement (%)**. The share of estimates with the same sign as the human estimate, where chance is 50%. Does the approach at least get the direction of each effect right?
- **Spearman $\rho$**. The rank correlation of the point estimates. Does the approach rank the interventions in the same order as humans? It is insensitive to scale or offset, so an approach that inflates all effects equally still scores $\rho$ = 1.
- **Pearson r**. The linear correlation of the point estimates. It captures whether predicted and human effects rise and fall together in proportion, but not whether their absolute size matches.
- **RMSE**. The root mean squared error, the average size of the prediction errors. In the cross-team analysis every estimate is expressed in **percentage points (pp) of its outcome's scale range** before scoring (see *Scoring principle*, below), so RMSE reads in pp and is reported both pooled across outcomes and per outcome, matching @ashokkumarLargeLanguageModels2026a, whose primary archive reports RMSE pooled across experiments in exactly these units. To judge whether an RMSE is good, compare it to the standard deviation of the human ATEs — how much the 16 human effects spread out around their mean (a spread across interventions, not a standard error). A submission that ignores the interventions entirely and predicts the same value for all of them can at best reach an RMSE equal to that SD, by predicting the mean ATE. A submission with intervention-specific skill must therefore come in below it. (In v1, which scores each outcome on its own raw scale, RMSE stays per outcome and is omitted from the pooled row.)

#### Adjusted correlation and RMSE (noise-corrected)

The effects in the human sample are noisy. That noise mechanically drags the correlation between effects of the human and synthetic sample down, because it inflates the observed spread of the reference effects. The raw Pearson r divides by that inflated spread and is biased toward zero. The same noise inflates the RMSE.

Wherever Pearson r appears in the cross-team analysis we therefore report, beside the raw value, a noise-corrected correlation **r~adj~** and a noise-corrected RMSE (`adjusted_metrics()`). The correction works by subtraction. Because the sampling noise is independent (that's an assumption we make), the true spread of the reference effects is their observed spread minus the average known sampling variance. The adjusted correlation divides the covariance by that corrected spread (Spearman disattenuation with known error variances). This method is similar to the meta-analytic r~adj~ of @ashokkumarLargeLanguageModels2026a, but uses fewer assumptions. As @ashokkumarLargeLanguageModels2026a, we will report r and r~adj~ side by side in the manuscript. This allows us to directly compare our results to their larger megastudy archive, the closest analogue to this benchmark^[We take Ashokkumar et al.'s megastudy figures, not their headline r = .85, as the reference because the two numbers measure different tasks. The .85 is a correlation over 469 effects pooled from 70 unrelated survey experiments, spanning many topics, treatments, and outcomes. Their megastudy figures instead measure how well the model distinguishes the several interventions of a single megastudy against that study's shared control, which is exactly this benchmark's setting (16 interventions, one control). Ashokkumar et al. report the megastudies as a separate secondary archive (15 megastudies, 606 effects) and note that these correlations run lower than the primary archive's.], survey experiments reach r = .43 (r~adj~ = .52) and text-based treatments r = .45 (r~adj~ = .54).

The adjusted RMSE applies the same logic to the RMSE. It subtracts the average reference sampling variance from the average squared prediction error and takes the square root.

```{r, results='asis'}
#| label: fn-adjusted-metrics
print_a_function_from_file("adjusted_metrics")
```

#### Distribution comparison metrics

`compare_distributions()` compares response distributions for a single condition and outcome using four metrics:

- **Variance ratio** (submission variance / human variance). It captures only relative spread. A ratio of 1 means equal spread and says nothing about shape. A ratio below 1 means the synthetic responses vary less than the human ones, the most widely reported problem with synthetic samples.
- **OVL (overlapping coefficient)**. The share of area the two kernel density estimates have in common (1 = identical distributions, 0 = no overlap). It summarizes overall shape similarity.
- **KS D-statistic**. The largest gap between the two empirical cumulative distributions (0 = indistinguishable everywhere). It is sensitive to local departures that OVL might average away.
- **Wasserstein-1 distance (W1)**, enabled with `include_w1 = TRUE` (new in this preregistration, not part of v1). The area between the two cumulative distributions, in outcome scale points (0 = identical). It is the bandwidth-free companion to OVL: kernel density estimates depend on bandwidths, which change with sample size and leak density past the 0/100 scale bounds, while W1 has no tuning parameter.

```{r, results='asis'}
#| label: fn-compare-distributions
print_a_function_from_file("compare_distributions")
```

```{r, results='asis'}
#| label: fn-compute-ovl
print_a_function_from_file("compute_ovl")
```

```{r, results='asis'}
#| label: fn-compute-w1
print_a_function_from_file("compute_w1")
```

#### Calibration regression

Correlation tells us only whether the human and predicted effects covary. It does not capture baseline differences or scaling. An approach could inflate every effect by a factor of two and still score Pearson r = 1. The calibration regression captures this.

`run_calibration_pooled()` regresses human ATEs on predicted ATEs, $\mathrm{ATE}_{human} = \alpha + \beta \times \mathrm{ATE}_{predicted}$, across all condition × outcome pairs pooled over the 13 outcomes (16 × 13 = 208 points, every estimate in pp of scale range). This is how @ashokkumarLargeLanguageModels2026a estimate their calibration slope, pooled across all effects of an archive; v1 instead ran the regression per outcome on 16 points, which left the tests with little power. The slope $\beta$ captures proportional scaling: $\beta = 1$ is perfect, $\beta < 1$ means predicted effects are exaggerated (a 1-pp change in prediction corresponds to less than 1 pp of human effect), $\beta > 1$ that they are compressed. The intercept $\alpha$ is a constant offset: the human effect expected when the predicted effect is zero, added on top of every prediction regardless of its size. An approach can be perfectly proportional ($\beta = 1$) and still have a constant bias ($\alpha \neq 0$), for example if its control condition sits at the wrong level. We report 95% CIs for both, and $R^2$ as the share of human-ATE variance linearly recoverable from the predictions. Calibration is then tested at two strictness levels. The joint F-test of ($\alpha = 0$, $\beta = 1$) tests perfect calibration, meaning the predictions can be read directly as human ATEs with no correction. The weaker slope-only F-test of ($\beta = 1$) asks whether the predictions are proportionally calibrated, permitting a constant offset. Because the human ATEs carry sampling error that differs across conditions, the tests use HC2 heteroskedasticity-robust standard errors (`robust = TRUE`). Non-rejection of these tests is weak evidence: it means the data cannot rule out calibration, not that calibration is demonstrated. Per approach, the leaderboard shows $\alpha$, $\beta$, and $R^2$; the CIs and the two F-test p-values are part of each approach's released results. The calibration lines drawn in @fig-headline are illustrative per-approach OLS fits (averaged for the field), not this HC2 regression.

```{r, results='asis'}
#| label: fn-calibration
print_a_function_from_file("run_calibration_pooled")
```

```{r, results='asis'}
#| label: fn-calibration-base
print_a_function_from_file("run_calibration")
```

#### Demographic baseline and predictability metrics

`compare_demographic_baselines()` computes, for each moderator, the Pearson correlation and RMSE between human and synthetic sample cell means in the control condition. `compare_demographic_predictability()` fits separate OLS regressions of the outcome on each moderator plus condition fixed effects, returning $R^2$ and dummy-coded coefficients for humans and for the submission. `demographic_parity_gap()` condenses the group-specific baseline errors into one fairness number per moderator, the gap between the worst- and best-served group. For example, if an approach's control-condition baseline is off by 6 pp for Republicans but only 1 pp for Democrats, the parity gap for partisan identity is 5 pp, with Republicans the worst-served group (see *Section 3*).

```{r, results='asis'}
#| label: fn-demographic-baselines
print_a_function_from_file("compare_demographic_baselines")
```

```{r, results='asis'}
#| label: fn-demographic-predictability
print_a_function_from_file("compare_demographic_predictability")
```

```{r, results='asis'}
#| label: fn-demographic-parity-gap
print_a_function_from_file("demographic_parity_gap")
```

### Section 1 — Average Treatment Effects

**ATE recovery.** For each intervention, the average treatment effect (ATE) is the difference in mean outcome between participants assigned to that intervention and those in the control condition. ATEs are estimated with `run_main_treatment_model()`, run on the human reference and on the submission separately for each preregistered outcome. For the binary `newsletter_signup`, `run_main_treatment_model_binary()` returns marginal effects on the probability scale. Before any cross-outcome scoring, every estimate and standard error is converted to percentage points of its outcome's scale range (see *Scoring principle*), so effects on the 0–100 sliders, the $0–10 donation, and the signup probability have comparable units. The comparison metrics (`pooled_metrics()`) and the calibration regression (`run_calibration_pooled()`) are then applied. The main analysis pools across all outcomes to maximize power.

**Response distributions.** Even an approach that matches every ATE may distribute responses differently. Two groups with the same mean can differ in spread, skewness, or tail shape. In the control condition, across all continuous outcomes, we compare the full response distribution (`compare_distributions()`, reporting the variance ratio, OVL, KS D, and W1). The **variance ratio** is the headline diagnostic of the four, because under-dispersion (synthetic respondents answering too much alike; ratio < 1) is a widely reported issue in the recent evaluation literature [@ashokkumarLargeLanguageModels2026a; @cumminsThreatAnalyticFlexibility2025b; @parkLLMAgentsGrounded2026; @arrudaCollectiveCooperationIndividual2026]. `newsletter_signup` is excluded, since kernel density is not meaningful for a binary outcome. This analysis requires respondent-level data and so is **Tier 1 only**.

### Section 2 — Subgroup Effects

**Heterogeneity in treatment effects.** Interventions may work differently across demographic groups. A consensus message might shift trust more among Republicans than Democrats, or more among younger than older respondents. For each of six categorical moderators (gender, age band, race, education, income, partisan identity), condition × moderator interactions are estimated with `run_moderator_model()` on the human reference and on the submission. For the binary outcome the same OLS fit is a linear probability model, so the interaction coefficient is the difference-in-differences of cell signup proportions on the probability scale (see *Estimation functions*). As in Section 1, all interaction estimates are converted to percentage points of each outcome's scale range before pooling. The interaction estimates are then compared with the same metrics as Section 1, except RMSE: directional agreement, Spearman $\rho$, and Pearson r (`signed_metrics()`), plus the adjusted correlation r~adj~ computed from the human interaction estimates' standard errors. RMSE is left out because interaction estimates are far noisier than ATEs, so an absolute-error comparison would mostly measure reference noise. The subgroup section is where the noise correction matters most, and r~adj~ is the emphasized value here. Small demographic cells make the reference noise large and uneven, so raw correlations are unevenly deflated across groups (see *Adjusted correlation and RMSE (noise-corrected)*).

One caveat is that if the human data turn out to contain little true moderation, subgroup recovery cannot separate approaches: a submission that predicts no heterogeneity at all then scores well, and a high correlation may reflect generalized predictions rather than subgroup-specific skill. @ashokkumarLargeLanguageModels2026a flag the same limitation and handle it the same way: they report that only 7.7%, 8.8%, and 18.3% of their archive's effects were significantly moderated by gender, ethnicity, and party, and read those shares against the 5% false-positive rate as showing virtually no heterogeneity. Following that practice, we report the share of human interaction estimates that are individually significant (normal approximation from the interaction standard errors, BH-adjusted within outcome × moderator) next to the subgroup results; when that share is near the 5% false-positive rate, the subgroup metrics are read as uninformative rather than as evidence for or against any approach. (The r~adj~ NA-guard flags the same condition from the noise side.) Eligible for Tiers 1–2.

**Response distributions by subgroup.** Mirroring Section 1, within each demographic subgroup in the control condition we compare response distributions (variance ratio, OVL, KS D, W1), to catch an approach that reproduces the overall distribution while miscalibrating within-group spreads. Tier 1 only.

### Section 3 — Demographic Baseline Calibration

Section 2 tests whether an approach exaggerates *effect* heterogeneity. Section 3 asks a separate question. Does the approach start from the right **baseline**? It uses only the control condition. An approach could assign far too little trust to Republicans, or too much to Democrats, before any treatment, a case of baseline stereotyping.

**Demographic baseline calibration.** For each moderator, we compute control condition group means for humans and for the submission, then correlate and compare them (`compare_demographic_baselines()`, which reports Pearson r and RMSE of the cell means). High correlation and low RMSE mean the approach starts from the right response level for each demographic group. This is a different question from the distribution comparisons of Sections 1 and 2: those ask whether the whole response distribution has the right shape and spread, while baseline calibration asks whether each group's *mean* sits at the right level. An approach can match every group mean while badly missing the spread within groups, and vice versa. Eligible for Tiers 1–2.

**Demographic parity gap (DPD).** The baseline-calibration metrics average over groups, and an average can hide a simulation approach that may serve most groups well while consistently missing one. Following the demographic-parity-difference summary of @parkLLMAgentsGrounded2026^[Park et al. compute the gap in per-group predictive accuracy over paired agent–person data, while this benchmark has no person-level pairing, so the gap is taken over group-mean baseline errors instead.], `demographic_parity_gap()` reports, per moderator, the gap between the worst- and best-served demographic group (the difference between the largest and smallest absolute cell-mean error in the control condition), together with which groups those are. The worst-served group's absolute error is reported alongside it, because a zero gap can also mean every group is served equally *badly*, which the worst-group error would reveal. Groups with fewer than 30 respondents on either side are skipped (`min_n = 30`), because a mean over a handful of respondents makes the worst-group error a maximum over noise. Eligible for Tiers 1–2; for Tier 2, the submitted moderator-cell means take the place of computed group means.

**Demographic predictability.** For each of the six moderators we fit a separate OLS regression of the outcome on that moderator plus condition fixed effects, on the human reference and on the submission independently (`compare_demographic_predictability()`). This answers questions such as how much does partisan identity alone explain outcome variance in humans versus in the synthetic samples? An $R^2$ substantially higher for the submission than for humans flags exaggeration of demographic variables (stereotyping). The dummy-coded coefficients identify which groups are over- or under-weighted. Tier 1 only.

### Tier-specific estimand recovery

The comparison metrics require one estimate per intervention × outcome. How it is obtained depends on the tier:

- **Tier 1.** The submission runs through `run_main_treatment_model()` (and `run_main_treatment_model_binary()` for the newsletter signup) exactly like the human data, which produces the ATE estimates.
- **Tier 2.** The ATE for each intervention × outcome is the submitted cell mean minus the submitted control mean. Subgroup estimates and demographic baselines are computed analogously from the moderator cell file.
- **Tier 3.** The submitted `ate` is the estimate.

Every tier thus reduces to the same set of point estimates, and the same metrics apply to all of them.

### Cross-team comparison

The primary goal of this project is to assess how well the research field of synthetic samples can predict the results of the behavioral experiment. It reports the distribution of predictions across submissions on each metric, its center and spread. The secondary goal is to rank individual approaches on the evaluation metrics.

**Scoring principle.** Each approach is scored on every comparison metric, with the intervention × outcome estimates pooled across all 13 outcomes into a single value per metric (the 11 continuous sliders plus the two behavioral outcomes, the donation via the OLS model and the newsletter signup via logistic marginal effects on the probability scale, as in v1's pooled row). Before pooling, every estimate and standard error is converted to percentage points (pp) of its outcome's scale range. This is the unit of the predicted-vs-observed analysis of @ashokkumarLargeLanguageModels2026a, whose primary archive pools effects only after this conversion and who avoid SD standardization because synthetic under-dispersion corrupts an SD denominator. The same rationale applies here. The conversion is a modeling choice which equates 3 points on a 0–100 slider with $0.30 on the donation and 3 pp of signup probability. We deliberately do not collapse the metrics into one composite "champion" score. An approach can recover the direction of effects well while mis-scaling them, or do well on the primary outcome while failing on the rest, and indeed this inconsistency across outcome metrics has been observed in other studies [@cumminsThreatAnalyticFlexibility2025b]. A pooled correlation can also be earned partly by predicting *which outcomes* are generally movable rather than which interventions work within an outcome; the per-outcome breakdown (see *Where the field predicts well*) is reported alongside every pooled value and shows this directly. With all estimates in pp, RMSE and its noise-corrected companion RMSE~adj~ are comparable across outcomes and are reported pooled as well as per outcome (in v1's raw-unit, per-outcome framing they could not be pooled). The per-intervention breakdown keeps its restriction to `outcomes_continuous`: the pp conversion removes the *unit* mix, but a per-intervention correlation across all 13 outcomes would still mix attitude and behavioral constructs.

Teams may enter up to two submissions per tier. For the main findings on the field (Finding 1) we use only one pre-designated primary entry per team. Every valid entry still appears on the leaderboard (Finding 2), labeled primary or secondary. Every submission must cover the full study, all 16 interventions and all 13 outcomes: a self-chosen, potentially easier-to-predict subset would not be comparable to the other entries, so partial submissions are not accepted. Submissions are validated before the lock, and cells that fail to parse at scoring time are dropped; because the share of usable output is itself an analytic choice that can vary widely across pipelines [@cumminsThreatAnalyticFlexibility2025b], the share of dropped cells is reported per submission on the leaderboard. Finally, the field distribution describes an invited, selected set of expert teams, a purposive sample of the current generation of serious approaches rather than a random sample of all possible ones. We state this wherever the field summary is interpreted.

**Presentation.** Every preregistered metric is computed and reported. However, for the key figures of the manuscript, we will pick one key metric per question. **Pearson *r*** is the key metric for ATE recovery (Section 1) and, on the same axis, for subgroup recovery (Section 2). The calibration slope $\beta$ and intercept $\alpha$ answer whether those effects are the right *size*. The **variance ratio** is the key metric for the response-distribution analyses of Sections 1 and 2. Every other preregistered metric (directional agreement, Spearman $\rho$, the adjusted correlations, RMSE, OVL, KS, W1) is still computed and scored for every submission, and reported in a companion table beside each figure.

The headline display (@fig-headline) puts the key measures in one figure with two panels. Panel A plots predicted against human ATEs across all approaches and outcomes, in percentage points of each outcome's scale range. This is the direct visual of the prediction task, in the units of the predicted-vs-observed figure of @ashokkumarLargeLanguageModels2026a. It is overlaid with the average field and human **calibration regressions** and a corner readout of their correlation and calibration coefficients (slope $\beta$, perfect = 1, and intercept $\alpha$, perfect = 0, with 95% CIs). Panel B **ranks** the approaches on the pooled Pearson *r*, each with a cluster-bootstrap **uncertainty interval** over interventions (new in this preregistration, since v1 reported point estimates only, and detailed under *Uncertainty intervals on the leaderboard* below). The human replication reference is pinned on top, and an **all-approaches summary row** at the bottom carries the field's density and mean with its CI, so this one panel shows both the field distribution (Finding 1) and the ranking (Finding 2). Three further figures break the key metrics down, by outcome and by intervention (@fig-field-by-group), by moderator for subgroup recovery (@fig-subgroup-headline), and as per-outcome response densities with the variance ratio and Wasserstein-1 distance stamped on each panel (@fig-distribution-headline). Every entry carries its disclosure-class badge (open / escrowed / sealed) in the leaderboard table, and the precise numbers accompany every figure as a companion table.

#### Uncertainty intervals on the leaderboard

Each point in the ranking panel is a single pooled score, one number per approach per metric. A bare point invites over-reading. If one approach scores Pearson *r* = 0.60 and another 0.55, is the first genuinely better, or is the gap an artifact of the particular interventions that happened to be tested? The interval answers that question. It guards against ranking approaches on differences that would not survive a different, equally reasonable test.

The pooled metric is computed once on the full data to give the point estimate. We then form a *cluster bootstrap over interventions* (`cluster_boot()` in `R/functions/statistics.R`, preregistered seed 2026). Each replicate draws 16 interventions with replacement (so some are drawn twice and some are dropped), and each drawn intervention brings along its rows for all outcomes, because rows that share an intervention are not independent. The pooled metric is recomputed on each replicate, and the 2.5th to 97.5th percentiles of those recomputed scores give the 95% interval. The 13 outcomes themselves are never resampled: they are a fixed design feature of the megastudy, so the interval reflects only uncertainty over which interventions were tested. How scores vary across outcomes is shown descriptively in the per-outcome breakdown (*Where the field predicts well*), whose per-outcome values rest on just 16 pairs each and therefore carry no intervals of their own.

No submitted prediction is ever altered. Each approach's predicted effect for each intervention stays fixed at the value the team submitted. What is resampled is which interventions enter the score. The interval therefore measures the spread of the recomputed score, not the spread of the raw effects. Averaging or correlating cancels much of the effect-to-effect spread, which is why the bar is typically far narrower than the range of the underlying estimates.

**Assumptions.** The interval treats the 16 interventions as a *test set* standing in for the broader space of interventions the benchmark cares about, just as a held-out test set stands in for future cases in predictive modeling. Read this way, the bar answers a generalization question. Would this approach's score, and its rank, survive on a different set of interventions of the same kind? Resampling does **not** assume that a team that predicted a given intervention well could have gotten it wrong, since the predictions are fixed. It assumes only that a replication might have tested a different set of interventions.

The metrics are bounded agreement scores with a ceiling at perfect agreement. An approach that tracks the human effects closely sits near that ceiling and stays there under resampling, because dropping or duplicating interventions barely moves a score already pinned near perfect agreement. Accurate approaches therefore carry *narrow* intervals. An approach whose predictions do not track the human effects has a score floating in the interior, free to swing with the resampled set, and so carries a *wide* interval. Interval width is entangled with accuracy, but in the helpful direction. A wide bar signals that a score is flimsy and may reflect test-set luck rather than skill, exactly the warning one wants attached to a weak entry. (Spread in the size of the true effects does not by itself widen the bar. The correlation and agreement metrics are scale-free, so what matters is whether predictions hug the line, not how far apart the interventions are. Width also partly reflects differences in per-intervention difficulty, so it is a flimsiness *signal*, not a pure standard error.)

The intervals are for reading single scores, not for comparing approaches. Because every submission is scored against the same reference, the intervals of two approaches share noise and cannot be compared like independent confidence intervals: overlap does not show that two approaches are equivalent, and non-overlap is not required to suspect a difference. We therefore present the ranking as descriptive and preregister no test of whether one approach outperforms another; among top performers, statistical near-ties are the most common outcome in comparable mass-collaboration benchmarks [@salganikMeasuringPredictabilityLife2020], and we expect them here. Two caveats apply to all bootstrap intervals. With 16 intervention clusters, percentile intervals are known to undercover somewhat. And the 16 ATEs share the single control arm, which induces a dependence the intervention-level resampling does not capture. The intervals are read as honest approximations, not exact coverage statements.

**Relation to the per-effect standard errors.** The bootstrap interval is a different layer of uncertainty from the sampling error of the individual human effects. A per-effect standard error is *within-intervention*: it says how precisely a single human effect is estimated, and it enters the scoring only through the noise-corrected r~adj~ and RMSE~adj~. The bootstrap interval is *across-intervention*: it says how much the pooled score depends on which interventions were tested, holding each estimate fixed at its point value.

#### Scoring functions

The cross-team analysis reuses the comparison functions specified under *Comparison functions*, above, and adds the scoring helpers below (in `R/functions/statistics.R`), each shown in full and explained before it.

`pooled_metrics()` computes the ATE-recovery metrics on intervention × outcome pairs pooled across all outcomes: directional agreement, Spearman $\rho$, Pearson r, and the noise-corrected companion r~adj~ (via `adjusted_metrics()`, see *Comparison functions*). The pairs it receives are already in pp of scale range, converted once at pair-building time, so `include_rmse = TRUE` adds RMSE and RMSE~adj~ in pp, used for the pooled leaderboard as well as the per-outcome breakdown. It expects pre-joined pairs with the reference (Human 1) columns `estimate_h`, `se_h` and the submission column `estimate_l`.

```{r, results='asis'}
#| label: fn-pooled-metrics
print_a_function_from_file("pooled_metrics")
```

`signed_metrics()` is the subgroup counterpart: directional agreement, Spearman $\rho$, and Pearson r (plus r~adj~ where the human interaction standard errors are available) on the interaction pairs, matching Section 2, above.

```{r, results='asis'}
#| label: fn-signed-metrics
print_a_function_from_file("signed_metrics")
```

`cluster_boot()` attaches the uncertainty interval described above. It resamples the levels of a clustering variable (here, the interventions) with replacement, recomputes the supplied metric function on each replicate, and returns the point estimate together with percentile confidence bounds in long form (one row per metric, with `value`, `lo`, `hi`). It is generic. The metric function is passed in, so the same bootstrap serves every panel.

```{r, results='asis'}
#| label: fn-cluster-boot
print_a_function_from_file("cluster_boot")
```

`metrics_by_group()` scores within each level of a grouping variable instead of pooling everything into one number. With `group = "outcome"` it returns one set of metrics per outcome (scored across the interventions), and with `group = "condition"` one per intervention (scored across the outcomes). It drives the by-outcome and by-intervention distribution figures. Per intervention the score runs across `outcomes_continuous` only (see *Scoring principle*).

```{r, results='asis'}
#| label: fn-metrics-by-group
print_a_function_from_file("metrics_by_group")
```

`summarise_field()` reduces the per-approach scores to the field-level summary printed beside the headline figure, giving the mean, median, SD, min, and max of each metric across approaches (excluding the reference row).

```{r, results='asis'}
#| label: fn-summarise-field
print_a_function_from_file("summarise_field")
```

The remainder of this section demonstrates the code on placeholder data with **30 mock "teams"**, each a stratified bootstrap resample (within condition) of the single individual-level placeholder dataset, enough approaches to give the field distributions a meaningful shape. To exercise all three tier code paths, 28 teams enter at Tier 1 (individual-level), one is collapsed into a Tier 2 cell-level file, and one into a Tier 3 effect-level file. All of this placeholder data, including the resampled teams and the collapsed Tier 2/3 files, is built in the simulation script (`R/simulate_data_preregistration.R`), not here. This document only reads it. Because every mock team is just a resample of the same dataset, all approaches are random noise relative to the human reference. On placeholder data the field distributions therefore cluster near the null, and the figures only show that the code runs. Real submissions will spread out. The adjusted correlation r~adj~ behaves accordingly. With no true signal, the reference effects' observed spread is all noise, so the NA guard binds for many mock teams and the surviving values are unstable. That is the intended behavior, demonstrated rather than hidden. The code applies to real submissions without modification.

```{r}
#| label: variables

outcomes_primary   <- "trust_multidimensional"
outcomes_secondary <- c("trust_post", "distrust_post",
                         "funding_perceptions", "policy_role_mean", "inst_trust_mean")
outcomes_tertiary  <- c("belief_post", "concern_mean", "policy_general",
                        "policy_specific_mean", "behavior_mean")
outcomes_continuous <- c(outcomes_primary, outcomes_secondary, outcomes_tertiary)

# The two behavioral outcomes enter the pooled cross-team metrics alongside
# the 0-100 sliders (as in v1's pooled ATE row): donation in dollars via the
# OLS model, newsletter signup via logistic marginal effects on the
# probability scale. The by-intervention breakdown stays on
# `outcomes_continuous` only.
outcomes_behavioral <- c("donation_ams", "newsletter_signup")
outcomes_pooled     <- c(outcomes_continuous, outcomes_behavioral)

# Scale ranges: all cross-team scoring runs in percentage points (pp) of each
# outcome's scale range — effects and SEs are converted once, at pair-building
# time, and every downstream metric, figure, and table reads the same unit.
# This follows Ashokkumar et al.'s primary archive, which pools effects only
# after expressing them in pp of the original scales, and deliberately avoids
# SD standardization (an under-dispersed submission would corrupt an SD
# denominator). Pooling raw mixed units instead would let the 0-100 sliders
# drown out the two behavioral outcomes (dollar and probability effects are
# 10-100x smaller numerically) and hand every submission 32 free near-origin
# points. The 0-100 sliders are already in pp; only the donation (0-10
# dollars) and signup probability (0-1) rescale.
scale_range <- c(
  set_names(rep(100, length(outcomes_continuous)), outcomes_continuous),
  donation_ams = 10, newsletter_signup = 1
)

# Bootstrap replicates: the PREREGISTERED value for the real-data run is
# 2,000. This placeholder render is a demonstration of the procedure, not of
# its precision, so it uses a small demo value to keep the render cheap —
# only this parameter changes when the submissions arrive, nothing else.
n_boot_real <- 2000
n_boot_demo <- 200

# Resplit sensitivity check (see "Robustness of the scoring pipeline"): same
# demo/real split as the bootstrap. The PREREGISTERED value for the real-data
# run is 200 resplits; the placeholder render uses 50 to keep the build cheap.
n_resplit_real <- 200
n_resplit_demo <- 50

# Demo-only outcome subset (see the callout at the top of the Analysis plan):
# the subgroup demo below runs these four outcomes plus the binary newsletter
# signup to keep the moderator-model fits fast in this render. The
# registered subgroup analysis runs all outcomes. The newsletter interaction
# difference-in-differences come from the linear probability model, so they
# live on the probability scale and pp-rescale like everything else.
outcomes_illustrative <- c("trust_multidimensional", "donation_ams",
                           "funding_perceptions", "policy_general")
outcomes_subgroup     <- c(outcomes_illustrative, "newsletter_signup")
moderators_cat <- c("gender", "age_band", "race", "education", "income", "party")

outcome_label_map <- c(
  trust_multidimensional = "Multidimensional trust (primary)",
  trust_post             = "Single-item trust",
  distrust_post          = "Distrust",
  funding_perceptions    = "Funding perceptions",
  policy_role_mean       = "Policy role",
  inst_trust_mean        = "Institutional trust",
  belief_post            = "Climate belief",
  concern_mean           = "Climate concern",
  policy_general         = "General climate policy",
  policy_specific_mean   = "Specific climate policy",
  behavior_mean          = "Pro-climate behavior",
  donation_ams           = "Donation to AMS ($)",
  newsletter_signup      = "Newsletter signup"
)
```

```{r}
#| label: leaderboard-helpers

# Extract ATE pairs (estimate, SE, adjusted p) — same helper as the original
# preregistration's rq1-all chunk.
prep_ate <- function(df) {
  df |>
    select(condition, estimate) |>
    mutate(
      se    = if ("std.error"        %in% names(df)) df$std.error        else NA_real_,
      p_adj = if ("p.value_adjusted" %in% names(df)) df$p.value_adjusted else NA_real_
    )
}

control_lab <- levels(human_data$condition)[1]
z95 <- qnorm(0.975)

# Reference (Human 1) ATE models on the continuous outcomes, fit once and
# reused.
human_models <- map(outcomes_continuous,
                    ~ run_main_treatment_model(human_data_1, outcome = .x)) |>
  set_names(outcomes_continuous)

# One ATE side (estimate, SE, BH-adjusted p) for an individual-level dataset,
# across all pooled outcomes: OLS with HC2 SEs for the continuous outcomes and
# the donation; logistic marginal effects on the probability scale for the
# binary newsletter signup (as in v1's pooled row). The binary SE is recovered
# from the marginal-effect CI.
ate_side <- function(dat, models = NULL) {
  cont <- map_dfr(c(outcomes_continuous, "donation_ams"), function(out) {
    res <- if (!is.null(models) && out %in% names(models)) models[[out]]
           else run_main_treatment_model(dat, outcome = out)
    prep_ate(res) |> mutate(outcome = out)
  })
  bin <- run_main_treatment_model_binary(dat,
                                         outcome = "newsletter_signup")$marginal_effects |>
    transmute(condition,
              estimate,
              se      = (conf.high - conf.low) / (2 * z95),
              p_adj   = p.value_adjusted,
              outcome = "newsletter_signup")
  bind_rows(cont, bin)
}

human_side <- ate_side(human_data_1, models = human_models) |>
  rename(estimate_h = estimate, se_h = se, p_adj_h = p_adj)

# Common unit for all scoring: estimates and SEs in pp of each outcome's scale
# range (see `scale_range` in the variables chunk). Applied once to every set
# of joined pairs — the three tier builders, the floor baseline, and the
# subgroup pairs — so all pooled metrics, figures, and tables read one unit.
# A positive multiplicative factor: signs and p-values are untouched.
pp_scale_pairs <- function(pairs) {
  pairs |>
    mutate(across(any_of(c("estimate_h", "se_h", "estimate_l", "se_l")),
                  ~ .x * 100 / scale_range[outcome]))
}

# Tier 1: ATE pairs from an individual-level dataset, refitting the
# preregistered treatment models on the submission.
build_pairs_t1 <- function(pred_data) {
  ate_side(pred_data) |>
    select(condition, outcome, estimate_l = estimate) |>
    inner_join(human_side, by = c("condition", "outcome")) |>
    pp_scale_pairs()
}

# Tier 2: ATEs from submitted cell means (mean per condition x outcome),
# treatment cell minus control cell.
build_pairs_t2 <- function(cells_main) {
  cells_main |>
    mutate(condition = as.character(condition)) |>
    group_by(outcome) |>
    group_modify(function(cells, key) {
      ctrl <- cells |> filter(condition == control_lab)
      cells |>
        filter(condition != control_lab) |>
        transmute(condition, estimate_l = mean - ctrl$mean)
    }) |>
    ungroup() |>
    inner_join(human_side, by = c("condition", "outcome")) |>
    pp_scale_pairs()
}

# Tier 3: submitted ATEs taken as-is.
build_pairs_t3 <- function(effects) {
  effects |>
    transmute(condition = as.character(condition), outcome, estimate_l = ate) |>
    inner_join(human_side, by = c("condition", "outcome")) |>
    pp_scale_pairs()
}

# `pooled_metrics()`, `signed_metrics()`, and `cluster_boot()` are defined in
# R/functions/statistics.R and shown under "Scoring functions" above.

# One leaderboard row: pooled metrics + the calibration intercept, slope, and
# R^2 from the pooled calibration regression (all outcomes, HC2-robust
# inference: the outcome side carries known, condition-varying sampling
# error) + the share of expected intervention x outcome cells that entered
# scoring (cells that fail to parse are dropped at pair-building, so this is
# the usable-output share reported per submission).
leaderboard_row <- function(pairs, submission, tier, disclosure) {
  calib <- run_calibration_pooled(pairs)
  pooled_metrics(pairs, include_rmse = TRUE) |>
    mutate(submission = submission, tier = tier, disclosure = disclosure,
           alpha = calib$alpha, beta = calib$beta, calib_r2 = calib$r_squared,
           usable_pct = 100 * nrow(pairs) / nrow(human_side))
}
```

```{r}
#| label: mock-submissions

# Mock Tier 2 (cell-level) and Tier 3 (effect-level) submissions, generated
# outside this script alongside the other placeholder data (see
# R/simulate_data_preregistration.R) so that no mock-data simulation lives in
# the preregistration itself. Tier 2 collapses one placeholder team to the
# cell-level schema (*_cells_main.csv); Tier 3 gives ATEs vs. control.
cells_t2   <- readRDS(here("data/simulation_preregistration/mock_submission_tier2_cells.rds"))
effects_t3 <- readRDS(here("data/simulation_preregistration/mock_submission_tier3_effects.rds"))
```

```{r}
#| label: leaderboard-build

# 30 mock submissions plus the fixed human replication reference (Human 1 vs. Human 2).
# Teams 1-28 enter as individual-level Tier 1 data; team 29 as a Tier 2
# cell-level file and team 30 as a Tier 3 effect-level file, so all three tier
# code paths are exercised. Each submission's ATE pairs are built once here and
# reused by the headline and breakdown figures, the leaderboard table, and the
# companion tables.
tier1_ids <- 1:28

# Cached: the 29 Tier-1 builds each refit the treatment models incl. the
# logistic marginal effects for the newsletter outcome, the slow step of a
# cold render.
pairs_list <- cached("amendment_pairs_list", {
  c(
    set_names(
      map(tier1_ids,
          ~ build_pairs_t1(llm_data_teams |> filter(team == paste0("team_", .x)))),
      paste0("Team ", tier1_ids)
    ),
    list(
      "Team 29"           = build_pairs_t2(cells_t2),
      "Team 30"           = build_pairs_t3(effects_t3),
      "Human replication" = build_pairs_t1(human_data_2)
    )
  )
})

set.seed(2026)
submission_meta <- tibble(
  submission = c(paste0("Team ", 1:30), "Human replication"),
  tier       = c(rep("1", 28), "2", "3", "—"),
  disclosure = c(sample(c("A", "B", "C"), 30, replace = TRUE), "—")
)

# Floor baseline: a naive reference row for the metrics without a natural
# null (chiefly RMSE). "No effect" predicts a zero ATE everywhere — a fully
# naive, blinded-achievable strategy. It is a reference row like the human
# replication, not a member of the field: excluded from the field violins and
# the field summary. (An oracle-mean floor — predict each outcome's mean human
# ATE everywhere — was considered and rejected: the mean is the in-sample
# least-squares constant fit to the very ATEs being scored, so it competes
# rather than floors; its interpretive content survives as the RMSE scale
# reference, the SD of the human ATEs.)
floor_pairs <- list(
  "Floor: no effect" = human_side |>
    mutate(estimate_l = 0) |>
    pp_scale_pairs()
)
floor_labels <- names(floor_pairs)

# Calibration is undefined for the floor (a constant prediction ->
# zero-variance regressor), so its row carries NA there. Its directional %
# reads 0, not 50: sign(0) matches neither sign, an artifact of the all-zero
# prediction, not below-chance skill (footnoted in the table).
floor_rows <- imap_dfr(floor_pairs, function(pr, sub) {
  pooled_metrics(pr, include_rmse = TRUE) |>
    mutate(submission = sub, tier = "—", disclosure = "—",
           alpha = NA_real_, beta = NA_real_, calib_r2 = NA_real_,
           usable_pct = 100 * nrow(pr) / nrow(human_side))
})

leaderboard <- bind_rows(
  imap_dfr(pairs_list, function(pr, sub) {
    m <- submission_meta |> filter(submission == sub)
    leaderboard_row(pr, sub, tier = m$tier, disclosure = m$disclosure)
  }),
  floor_rows
) |>
  arrange(desc(submission == "Human replication"), desc(pearson_r))
```

```{r}
#| label: ate-ladder-labels

# Display labels / panel order for the metrics.
ladder_labels <- c(
  directional_pct = "Directional %",
  spearman_rho    = "Spearman ρ",
  pearson_r       = "Pearson r",
  pearson_adj     = "Pearson r (adj.)",
  rmse            = "RMSE (pp)",
  rmse_adj        = "RMSE (adj., pp)"
)
relabel_metrics <- function(df, labels) {
  df |> mutate(metric = factor(labels[metric], levels = labels))
}
```

#### Finding 1 — How well can the field predict the experiment?

The headline question is not *who wins* but how well the current generation of approaches can predict a novel experiment at all. To keep that readable, each question gets **one key metric in the figures**. Pearson *r* serves ATE recovery and, on the interaction estimates, subgroup recovery, and the variance ratio serves response shape. The correlation is the literature's common currency [@ashokkumarLargeLanguageModels2026a], and one shared axis across the ATE and subgroup sections makes them directly comparable. The variance ratio carries the field's one directional expectation (under-dispersion). All preregistered metrics are computed for everything and reported in the companion tables.

@fig-headline shows the raw prediction task with the field's calibration lines and coefficients (A) and the ranking of the field on the pooled Pearson *r* (B). One rendering note on panel B. This placeholder demonstration runs the bootstrap with a small replicate count (`n_boot_demo = 200`) to keep the document cheap to build. The **preregistered value for the real-data run is `n_boot_real = 2,000`**. When the submissions arrive, only that parameter changes.

```{r}
#| label: fig-headline
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "Headline cross-team figure (placeholder data). (A) Predicted against human ATEs across all outcomes and approaches, in percentage points of each outcome's scale range (as in the predicted-vs-observed figure of Ashokkumar et al.), one point per approach × intervention × outcome. Axes are centered on zero, the dashed line is the identity, and red crosses mark the human replication reference (Human 2 predicting Human 1). The solid lines are the average calibration regression of human on predicted ATEs, pooled across all outcomes, dark blue for the field (per-approach fits averaged) and red for the human replication reference. The corner prints the mean field Pearson r (with its noise-corrected value) and the human replication's, the same pooled metric ranked in B, together with the field and human calibration coefficients (slope β, perfect = 1, and intercept α, perfect = 0) with 95% CIs. (B) Approaches ranked by Pearson r pooled across all outcomes against Human 1. Filled points carry 95% cluster-bootstrap intervals over interventions, open points show the noise-corrected r (adj.) where its guard allows a value, and the human replication reference is pinned on top with its value dropped as a dotted line. The bottom row summarizes the field, with the density of the approaches' scores, a diamond at their mean with its 95% CI, and an open point at the mean noise-corrected r. On placeholder data every approach is a noise resample of the same dataset, so the field clusters near r = 0 in panel B, as expected of a random resample."

# (A) the raw prediction task: all approaches, all outcomes. The pairs are
# already in pp of scale range (pp_scale_pairs() at pair-building), the same
# unit as every pooled metric.
headline_pairs <- imap_dfr(pairs_list, ~ mutate(.x, submission = .y))
p_scatter <- plot_pred_scatter(
  headline_pairs,
  x_lab       = "Predicted ATE (pp of scale range)",
  y_lab       = "Human ATE (pp of scale range)",
  tag         = "A",
  point_alpha = 0.15, point_size = 0.5,
  ref_alpha   = 0.6,  ref_size   = 1,
  symmetric   = TRUE
)

# Average field / human calibration lines + a corner readout of correlation
# (r, adj.) and coefficients (β, α with CI). Mean field Pearson r and its
# noise-corrected value come from the pooled leaderboard — the identical pp
# quantity ranked in panel B; the coefficients are fit from the same pp pairs
# by calibration_overlay().
field_ld <- leaderboard |> filter(!submission %in% floor_labels,
                                  submission != "Human replication")
human_ld <- leaderboard |> filter(submission == "Human replication")
p_scatter <- p_scatter +
  calibration_overlay(headline_pairs,
                      r_field    = mean(field_ld$pearson_r,   na.rm = TRUE),
                      radj_field = mean(field_ld$pearson_adj, na.rm = TRUE),
                      r_human    = human_ld$pearson_r,
                      radj_human = human_ld$pearson_adj)

# (B) ranked field on the key metric, with cluster-bootstrap intervals.
# n_boot_demo here; the real-data run uses n_boot_real = 2,000 (see variables).
pooled_f <- \(p) pooled_metrics(p)
ate_forest <- cached("amendment_ate_forest", {
  imap_dfr(pairs_list, function(pr, sub) {
    cluster_boot(pr, pooled_f, cluster = "condition", n_boot = n_boot_demo) |>
      mutate(submission = sub)
  }) |>
    relabel_metrics(ladder_labels)
})
p_forest <- plot_field_forest_single(ate_forest, tag = "B")

(p_scatter | p_forest) + plot_layout(widths = c(1, 1.15))
```

Every pooled metric stays scored for every approach. @tbl-field-summary gives the field's center and spread on each.

```{r}
#| label: field-pooled-compute

field_pooled <- leaderboard |>
  filter(!submission %in% floor_labels) |>   # floors are reference rows, not field members
  select(submission, all_of(names(ladder_labels))) |>
  pivot_longer(-submission, names_to = "metric", values_to = "value") |>
  relabel_metrics(ladder_labels)
```

```{r}
#| label: tbl-field-summary
#| tbl-cap: "Field summary (placeholder data). Center and spread of each pooled ATE-recovery metric across the 30 approaches (the human replication reference excluded)."

summarise_field(field_pooled) |>
  gt() |>
  fmt_number(columns = c(mean, median, sd, min, max), decimals = 2) |>
  cols_label(metric = "Metric", mean = "Mean", median = "Median",
             sd = "SD", min = "Min", max = "Max")
```

##### Where the field predicts well

Split **by outcome** (one distribution per outcome, scored across the interventions) and **by intervention** (one per intervention, scored across the continuous outcomes, see *Scoring principle*), the key metric shows *where* the field predicts well and what resists prediction (@fig-field-by-group). The full metrics per outcome and per intervention are in @tbl-field-by-outcome and @tbl-field-by-intervention.

```{r}
#| label: field-by-group-compute

# include_rmse: with all pairs in pp, RMSE and its noise-corrected companion
# RMSE_adj are defined per outcome and pooled alike; here they localise where
# the absolute error concentrates.
field_by_outcome <- imap_dfr(pairs_list, function(pr, sub) {
  metrics_by_group(pr, "outcome",
                   f = \(p) pooled_metrics(p, include_rmse = TRUE)) |>
    mutate(submission = sub)
}) |>
  mutate(outcome = factor(unname(outcome_label_map[outcome]),
                          levels = unname(outcome_label_map))) |>
  pivot_longer(c(directional_pct, spearman_rho, pearson_r, pearson_adj,
                 rmse, rmse_adj),
               names_to = "metric", values_to = "value") |>
  relabel_metrics(ladder_labels)

# Restricted to the 0-100 continuous outcomes: with pp units the scale
# confound is gone, but a per-intervention correlation across all 13 outcomes
# would still mix constructs (attitude sliders vs. behaviors), so the
# conservative restriction is kept.
field_by_intervention <- imap_dfr(pairs_list, function(pr, sub) {
  metrics_by_group(pr |> filter(outcome %in% outcomes_continuous), "condition") |>
    mutate(submission = sub)
}) |>
  pivot_longer(c(directional_pct, spearman_rho, pearson_r, pearson_adj),
               names_to = "metric", values_to = "value") |>
  relabel_metrics(ladder_labels)

# Companion-table helper: field median per metric within each group level,
# with the human replication reference value in parentheses.
field_group_table <- function(field_long, group,
                              pct_metrics = "Directional %") {
  fmt <- function(v, pct) if_else(is.na(v), "—",
                                  if_else(pct, sprintf("%.0f", v),
                                          sprintf("%.2f", v)))
  med <- field_long |>
    filter(submission != "Human replication") |>
    group_by(across(all_of(group)), metric) |>
    summarise(med = median(value, na.rm = TRUE), .groups = "drop")
  ref <- field_long |>
    filter(submission == "Human replication") |>
    select(all_of(group), metric, ref = value)
  med |>
    left_join(ref, by = c(group, "metric")) |>
    mutate(pct  = metric %in% pct_metrics,
           cell = paste0(fmt(med, pct), " (", fmt(ref, pct), ")")) |>
    select(all_of(group), metric, cell) |>
    pivot_wider(names_from = metric, values_from = cell)
}
```

```{r}
#| label: fig-field-by-group
#| fig-width: 11
#| fig-height: 6.5
#| fig-cap: "Where the field predicts well (placeholder data). Distribution of each approach's Pearson r by outcome (A, scored across the interventions) and by intervention (B, scored across the continuous outcomes). One dot per approach over the field's half-violin, with the field mean and its 95% CI (white diamond) on each row. Red crosses mark the human replication reference, the dotted line the null."

by_outcome_r <- field_by_outcome |>
  filter(metric == "Pearson r") |>
  mutate(metric = factor("Outcomes"))
by_intervention_r <- field_by_intervention |>
  filter(metric == "Pearson r") |>
  mutate(metric = factor("Interventions"))

p_by_outcome <- plot_metric_distribution(
  by_outcome_r, x = "outcome", mean_ci = TRUE,
  ref_lines = c("Outcomes" = 0)
) + labs(tag = "A", y = "Pearson r")
p_by_intervention <- plot_metric_distribution(
  by_intervention_r, x = "condition", mean_ci = TRUE,
  ref_lines = c("Interventions" = 0)
) + labs(tag = "B", y = "Pearson r")

p_by_outcome | p_by_intervention
```

```{r}
#| label: tbl-field-by-outcome
#| tbl-cap: "Field summary by outcome (placeholder data). Median across approaches of each metric, scored per outcome across the interventions, with the human replication reference in parentheses. RMSE and RMSE (adj.) are in pp of scale range, the same unit as the pooled row. Per outcome they localize where the absolute error concentrates."

field_group_table(field_by_outcome, "outcome") |>
  gt() |>
  sub_missing(missing_text = "—") |>
  cols_label(outcome = "Outcome")
```

```{r}
#| label: tbl-field-by-intervention
#| tbl-cap: "Field summary by intervention (placeholder data). Median across approaches of each metric, scored per intervention across the continuous outcomes, with the human replication reference in parentheses."

field_group_table(field_by_intervention, "condition") |>
  gt() |>
  sub_missing(missing_text = "—") |>
  cols_label(condition = "Intervention")
```

##### Subgroup recovery

Whether an approach reproduces *who* an intervention moves, not just its average effect. For each Tier-1 submission we fit the preregistered condition × moderator interaction model (`run_moderator_model()`) across the demo outcome set (four outcomes plus the binary newsletter signup, the latter as a linear probability model, see *Estimation functions*; the registered analysis runs all outcomes — see the callout above) over the six demographic moderators, then pool the interaction estimates against Human 1. All interaction estimates are placed in percentage points of each outcome's scale range before pooling, the same unit as Section 1. Because interaction estimates require respondent-level data, this view covers Tier 1 only. As specified in Section 2 above, the scored metrics are directional agreement, Spearman $\rho$, and Pearson r, plus the adjusted correlation r~adj~, emphasized here because small subgroup cells make the reference noise uneven across cells (see *Adjusted correlation and RMSE (noise-corrected)*).

```{r}
#| label: subgroup-pairs

# Human-1 interaction estimates, computed once and reused for every submission.
# The cached() keys keep their historical "amendment_" prefix (the analysis is
# unchanged); they are internal identifiers for analysis/results/*.rds, not tied
# to this file's name, so renaming the document does not require a recompute.
# The binary newsletter outcome runs through the same OLS machinery as a
# linear probability model: in the saturated condition × moderator
# specification the interaction coefficient is exactly the
# difference-in-differences of cell signup proportions — the interaction on
# the probability scale — and the HC2 robust SEs are valid for a binary
# response. It then pp-rescales (× 100) like every other outcome.
human_mod_side <- cached("amendment_human_mod_side", {
  map_dfr(outcomes_subgroup, function(out) {
    map_dfr(moderators_cat, function(mod) {
      run_moderator_model(human_data_1, outcome = out, moderator = mod,
                          compute_predicted = FALSE)$interaction_effects |>
        # se_h feeds the adjusted correlation; only the human reference side
        # is corrected (see "Adjusted correlation and RMSE").
        select(condition, moderator_level,
               estimate_h = estimate, se_h = std.error) |>
        mutate(outcome = out, moderator = mod)
    })
  })
})

build_subgroup_pairs <- function(pred_data) {
  pred_side <- map_dfr(outcomes_subgroup, function(out) {
    map_dfr(moderators_cat, function(mod) {
      run_moderator_model(pred_data, outcome = out, moderator = mod,
                          compute_predicted = FALSE)$interaction_effects |>
        select(condition, moderator_level, estimate_l = estimate) |>
        mutate(outcome = out, moderator = mod)
    })
  })
  inner_join(human_mod_side, pred_side,
             by = c("condition", "moderator_level", "outcome", "moderator")) |>
    pp_scale_pairs()
}

# Tier-1 submissions plus the reference are eligible. Interaction pairs are built
# once per submission and reused by the subgroup figure and its companion table.
subgroup_inputs <- c(
  set_names(
    map(tier1_ids, ~ llm_data_teams |> filter(team == paste0("team_", .x))),
    paste0("Team ", tier1_ids)
  ),
  list("Human replication" = human_data_2)
)
subgroup_pairs_list <- cached("amendment_subgroup_pairs", {
  map(subgroup_inputs, build_subgroup_pairs)
})

# Context statistic for reading the subgroup scores (see Section 2): the share
# of human interaction estimates that are individually significant, from the
# normal approximation on the interaction SEs, BH-adjusted within outcome ×
# moderator. Near 5% (the false-positive rate) = the human data show little
# true moderation, and the subgroup metrics are read as uninformative.
human_int_sig_share <- human_mod_side |>
  group_by(outcome, moderator) |>
  mutate(p_adj = p.adjust(2 * pnorm(-abs(estimate_h / se_h)), method = "BH")) |>
  ungroup() |>
  summarise(share = mean(p_adj < .05, na.rm = TRUE)) |>
  pull(share)
```

@fig-subgroup-headline shows the moderation analogue of the headline figure. Panel A shows the raw prediction task on the interaction estimates, and panel B the key metric per moderator, showing which kinds of heterogeneity the field can predict at all. The full signed-effect metrics, pooled, are in @tbl-field-subgroup. As preregistered in Section 2, the share of human interaction estimates that are individually significant is reported here as context: on the placeholder data it is `r sprintf("%.0f%%", 100 * human_int_sig_share)`, and a value near the 5% false-positive rate on the real data will mean the human sample shows little true moderation, so the subgroup metrics cannot separate approaches and are read as uninformative rather than as evidence about any submission.

```{r}
#| label: fig-subgroup-headline
#| fig-width: 11
#| fig-height: 4.2
#| fig-cap: "Subgroup recovery (placeholder data, Tier 1). (A) Predicted against human condition × moderator interaction estimates pooled across all approaches and the demo outcome set (four outcomes plus the binary newsletter signup; the registered analysis runs all outcomes), in percentage points of each outcome's scale range. Axes are centered on zero, the dashed line is the identity, and red crosses mark the human replication reference. The solid lines are the average calibration regression of human on predicted estimates (dark blue the field, red the human replication), with a corner readout of the mean field and human Pearson r (with its noise-corrected value) and calibration coefficients (β and α with 95% CIs). (B) Distribution of each approach's Pearson r on the interaction estimates, per moderator (pooled across the demo outcome set), with the field mean and its 95% CI (white diamond) per row. Red crosses mark the reference, the dotted line the null."

moderator_label_map <- c(gender = "Gender", age_band = "Age",
                         race = "Race/ethnicity", education = "Education",
                         income = "Income", party = "Party ID")

# Pairs are already in pp of scale range (pp_scale_pairs() in the builder).
subgroup_scatter_pairs <- imap_dfr(subgroup_pairs_list,
                                   ~ mutate(.x, submission = .y))
# ~37,000 pooled pairs: much lighter marks than the ATE scatter, or it reads
# as a solid blob.
p_sub_scatter <- plot_pred_scatter(
  subgroup_scatter_pairs,
  x_lab       = "Predicted interaction estimate (pp)",
  y_lab       = "Human interaction estimate (pp)",
  tag         = "A",
  point_alpha = 0.05, point_size = 0.3,
  ref_alpha   = 0.2,  ref_size   = 0.5,
  symmetric   = TRUE
)

# Mean field / human Pearson r on the interaction estimates (from the pooled
# signed-effect ladder, matching panel B), and the calibration lines + readout.
subgroup_ld <- imap_dfr(subgroup_pairs_list,
                        ~ signed_metrics(.x) |> mutate(submission = .y))
sub_field_ld <- subgroup_ld |> filter(submission != "Human replication")
sub_human_ld <- subgroup_ld |> filter(submission == "Human replication")
p_sub_scatter <- p_sub_scatter +
  calibration_overlay(subgroup_scatter_pairs,
                      r_field    = mean(sub_field_ld$pearson_r,   na.rm = TRUE),
                      radj_field = mean(sub_field_ld$pearson_adj, na.rm = TRUE),
                      r_human    = sub_human_ld$pearson_r,
                      radj_human = sub_human_ld$pearson_adj)

field_subgroup_by_mod <- imap_dfr(subgroup_pairs_list, function(pr, sub) {
  pr |>
    group_by(moderator) |>
    group_modify(~ signed_metrics(.x)) |>
    ungroup() |>
    mutate(submission = sub)
}) |>
  transmute(submission,
            value      = pearson_r,
            moderator  = factor(unname(moderator_label_map[moderator]),
                                levels = unname(moderator_label_map)),
            metric     = factor("Moderators"))
p_sub_field <- plot_metric_distribution(
  field_subgroup_by_mod, x = "moderator", mean_ci = TRUE,
  ref_lines = c("Moderators" = 0)
) + labs(tag = "B", y = "Pearson r")

(p_sub_scatter | p_sub_field) + plot_layout(widths = c(1, 1.3))
```

```{r}
#| label: tbl-field-subgroup
#| tbl-cap: "Field summary of subgroup recovery (placeholder data, Tier 1). Center and spread across approaches of each signed-effect metric on the condition × moderator interaction estimates (in pp of scale range), pooled across the demo outcome set (four outcomes plus newsletter signup; the registered analysis runs all outcomes) and six moderators (human replication reference excluded)."

field_subgroup <- imap_dfr(subgroup_pairs_list, function(pr, sub) {
  signed_metrics(pr) |>
    pivot_longer(everything(), names_to = "metric", values_to = "value") |>
    mutate(submission = sub)
}) |>
  relabel_metrics(ladder_labels[c("directional_pct", "spearman_rho",
                                  "pearson_r", "pearson_adj")])

summarise_field(field_subgroup) |>
  gt() |>
  fmt_number(columns = c(mean, median, sd, min, max), decimals = 2) |>
  cols_label(metric = "Metric", mean = "Mean", median = "Median",
             sd = "SD", min = "Min", max = "Max")
```

##### Response-shape recovery

How well an approach reproduces the **shape** of the human response distribution, not only its mean. Using the preregistered `compare_distributions()` in the shared control condition, we score the overlap coefficient (OVL, 1 = identical densities), the **Wasserstein-1 distance** (W1, 0 = identical, computed as the integral of the absolute ECDF difference, in scale points), the Kolmogorov–Smirnov distance (KS, 0 = indistinguishable), and the variance ratio (clone/human, 1 = matched spread) for each continuous outcome (see *Distribution comparison metrics* for the definitions). The **variance ratio stays the headline diagnostic**, with the directional expectation of under-dispersion (ratio < 1). These are inherently per-outcome, control-condition quantities. @fig-distribution-headline therefore shows the raw response densities for every continuous outcome, one facet per outcome. Each facet stamps the field's median variance ratio and Wasserstein-1 distance next to the human replication's, so spread (VR) and overall distributional distance (W1) can be read at a glance for both the synthetic field and the human reference. All four metrics per outcome are in @tbl-field-distribution.

```{r}
#| label: distribution-pairs

dist_per_outcome <- function(pred_data) {
  map_dfr(outcomes_continuous, function(out) {
    compare_distributions(human_data_1, pred_data, out, control_lab,
                          include_w1 = TRUE) |>
      mutate(outcome = out)
  })
}

dist_labels <- c(ovl = "Overlap (OVL)", w1 = "Wasserstein-1",
                 ks_d = "KS distance", variance_ratio = "Variance ratio")

# Per-outcome shape metrics built once per submission, reused by the
# distribution figure and its companion table (cached: 29 x 11 KDE/ECDF
# passes over full respondent-level data).
dist_per_outcome_list <- cached("amendment_dist_per_outcome", {
  map(subgroup_inputs, dist_per_outcome)
})
```

```{r}
#| label: fig-distribution-headline
#| fig-width: 11
#| fig-height: 7
#| fig-cap: "Response-shape recovery (placeholder data, Tier 1). Control-condition response densities for every continuous outcome. In each facet, every Tier-1 approach is a thin line, Human 1 (the reference half) the bold line, and Human 2 (the replication half) the dashed red line. The corner stamp reports two numbers for that outcome, the variance ratio (VR, approach ÷ human, 1 = matched spread, below 1 = under-dispersion) and the Wasserstein-1 distance (W1, in scale points, 0 = identical distributions, larger = more different), for both the synthetic field (median across approaches, 'Field') and the human replication ('Human', Human 2 vs Human 1). VR captures spread, W1 captures how far the whole distribution sits from the human one in both location and shape. The full shape metrics per outcome are in the companion table."

# Per facet, stamp the field's median variance ratio and Wasserstein-1 distance
# alongside the human replication's (Human 2 vs Human 1), so each panel shows
# both spread (VR) and overall distributional distance (W1) for the synthetic
# field and the human reference.
field_dist_ann <- imap_dfr(dist_per_outcome_list,
                           ~ mutate(.x, submission = .y)) |>
  filter(submission != "Human replication") |>
  group_by(outcome) |>
  summarise(vr = median(variance_ratio, na.rm = TRUE),
            w1 = median(w1, na.rm = TRUE), .groups = "drop")

human_dist_ann <- dist_per_outcome_list[["Human replication"]] |>
  select(outcome, w1, vr = variance_ratio)

dist_ann <- left_join(field_dist_ann, human_dist_ann,
                      by = "outcome", suffix = c("_field", "_human")) |>
  mutate(label = sprintf("Field   VR %.2f   W1 %.2f\nHuman  VR %.2f   W1 %.2f",
                         vr_field, w1_field, vr_human, w1_human)) |>
  select(outcome, label)

# wrapped primary-outcome label: the full one clips its facet strip
density_labels <- outcome_label_map
density_labels["trust_multidimensional"] <- "Multidimensional trust\n(primary)"

plot_density_overlay(
  human_data_1, human_data_2,
  # Tier-1 entries only, matching the caption and the field stamps (teams 29
  # and 30 exist as individual-level placeholder data but enter as Tier 2/3).
  llm_data_teams |> filter(!team %in% c("team_29", "team_30")),
  outcomes       = outcomes_continuous,
  condition_val  = control_lab,
  outcome_labels = density_labels,
  annotations    = dist_ann,
  x_lab          = "Response (0–100, control condition)"
)
```

```{r}
#| label: tbl-field-distribution
#| tbl-cap: "Field summary of response-shape recovery by outcome (placeholder data, Tier 1). Median across approaches of each shape metric in the control condition, per continuous outcome, with the human replication reference in parentheses. Ideal values are OVL = 1, W1 = 0, KS = 0, and variance ratio = 1."

field_distribution <- imap_dfr(dist_per_outcome_list, function(d, sub) {
  d |>
    pivot_longer(c(ovl, w1, ks_d, variance_ratio),
                 names_to = "metric", values_to = "value") |>
    mutate(submission = sub)
}) |>
  mutate(outcome = factor(unname(outcome_label_map[outcome]),
                          levels = unname(outcome_label_map))) |>
  relabel_metrics(dist_labels)

field_group_table(field_distribution, "outcome", pct_metrics = character()) |>
  gt() |>
  sub_missing(missing_text = "—") |>
  cols_label(outcome = "Outcome")
```

#### Finding 2 — Ranking approaches (secondary)

Having characterized the field, we rank individual approaches. This is the secondary read. Each metric is its own ranking, and the interesting result is *which* approaches lead on *which* metric. The leaderboard **table** is the main summary. The ranking *visual* for the key metric, with its cluster-bootstrap uncertainty intervals, is panel B of @fig-headline.

```{r}
#| label: tbl-leaderboard
#| tbl-cap: "Cross-team leaderboard (placeholder data). Pooled ATE-recovery metrics across all outcomes (every estimate in percentage points of its outcome's scale range), plus the intercept, slope, and R² of the pooled calibration regression and the share of usable cells, scored against the Human 1 reference half. The Human replication row (Human 1 vs. Human 2) shows how well two human halves of this size agree. Metrics run least to most strict, left to right. Class is the disclosure-class badge (A open, B escrowed, C sealed)."

leaderboard |>
  select(submission, tier, disclosure, usable_pct, directional_pct,
         spearman_rho, pearson_r, pearson_adj, rmse, rmse_adj,
         alpha, beta, calib_r2) |>
  gt() |>
  fmt_number(columns = c(spearman_rho, pearson_r, pearson_adj,
                         rmse, rmse_adj, alpha, beta, calib_r2),
             decimals = 2) |>
  fmt_number(columns = c(usable_pct, directional_pct), decimals = 0) |>
  sub_missing(missing_text = "—") |>
  cols_label(
    submission      = "Submission",
    tier            = "Tier",
    disclosure      = "Class",
    usable_pct      = "Usable cells %",
    directional_pct = "Directional %",
    spearman_rho    = "Spearman ρ",
    pearson_r       = "Pearson r",
    pearson_adj     = "Pearson r (adj.)",
    rmse            = "RMSE (pp)",
    rmse_adj        = "RMSE (adj., pp)",
    alpha           = "Calib. α",
    beta            = "Calib. β",
    calib_r2        = "Calib. R²"
  ) |>
  tab_footnote(
    footnote = "Share of expected intervention × outcome cells that parsed and entered scoring. Cells that fail to parse are dropped at pair-building.",
    locations = cells_column_labels(columns = usable_pct)
  ) |>
  tab_footnote(
    footnote = "The floor's directional % reads 0, not 50, because sign(0) matches neither sign. This is an artifact of the all-zero prediction, not below-chance skill.",
    locations = cells_body(columns = directional_pct,
                           rows = submission %in% floor_labels)
  ) |>
  tab_style(
    style     = cell_fill(color = "#eef3f7"),
    locations = cells_body(rows = submission == "Human replication")
  ) |>
  tab_style(
    style     = cell_fill(color = "#f4f4f4"),
    locations = cells_body(rows = submission %in% floor_labels)
  )
```

##### Cross-metric consistency

A single-metric verdict can mislead. An approach that leads on one metric may trail on another, and @cumminsThreatAnalyticFlexibility2025b show empirically that performance on one data feature routinely fails to transfer to others. Rather than leave that to a reader's cross-panel comparison, we report it directly as the Spearman correlation between the approaches' rankings on each pair of metrics (reference and floor rows excluded). High values mean the metrics sort the field similarly, so a summary rank would be defensible. Low or negative values mean each metric tells its own story and no single number should be trusted alone.

```{r}
#| label: tbl-metric-consistency
#| tbl-cap: "Cross-metric rank consistency (placeholder data). Spearman correlations between the approaches' rankings on each pair of pooled metrics (reference and floor rows excluded). On placeholder data all scores are noise, so consistency is expected to be near zero."

metric_cols <- names(ladder_labels)

rank_consistency <- leaderboard |>
  filter(submission != "Human replication", !submission %in% floor_labels) |>
  select(all_of(metric_cols)) |>
  cor(method = "spearman", use = "pairwise.complete.obs")

rank_consistency |>
  as_tibble(rownames = "metric") |>
  mutate(metric = unname(ladder_labels[metric])) |>
  rename_with(~ unname(ladder_labels[.x]), all_of(metric_cols)) |>
  gt() |>
  fmt_number(columns = -metric, decimals = 2) |>
  sub_missing(missing_text = "—") |>
  cols_label(metric = "")
```

#### Design analysis of what the metrics can resolve at this sample size

Before locking the metrics, we check what they can resolve at this design, meaning how well even a genuinely good submission could score. The check uses design constants and literature priors only and touches no data. The inputs are the design's cell sizes (≈500 per intervention and ≈1,000 in control per human half), a typical 0–100 slider SD of ≈25 points, and plausible spreads of true megastudy ATEs (SDs of roughly 0.5–2 points, since large message-based megastudies rarely move attitude scales by more). Two conclusions follow. First, the reference half-sample ATE SE is ≈1.4 points, the noise floor under every raw metric. Second, we can compute the expected human-replication correlation and the expected correlation of a *perfect*, noiseless predictor for each plausible true-ATE spread. The perfect predictor's is the square root of the replication row's. When the true spread is small relative to 1.4, even a perfect predictor's raw correlation is modest, and the r~adj~ NA-guard is expected to bind for many entries.

```{r}
#| label: tbl-design-analysis
#| tbl-cap: "Blinded design analysis (no data touched). Expected values of key scoring quantities from design constants (cell sizes ≈500/1,000 per half, outcome SD ≈25) under three plausible spreads of true ATEs. r (replication) is the expected human-replication correlation. r (perfect) is the expected raw correlation of an error-free predictor, the square root of the replication value."

sd_y     <- 25
n_arm_h  <- 500
n_ctrl_h <- 1000
se_ate_h <- sd_y * sqrt(1 / n_arm_h + 1 / n_ctrl_h)

tibble(tau = c(0.5, 1, 2)) |>
  mutate(
    se_reference    = se_ate_h,
    r_replication   = tau^2 / (tau^2 + se_ate_h^2),
    r_perfect       = sqrt(r_replication)
  ) |>
  gt() |>
  fmt_number(columns = -tau, decimals = 2) |>
  cols_label(
    tau           = "SD of true ATEs (pts)",
    se_reference  = "Reference SE (pts)",
    r_replication = "r (replication)",
    r_perfect     = "r (perfect predictor)"
  )
```

#### Robustness of the scoring pipeline

The scoring pipeline has researcher degrees of freedom of its own, chiefly the preregistered random seed that splits the human sample into Human 1 and Human 2. @cumminsThreatAnalyticFlexibility2025b show that such analytic choices can move fidelity conclusions substantially, and a skeptic could reasonably ask whether the reported reference value reflects the split we happened to draw. We therefore run a **resplit sensitivity check**. The half-split is redrawn many times with different seeds (`n_resplit_real = 200` for the real-data run; this placeholder render uses `n_resplit_demo = 50` to keep the build cheap, and as with the bootstrap only that parameter changes when the submissions arrive), the human replication reference is recomputed on each with the same 13-outcome pipeline as the leaderboard, and the spread of the reference's metrics across resplits is reported alongside the primary (preregistered-seed) values. A narrow spread shows that the seed is not driving the results. When real data arrive, the same check is run for the full field. Every submission is rescored against each resplit's Human 1, and the stability of the field summary (and of leaderboard ranks) across resplits is reported.

```{r}
#| label: tbl-resplit
#| tbl-cap: "Split-seed sensitivity of the human replication reference (placeholder data, 50 resplits; the real-data run uses the preregistered 200). Distribution of each pooled ATE-recovery metric for the Human 1 vs. Human 2 comparison across the alternative random half-splits, next to the value under the preregistered seed. On placeholder data the reference is itself noise-level, so the check demonstrates the code, not a substantive reference value. The resplits are drawn from the same sample and so are mutually dependent. The spread quantifies split-conditional variability, not independent replications."

resplit_ceiling <- cached("amendment_resplit_ceiling", {
  all_ids <- unique(human_data$id)

  map_dfr(seq_len(n_resplit_demo), function(i) {
    set.seed(1000 + i)
    ids_i <- sample(all_ids, size = floor(length(all_ids) / 2))
    h1_i  <- human_data |> filter(id %in% ids_i)
    h2_i  <- human_data |> filter(!id %in% ids_i)

    # Same 13-outcome pipeline as the leaderboard: ate_side() covers the
    # continuous outcomes, the donation, and the binary newsletter signup;
    # pp_scale_pairs() puts every estimate in pp of scale range.
    inner_join(
      ate_side(h1_i) |>
        rename(estimate_h = estimate, se_h = se, p_adj_h = p_adj),
      ate_side(h2_i) |>
        select(condition, outcome, estimate_l = estimate),
      by = c("condition", "outcome")
    ) |>
      pp_scale_pairs() |>
      pooled_metrics(include_rmse = TRUE) |>
      mutate(resplit = i)
  })
})

resplit_ceiling |>
  pivot_longer(-resplit, names_to = "metric", values_to = "value") |>
  filter(metric %in% names(ladder_labels)) |>
  group_by(metric) |>
  summarise(mean = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE),
            min = min(value, na.rm = TRUE),  max = max(value, na.rm = TRUE),
            n_defined = sum(!is.na(value)), .groups = "drop") |>
  left_join(
    pairs_list[["Human replication"]] |>
      pooled_metrics(include_rmse = TRUE) |>
      pivot_longer(everything(), names_to = "metric",
                   values_to = "preregistered_seed"),
    by = "metric"
  ) |>
  mutate(metric = factor(unname(ladder_labels[metric]),
                         levels = unname(ladder_labels))) |>
  arrange(metric) |>
  gt() |>
  fmt_number(columns = c(mean, sd, min, max, preregistered_seed), decimals = 2) |>
  sub_missing(missing_text = "—") |>
  cols_label(metric = "Metric", mean = "Mean", sd = "SD", min = "Min",
             max = "Max", n_defined = "N defined",
             preregistered_seed = "Preregistered seed")
```

When real submissions arrive, both findings are recomputed by replacing the placeholder datasets with the submitted prediction files routed through their tier's builder (`build_pairs_t1()` / `build_pairs_t2()` / `build_pairs_t3()`). The subgroup and distribution figures are produced for the tiers eligible for them, alongside the demographic-baseline and stereotyping diagnostics of Section 3, including the demographic parity gap (`demographic_parity_gap()`).

