---
title: "The Silicon Sample Benchmark: A multi-team preregistration for predicting behavioral-experiment effects"

abstract: |
    This is the preregistration for the **Silicon Sample Benchmark**, a multi-team prediction benchmark. We invite independent research teams to predict the results of a novel behavioral megastudy (16 message interventions to increase trust in climate scientists, 13 preregistered outcomes, a representative U.S. sample of ~18,000) using any AI-based approach they choose—behavioral clones, group-level simulation, or direct effect forecasting. Submissions enter at one of **three tiers** defined by the granularity of the prediction (individual-level synthetic data, cell-level statistics, or effect-level estimates with uncertainty intervals), and every approach is scored on the same preregistered metric ladder against the same human replication reference. This document is **self-contained**: it restates the full analysis plan so that a participating team never needs to consult the project's earlier, single-team preregistration (v1); the section *About this preregistration* narrates what it keeps from v1 and what it changes. At the time of writing, human data are partially collected (~10,000 of the ~18,000 responses); none of these data have been shared, published, or made accessible to any prospective team, and no team will be granted access before the prediction lock. Before any human data are released, every team registers its approach using a standardized method-registration form (an extended version of the GUIDE-LLM reporting checklist) and deposits its predictions on Zenodo—publicly, or partly escrowed under a disclosure policy that keeps every entry locked and verifiable even where it cannot be fully public.

date: "2026-06-11"

format:
  html:
    code-fold: true
    toc: true
    toc-depth: 4
  # PDF disabled for now to speed up iteration — re-enable for the final render.
  # pdf:
  #   toc: true
  #   toc-depth: 5
  #   include-before-body:
  #     text: |
  #       \clearpage
  #   colorlinks: true
  #   tbl-colwidths: true
  #   header-includes:
  #     - \usepackage{float}
  #     - \floatplacement{figure}{H}
  #     - \floatplacement{table}{H}
  #     - \usepackage{listings}
  #     - \lstset{breaklines=true, basicstyle=\small\ttfamily}

execute:
  echo: true
  warning: false
  message: false
  fig-width: 8
  fig-height: 5
  fig-dpi: 300
---

```{r}
#| label: setup
#| output: false

# packages
library(tidyverse)
library(sandwich)
library(lmtest)
library(marginaleffects)
library(broom)
library(MetBrewer)
library(gghalves)
library(patchwork)
library(gt)
library(tinytable)
library(here)

# custom functions (reused unchanged from the original preregistration)
source(here("R/functions/statistics.R"))
source(here("R/functions/plots.R"))
source(here("R/functions/tables.R"))
source(here("R/functions/reporting.R"))

# simulated (random) placeholder data
human_data     <- readRDS(here("data/simulation_preregistration/human_data_preregistration.rds"))
llm_data_teams <- readRDS(here("data/simulation_preregistration/llm_data_preregistration_teams.rds"))

# Restrict the human data to the scored conditions (the 16 non-interactive
# interventions + control) BEFORE any model is fit. The real cleaned human
# data will additionally contain the three chatbot arms and the value-
# similarity arm, which no submission predicts; fitting the human models on
# those arms would put the human side's BH adjustment over a different test
# family than the submissions'. A no-op on the placeholder data (which has
# scored conditions only), load-bearing when the real loader is wired in.
scored_conditions <- levels(droplevels(factor(llm_data_teams$condition)))
human_data <- human_data |>
  filter(condition %in% scored_conditions) |>
  mutate(condition = droplevels(condition))

# Same preregistered random split of the human sample as in the original
# preregistration. Human 1 is the reference half all submissions are scored
# against; Human 2 is the human replication reference.
set.seed(42)
split_ids    <- sample(unique(human_data$id), size = floor(length(unique(human_data$id)) / 2))
human_data_1 <- human_data |> filter(id %in% split_ids)
human_data_2 <- human_data |> filter(!id %in% split_ids)
```


::: {.callout-warning}
**This is an ongoing draft, to share with collaborators. This preregistration is not yet registered and will change still.**
:::

::: {.callout-important}
**This is the canonical preregistration for the benchmark.** It is self-contained: the full analysis plan is restated here (see *Preregistered analysis plan*), so a participating team does not need to open the project's earlier single-team preregistration ([v1](preregistration.qmd)). The section *About this preregistration* explains why this second preregistration exists, what it keeps from v1, and every way it changes.

The benchmark's public-facing design—motivation, rules, the three tiers, the full method-registration form, the disclosure policy, scoring, and the timeline—is also described in the [call for participation](index.html). To avoid duplication, this document summarizes that design only as far as the analysis needs it, and concentrates on the **preregistered analysis plan**. Where the two overlap, the call is the canonical description of the procedure and this document is the canonical description of the analysis.

All datasets used for demonstration here are simulated random noise with the correct structure.
:::

## About this preregistration

**Why a second preregistration.** The project began as a single-team simulation study, preregistered as *"Predicting the effects of behavioral interventions on humans with LLMs"* (the original preregistration, **v1**). That plan asked how well one approach—a group-level simulation built by the core team from demographically representative U.S. profiles with a particular model and prompt—recovers a novel megastudy. A single team's result, however, is only one point in a large design space on how to build a synthetic sample. This document reconceives the project as a **multi-team benchmark**: many independent teams predict the same novel experiment with any AI-based approach, and every approach is scored identically. It is now the **canonical, self-contained** plan for the project, superseding the original **v1** preregistration.

**What it keeps from v1.** The scientific target and the analytical backbone are unchanged. The estimands (average treatment effects per intervention × outcome, subgroup interactions, control-condition response distributions, and demographic baselines), the three-section analysis structure, the least-to-most-strict metric ladder, and the preregistered split-half **human replication reference** are all carried over from v1 and **restated in full below** (see *Preregistered analysis plan*) rather than referenced by pointer. The model and comparison functions in `R/functions/statistics.R` are reused unchanged.

**What it changes and adds.** This preregistration is written for teams who are **blinded to the human outcome data**. Revising the analysis plan therefore does not carry the usual risk of data-contingent researcher degrees of freedom. The changes fall into three groups — how the study is reconceived, what it now reports, and a set of scoring add-ons drawn from the recent silicon-sample literature. Each is summarized here at a high level, with a pointer to the section that specifies it in full.

**A. From one approach to a field.** These reconceive the *design*; they leave v1's analysis pipeline untouched.

1. **Single team → multi-team benchmark.** The project now scores a *field* of independent submissions rather than one core-team approach, through three submission tiers defined by prediction granularity (Tier 1 individual-level synthetic data, Tier 2 cell-level statistics, Tier 3 effect-level estimates with uncertainty intervals). See *Benchmark design*.
2. **Method registration and disclosure.** Every team documents its pipeline on a standardized method-registration form (an extended GUIDE-LLM checklist) and deposits its predictions under a disclosure policy that keeps even proprietary entries verifiable. See *Method registration and disclosure*.
3. **Per-tier submission specifications.** Each tier has an exact file schema, and the shared estimands are recovered from whatever a tier supplies. For instance, effect standard errors are reconstructed from the uncertainty intervals that Tiers 2–3 submit. See *Submission specification* and *Tier-specific estimand recovery*.

**B. What the benchmark reports.** These sit on top of v1's pipeline; no estimand, model, or metric inherited from v1 is redefined.

4. **The headline result is the field, not a single scorecard.** The primary result is no longer one approach's report card but how the *whole field* of approaches fares at predicting human behavior. The main purpose is to present the distribution of predictive quality across the field. Every approach is scored against a split-half **human replication reference** (how well one half of the human sample predicts the other). Because teams may enter more than once, the field summary counts one primary entry per team; and because a team may choose to predict only part of the study rather than every intervention and outcome, the main leaderboard is limited to *full-coverage* entries — those that predicted the complete set — so that its scores are all measured on the same basis, with partial entries reported separately. See *Cross-team comparison*.
5. **One score per metric, pooled across outcomes.** A single data point is one predicted effect placed next to the matching human effect — for example, how much a particular message shifted donations to the science organization. For each metric on the ladder, all such comparisons — every intervention paired with every outcome — are pooled into one number, which uses all of them at once rather than scoring each outcome on its own. We deliberately do *not* collapse the metrics into one overall "champion" score: an approach can get the direction of effects right while mis-scaling their size, or nail the main outcome while failing on the rest, and a single number would hide exactly the structure the benchmark exists to reveal. Each metric is therefore read on its own. See *Cross-team comparison* (*Scoring principle*).
6. **An uncertainty interval on every score.** The benchmark tests 16 interventions, but these are only a sample of the messages one could have tested — a different study could have used 16 others of the same kind. An approach's score is therefore an estimate, and like any estimate it would move somewhat under a different draw of interventions. We attach an interval to every score showing how much it would move, so that a small gap between two approaches is not mistaken for a real difference when it sits within that wobble. For the same reason, we never call one approach better than another from the ranking alone: the claim has to pass a preregistered test that compares the two directly. See *Uncertainty intervals on the leaderboard*.

**C. Evaluation add-ons from the silicon-sample literature.** The remaining additions extend *how* an approach is scored, each motivated by how recent work validates synthetic samples. Every one is new to this preregistration, supplements rather than replaces an inherited metric, and points to the section where it is specified in full.

7. **A noise-corrected correlation.** The human effects an approach is scored against are themselves estimated from a finite (here, half-split) sample, so they carry sampling noise that mechanically drags every correlation down — penalizing all approaches, and the reference itself, for something unrelated to predictive skill. Following @hewittPredictingResultsSocial2025, we report a noise-corrected correlation *beside* the raw one (the raw value stays the ranked metric). See *Comparison functions*, and *Section 2*, where small demographic cells make the correction matter most.
8. **Did it pick the right interventions?** The correlation rungs summarize the whole effect pattern, but the decision a many-arm megastudy exists to inform is narrower — *which* interventions are worth acting on. We score that directly: are an approach's top-ranked interventions the ones with the largest true effects [@hewittPredictingResultsSocial2025]? See *Section 1*.
9. **A floor for the rungs without a natural zero.** Some rungs — inferential agreement, equivalence, and RMSE — have no natural zero to read them against, and on inferential agreement a lazy "predict no effect for everything" strategy scores deceptively well when most true effects are small. Adding a naive "do-nothing" row makes these scores interpretable: each approach is then read as the distance between that floor and the human reference [@hewittPredictingResultsSocial2025; @cumminsThreatAnalyticFlexibility2025b]. See *Comparison functions*.
10. **A fixed bound for the equivalence test.** One rung asks whether a predicted effect is *close enough* to the human effect to count as equivalent, which needs a bound for "close enough." v1 set that bound as a fraction of the human effect being tested — but that effect is itself measured with error, so the bound moved with the noise and the test was not valid. We now fix the bound in advance, at a set fraction of each outcome's scale, so it no longer depends on the numbers under test. See *Comparison functions*.
11. **Is an approach consistent across metrics?** Cummins' central finding is that a configuration strong on one measure is routinely weak on another, so any single-metric verdict misleads; we report how stable each approach's rank is across the whole ladder [@cumminsThreatAnalyticFlexibility2025b]. See *Cross-team comparison*.
12. **A fairness gap, and under-dispersion foregrounded.** Averaging baseline accuracy over demographic groups can hide an approach that serves most groups well but consistently misses one, so we report the gap between the worst- and best-served group [@parkLLMAgentsGrounded2026]. And because too-similar synthetic responses (under-dispersion) are the failure mode the literature flags most often, we promote it to the headline distribution diagnostic. See *Section 3*, and *Section 1* (response distributions).
13. **Does the result depend on an arbitrary split?** The human sample is divided in half at random to build the reference. That split is fixed before any data are seen, so it is not a researcher degree of freedom — but the particular split we drew is still arbitrary, and another random split would serve just as well. To show the results do not hinge on the one we happened to draw, we recompute them across many alternative splits and report how stable they are [@cumminsThreatAnalyticFlexibility2025b]. See *Cross-team comparison*.
14. **A stated scope: populations, not individuals.** The benchmark cannot assess individual-level fidelity — teams build their own synthetic respondents with no pairing to specific humans — so its distribution metrics are claims about the *population* response shape, not about matching any person. We say so plainly, because aggregate agreement can coexist with wrong individual behavior [@parkLLMAgentsGrounded2026; @arrudaCollectiveCooperationIndividual2026]. See *Overview* (scope statement).
15. **Scoring the uncertainty intervals themselves.** Teams in Tiers 2–3 submit not only a predicted effect but an uncertainty interval around it, and we use that interval to decide whether the effect counts as statistically significant (does the interval exclude zero?). This opens a loophole: a team that submits very wide intervals will have every interval include zero, so every prediction reads "not significant" — and since most effects in behavioral experiments are expected to be relatively small, matching "not significant" everywhere might earn a high agreement score with no real skill, just by being vague. To close the loophole we also score the intervals directly: whether they are well-calibrated (containing the true effect about as often as they claim to) and appropriately narrow. Being vague then carries its own cost. See *Tier-specific estimand recovery*.

Everything in groups B and C is **additive**: each item adds to an inherited metric without removing or redefining one. Two things do change from v1, both only for the cross-team comparison and both explained above. First, the equivalence test now uses a bound fixed in advance (item 10). Second, we keep v1's split-half human comparison exactly as it was, and change only how we describe it: v1 called it a "ceiling," implying no approach could do better, whereas we treat it as a reference an approach can beat, and rename it accordingly (item 4).


## Motivation

Whether large language models can stand in for human participants is one of the most consequential open questions in behavioral science, and the evidence so far is mixed and hard to interpret: most existing evaluations target experiments that were already published—and plausibly in the models' training data—and almost none are preregistered, so reported successes are entangled with researcher degrees of freedom in model and prompt choice [@cumminsThreatAnalyticFlexibility2025b]. 

We therefore aim to provide a preregistered, multi-team prediction benchmark, the *Silicon Sample Benchmark*: independent teams predict the same novel megastudy (N ≈ 18,000) with any AI-based approach, and every approach is scored on the same preregistered metric ladder. The design removes issues of existing evaluations: (i) the target is genuinely novel (none of the interventions has ever been tested in this form), and (ii) every prediction is deposited and time-stamped before any human data are released. The full rationale, rules, and incentives for participating teams are laid out in the [call for participation](index.html).

The benchmark scores the 16 non-interactive text interventions of the megastudy^[The original experiment contains 20 conditions; the benchmark scores only the 16 non-interactive text interventions. The four left out are the three interactive LLM-chatbot arms and the "Value similarity" quiz, all of which depend on a live back-and-forth that would substantially complicate the simulation task.], across self-reported attitudes (e.g., trust in climate scientists, support for funding of climate science) and behavioral measures (a donation to a scientific organization, a newsletter sign-up).

@tbl-interventions gives an overview of the 16 interventions, grouped by their content. All are short text messages (e.g., an article, an interview excerpt, corrected statistics, a portrait) shown once before the outcome battery. Together with the shared control condition they make up the 17 conditions every submission needs to cover.

```{r}
#| label: tbl-interventions
#| tbl-cap: "The 16 text-based interventions scored in the benchmark, grouped by the broad angle each takes on trust in climate scientists. The **Code name** column gives the internal condition identifier used in the raw survey files (`survey.qsf` / `survey_definition.json`), for teams that parse those directly. Each intervention has exactly one code name; where intervention-author teams were combined, the code name is a semicolon-joined list of animal pairs, and the semicolons are part of the name — match on the full string, never split it (see `survey/condition_codenames.csv` in the benchmark repository). The original megastudy also fields three interactive LLM-chatbot arms and a 'Value similarity' quiz; these four are interactive and excluded from the simulation benchmark, leaving these 16 plus a shared control condition (17 conditions in total)."

interventions <- read_csv(here("data/interventions.csv"), show_col_types = FALSE)

tag_order <- c(
  "Collaboration and peer-review",
  "Scientific methods and results",
  "Applications and impact",
  "Others' endorsement",
  "Values",
  "Other"
)

# Same table layout as v1: category headers inserted
# as bold full-width rows, with a running intervention number.
intervention_table_data <- interventions |>
  mutate(across(c(title, summary, tag, code_name), str_trim)) |>
  filter(tag != "LLM-chatbot", title != "Value similarity") |>
  mutate(tag = factor(tag, levels = tag_order)) |>
  arrange(tag, title) |>
  mutate(number = as.character(row_number())) |>
  # internal condition code names exactly as in the raw survey files
  # (survey.qsf / survey_definition.json); semicolon-joined names are a
  # single identifier — semicolons are part of the name, never split.
  mutate(codename = code_name) |>
  group_by(tag) |>
  group_modify(~ bind_rows(
    tibble(title = as.character(.y$tag), codename = "", summary = ""),
    .x
  )) |>
  ungroup() |>
  mutate(` ` = ifelse(title %in% levels(tag), "", number)) |>
  select(` `, title, codename, summary) |>
  rename(`Intervention Title` = title, `Code name` = codename, Summary = summary)

# identify header rows
header_rows <- which(intervention_table_data$` ` == "")

intervention_table_data |>
  tt(width = c(0.04, 0.19, 0.17, 0.60)) |>
  format_tt(escape = TRUE) |>
  style_tt(align = "l", alignv = "t") |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = header_rows, bold = TRUE, j = 2, colspan = 3) |>
  style_tt(j = 3, monospace = TRUE) |>
  style_tt(fontsize = 0.85)
```

The first goal of the benchmark is *not* to crown a single champion team, but to map the current design space of synthetic sample approaches, and evaluating their capacity in predicting human behavior as a field. This benchmark study will provide a **catalogue**: a structured, comparable description of how research teams currently approach AI prediction of behavioral experiments—open- vs. closed-weight models, persona sources, administration protocols, fine-tuning, in-context learning—and which of these choices appear to matter. We will evaluate approaches on a ladder of metrics of increasing strictness—from directional agreement to formal equivalence—plus distributional fidelity and stereotyping diagnostics (all specified in full under *Preregistered analysis plan*, below). We expect different approaches to win on different rungs.

## The human target sample

Predictions are scored against the human data from the megastudy itself. Participants are recruited from a non-probability opt-in panel, quota-matched to the population of U.S. residents, all aged 18 or older, who pass a series of attention and bot-detection checks before treatment assignment.

The benchmark targets **N ≈ 18,000 complete responses**: 1,000 per intervention across the 16 text-based interventions, plus 2,000 in the shared control condition. Recruitment uses **census-based cross quotas** on gender × age and gender × race/ethnicity to approximate the U.S. adult population; the targets below match the proportions of the parent megastudy, with counts simply rescaled to a total N ≈ 18,000. Participants selecting "Other" as their gender are not quota-constrained.

The targets below are drawn from the 2024 U.S. Census Bureau Population Estimates Program. *Total* is the share within each category; *Male* and *Female* give the within-category gender split. The table describes the key demographics of the human sample — participating teams are not required to use this information for their own synthetic sample.

```{r}
#| label: tbl-quotas
#| tbl-cap: "Sampling quota targets (N ≈ 18,000), from the 2024 U.S. Census Bureau Population Estimates Program. Proportions match the parent megastudy, with counts rescaled to N ≈ 18,000, while the target of the megastudy is N≈ 22,000."

# Same layout as the interventions table: category headers inserted as bold
# full-width rows above their levels (tinytable, for a consistent look).
quotas <- read_csv(here("data/quotas_18000.csv"), show_col_types = FALSE)

quota_table_data <- quotas |>
  group_by(Variable) |>
  group_modify(~ bind_rows(
    tibble(Category = as.character(.y$Variable), Total = "", Male = "", Female = ""),
    .x
  )) |>
  ungroup() |>
  select(Category, Total, Male, Female) |>
  rename(` ` = Category)

header_rows_q <- which(quota_table_data$Total == "")

quota_table_data |>
  tt(width = c(0.40, 0.20, 0.20, 0.20)) |>
  format_tt(escape = TRUE) |>
  style_tt(align = "l", alignv = "t") |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = header_rows_q, bold = TRUE) |>
  style_tt(fontsize = 0.85)
```

Gender, age, and race are the only mandatory survey items (needed to fill the quotas) and so enter every analysis model as covariates with no missing data. The six demographic moderators used for the subgroup and calibration analyses—gender, age, race, education, income, and partisan identity—are specified under *Preregistered analysis plan* (below) and in the codebook; full sampling and exclusion detail is in the [human study preregistration](https://doi.org/10.5281/zenodo.20160212).

## Benchmark design

Participation is by invitation. Each team builds an AI-based prediction pipeline from the shipped materials (intervention texts, the full survey instrument, the codebook, the registration template, and the validator). 

Teams can use any models, personas, prompts, fine-tuning, retrieval, or prior literature they want. They will receive **no participant pool** (each team builds its own profiles) and **never any human outcome data**. Before the deadline for the prediction lock, each team registers its approach by depositing their synthetic sample/predictions and materials on Zenodo, hashed and time-stamped. After the prediction lock deadline (and once the human data collection is finished), every submission runs through the preregistered analysis pipeline unchanged and is reported on a public leaderboard alongside the human replication reference, and in a joint manuscript with all participating teams as co-authors. Three ground rules define a valid entry and matter for the analysis: **blinding is absolute**, i.e. no team accesses, solicits, or is shown any human outcome data before the lock (at registration time only ~10,000 of the planned ~18,000 responses had been collected and none released), pipelines are **AI-based and automated**, and **everything is registered though not necessarily public** (see the disclosure policy below). The full procedure, ground rules, timeline, and incentives are in the [call for participation](index.html).

**Organizer access and firewall.** Blinding constrains the organizers too, so we state who has seen what. Interim human data (the ~10,000 responses collected at the time of writing) have been accessed only by the research lead, Jan Pfänder, for data-quality monitoring. As a result, members or affiliates of the core team who have had **no access to any human outcome data** may enter the benchmark as regular teams under the same rules as everyone else (same registration form, same blinding attestation, same deadlines). 

For full disclosure, Jan Pfänder presented some very preliminary (human megastudy sample at the time N ≈ 3000) results from an initial simulation attempt (following preregistration **v1**) in a 5min talk at a closed academic workshop at the Max Planck Institute for Human Development in Berlin (May 2026). This is the only presentation that has targeted an audience of potential collaborators on the benchmark study, and it has focused on prediction abilites of the simulation approach, without revealing anything about how humans behaved in the megastudy. Two other talks were explictly about the human megastudy results: Jan Pfänder presented preliminary results from the human megastudy at an invited talk at the Max Planck Institute for Human Development in Berlin (May 2026, N at the time ≈ 3000 of 22000), and at a small conference at the University of Hohenheim (June 2026, N at the time ≈ 13000 of 22000). No one in the audience of these talks will be invited to participate in the benchmark study. In additon, as we ask for extensive disclosure of materials, a "cheating" prompt (e.g., "make sure intervention XX does best) would surface in the reporting framework that teams are required to fill out. 

### Three ways to compete

We deliberately welcome very different approaches. Submissions enter at one of **three tiers**, defined by the granularity of the prediction. The common currency across tiers is the set of preregistered estimands—the same quantities the human analysis produces—so every approach is scored on the rungs it can reach. Teams choose the highest tier their approach supports; lower tiers are simply eligible for fewer analyses.

- **Tier 1 · Full — individual-level synthetic data.** A respondent-per-row dataset of behavioral clones: each synthetic participant completes the experiment in one condition. Flows through the preregistered analysis pipeline completely unchanged. Eligible for everything: ATE recovery, calibration regression, response distributions, subgroup heterogeneity, demographic baselines, and stereotyping diagnostics. *Grain: 1 row = 1 synthetic respondent.*
- **Tier 2 · Cell-level — group-level predictions.** A predicted mean and a **95% uncertainty interval** on that mean per cell—per condition × outcome, and per condition × moderator-level × outcome. For approaches that simulate or reason about groups rather than individuals. Eligible for ATEs, calibration, subgroup heterogeneity, and demographic baselines; not eligible for distribution-shape metrics (OVL, KS) or predictability regressions. *Grain: 1 row = 1 cell (condition × [moderator level ×] outcome).*
- **Tier 3 · Effect-level — direct effect predictions.** One predicted ATE per intervention × outcome, in original outcome units, with a **95% uncertainty interval** expressing the team's epistemic uncertainty. For zero-shot forecasting, in-context learning over the literature, meta-analytic priors, or hybrid pipelines. Eligible for directional agreement, Spearman ρ, Pearson r, RMSE, and calibration regression; inferential agreement and TOST are computed from the submitted intervals. *Grain: 1 row = 1 intervention × outcome estimate.*

### Analysis eligibility by tier

| Preregistered analysis | Tier 1 | Tier 2 | Tier 3 |
|---|:-:|:-:|:-:|
| ATE recovery — six-metric ladder (Section 1) | ✓ | ✓ | ✓ * |
| Calibration regression α / β (Section 1) | ✓ | ✓ | ✓ |
| Response distributions — variance ratio, OVL, KS (Section 1) | ✓ | — | — |
| Subgroup heterogeneity — condition × moderator (Section 2) | ✓ | ✓ | — |
| Within-subgroup distributions (Section 2) | ✓ | — | — |
| Demographic baseline calibration (Section 3) | ✓ | ✓ | — |
| Demographic predictability / stereotyping (Section 3) | ✓ | — | — |

\* For Tiers 2 and 3, inferential agreement and equivalence (TOST) are evaluated against the submitted 95% uncertainty intervals rather than model-based standard errors. For Tier 1, standard errors fall out of the fitted models, so no separate uncertainty input is required.

## Method registration and disclosure

Before the prediction lock, every team documents its pipeline on a structured **method-registration form**—built on the GUIDE-LLM reporting checklist [@feuerriegelReportingChecklistLarge2026] and extended where a reporting checklist alone does not guarantee, in principle, full replicability of a prediction pipeline—and deposits it alongside the predictions. The full item-by-item form and the **disclosure policy** (which lets proprietary teams seal items under escrow while keeping the lock publicly verifiable) are published in the [call for participation](index.html#registration-form); a fillable template ships in the [benchmark repository](https://github.com/janpfander/silicon-sample-submission). Two aspects of the registration feed the analysis directly:

- **Approach characteristics** recorded on the form—open- vs. closed-weight model, persona source, administration protocol, fine-tuning, in-context learning, and so on—are the features joined to the leaderboard for the exploratory design-choice analysis (see *Cross-team comparison*).
- **Disclosure class** is a per-entry badge the policy assigns from the entry's most restricted item: **Class A · Open** (all items public), **Class B · Escrowed** (some items sealed but available to the core team and auditors under confidentiality, so compliance stays verifiable), and **Class C · Sealed** (one or more items withheld even from escrow). All three classes receive full leaderboard standing carrying their badge; only publicly disclosed features enter the design-choice analysis, and Class C entries are excluded from that analysis only—not from the leaderboard.

## Submission specification

Each tier defines the exact prediction file(s) a submission provides; the schemas below are the contract the scoring code consumes (`build_pairs_t1()` / `build_pairs_t2()` / `build_pairs_t3()`, see *Cross-team comparison*). The deposit logistics—Zenodo record structure, the `metadata.json` file, file naming, the self-check validator, and the public/escrowed split—are specified on the [materials page](materials.html) and in the [benchmark repository](https://github.com/janpfander/silicon-sample-submission). Two requirements bear directly on the analysis: full **coverage** is all 16 interventions plus control across all 13 outcomes (declared subsets are permitted and scored pairwise, with coverage reported on the leaderboard), and **condition and column labels must match the published codebook exactly** (submissions are validated before the lock; cells that fail to parse are dropped).

### Tier 1 — individual-level dataset

One row per synthetic respondent, matching the structure of the cleaned analysis data (`cleaned_llm.rds` schema). Each team supplies its own synthetic respondents: `profile_id` is a team-assigned identifier, unique within the submission, and assignment of respondents to conditions (between-subjects, covering all 17 conditions) is the team's responsibility, registered under C.4/C.6.

| Column(s) | Type | Specification |
|---|---|------|
| `profile_id` | string | A unique respondent identifier you assign (unique within the submission). |
| `condition` | factor | Control or one of the 16 intervention labels, exactly as in the codebook. |
| `gender`, `age_band`, `race`, `education`, `income`, `party` | factor | The six moderators, levels per codebook (from your own profiles). |
| `trust_multidimensional` + 12 trust items | numeric | Primary composite 0–100; the 12 underlying items per codebook scale. |
| `trust_post`, `distrust_post`, `funding_perceptions`, `policy_role_mean`, `inst_trust_mean` | numeric | Secondary outcomes, original scales per codebook. |
| `belief_post`, `concern_mean`, `policy_general`, `policy_specific_mean`, `behavior_mean` | numeric | Tertiary outcomes, original scales per codebook. |
| `donation_ams` | numeric | Donation to AMS in $, per codebook. |
| `newsletter_signup` | 0 / 1 | Binary behavioral outcome. |

Because a Tier 1 submission has the same shape as the cleaned analysis data, it flows through the original analysis pipeline unchanged and is eligible for **all** analyses. The model functions in `R/functions/statistics.R` are agnostic to the specific condition labels and to the data source, so the same code path applies to the human data and to any team's submission.

**Precision requirement.** An approach's estimated effects are only as precise as the synthetic sample behind them: the fewer synthetic respondents, the noisier each estimated effect. That noise pulls the approach's correlation with the human effects toward zero and inflates its error, so the *same* method scores worse on a small synthetic sample than on a large one.

We could let teams submit small, noisy samples and simply take the resulting hit — nothing about the pipeline breaks, and we would not need to correct for it. We require a minimum instead because the benchmark exists to compare *methods*, and if the noise varied with each team's sample size, the scores would partly reflect how many synthetic respondents a team chose to generate rather than how good its method is: a strong method run on a small sample could land below a weak method run on a large one. Generating more synthetic respondents is cheap, so rather than let this arbitrary factor blur the comparison, we remove it at the source and require a sample large enough that submission noise is small for everyone.

Concretely, a Tier 1 submission must have at least **1,000 synthetic respondents per intervention and 2,000 in the control condition**, with several times more encouraged — the full human study's cell size, and twice the ≈ 500 / ≈ 1,000 half-sample each submission is scored against, so the submission is already more precise than the human reference and its own noise is not the bottleneck. The other route — letting teams submit small samples and then statistically correcting each score for its own noise, the way we *do* correct the fixed, unavoidable noise on the human side — we reject on the submission side: it would hand an under-sized entry the score of a precise one, rewarding teams for skimping. @hewittPredictingResultsSocial2025 likewise kept their predictions precise by averaging many simulated responses per condition rather than correcting noisy ones.

### Tier 2 — cell-level file(s)

Two long-format files. `*_cells_main.csv`: one row per condition × outcome. `*_cells_moderator.csv`: one row per condition × moderator × moderator-level × outcome.

| Column | Type | Specification |
|---|---|------|
| `condition` | factor | Per codebook. |
| `moderator`, `moderator_level` | factor | Moderator file only; per codebook levels. |
| `outcome` | string | Outcome variable name per codebook. |
| `mean` | numeric | Predicted cell mean on the original scale (proportion for `newsletter_signup`). |
| `pi_lower`, `pi_upper` | numeric | 95% uncertainty interval bounds on the cell mean — the team's epistemic uncertainty about the *human* cell mean, not a within-simulation SD. The ATE interval is derived by combining the treatment- and control-cell intervals; used for inferential-agreement and equivalence scoring. |

### Tier 3 — effect-level file

One row per intervention × outcome (16 × 13 = 208 rows for full coverage). Estimates are ATEs versus control on the original outcome scale; the binary outcome is on the probability scale.

| Column | Type | Specification |
|---|---|------|
| `condition` | factor | Intervention label per codebook. |
| `outcome` | string | Outcome variable name per codebook. |
| `ate` | numeric | Predicted ATE, original units. |
| `pi_lower`, `pi_upper` | numeric | 95% uncertainty interval bounds — the team's epistemic uncertainty about the *human* ATE, not a within-simulation standard error. Used for inferential-agreement and equivalence scoring. |

## Preregistered analysis plan

### Overview and the human replication reference

Each submission is scored against the human data with the same three-section plan, restated in full in the subsections below so that this document stands alone. The plan is applied **symmetrically** to the human sample and to every submission: the same model and comparison functions (`R/functions/statistics.R`) run on both, so a submission is never held to a standard the human data is not.

The human sample is split into two halves by a preregistered random seed. **Human 1** is the reference half every submission is scored against. The agreement between **Human 1 and Human 2** defines the **human replication reference**—how closely two independent halves of the human sample agree on the same estimands. We deliberately do **not** call this a ceiling: it is *not* an upper bound. The replication row carries sampling noise on *both* of its sides, while a submission's score carries it only on the reference side, so a sufficiently precise method (Tier 1 at the required synthetic sample sizes) is *expected* to score above the replication row when it genuinely tracks the true effects — under a fixed-true-effects noise model, a perfect predictor's expected raw correlation with Human 1 is the square root of the replication row's. Exceeding the replication reference is therefore a meaningful achievement, not an anomaly; the reference answers "how well do humans replicate humans at this sample size," which is the natural yardstick, not a bound. The reference is computed with the identical code path and reported in the same row format as every team; it is the one fixed reference row in every results table. A **floor baseline** ("no effect") anchors the other end for the metrics without a natural null — see *Cross-team comparison*. Because the split seed is itself an analytic choice, its influence is checked and reported: the reference (and, with real data, the full field) is recomputed across many alternative half-splits, so that results demonstrably do not hinge on the seed we happened to draw (*Robustness of the scoring pipeline*, below). Every submission is scored under identical rules and appears on the leaderboard like any other; no entry receives a special reference row beyond the reference and floor.

**Scope: aggregate and distributional fidelity, not individual fidelity.** The benchmark tests whether an approach reproduces the megastudy's *estimands* — treatment effects, subgroup patterns, and population response distributions — not whether any synthetic respondent matches any real person. It cannot test the latter by design: teams build their own synthetic respondents, with no pairing between a synthetic and a human individual, so an individual-level claim is never made or scored. The distinction matters beyond bookkeeping. Recent work shows that aggregate agreement can coexist with divergent individual-level behaviour — simulated populations that reproduce collective outcomes while getting individual heterogeneity and decision rules wrong [a "macro–micro dissociation"; @arrudaCollectiveCooperationIndividual2026]; and benchmarks that do score individual fidelity require paired designs with per-person consistency ceilings [@parkLLMAgentsGrounded2026]. Our distribution metrics (variance ratio, OVL, KS) are the *population-level* response to that warning — they catch a synthetic crowd whose members answer too much alike — but they should not be read as evidence about individuals.

Which of the analyses below applies to a given submission follows from its tier (see *Analysis eligibility by tier*): Tier 1 is eligible for all of them, Tiers 2–3 for the estimate-based subset.

### Comparison functions

The comparison metrics are operationalized by a small set of helper functions, applied symmetrically to the human reference and to each submission. They are unchanged from v1 and shown here in full.

#### Comparison metrics

`compare_estimates()` is used throughout—for ATEs (Section 1), moderator interaction estimates (Section 2), and secondary analyses. It computes six metrics, reported in increasing order of strictness. The rationale for the ordering is to locate precisely *where* an approach breaks down: a submission might pass the weakest test (getting directions right) while failing a stricter one (magnitude equivalence), and knowing at which rung the ladder breaks is more informative than any single pass/fail verdict.

- **Directional agreement (%)** — *least strict.* % of estimates with the same sign as the human estimate. Chance is 50%. Do approaches at least identify the direction of each effect?
- **Spearman ρ** — rank correlation of point estimates. Do approaches rank interventions in the same order as humans? Insensitive to scale or offset; an approach that inflates all effects equally still scores ρ = 1.
- **Pearson r** — linear correlation of point estimates. Captures proportionality between predicted and human estimates but not absolute level.
- **Inferential agreement (%)** — % of estimates with the same BH-adjusted conclusion (significantly positive, significantly negative, or non-significant at α = .05). Requires getting both magnitude and uncertainty right; would a researcher using the submission reach the same inferential decision as one using human data?
- **RMSE** — root mean squared error in outcome units. Average absolute magnitude error; scale-dependent and therefore not comparable across outcomes (hence omitted from pooled-across-outcome scores; see *Scoring principle*, below). As a scale reference, RMSE is read against the **spread (SD) of the human ATEs** for that outcome: the spread is the error of the best possible constant prediction, so a submission with intervention-specific skill must come in below it.
- **TOST (%)** — *most strict.* % of estimate pairs passing an equivalence test (two one-sided tests, α = .05) with a **fixed, prespecified bound of 3% of each outcome's scale range** (3 points on the 0–100 sliders, $0.30 on the donation, 3 percentage points on the signup proportion). This replaces v1's relative bound Δ_k = 0.5 × |ATE_human_k| for two preregistered reasons: a bound computed from the noisy estimate under test is not a valid α-level equivalence procedure (equivalence bounds must be set independently of the observed data), and the design analysis (see *Cross-team comparison*) shows the relative bound is smaller than the smallest difference the test could confirm at this design's reference SEs — a rung that would read ~0% for every entry. The v1 analysis (locked) retains its original bound; the fixed bound governs all cross-team scoring.

```{r, results='asis'}
#| label: fn-compare-estimates
print_a_function_from_file("compare_estimates")
```

#### Adjusted correlation and RMSE (noise-corrected)

The reference effects any submission is scored against are themselves noisy: each human ATE is estimated from a finite sample (and, here, from a *half*-split of it), so it carries a known standard error. That noise mechanically dilutes the correlation rungs — it cannot add to the covariance between predictions and effects (it is random, hence unrelated to the predictions), but it inflates the observed spread of the reference effects, so the raw Pearson r is biased toward zero. The same noise inflates the RMSE.

We therefore report, as a **companion to the raw values wherever Pearson r appears in the cross-team analysis**, a measurement-error-adjusted correlation **r~adj~** and an adjusted RMSE (`adjusted_metrics()`). The estimand is deliberately framed in **fixed-effects** terms: *the correlation between the predictions and these same 16 true effects, free of the reference sample's measurement noise* — no assumption that the interventions' effects are draws from a common distribution is needed. The computation is the variance-decomposition identity (Spearman disattenuation with known error variances): the reference effects' true spread is their observed spread minus the average known sampling variance, and the adjusted correlation divides the (noise-immune) covariance by that corrected spread. This is the fixed-effects, method-of-moments analogue of the two-stage meta-analytic r~adj~ of @hewittPredictingResultsSocial2025; with the benchmark's near-equal per-arm sample sizes the precision weighting their likelihood machinery adds is immaterial, and our approach avoids its bivariate-normal assumption. Adjusted RMSE applies the same measurement-error logic to the RMSE rung: the average squared prediction error minus the average reference sampling variance, square-rooted.

Three disciplining rules govern its use. **Raw r remains the primary, ranked metric** — it is assumption-free, and because reference noise is roughly uniform across interventions it deflates every submission *and the human replication reference* together without reordering them. **r~adj~ is always shown with its uncertainty interval** (the same cluster bootstrap as every other metric): when true effect heterogeneity is small relative to the noise, the adjustment is intrinsically fragile, and the interval announces that rather than hiding it. **The adjustment is undefined when noise swamps signal**: if the estimated true variance is not positive, `adjusted_metrics()` returns NA instead of a fabricated value (on pure-noise placeholder data this guard binds by construction). Only the reference (Human 1) side is corrected — mirroring @hewittPredictingResultsSocial2025, who stabilized the prediction side at the source by ensemble averaging rather than by correction; the benchmark's analogue is the Tier-1 precision requirement (see *Submission specification*). The replication row's own r~adj~ corrects the shared reference (Human 1) noise only — the Human 2 side stays noisy, which is one of the two reasons the replication reference is beatable rather than a bound (see *Overview*): a precise submission's r~adj~ can approach 1 while the replication row's cannot. The adjustment matters most in the **subgroup analyses**, where small demographic cells make reference noise non-uniform and raw correlations unevenly deflated across exactly the stereotyping-relevant comparisons; there r~adj~ is the emphasized value, with raw r beside it. (Spearman ρ is attenuated by reference noise in the same direction but has no analogous closed-form correction, so it is reported raw only; the asymmetry is one more reason the ladder is read as a whole rather than any single rung.)

A companion benefit is comparability: @hewittPredictingResultsSocial2025 report r and r~adj~ side by side throughout. Across their megastudies — the closest analogue to this benchmark^[We take Hewitt et al.'s megastudy figures, not their headline r = .85, as the reference because the two numbers measure different tasks. The .85 is an *across-study* correlation over 476 effects pooled from 70 unrelated survey experiments, spanning many topics, treatments, and outcomes. Their megastudy figures instead come from a *within-study* analysis — how well the model distinguishes the several interventions of a single megastudy against that study's shared control — which is exactly this benchmark's setting (16 interventions, one control). Hewitt et al. draw this within- vs across-study distinction themselves.] — survey experiments reach r = .47 (r~adj~ = .61) and text-based treatments r = .46 (r~adj~ = .59), well below the r = .85 (r~adj~ = .91) of their primary 70-experiment archive; reporting both metrics places every submission, and the field as a whole, directly against these published reference points.

```{r, results='asis'}
#| label: fn-adjusted-metrics
print_a_function_from_file("adjusted_metrics")
```

#### Distribution comparison metrics

`compare_distributions()` compares response distributions for a single condition and outcome using three metrics, again least-to-most strict:

- **Variance ratio** (submission / human variance) — *least strict.* Captures only relative spread; a ratio of 1 means equal dispersion and does not assess shape. A ratio below 1 is the widely reported failure mode of synthetic samples: under-dispersed responses.
- **OVL (overlapping coefficient)** — proportion of area shared between the two kernel density estimates; 1 = identical distributions, 0 = no overlap. Summarises overall shape similarity.
- **KS D-statistic** — *most strict.* Maximum absolute gap between the two empirical CDFs; 0 = indistinguishable CDFs everywhere. Sensitive to local departures OVL might average away.

```{r, results='asis'}
#| label: fn-compare-distributions
print_a_function_from_file("compute_ovl")
```

#### Calibration regression

Correlation tells us only whether the human and predicted effect-size vectors *covary*; it does not capture baseline differences or multiplicative scaling. An approach could inflate every effect by a factor of two and still score Pearson r = 1. Calibration regression separates two failure modes—additive bias and proportionality.

`run_calibration()` regresses human ATEs on predicted ATEs: `ATE_human = α + β × ATE_predicted`. The intercept α captures additive bias (α > 0: the approach systematically underestimates; α < 0: overestimates). The slope β captures proportionality (β = 1 is perfect scaling; β > 1 means predicted effects are compressed, β < 1 means exaggerated). We report 95% CIs for both and R² as the share of human-ATE variance linearly recoverable from the predictions. We then test calibration at two strictness levels: the **joint F-test of (α = 0, β = 1)**—the omnibus test of perfect calibration, whether the predictions can be read directly as human ATEs with no correction—and the weaker **slope-only F-test of (β = 1)**, whether the predictions are *proportionally* calibrated while permitting a constant additive offset. Proportional calibration is the more practically relevant criterion: additive bias is easy to correct with a single human anchor, whereas a slope distortion reflects a deeper failure. In the benchmark the regression runs with **HC2 heteroskedasticity-robust inference** (`robust = TRUE`): the outcome side carries known, condition-varying sampling error that classical F-tests ignore; a precision-weighted variant (weights 1/SE²~h~, `weight_by_precision = TRUE`) is reported as a sensitivity check with the real data. Three honest caveats travel with every calibration verdict: the regression has only 16 points, so the tests have little power and **non-rejection is not evidence of calibration**; reference-side noise in the regressor direction is absent here (the human ATEs are the outcome), but for Tiers 2–3 the *predicted* ATEs carry the team's epistemic uncertainty as regressor noise, attenuating β̂ toward zero and biasing the verdict toward "exaggerated"; and the leaderboard's β and R² columns are computed on the primary outcome only (stated in the table caption).

```{r, results='asis'}
#| label: fn-calibration
print_a_function_from_file("run_calibration")
```

#### Demographic baseline and predictability metrics

`compare_demographic_baselines()` computes, for each moderator, the Pearson correlation and RMSE between human and predicted cell means in the control condition. `compare_demographic_predictability()` fits **separate** OLS regressions of the outcome on each moderator plus condition fixed effects, returning R² and dummy-coded coefficients for humans and for the submission. `demographic_parity_gap()` condenses the per-group baseline errors into one fairness number per moderator — the gap between the worst- and best-served group (see *Section 3*).

```{r, results='asis'}
#| label: fn-demographic-baselines
print_a_function_from_file("compare_demographic_baselines")
```

```{r, results='asis'}
#| label: fn-demographic-predictability
print_a_function_from_file("compare_demographic_predictability")
```

```{r, results='asis'}
#| label: fn-demographic-parity-gap
print_a_function_from_file("demographic_parity_gap")
```

### Section 1 — Average Treatment Effects

**ATE recovery.** For each intervention, the average treatment effect (ATE) is the difference in mean outcome between participants assigned to that intervention and those in the control condition. ATEs are estimated with `run_main_treatment_model()`—OLS with HC2 heteroskedasticity-robust standard errors and BH-adjusted p-values—run on the human reference and on the submission separately for each preregistered outcome. For `newsletter_signup` (binary), `run_main_treatment_model_binary()` returns marginal effects on the probability scale. The full six-metric ladder (`compare_estimates()`) and the calibration regression (`run_calibration()`) are then applied. Following v1, the ladder is reported pooled across all outcomes to maximize power, then broken down by outcome; a pooled row prepended to the ATE table carries all outcomes at once (RMSE omitted there, as it is not comparable across scales).

**Top-intervention selection.** The correlation rungs summarize the whole effect pattern; the decision a many-arm megastudy exists to inform is narrower — *which interventions should one act on?* `topk_selection()` scores it directly: per outcome, an approach's top **k = 3** interventions by predicted ATE (≈ the top 20% of the 16 arms, in the spirit of the intervention-selection use case that motivates @hewittPredictingResultsSocial2025) are compared with the truly best three by the human reference, via the **hit rate** (share of picks in the true best set; chance = 3/16) and the **regret ratio** (mean true ATE of the picked set / mean true ATE of the best set; 1 = perfect pick, scale-free within outcome, undefined when the best achievable mean ATE is not positive). Available to **all tiers** (it needs only point predictions).

**Response distributions.** Even an approach that matches every ATE may distribute responses differently: two groups with the same mean can differ in spread, skewness, or tail shape. In the **control condition**, across all continuous outcomes, we compare the full response distribution (`compare_distributions()`: variance ratio, OVL, KS D). The **variance ratio is the headline diagnostic** of the three, with a stated directional expectation: the recurring failure mode of synthetic samples — reported independently across the recent evaluation literature [@hewittPredictingResultsSocial2025; @cumminsThreatAnalyticFlexibility2025b; @parkLLMAgentsGrounded2026; @arrudaCollectiveCooperationIndividual2026] — is **under-dispersion**, synthetic respondents answering too much alike (ratio < 1). `newsletter_signup` is excluded (kernel density is not meaningful for a binary outcome). This analysis requires respondent-level data and so is **Tier 1 only**. By-outcome breakdowns use the illustrative subset (multidimensional trust, donation to AMS, funding perceptions, general climate policy).

### Section 2 — Subgroup Effects

**Heterogeneity in treatment effects.** Interventions may work differently across demographic groups—a consensus message might shift trust more among Republicans than Democrats, or more among younger than older respondents. For each of six categorical moderators (gender, age band, race, education, income, partisan identity), condition × moderator interactions are estimated with `run_moderator_model()` on the human reference and on the submission, across the illustrative outcomes (`compute_predicted = FALSE` skips the marginal-effects step the leaderboard does not use). The full set of interaction estimates is compared with the same ladder ordering as Section 1. Because these pairs carry no inferential category per interaction cell, the pooled subgroup score uses the reduced signed-effect ladder—directional agreement, Spearman ρ, Pearson r (`signed_metrics()`)—plus the adjusted correlation r~adj~ computed from the human interaction estimates' standard errors. The subgroup section is where the adjustment matters most and is the emphasized value there: small demographic cells make the reference noise non-uniform, so raw correlations are unevenly deflated across exactly the comparisons the stereotyping question cares about (see *Adjusted correlation and RMSE*). Eligible for Tiers 1–2.

**Response distributions by subgroup.** Mirroring Section 1, within each demographic subgroup in the control condition we compare response distributions (OVL, KS D, variance ratio), to catch an approach that reproduces the overall distribution while miscalibrating within-group spreads. Run for all outcomes, demonstrated on the primary outcome (`trust_multidimensional`); a minimum group size of 10 is required for stable kernel density estimation. **Tier 1 only.**

### Section 3 — Demographic Baseline Calibration

Section 2 tests whether an approach exaggerates *effect* heterogeneity; Section 3 asks the orthogonal question of whether it starts from the right **baseline**, using only the control condition. An approach could assign far too little trust to Republicans, or too much to Democrats, before any treatment—a distinct failure mode (baseline stereotyping rather than effect stereotyping).

**Demographic baseline calibration.** For each moderator, we compute control-condition group means for humans and for the submission, then correlate and compare them (`compare_demographic_baselines()`: Pearson r and RMSE of cell means). High correlation and low RMSE mean the approach starts from the right response level for each demographic type. Eligible for Tiers 1–2.

**Demographic parity gap (DPD).** The baseline-calibration metrics average over groups, and an average can hide exactly what a fairness question asks about: an approach may serve most groups well while consistently missing one. Following the demographic-parity-difference summary of @parkLLMAgentsGrounded2026, `demographic_parity_gap()` reports, per moderator, the gap between the **worst- and best-served demographic group** — the difference between the largest and smallest absolute cell-mean error in the control condition — together with which groups those are, and the **worst-served group's absolute error** alongside it (a zero gap can also mean every group is served equally *badly*, which the worst-group error catches). Groups with fewer than 30 respondents on either side are skipped: a small-cell group mean has an SE of several scale points, and a max-over-groups error is a max-of-noise statistic that is mechanically positive even for a perfect approach — DPD values are read against the human-replication reference row, which carries the same affliction. A large gap flags that baseline accuracy is concentrated in some groups, the pattern prior work finds for populations under-represented in training data. Eligible for Tiers 1–2.

**Demographic predictability.** For each of the six moderators we fit a **separate** OLS regression of the outcome on that moderator plus condition fixed effects, on the human reference and on the submission independently (`compare_demographic_predictability()`). Separate per-moderator regressions avoid conflating correlated predictors (race, income, education are themselves related) and give a clean answer: how much does, say, partisan identity alone explain outcome variance in humans versus in the submission? An R² substantially higher for the submission than for humans flags over-reliance on demographic stereotypes; the dummy-coded coefficients identify which groups are over- or under-weighted. Run for all outcomes, demonstrated on the primary outcome. **Tier 1 only.**

### Tier-specific estimand recovery

The comparison-metric ladder requires, per intervention × outcome, an estimate, a standard error, and an inferential category. How these are obtained depends on the tier:

- **Tier 1:** the submission runs through `run_main_treatment_model()` exactly like the human data; estimates and HC2 robust standard errors fall out of the fitted models. Submission-side **inference is then standardized to the human design's precision**: each SE is rescaled by the two-cell design factor $\sqrt{(1/n_{h,k} + 1/n_{h,ctrl})\,/\,(1/n_{l,k} + 1/n_{l,ctrl})}$ and p-values recomputed (normal approximation, BH within outcome, mirroring the Tier-2 path). A team may simulate any pool size, so unstandardized significance would measure that choice rather than the method — and the Tier-1 precision requirement would mechanically hurt precise teams wherever a human effect is n.s. Point estimates are untouched, so the correlation, directional, and RMSE rungs are unaffected; the standardization makes the inferential and TOST columns comparable across tiers.
- **Tier 2:** the ATE for each intervention × outcome is the submitted cell mean minus the submitted control mean. Because only cell summaries are available, the preregistered OLS cannot be refit on a Tier 2 submission; instead we compute the ATE's standard error from the submitted cell uncertainty intervals. Each cell interval implies a standard error on that cell mean, $\mathrm{SE}_{cell} = (\mathrm{pi}_{upper} - \mathrm{pi}_{lower}) / (2 \times z_{0.975})$, and the treatment and control cells combine in the usual way for a between-subjects contrast, $\mathrm{SE} = \sqrt{\mathrm{SE}_t^2 + \mathrm{SE}_c^2}$. The test statistic $z = \mathrm{ATE} / \mathrm{SE}$ is referred to the standard normal distribution for a two-sided p-value, $p = 2\,\Phi(-|z|)$. These p-values are BH-adjusted within outcome, mirroring the Tier 1 models. Subgroup ATEs and demographic baselines are computed analogously from the moderator cell file.
- **Tier 3:** the submitted `ate` is the estimate. Inferential agreement uses the submitted 95% uncertainty interval: an effect counts as significantly positive (negative) if the interval lies entirely above (below) zero—the direct analog of a two-sided test at α = .05, so the significance criterion matches the one applied to Tiers 1–2. For the TOST equivalence test, the interval implies a standard error, $\mathrm{SE} = (\mathrm{pi}_{upper} - \mathrm{pi}_{lower}) / (2 \times z_{0.975})$.

**The submitted intervals are also scored directly (Tiers 2–3).** An uncertainty interval that only feeds the inferential rung would create a gaming channel: arbitrarily wide intervals turn every prediction "n.s.", which matches a mostly-null world for free. Two preregistered scores close it (`pi_metrics()`): **PI coverage** — the share of reference ATEs falling inside the submitted (Tier 3) or implied (Tier 2) 95% intervals; honest intervals land near 95% — and the **Winkler interval score** — the interval's width plus $2/\alpha$ times the distance by which the reference value misses it, averaged over pairs. The Winkler score is a *proper* score, minimized in expectation by reporting one's true predictive interval, so neither padding intervals wide nor squeezing them tight pays. It is scale-bound and therefore computed on the 0–100 outcomes only. Both appear as leaderboard columns for Tiers 2–3; they are undefined for Tier 1 (whose SEs are sampling, not epistemic, uncertainty) and for the reference rows.

**Human-side missing data** are handled by complete-case estimation within each model (the `lm`/`glm` default), exactly as in v1 and the main megastudy's preregistered analysis; no imputation is used anywhere in the scoring pipeline.

### Cross-team comparison (new)

This is the new analytical contribution of this preregistration, and it reports **two findings**. The **headline finding** asks how well the *field* of approaches predicts the experiment at all: across every submission, what is the **distribution of predictive quality** on each metric — its centre (mean, median), its spread, and how close the field comes to the human replication reference. The **secondary finding** ranks individual approaches against one another; it is a structural read of *which* approaches lead on *which* rung, deliberately not a hunt for a single champion.

Both findings score every submission with the same metric ladder specified under *Comparison functions*, above—in increasing order of strictness: directional agreement, Spearman ρ, Pearson r (with the noise-corrected r~adj~ as its companion), inferential agreement, RMSE, and TOST equivalence—so that approaches can be compared directly and against the human replication reference. Knowing on which rung an approach breaks is more informative than any single pass/fail verdict. On top of the ladder: the calibration regression (human ATEs on predicted ATEs) separates additive bias (α) from proportionality (β); distributional metrics (variance ratio, overlap coefficient, KS distance) test whether Tier 1 submissions reproduce response *shapes*, not just means; and the stereotyping diagnostics ask whether demographic variables predict synthetic responses with the same magnitude as in humans—or whether simulated respondents over-rely on demographic cues.

**Scoring principle: one score per metric, pooled across outcomes.** Each approach is scored on every rung of the ladder, with the intervention × outcome estimates **pooled across all 13 outcomes** — the 11 continuous sliders plus the two behavioral outcomes (donation via the OLS model; newsletter signup via logistic marginal effects on the probability scale, as in v1's pooled row) — into a single value per metric. We deliberately do *not* collapse the rungs into one composite "champion" score: an approach can recover the *direction* of effects well while mis-scaling them, or nail the primary outcome while failing on the periphery, and a single number would hide exactly the structure the benchmark exists to reveal. So results are read **per metric** — each panel is its own story. Pooling across outcomes (rather than averaging per-outcome scores) keeps every intervention × outcome comparison as one data point and matches the pooled-row logic of the ATE analysis in Section 1. Pooling has one known artifact, and we report its diagnostic: a pooled correlation can be inflated by a submission that predicts *which outcomes* are generally movable while misranking the interventions within every outcome. The leaderboard therefore carries the **Fisher-z average of the per-outcome correlations** beside the pooled value (`fisher_z_avg()`); when the two diverge, that is the signature of exactly this artifact, and the per-outcome breakdown shows where. The one metric that cannot be pooled across outcomes is RMSE (not comparable across outcome scales); it — and its noise-corrected companion RMSE~adj~ — is therefore reported only in the per-outcome and per-intervention breakdowns, where the scale is held constant. The same scale constancy is what licenses the correlation rungs in the per-intervention breakdown: every intervention is scored across `outcomes_continuous`, which are all 0–100 slider scales, so a per-intervention Spearman ρ or Pearson r reflects the effect pattern rather than a mix of measurement units.

**Unit of analysis and coverage rules.** Teams may enter up to two submissions per tier, so submissions are not independent draws of "an approach": sibling entries from one team would overweight prolific teams in the field distribution and inflate its maximum. The **field statistics (Finding 1) therefore use one pre-designated primary entry per team**; every valid entry still appears on the leaderboard (Finding 2), labelled primary or secondary. Declared coverage subsets are permitted, but a self-chosen easy subset could top a pooled correlation without being comparable to full-coverage entries; the **primary leaderboard is therefore restricted to full-coverage entries**, with partial-coverage entries reported in a separate annex, scored pairwise on what they cover, with their coverage stated. Finally, the field distribution describes an invited, self-selected set of expert teams — a purposive sample of the current generation of serious approaches, not a random sample of all possible ones; we state this wherever the field summary is interpreted, and we do not display a field summary at all if fewer than **five** primary entries are scored.

**Presentation.** Every preregistered metric is computed and reported, but a figure that shows six metrics at once shows none of them well, so the figures carry only **one key metric per question**, stated here once and used consistently throughout: **Pearson *r*** for ATE recovery (Section 1) and, on the same axis, for subgroup recovery (Section 2); the **calibration slope β and intercept α** for whether those effects are the right *size*, not only correctly ranked; and the **variance ratio** for response-shape recovery (Section 1b/2b). Every other rung of the preregistered ladder — directional agreement, Spearman ρ, the adjusted correlations, inferential and TOST agreement, RMSE, OVL, KS, W1 — is still computed and scored for every submission, and reported in a companion table beside each figure; nothing is dropped, only de-emphasized in the plots.

The headline display (@fig-headline) puts the key measures in one figure, two panels: (A) predicted against human ATEs across all approaches and outcomes, in percentage points of each outcome's scale range — the direct visual of the prediction task, in the units of the predicted-vs-observed figure of @hewittPredictingResultsSocial2025 (effects on the original scales; no SD standardization, which an under-dispersed submission would corrupt) — overlaid with the average field and human **calibration regressions** and a corner readout of their correlation and calibration coefficients (slope β, perfect = 1; intercept α, perfect = 0, with 95% CIs); and (B) approaches **ranked** on the pooled Pearson *r*, each with a cluster-bootstrap **uncertainty interval** over interventions (introduced by this preregistration — v1 reported point estimates only; detailed under *Uncertainty intervals* below), the human replication reference pinned on top and an **all-approaches summary row** at the bottom carrying the field's density and mean with its CI — this one panel carries both the field distribution (Finding 1) and the ranking (Finding 2). Three further figures break the ATE and shape key metrics down: by outcome and by intervention (@fig-field-by-group), by moderator for subgroup recovery (@fig-subgroup-headline), and per-outcome response densities with the variance ratio stamped on each panel (@fig-distribution-headline). Every entry carries its disclosure-class badge (open / escrowed / sealed) in the leaderboard table, and the precise numbers accompany every figure as a companion table.

#### Uncertainty intervals on the leaderboard

**Why the leaderboard carries intervals at all.** Each point in the ranking panel is a single pooled score — one number per approach per metric. A bare point invites over-reading: if one approach scores Pearson *r* = 0.60 and another 0.55, is the first genuinely better, or is the gap an artifact of the particular interventions that happened to be tested? The interval answers exactly that question. It is the guard against ranking approaches on differences that would not survive a different, equally reasonable test.

**What the interval is, mechanically.** The pooled metric is computed once on the full data to give the point estimate. We then form a *cluster bootstrap over interventions*: the 16 interventions are resampled with replacement (some drawn twice, some dropped), the pooled metric is recomputed on each resampled set, and the 2.5th–97.5th percentiles of those recomputed scores give the 95% interval (`cluster_boot()` in `R/functions/statistics.R`). Clustering is *by intervention* — each intervention's full block of outcome rows is drawn or dropped together — because the outcome rows belonging to one intervention are not independent. Crucially, **no submitted prediction is ever altered**: each approach's predicted effect for each intervention stays fixed at the value the team submitted. What is resampled is *which interventions enter the score*, not how well anyone did on them. The interval therefore measures the spread of the recomputed *score*, not the spread of the raw effects: averaging or correlating cancels much of the effect-to-effect spread, which is why the bar is typically far narrower than the range of the underlying estimates.

**What the interval means, and the one assumption it rests on.** The interval treats the 16 interventions as a *test set* standing in for the broader space of interventions the benchmark cares about, exactly as a held-out test set stands in for future cases in predictive modelling. Read this way, the bar answers a generalisation question: *would this approach's score, and its rank, survive on a different set of interventions of the same kind?* This is the appropriate question for a benchmark, whose purpose is to identify general predictive skill rather than performance on sixteen specific stimuli. The assumption is therefore about the *interventions*, not the predictions: that the tested set is a sample from a population of possible interventions. Resampling does **not** assume that a team that predicted a given intervention well "could have" gotten it wrong — the predictions are fixed; it assumes only that a replication might have tested a different set of interventions. (If one instead regarded these exact sixteen as the entire universe of interest, there would be no sampling and hence no uncertainty to display — the scores would simply be exact. We do not adopt that reading: the benchmark exists to compare *approaches*, not to characterise these particular stimuli.)

**Interval width is informative, not a penalty on accuracy.** Because the metrics are bounded agreement scores with a ceiling at perfect agreement, an approach that tracks the human effects closely sits on the diagonal and stays there under resampling — dropping or duplicating interventions barely moves a score already pinned near perfect agreement — so accurate approaches carry *narrow* intervals. An approach whose predictions do not track the human effects has a score floating in the interior, free to swing with the resampled set, and so carries a *wide* interval. Interval width is thus entangled with accuracy, but in the helpful direction: a wide bar is the leaderboard signalling that a score is flimsy and may reflect test-set luck rather than skill — precisely the warning one wants attached to a weak entry. (Spread in the size of the true effects does not by itself widen the bar; the correlation and agreement metrics are scale-free, so what matters is whether predictions hug the line, not how far apart the interventions are. Width also partly reflects heterogeneity in per-intervention difficulty, so it is a flimsiness *signal*, not a pure standard error.)

**Decision rule for ranking claims: paired differences on shared replicates.** Overlapping marginal bootstrap intervals do **not** license "A and B are indistinguishable", and non-overlap is not required to claim a difference — the standard misreading. Because every submission is scored against the *same* reference on the *same* intervention clusters with the *same* preregistered bootstrap seed, the correct comparison is paired: draw one set of clusters per replicate, score both submissions on it, and take the difference, so the shared reference noise cancels within each replicate (`cluster_boot_diff()`). **We preregister the rule that any claim that one approach outperforms another on a metric requires the paired-difference 95% interval to exclude zero**; otherwise the two are reported as statistically indistinguishable on that metric — the modal outcome among top performers in comparable mass-collaboration benchmarks (Salganik et al. 2020), and one we expect here. Two caveats on all bootstrap intervals: with 16 intervention clusters, percentile intervals are known to undercover somewhat, and the 16 ATEs share the single control arm, which induces positive dependence the intervention-level resampling does not capture; the intervals are read as honest approximations, not exact coverage statements.

**Relation to the per-effect confidence and uncertainty intervals.** The bootstrap interval is a *different layer* of uncertainty from the confidence intervals (Tier 1) and uncertainty intervals (Tiers 2–3) attached to individual treatment effects, and the two should not be conflated. A per-effect CI or PI is *within-intervention*: it says how precisely a single effect is known — finite-respondent sampling noise for a Tier 1 model fit, or the team's own stated uncertainty for a Tier 2/3 submission. The bootstrap interval is *across-intervention*: it says how much the pooled score depends on which interventions were tested. The two are **nested, not competing**. The per-effect intervals enter the *value* of each row: for the correlation and directional rungs only the point estimates are used, so the per-effect intervals do not affect those scores at all and the bootstrap is their sole source of an error bar; for the inferential-agreement and TOST rungs the per-effect intervals set each row's significance or equivalence label, which the bootstrap then pools and resamples. In all cases the bootstrap holds each intervention's estimate fixed at its point value: it propagates the *which-interventions* uncertainty into the bars but **not** the within-effect sampling noise.

**A noted alternative.** Folding both layers into a single bar would require a *two-level* bootstrap — in each replicate, resample the interventions *and* perturb each drawn effect by its standard error (drawing the effect from a normal centred on the estimate with that SE). We adopt the single-level (point-estimate) cluster bootstrap as the default because, with only sixteen interventions, the *which-interventions* term dominates the uncertainty in any ranking, while the human study is powered such that each effect's standard error is comparatively small relative to the treatment-to-treatment spread. The per-effect CIs and PIs remain reported in the per-outcome breakdowns as the finer, within-effect layer.

**Scope across panels.** The ATE leaderboard scores interventions, so its bootstrap resamples interventions and is available to **every tier** — the resampling needs only one predicted effect per intervention, which all tiers supply (including effect-only Tier 3). The subgroup and distributional leaderboards are restricted to Tier 1, but for a different reason: their *metrics* (interaction models, response-shape comparisons) require respondent-level data, not the bootstrap. The distributional metrics are computed in the control condition only and have no intervention contrast, so that panel necessarily resamples outcomes rather than interventions; with a handful of outcomes those intervals are correspondingly coarser and are read with that caveat.

Registered approach characteristics (open- vs. closed-weight, persona source, administration protocol, fine-tuned vs. not, …) are joined to the leaderboard for a descriptive, hypothesis-generating analysis of which design choices accompany better predictions—preregistered as exploratory, and restricted to publicly disclosed features (Class C entries are excluded from this analysis, not from the leaderboard).

::: {.callout-note}
**Exploratory by design.** With a modest number of teams, the relationship between approach characteristics and predictive performance is reported **descriptively** (rankings, group comparisons, visual patterns). We do not preregister inferential tests of feature effects; the cross-team feature analysis is hypothesis-generating, intended to surface which design choices appear to matter for future, larger benchmarks.
:::

#### Scoring functions

The cross-team analysis reuses the comparison functions specified under *Comparison functions*, above, and adds the helpers below. The scoring helpers (in `R/functions/statistics.R`) are shown in full and explained before each; the two plotting helpers (in `R/functions/plots.R`) are described but not reproduced here.

`pooled_metrics()` computes the ATE-recovery ladder on intervention × outcome pairs **pooled across all outcomes**: directional agreement, Spearman ρ, Pearson r, the noise-corrected companion r~adj~ (via `adjusted_metrics()`, see *Comparison functions*), the Fisher-z per-outcome-averaged correlations (via `fisher_z_avg()`; the pooled-vs-averaged divergence diagnostic, see *Scoring principle*), inferential agreement, and TOST equivalence with the fixed per-outcome bounds (`tost_delta`). It mirrors the original `compare_estimates()` but operates on the pooled set of pairs rather than a single outcome; pooled RMSE is deliberately omitted, as it is not comparable across outcome scales, but `include_rmse = TRUE` adds RMSE and RMSE~adj~ for the per-outcome and per-intervention breakdowns where scale is constant. It expects pre-joined pairs with the reference (Human 1) columns `estimate_h`, `se_h`, `p_adj_h` and the submission columns `estimate_l`, `se_l`, `infer_l`. `pi_metrics()` scores the Tier-2/3 uncertainty intervals (coverage + Winkler interval score); `cluster_boot_diff()` implements the paired-replicate decision rule for ranking claims.

```{r, results='asis'}
#| label: fn-pooled-metrics
print_a_function_from_file("pooled_metrics")
```

`signed_metrics()` is the reduced ladder for the subgroup leaderboard, where only the signed-effect rungs are pooled (directional agreement, Spearman ρ, Pearson r). This matches Section 2, above, which reports these three on estimate-only pairs (no standard error or inferential category).

```{r, results='asis'}
#| label: fn-signed-metrics
print_a_function_from_file("signed_metrics")
```

`cluster_boot()` attaches the uncertainty interval described above. It resamples the levels of a clustering variable (interventions for the ATE and subgroup panels, outcomes for the distributional panel) with replacement, recomputes the supplied metric function on each replicate, and returns the point estimate together with percentile confidence bounds in long form (one row per metric, with `value`, `lo`, `hi`). It is generic: the metric function is passed in, so the same bootstrap serves every panel.

```{r, results='asis'}
#| label: fn-cluster-boot
print_a_function_from_file("cluster_boot")
```

`metrics_by_group()` scores within each level of a grouping variable instead of pooling everything into one number: with `group = "outcome"` it returns one ladder per outcome (scored across the interventions), and with `group = "condition"` one ladder per intervention (scored across the outcomes). It drives the by-outcome and by-intervention distribution figures. Per intervention the score runs across `outcomes_continuous`, all on the same 0–100 slider scale, so the correlation rungs are not scale-confounded.

```{r, results='asis'}
#| label: fn-metrics-by-group
print_a_function_from_file("metrics_by_group")
```

`summarise_field()` reduces the per-approach scores to the field-level summary printed beside the headline figure: mean, median, sd, min, and max of each metric across approaches (excluding the reference row).

```{r, results='asis'}
#| label: fn-summarise-field
print_a_function_from_file("summarise_field")
```

The plotting helpers in `R/functions/plots.R` render these scores on a shared MetBrewer *Juarez* palette. `plot_pred_scatter()` draws predicted against human estimates pooled over approaches, with the identity line, the human replication reference marked, and optionally zero-centred symmetric axes; `calibration_overlay()` adds the average field and human calibration regressions and a corner readout of correlation and calibration coefficients on top of it. `plot_field_forest_single()` draws the single-metric ranking: approaches ordered by the metric with cluster-bootstrap intervals as horizontal bars, the noise-corrected companion as open points, the reference pinned on top, and an all-approaches summary row at the bottom (field density, mean with its 95% CI, and the mean of the noise-corrected companion as an open point). `plot_metric_distribution()` draws the distribution of approaches as a half-violin of the field over a dot strip — pooled (with mean, median, and optionally the mean's CI) or one row per group (outcome, intervention, or moderator, with the field mean and 95% CI per row). `plot_density_overlay()` overlays every Tier-1 approach's response density with the human reference and replication halves, one facet per outcome, with an optional per-facet metric stamp.

The remainder of this section demonstrates the code on placeholder data: **30 mock "teams"**, each a stratified bootstrap resample (within condition) of the single individual-level placeholder dataset, stacked with a `team` label — enough approaches to give the field distributions a meaningful shape. To exercise all three tier code paths, 28 teams enter at Tier 1 (individual-level), one is collapsed into a Tier 2 cell-level file, and one into a Tier 3 effect-level file. All of this placeholder data—including the resampled teams and the collapsed Tier 2/3 files—is built in the simulation script (`R/simulate_data_preregistration.R`), not here; this document only reads it. Because every mock team is just a resample of the same dataset, all approaches are random noise relative to the human reference: on placeholder data the field distributions therefore cluster near the null, and the figures only show that the code runs — real submissions will spread out. The adjusted correlation r~adj~ behaves accordingly: with no true signal, the reference effects' observed spread is all noise, so its guard (estimated true variance ≤ 0 → NA) binds for many mock teams and the surviving values are unstable — the intended behaviour, demonstrated rather than hidden. The code applies to real submissions without modification.

```{r}
#| label: variables

outcomes_primary   <- "trust_multidimensional"
outcomes_secondary <- c("trust_post", "distrust_post",
                         "funding_perceptions", "policy_role_mean", "inst_trust_mean")
outcomes_tertiary  <- c("belief_post", "concern_mean", "policy_general",
                        "policy_specific_mean", "behavior_mean")
outcomes_continuous <- c(outcomes_primary, outcomes_secondary, outcomes_tertiary)

# The two behavioral outcomes enter the pooled cross-team ladder alongside the
# 0-100 sliders (as in v1's pooled ATE row): donation in dollars via the OLS
# model, newsletter signup via logistic marginal effects on the probability
# scale. Scale-bound quantities (RMSE, the Winkler interval score, the
# by-intervention ladder) stay on `outcomes_continuous` only.
outcomes_behavioral <- c("donation_ams", "newsletter_signup")
outcomes_pooled     <- c(outcomes_continuous, outcomes_behavioral)

# Fixed, prespecified TOST equivalence bounds in outcome units — 3% of each
# scale's range: 3 points on the 0-100 sliders, $0.30 on the 0-10 donation,
# 3 percentage points on the signup proportion. Fixed rather than
# |ATE_h|-relative (v1's bound): a bound computed from the noisy estimate
# under test is not a valid alpha-level procedure, and the design analysis
# below shows the relative bound has no dynamic range at this design's
# reference SEs.
tost_deltas <- c(
  set_names(rep(3, length(outcomes_continuous)), outcomes_continuous),
  donation_ams = 0.3, newsletter_signup = 0.03
)

# Scale ranges, for expressing effects in percentage points of each outcome's
# range in the figures — following @hewittPredictingResultsSocial2025, whose
# predicted-vs-observed figure reports effects in pp on the original scales
# and deliberately avoids SD standardization (an under-dispersed submission
# would corrupt an SD denominator). The 0-100 sliders are already in pp; only
# the donation (0-10 dollars) and signup probability (0-1) rescale.
scale_range <- c(
  set_names(rep(100, length(outcomes_continuous)), outcomes_continuous),
  donation_ams = 10, newsletter_signup = 1
)

# Bootstrap replicates: the PREREGISTERED value for the real-data run is
# 2,000. This placeholder render is a demonstration of the procedure, not of
# its precision, so it uses a small demo value to keep the render cheap —
# only this parameter changes when the submissions arrive, nothing else.
n_boot_real <- 2000
n_boot_demo <- 200

# Illustrative outcome subset and categorical moderators — same definitions as
# Section 2 above; used for the subgroup-effects leaderboard.
outcomes_illustrative <- c("trust_multidimensional", "donation_ams",
                           "funding_perceptions", "policy_general")
moderators_cat <- c("gender", "age_band", "race", "education", "income", "party")

outcome_label_map <- c(
  trust_multidimensional = "Multidimensional trust (primary)",
  trust_post             = "Overall trust",
  distrust_post          = "Distrust",
  funding_perceptions    = "Funding perceptions",
  policy_role_mean       = "Policy role",
  inst_trust_mean        = "Institutional trust",
  belief_post            = "Climate belief",
  concern_mean           = "Climate concern",
  policy_general         = "General climate policy",
  policy_specific_mean   = "Specific climate policy",
  behavior_mean          = "Pro-climate behavior",
  donation_ams           = "Donation to AMS ($)",
  newsletter_signup      = "Newsletter signup"
)
```

```{r}
#| label: leaderboard-helpers

# Extract ATE pairs (estimate, SE, adjusted p) — same helper as the original
# preregistration's rq1-all chunk.
prep_ate <- function(df) {
  df |>
    select(condition, estimate) |>
    mutate(
      se    = if ("std.error"        %in% names(df)) df$std.error        else NA_real_,
      p_adj = if ("p.value_adjusted" %in% names(df)) df$p.value_adjusted else NA_real_
    )
}

# Inferential category from estimate sign + adjusted p-value (Tiers 1-2)
classify_sig <- function(estimate, p_adj) {
  case_when(
    !is.na(p_adj) & p_adj < .05 & estimate > 0 ~ "pos",
    !is.na(p_adj) & p_adj < .05 & estimate < 0 ~ "neg",
    TRUE                                       ~ "ns"
  )
}

control_lab <- levels(human_data$condition)[1]
z95 <- qnorm(0.975)

# Reference (Human 1) ATE models on the continuous outcomes, fit once and
# reused (the calibration regression reads the primary outcome's model).
human_models <- map(outcomes_continuous,
                    ~ run_main_treatment_model(human_data_1, outcome = .x)) |>
  set_names(outcomes_continuous)

# One ATE side (estimate, SE, BH-adjusted p) for an individual-level dataset,
# across all pooled outcomes: OLS with HC2 SEs for the continuous outcomes and
# the donation; logistic marginal effects on the probability scale for the
# binary newsletter signup (as in v1's pooled row). The binary SE is recovered
# from the marginal-effect CI.
ate_side <- function(dat, models = NULL) {
  cont <- map_dfr(c(outcomes_continuous, "donation_ams"), function(out) {
    res <- if (!is.null(models) && out %in% names(models)) models[[out]]
           else run_main_treatment_model(dat, outcome = out)
    prep_ate(res) |> mutate(outcome = out)
  })
  bin <- run_main_treatment_model_binary(dat,
                                         outcome = "newsletter_signup")$marginal_effects |>
    transmute(condition,
              estimate,
              se      = (conf.high - conf.low) / (2 * z95),
              p_adj   = p.value_adjusted,
              outcome = "newsletter_signup")
  bind_rows(cont, bin)
}

human_side <- ate_side(human_data_1, models = human_models) |>
  rename(estimate_h = estimate, se_h = se, p_adj_h = p_adj)

# Human per-condition cell sizes (named vector), for standardizing Tier-1
# submission-side inference to the human design's precision.
n_cells_h <- human_data_1 |> count(condition) |> deframe()

# Tier 1: pool ATE pairs from an individual-level dataset by refitting the
# preregistered treatment models on the submission. Submission-side INFERENCE
# (the inferential-agreement category and the TOST SE) is evaluated at the
# HUMAN design's per-arm precision: a team may simulate any pool size, so
# model-based significance would otherwise measure that choice rather than the
# method — and the Tier-1 precision requirement would mechanically hurt
# precise teams wherever a human effect is n.s. Each submission SE is rescaled
# by the two-cell design factor sqrt((1/n_h,k + 1/n_h,ctrl) /
# (1/n_l,k + 1/n_l,ctrl)) and p-values recomputed (normal approximation, BH
# within outcome, mirroring the Tier-2 path). Point estimates are untouched,
# so the correlation, directional, and RMSE rungs are unaffected. This also
# makes the inferential and TOST columns comparable across tiers: every
# submission's "significance" is evaluated at the same design precision.
build_pairs_t1 <- function(pred_data) {
  n_cells_l <- pred_data |> count(condition) |> deframe()
  fac <- sqrt((1 / n_cells_h + 1 / n_cells_h[[control_lab]]) /
              (1 / n_cells_l[names(n_cells_h)] + 1 / n_cells_l[[control_lab]]))
  ate_side(pred_data) |>
    mutate(se_std = se * unname(fac[condition])) |>
    group_by(outcome) |>
    mutate(p_adj = p.adjust(2 * pnorm(-abs(estimate / se_std)), method = "BH")) |>
    ungroup() |>
    mutate(infer_l = classify_sig(estimate, p_adj)) |>
    select(condition, outcome, estimate_l = estimate, se_l = se_std, infer_l) |>
    inner_join(human_side, by = c("condition", "outcome"))
}

# Tier 2: ATEs and SEs from submitted cell means and their 95% prediction
# intervals (mean, pi_lower, pi_upper per condition x outcome). Each cell
# interval implies an SE on the cell mean; the treatment and control cells
# combine for the between-subjects contrast. Normal-approximation p-values,
# BH-adjusted within outcome to mirror the Tier 1 models.
build_pairs_t2 <- function(cells_main) {
  cells_main |>
    mutate(
      condition = as.character(condition),
      se_cell   = (pi_upper - pi_lower) / (2 * z95)
    ) |>
    group_by(outcome) |>
    group_modify(function(cells, key) {
      ctrl <- cells |> filter(condition == control_lab)
      cells |>
        filter(condition != control_lab) |>
        mutate(
          estimate_l = mean - ctrl$mean,
          se_l       = sqrt(se_cell^2 + ctrl$se_cell^2),
          p_adj      = p.adjust(2 * pnorm(-abs(estimate_l / se_l)), method = "BH"),
          infer_l    = classify_sig(estimate_l, p_adj)
        ) |>
        select(condition, estimate_l, se_l, infer_l)
    }) |>
    ungroup() |>
    inner_join(human_side, by = c("condition", "outcome"))
}

# Tier 3: submitted ATEs taken as-is; inference from the submitted 95%
# uncertainty interval (significant iff the interval excludes zero, the direct
# analog of a two-sided test at alpha = .05); implied SE from the interval
# width for the TOST computation.
build_pairs_t3 <- function(effects) {
  effects |>
    mutate(
      condition  = as.character(condition),
      estimate_l = ate,
      se_l       = (pi_upper - pi_lower) / (2 * z95),
      infer_l    = case_when(
        pi_lower > 0 ~ "pos",
        pi_upper < 0 ~ "neg",
        TRUE         ~ "ns"
      )
    ) |>
    inner_join(human_side, by = c("condition", "outcome")) |>
    select(condition, outcome, estimate_l, se_l, infer_l,
           estimate_h, se_h, p_adj_h)
}

# `pooled_metrics()`, `signed_metrics()`, and `cluster_boot()` are defined in
# R/functions/statistics.R and shown under "Scoring functions" above.

# One leaderboard row: pooled ladder + calibration slope and R^2 on the
# primary outcome (HC2-robust inference: the outcome side of the calibration
# regression carries known, condition-varying sampling error) + the two
# interval scores for Tiers 2-3 (pi_metrics(); the Winkler score is
# scale-bound, so it runs on the 0-100 outcomes only). Interval scores are NA
# for Tier 1 (its SEs are sampling, not epistemic, uncertainty) and for the
# reference rows.
leaderboard_row <- function(pairs, submission, tier, disclosure) {
  calib <- run_calibration(
    human_models[[outcomes_primary]],
    pairs |> filter(outcome == outcomes_primary) |>
      select(condition, estimate = estimate_l),
    robust = TRUE
  )
  pi <- if (tier %in% c("2", "3"))
    pi_metrics(pairs |> filter(outcome %in% outcomes_continuous))
  else tibble(pi_coverage_pct = NA_real_, interval_score = NA_real_)
  pooled_metrics(pairs, tost_delta = tost_deltas) |>
    mutate(submission = submission, tier = tier, disclosure = disclosure,
           beta = calib$beta, calib_r2 = calib$r_squared,
           pi_coverage_pct = pi$pi_coverage_pct,
           interval_score  = pi$interval_score)
}
```

```{r}
#| label: mock-submissions

# Mock Tier 2 (cell-level) and Tier 3 (effect-level) submissions, generated
# outside this script alongside the other placeholder data (see
# R/simulate_data_preregistration.R) so that no mock-data simulation lives in
# the preregistration itself. Tier 2 collapses one placeholder team to the
# cell-level schema (*_cells_main.csv); Tier 3 gives ATEs vs. control with a
# widened 95% uncertainty interval standing in for a team's epistemic uncertainty.
cells_t2   <- readRDS(here("data/simulation_preregistration/mock_submission_tier2_cells.rds"))
effects_t3 <- readRDS(here("data/simulation_preregistration/mock_submission_tier3_effects.rds"))
```

```{r}
#| label: leaderboard-build

# 30 mock submissions plus the fixed human replication reference (Human 1 vs. Human 2).
# Teams 1-28 enter as individual-level Tier 1 data; team 29 as a Tier 2
# cell-level file and team 30 as a Tier 3 effect-level file, so all three tier
# code paths are exercised. Each submission's ATE pairs are built once here and
# reused by the headline and breakdown figures, the leaderboard table, and the
# companion tables.
tier1_ids <- 1:28

# Cached: the 29 Tier-1 builds each refit the treatment models incl. the
# logistic marginal effects for the newsletter outcome, the slow step of a
# cold render.
pairs_list <- cached("amendment_pairs_list", {
  c(
    set_names(
      map(tier1_ids,
          ~ build_pairs_t1(llm_data_teams |> filter(team == paste0("team_", .x)))),
      paste0("Team ", tier1_ids)
    ),
    list(
      "Team 29"           = build_pairs_t2(cells_t2),
      "Team 30"           = build_pairs_t3(effects_t3),
      "Human replication" = build_pairs_t1(human_data_2)
    )
  )
})

set.seed(2026)
submission_meta <- tibble(
  submission = c(paste0("Team ", 1:30), "Human replication"),
  tier       = c(rep("1", 28), "2", "3", "—"),
  disclosure = c(sample(c("A", "B", "C"), 30, replace = TRUE), "—")
)

# Floor baseline: a naive reference row for the rungs without a natural null
# (inferential %, TOST %, RMSE). "No effect" predicts a zero ATE everywhere —
# a fully naive, blinded-achievable strategy whose inferential agreement is
# whatever the share of truly null human effects hands it for free. It is a
# reference row like the reference, not a member of the field: excluded from the
# field violins and the field summary. (An oracle-mean floor — predict each
# outcome's mean human ATE everywhere — was considered and rejected: the mean
# is the in-sample least-squares constant fit to the very ATEs being scored,
# so it competes rather than floors; its interpretive content survives as the
# RMSE scale reference, the SD of the human ATEs.)
floor_pairs <- list(
  "Floor: no effect" = human_side |>
    mutate(estimate_l = 0, se_l = 0, infer_l = "ns")
)
floor_labels <- names(floor_pairs)

# Calibration is undefined for the floor (a constant prediction within the
# primary outcome -> zero-variance regressor), so their rows carry NA there.
# Its directional % reads 0, not 50: sign(0) matches neither sign, an artifact
# of the all-zero prediction, not below-chance skill (footnoted in the table).
floor_rows <- imap_dfr(floor_pairs, function(pr, sub) {
  pooled_metrics(pr, tost_delta = tost_deltas) |>
    mutate(submission = sub, tier = "—", disclosure = "—",
           beta = NA_real_, calib_r2 = NA_real_,
           pi_coverage_pct = NA_real_, interval_score = NA_real_)
})

leaderboard <- bind_rows(
  imap_dfr(pairs_list, function(pr, sub) {
    m <- submission_meta |> filter(submission == sub)
    leaderboard_row(pr, sub, tier = m$tier, disclosure = m$disclosure)
  }),
  floor_rows
) |>
  arrange(desc(submission == "Human replication"), desc(pearson_r))
```

```{r}
#| label: ate-ladder-labels

# Display labels / panel order for the metric ladder, and the null reference
# line per metric (50 = chance for directional agreement, 0 for the
# correlations; omitted for the percentage-agreement metrics where no clean
# null exists).
ladder_labels <- c(
  directional_pct = "Directional %",
  spearman_rho    = "Spearman ρ",
  pearson_r       = "Pearson r",
  pearson_adj     = "Pearson r (adj.)",
  inferential_pct = "Inferential %",
  tost_pct        = "TOST %"
)
ladder_nulls <- c("Directional %" = 50, "Spearman ρ" = 0, "Pearson r" = 0,
                  "Pearson r (adj.)" = 0)

relabel_metrics <- function(df, labels) {
  df |> mutate(metric = factor(labels[metric], levels = labels))
}
```

#### Finding 1 — How well can the field predict the experiment?

The headline question is not *who wins* but how well the current generation of approaches can predict a novel experiment at all. To keep that readable, each question gets **one key metric in the figures** — Pearson *r* for ATE recovery, Pearson *r* on the interaction estimates for subgroup recovery, and the variance ratio for response shape. The correlation is the literature's common currency [@hewittPredictingResultsSocial2025], and one shared axis across the ATE and subgroup sections makes them directly comparable; the variance ratio carries the field's one directional expectation (under-dispersion). The full preregistered ladders are computed for everything and reported in the companion tables.

@fig-headline shows the headline: the raw prediction task with the field's calibration lines and coefficients (A), and the field and its ranking on the pooled Pearson *r* (B). One rendering note on panel B: this placeholder demonstration runs the bootstrap with a small replicate count (`n_boot_demo = 200`) to keep the document cheap to build; the **preregistered value for the real-data run is `n_boot_real = 2,000`** — when the submissions arrive, only that parameter changes, nothing else about the procedure.

```{r}
#| label: fig-headline
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "Headline cross-team figure (placeholder data). (A) Predicted against human ATEs across all outcomes and approaches, in percentage points of each outcome's scale range (as in the predicted-vs-observed figure of Hewitt et al.; the 0–100 sliders are already in pp, the behavioral outcomes rescale) — one point per approach × intervention × outcome, axes centred on zero, the dashed line the identity, red crosses the human replication reference (Human 2 predicting Human 1). The solid lines are the average calibration regression of human on predicted, pooled across all outcomes: dark blue the field (per-approach fits averaged), red the human replication reference. The corner prints the mean field Pearson r (and its noise-corrected value) and the human replication's — the same pooled metric ranked in B — together with the field and human calibration coefficients (slope β, perfect = 1; intercept α, perfect = 0) with 95% CIs. (B) Approaches ranked by Pearson r pooled across all outcomes against Human 1: filled points with 95% cluster-bootstrap intervals over interventions, open points the noise-corrected r (adj.) where its guard allows a value, the human replication reference pinned on top (its value dropped as a dotted line). The bottom row summarises the field meta-analytically: the density of the approaches' scores, the diamond their mean with its 95% CI, and the open point the mean noise-corrected r. On placeholder data every approach is a noise resample of the same dataset, so the field clusters near r = 0 in panel B — no relationship to the truth, as expected of a random resample."

# (A) the raw prediction task: all approaches, all outcomes, in percentage
# points of each outcome's scale range
headline_pairs <- imap_dfr(pairs_list, ~ mutate(.x, submission = .y)) |>
  mutate(estimate_l = estimate_l * 100 / scale_range[outcome],
         estimate_h = estimate_h * 100 / scale_range[outcome])
p_scatter <- plot_pred_scatter(
  headline_pairs,
  x_lab       = "Predicted ATE (pp of scale range)",
  y_lab       = "Human ATE (pp of scale range)",
  tag         = "A",
  point_alpha = 0.15, point_size = 0.5,
  ref_alpha   = 0.6,  ref_size   = 1,
  symmetric   = TRUE
)

# Average field / human calibration lines + a corner readout of correlation
# (r, adj.) and coefficients (β, α with CI). Mean field Pearson r and its
# noise-corrected value come from the pooled leaderboard — the same metric
# ranked in panel B — so the raw task carries its one-number summary; the
# coefficients are fit from the plotted pp pairs by calibration_overlay().
field_ld <- leaderboard |> filter(!submission %in% floor_labels,
                                  submission != "Human replication")
human_ld <- leaderboard |> filter(submission == "Human replication")
p_scatter <- p_scatter +
  calibration_overlay(headline_pairs,
                      r_field    = mean(field_ld$pearson_r,   na.rm = TRUE),
                      radj_field = mean(field_ld$pearson_adj, na.rm = TRUE),
                      r_human    = human_ld$pearson_r,
                      radj_human = human_ld$pearson_adj)

# (B) ranked field on the key metric, with cluster-bootstrap intervals.
# n_boot_demo here; the real-data run uses n_boot_real = 2,000 (see variables).
pooled_f <- \(p) pooled_metrics(p, tost_delta = tost_deltas)
ate_forest <- cached("amendment_ate_forest", {
  imap_dfr(pairs_list, function(pr, sub) {
    cluster_boot(pr, pooled_f, cluster = "condition", n_boot = n_boot_demo) |>
      mutate(submission = sub)
  }) |>
    relabel_metrics(ladder_labels)
})
p_forest <- plot_field_forest_single(ate_forest, tag = "B")

(p_scatter | p_forest) + plot_layout(widths = c(1, 1.15))
```

The full pooled ladder stays scored for every approach; the field's centre and spread on each rung:

```{r}
#| label: field-pooled-compute

field_pooled <- leaderboard |>
  filter(!submission %in% floor_labels) |>   # floors are reference rows, not field members
  select(submission, all_of(names(ladder_labels))) |>
  pivot_longer(-submission, names_to = "metric", values_to = "value") |>
  relabel_metrics(ladder_labels)
```

```{r}
#| label: tbl-field-summary
#| tbl-cap: "Field summary (placeholder data): centre and spread of each pooled ATE-recovery metric across the 30 approaches (the human replication reference excluded)."

summarise_field(field_pooled) |>
  gt() |>
  fmt_number(columns = c(mean, median, sd, min, max), decimals = 2) |>
  cols_label(metric = "Metric", mean = "Mean", median = "Median",
             sd = "SD", min = "Min", max = "Max")
```

##### Selecting the best interventions

The correlation rungs summarize the whole effect pattern; a decision-maker's question is narrower — *if I trust this approach and act on its top picks, do I get the interventions that truly work best?* `topk_selection()` scores exactly that (see *Section 1*): per outcome, each approach's top **k = 3** interventions by predicted ATE (≈ the top 20% of 16 arms, in the spirit of the intervention-selection use case that motivates @hewittPredictingResultsSocial2025) are compared with the truly best three by the human reference — the **hit rate** (share of picks that are in the true best set; chance = 3/16 ≈ 0.19) and the **regret ratio** (mean true ATE of the picked set as a share of the best achievable; 1 = perfect pick). The table shows both on the primary outcome and averaged across the continuous outcomes.

```{r}
#| label: tbl-topk
#| tbl-cap: "Top-3 intervention selection (placeholder data): per approach, the hit rate (share of its top-3 picks that are among the truly best 3; chance ≈ 0.19) and the regret ratio (mean true ATE of the picked set / best achievable; 1 = perfect), on the primary outcome and averaged across the continuous outcomes. On placeholder data picks are random, so hit rates sit near chance."

topk_table <- imap_dfr(pairs_list, function(pr, sub) {
  scores <- topk_selection(pr, k = 3)
  bind_cols(
    scores |>
      filter(outcome == outcomes_primary) |>
      transmute(hit_primary = hit_rate, regret_primary = regret_ratio),
    scores |>
      summarise(hit_mean    = mean(hit_rate, na.rm = TRUE),
                regret_mean = mean(regret_ratio, na.rm = TRUE))
  ) |>
    mutate(submission = sub, .before = 1)
}) |>
  arrange(desc(submission == "Human replication"), desc(hit_mean))

topk_table |>
  gt() |>
  fmt_number(columns = c(hit_primary, regret_primary, hit_mean, regret_mean),
             decimals = 2) |>
  sub_missing(missing_text = "—") |>
  cols_label(
    submission     = "Submission",
    hit_primary    = "Hit rate (primary)",
    regret_primary = "Regret ratio (primary)",
    hit_mean       = "Hit rate (mean)",
    regret_mean    = "Regret ratio (mean)"
  ) |>
  tab_style(
    style     = cell_fill(color = "#eef3f7"),
    locations = cells_body(rows = submission == "Human replication")
  )
```

##### Where the field predicts well

Split **by outcome** (one distribution per outcome, scored across the interventions) and **by intervention** (one per intervention, scored across the 0–100 continuous outcomes, so the correlations are not scale-confounded by the donation and newsletter outcomes), the key metric shows *where* the field predicts well and what resists prediction (@fig-field-by-group). The full ladders per outcome and per intervention are in @tbl-field-by-outcome and @tbl-field-by-intervention.

```{r}
#| label: field-by-group-compute

# include_rmse: scale is constant within each outcome, so RMSE and its
# noise-corrected companion RMSE_adj are defined here (not in the pooled view).
field_by_outcome <- imap_dfr(pairs_list, function(pr, sub) {
  metrics_by_group(pr, "outcome",
                   f = \(p) pooled_metrics(p, tost_delta = tost_deltas,
                                           include_rmse = TRUE)) |>
    mutate(submission = sub)
}) |>
  mutate(outcome = factor(unname(outcome_label_map[outcome]),
                          levels = unname(outcome_label_map))) |>
  pivot_longer(c(directional_pct, spearman_rho, pearson_r, pearson_adj,
                 inferential_pct, tost_pct, rmse, rmse_adj),
               names_to = "metric", values_to = "value") |>
  relabel_metrics(c(ladder_labels, rmse = "RMSE", rmse_adj = "RMSE (adj.)"))

# Scale-bound view: restricted to the 0-100 continuous outcomes (the donation
# and newsletter outcomes, on other scales, would confound the correlation
# rungs here).
field_by_intervention <- imap_dfr(pairs_list, function(pr, sub) {
  metrics_by_group(pr |> filter(outcome %in% outcomes_continuous), "condition") |>
    mutate(submission = sub)
}) |>
  pivot_longer(c(directional_pct, spearman_rho, pearson_r, pearson_adj,
                 inferential_pct, tost_pct),
               names_to = "metric", values_to = "value") |>
  relabel_metrics(ladder_labels)

# Companion-table helper: field median per metric within each group level,
# with the human replication reference value in parentheses.
field_group_table <- function(field_long, group,
                              pct_metrics = c("Directional %", "Inferential %",
                                              "TOST %")) {
  fmt <- function(v, pct) if_else(is.na(v), "—",
                                  if_else(pct, sprintf("%.0f", v),
                                          sprintf("%.2f", v)))
  med <- field_long |>
    filter(submission != "Human replication") |>
    group_by(across(all_of(group)), metric) |>
    summarise(med = median(value, na.rm = TRUE), .groups = "drop")
  ref <- field_long |>
    filter(submission == "Human replication") |>
    select(all_of(group), metric, ref = value)
  med |>
    left_join(ref, by = c(group, "metric")) |>
    mutate(pct  = metric %in% pct_metrics,
           cell = paste0(fmt(med, pct), " (", fmt(ref, pct), ")")) |>
    select(all_of(group), metric, cell) |>
    pivot_wider(names_from = metric, values_from = cell)
}
```

```{r}
#| label: fig-field-by-group
#| fig-width: 11
#| fig-height: 6.5
#| fig-cap: "Where the field predicts well (placeholder data): distribution of each approach's Pearson r by outcome (A; scored across the interventions) and by intervention (B; scored across the 0–100 continuous outcomes). One dot per approach over the field's half-violin, with the field mean and its 95% CI (white diamond) on each row; red crosses mark the human replication reference, the dotted line the null."

by_outcome_r <- field_by_outcome |>
  filter(metric == "Pearson r") |>
  mutate(metric = factor("Outcomes"))
by_intervention_r <- field_by_intervention |>
  filter(metric == "Pearson r") |>
  mutate(metric = factor("Interventions"))

p_by_outcome <- plot_metric_distribution(
  by_outcome_r, x = "outcome", mean_ci = TRUE,
  ref_lines = c("Outcomes" = 0)
) + labs(tag = "A", y = "Pearson r")
p_by_intervention <- plot_metric_distribution(
  by_intervention_r, x = "condition", mean_ci = TRUE,
  ref_lines = c("Interventions" = 0)
) + labs(tag = "B", y = "Pearson r")

p_by_outcome | p_by_intervention
```

```{r}
#| label: tbl-field-by-outcome
#| tbl-cap: "Field summary by outcome (placeholder data): median across approaches of each ladder metric, scored per outcome across the interventions; the human replication reference in parentheses. RMSE and RMSE (adj.) are defined here because scale is constant within an outcome."

field_group_table(field_by_outcome, "outcome") |>
  gt() |>
  sub_missing(missing_text = "—") |>
  cols_label(outcome = "Outcome")
```

```{r}
#| label: tbl-field-by-intervention
#| tbl-cap: "Field summary by intervention (placeholder data): median across approaches of each ladder metric, scored per intervention across the 0–100 continuous outcomes; the human replication reference in parentheses."

field_group_table(field_by_intervention, "condition") |>
  gt() |>
  sub_missing(missing_text = "—") |>
  cols_label(condition = "Intervention")
```

##### Subgroup recovery

Whether an approach reproduces *who* an intervention moves, not just its average effect. For each Tier-1 submission we fit the preregistered condition × moderator interaction model (`run_moderator_model()`) across the illustrative outcomes and the six demographic moderators, then pool the interaction estimates against Human 1 (`compute_predicted = FALSE` skips the marginal-effects step, which the leaderboard does not use). Because interaction estimates require respondent-level data, this view covers Tier 1 only. As specified in Section 2 above, the signed-effect rungs are scored (directional agreement, Spearman ρ, Pearson r) plus the adjusted correlation r~adj~ — emphasized here, because small subgroup cells make the reference noise non-uniform across cells (see *Adjusted correlation and RMSE*).

```{r}
#| label: subgroup-pairs

# Human-1 interaction estimates, computed once and reused for every submission.
# The cached() keys keep their historical "amendment_" prefix (the analysis is
# unchanged); they are internal identifiers for analysis/results/*.rds, not tied
# to this file's name, so renaming the document does not require a recompute.
human_mod_side <- cached("amendment_human_mod_side", {
  map_dfr(outcomes_illustrative, function(out) {
    map_dfr(moderators_cat, function(mod) {
      run_moderator_model(human_data_1, outcome = out, moderator = mod,
                          compute_predicted = FALSE)$interaction_effects |>
        # se_h feeds the adjusted correlation; only the human reference side
        # is corrected (see "Adjusted correlation and RMSE").
        select(condition, moderator_level,
               estimate_h = estimate, se_h = std.error) |>
        mutate(outcome = out, moderator = mod)
    })
  })
})

build_subgroup_pairs <- function(pred_data) {
  pred_side <- map_dfr(outcomes_illustrative, function(out) {
    map_dfr(moderators_cat, function(mod) {
      run_moderator_model(pred_data, outcome = out, moderator = mod,
                          compute_predicted = FALSE)$interaction_effects |>
        select(condition, moderator_level, estimate_l = estimate) |>
        mutate(outcome = out, moderator = mod)
    })
  })
  inner_join(human_mod_side, pred_side,
             by = c("condition", "moderator_level", "outcome", "moderator"))
}

# Tier-1 submissions plus the reference are eligible. Interaction pairs are built
# once per submission and reused by the subgroup figure and its companion table.
subgroup_inputs <- c(
  set_names(
    map(tier1_ids, ~ llm_data_teams |> filter(team == paste0("team_", .x))),
    paste0("Team ", tier1_ids)
  ),
  list("Human replication" = human_data_2)
)
subgroup_pairs_list <- cached("amendment_subgroup_pairs", {
  map(subgroup_inputs, build_subgroup_pairs)
})
```

@fig-subgroup-headline shows the moderation analogue of the headline figure: the raw prediction task on the interaction estimates (A), and the key metric per moderator (B) — which kinds of heterogeneity the field can predict at all. The full signed-effect ladder, pooled, is in @tbl-field-subgroup.

```{r}
#| label: fig-subgroup-headline
#| fig-width: 11
#| fig-height: 4.2
#| fig-cap: "Subgroup recovery (placeholder data, Tier 1). (A) Predicted against human condition × moderator interaction estimates pooled across all approaches and the four illustrative outcomes (Section 2 scope), in percentage points of each outcome's scale range, axes centred on zero; dashed identity line, red crosses the human replication reference. The solid lines are the average calibration regression of human on predicted (dark blue the field, red the human replication), with a corner readout of the mean field and human Pearson r (and its noise-corrected value) and calibration coefficients (β, α with 95% CI). (B) Distribution of each approach's Pearson r on the interaction estimates, per moderator (pooled across the illustrative outcomes), with the field mean and its 95% CI (white diamond) per row; red crosses the reference, dotted line the null."

moderator_label_map <- c(gender = "Gender", age_band = "Age",
                         race = "Race/ethnicity", education = "Education",
                         income = "Income", party = "Party ID")

subgroup_scatter_pairs <- imap_dfr(subgroup_pairs_list,
                                   ~ mutate(.x, submission = .y)) |>
  mutate(estimate_l = estimate_l * 100 / scale_range[outcome],
         estimate_h = estimate_h * 100 / scale_range[outcome])
# ~37,000 pooled pairs: much lighter marks than the ATE scatter, or it reads
# as a solid blob.
p_sub_scatter <- plot_pred_scatter(
  subgroup_scatter_pairs,
  x_lab       = "Predicted interaction estimate (pp)",
  y_lab       = "Human interaction estimate (pp)",
  tag         = "A",
  point_alpha = 0.05, point_size = 0.3,
  ref_alpha   = 0.2,  ref_size   = 0.5,
  symmetric   = TRUE
)

# Mean field / human Pearson r on the interaction estimates (from the pooled
# signed-effect ladder, matching panel B), and the calibration lines + readout.
subgroup_ld <- imap_dfr(subgroup_pairs_list,
                        ~ signed_metrics(.x) |> mutate(submission = .y))
sub_field_ld <- subgroup_ld |> filter(submission != "Human replication")
sub_human_ld <- subgroup_ld |> filter(submission == "Human replication")
p_sub_scatter <- p_sub_scatter +
  calibration_overlay(subgroup_scatter_pairs,
                      r_field    = mean(sub_field_ld$pearson_r,   na.rm = TRUE),
                      radj_field = mean(sub_field_ld$pearson_adj, na.rm = TRUE),
                      r_human    = sub_human_ld$pearson_r,
                      radj_human = sub_human_ld$pearson_adj)

field_subgroup_by_mod <- imap_dfr(subgroup_pairs_list, function(pr, sub) {
  pr |>
    group_by(moderator) |>
    group_modify(~ signed_metrics(.x)) |>
    ungroup() |>
    mutate(submission = sub)
}) |>
  transmute(submission,
            value      = pearson_r,
            moderator  = factor(unname(moderator_label_map[moderator]),
                                levels = unname(moderator_label_map)),
            metric     = factor("Moderators"))
p_sub_field <- plot_metric_distribution(
  field_subgroup_by_mod, x = "moderator", mean_ci = TRUE,
  ref_lines = c("Moderators" = 0)
) + labs(tag = "B", y = "Pearson r")

(p_sub_scatter | p_sub_field) + plot_layout(widths = c(1, 1.3))
```

```{r}
#| label: tbl-field-subgroup
#| tbl-cap: "Field summary of subgroup recovery (placeholder data, Tier 1): centre and spread across approaches of each signed-effect metric on the condition × moderator interaction estimates, pooled across the illustrative outcomes and six moderators (human replication reference excluded)."

field_subgroup <- imap_dfr(subgroup_pairs_list, function(pr, sub) {
  signed_metrics(pr) |>
    pivot_longer(everything(), names_to = "metric", values_to = "value") |>
    mutate(submission = sub)
}) |>
  relabel_metrics(ladder_labels[c("directional_pct", "spearman_rho",
                                  "pearson_r", "pearson_adj")])

summarise_field(field_subgroup) |>
  gt() |>
  fmt_number(columns = c(mean, median, sd, min, max), decimals = 2) |>
  cols_label(metric = "Metric", mean = "Mean", median = "Median",
             sd = "SD", min = "Min", max = "Max")
```

##### Response-shape recovery

How well an approach reproduces the **shape** of the human response distribution, not only its mean. Using the preregistered `compare_distributions()` in the shared control condition, we score the overlap coefficient (OVL, 1 = identical densities), the **Wasserstein-1 distance** (W1, 0 = identical; the integral of the absolute ECDF difference, in scale points), the Kolmogorov–Smirnov distance (KS, 0 = indistinguishable), and the variance ratio (clone/human, 1 = matched spread) for each continuous outcome. W1 is the field's current standard distributional metric for bounded ordinal scales and is added here as the bandwidth-free companion to OVL: the KDE-based OVL depends on kernel bandwidths, which scale with the two samples' sizes (a large synthetic sample is smoothed differently from the human half) and leak density past the 0/100 scale bounds, whereas W1 has no tuning parameter and reads directly in scale points. The **variance ratio stays the headline diagnostic** (directional expectation: under-dispersion, ratio < 1). These are inherently per-outcome, control-condition quantities. @fig-distribution-headline therefore shows the raw response densities for every continuous outcome, one facet per outcome, with the field's median variance ratio stamped on each facet; all four metrics per outcome are in @tbl-field-distribution.

```{r}
#| label: distribution-pairs

dist_per_outcome <- function(pred_data) {
  map_dfr(outcomes_continuous, function(out) {
    compare_distributions(human_data_1, pred_data, out, control_lab,
                          include_w1 = TRUE) |>
      mutate(outcome = out)
  })
}

dist_labels <- c(ovl = "Overlap (OVL)", w1 = "Wasserstein-1",
                 ks_d = "KS distance", variance_ratio = "Variance ratio")

# Per-outcome shape metrics built once per submission, reused by the
# distribution figure and its companion table (cached: 29 x 11 KDE/ECDF
# passes over full respondent-level data).
dist_per_outcome_list <- cached("amendment_dist_per_outcome", {
  map(subgroup_inputs, dist_per_outcome)
})
```

```{r}
#| label: fig-distribution-headline
#| fig-width: 11
#| fig-height: 7
#| fig-cap: "Response-shape recovery (placeholder data, Tier 1): control-condition response densities for every continuous outcome. In each facet, every Tier-1 approach is a thin line, Human 1 (the reference half) the bold line, Human 2 (the replication half) dashed red; the corner stamp is the field's median variance ratio for that outcome (VR; approach ÷ human, 1 = matched spread, below 1 = under-dispersion — the expected failure mode of synthetic samples). The full shape metrics per outcome are in the companion table."

vr_ann <- imap_dfr(dist_per_outcome_list,
                   ~ mutate(.x, submission = .y)) |>
  filter(submission != "Human replication") |>
  group_by(outcome) |>
  summarise(label = sprintf("VR = %.2f", median(variance_ratio, na.rm = TRUE)),
            .groups = "drop")

# wrapped primary-outcome label: the full one clips its facet strip
density_labels <- outcome_label_map
density_labels["trust_multidimensional"] <- "Multidimensional trust\n(primary)"

plot_density_overlay(
  human_data_1, human_data_2, llm_data_teams,
  outcomes       = outcomes_continuous,
  condition_val  = control_lab,
  outcome_labels = density_labels,
  annotations    = vr_ann,
  x_lab          = "Response (0–100, control condition)"
)
```

```{r}
#| label: tbl-field-distribution
#| tbl-cap: "Field summary of response-shape recovery by outcome (placeholder data, Tier 1): median across approaches of each shape metric in the control condition, per continuous outcome; the human replication reference in parentheses. Ideal values: OVL = 1, W1 = 0, KS = 0, variance ratio = 1."

field_distribution <- imap_dfr(dist_per_outcome_list, function(d, sub) {
  d |>
    pivot_longer(c(ovl, w1, ks_d, variance_ratio),
                 names_to = "metric", values_to = "value") |>
    mutate(submission = sub)
}) |>
  mutate(outcome = factor(unname(outcome_label_map[outcome]),
                          levels = unname(outcome_label_map))) |>
  relabel_metrics(dist_labels)

field_group_table(field_distribution, "outcome", pct_metrics = character()) |>
  gt() |>
  sub_missing(missing_text = "—") |>
  cols_label(outcome = "Outcome")
```

#### Finding 2 — Ranking approaches (secondary)

Having characterised the field, we rank individual approaches. This is the secondary, structural read: each metric is its own ranking, and the interesting result is *which* approaches lead on *which* rung. The leaderboard **table** is the main summary; the ranking *visual* for the key metric — with its cluster-bootstrap uncertainty intervals — is panel B of @fig-headline.

```{r}
#| label: tbl-leaderboard
#| tbl-cap: "Cross-team leaderboard (placeholder data). Pooled ATE-recovery metrics across all outcomes, plus calibration slope and R² on the primary outcome and interval scores for Tiers 2–3, scored against the Human 1 reference half. The Human replication row (Human 1 vs. Human 2) shows how well humans replicate humans at this sample size; it is a reference, not an upper bound — a precise submission can legitimately exceed it. Metrics run least-to-most strict, left to right; Class is the disclosure-class badge (A open / B escrowed / C sealed)."

leaderboard |>
  select(submission, tier, disclosure, directional_pct, spearman_rho, pearson_r,
         pearson_adj, pearson_fz, inferential_pct, tost_pct, beta, calib_r2,
         pi_coverage_pct, interval_score) |>
  gt() |>
  fmt_number(columns = c(spearman_rho, pearson_r, pearson_adj, pearson_fz,
                         beta, calib_r2, interval_score),
             decimals = 2) |>
  fmt_number(columns = c(directional_pct, inferential_pct, tost_pct,
                         pi_coverage_pct), decimals = 0) |>
  sub_missing(missing_text = "—") |>
  cols_label(
    submission      = "Submission",
    tier            = "Tier",
    disclosure      = "Class",
    directional_pct = "Directional %",
    spearman_rho    = "Spearman ρ",
    pearson_r       = "Pearson r",
    pearson_adj     = "Pearson r (adj.)",
    pearson_fz      = "Pearson r (per-outcome)",
    inferential_pct = "Inferential %",
    tost_pct        = "TOST %",
    beta            = "Calib. β",
    calib_r2        = "Calib. R²",
    pi_coverage_pct = "PI coverage %",
    interval_score  = "Interval score"
  ) |>
  tab_footnote(
    footnote = "Fisher-z average of the per-outcome correlations. Divergence from the pooled Pearson r flags a submission that predicts which outcomes move without ranking the interventions within them.",
    locations = cells_column_labels(columns = pearson_fz)
  ) |>
  tab_footnote(
    footnote = "Tiers 2–3 only: coverage of the submitted 95% uncertainty intervals (honest intervals land near 95) and the mean Winkler interval score on the 0–100 outcomes (lower = better; a proper score, so neither padding intervals wide nor squeezing them tight pays).",
    locations = cells_column_labels(columns = pi_coverage_pct)
  ) |>
  tab_footnote(
    footnote = "The floor's directional % reads 0, not 50: sign(0) matches neither sign — an artifact of the all-zero prediction, not below-chance skill.",
    locations = cells_body(columns = directional_pct,
                           rows = submission %in% floor_labels)
  ) |>
  tab_style(
    style     = cell_fill(color = "#eef3f7"),
    locations = cells_body(rows = submission == "Human replication")
  ) |>
  tab_style(
    style     = cell_fill(color = "#f4f4f4"),
    locations = cells_body(rows = submission %in% floor_labels)
  )
```

##### Cross-metric consistency

A single-metric verdict can mislead: an approach that leads on one rung may trail on another, and @cumminsThreatAnalyticFlexibility2025b show empirically that performance on one data feature routinely fails to transfer to others. Rather than leave that to a reader's cross-panel comparison, we report it directly: the Spearman correlation between the approaches' rankings on each pair of ladder rungs (reference and floor rows excluded). High values mean the rungs sort the field similarly — a summary rank would be defensible; low or negative values mean each rung tells its own story and no single number should be trusted alone.

```{r}
#| label: tbl-metric-consistency
#| tbl-cap: "Cross-metric rank consistency (placeholder data): Spearman correlations between the approaches' rankings on each pair of pooled ladder metrics (reference and floor rows excluded). On placeholder data all scores are noise, so consistency is expected to be near zero."

metric_cols <- names(ladder_labels)

rank_consistency <- leaderboard |>
  filter(submission != "Human replication", !submission %in% floor_labels) |>
  select(all_of(metric_cols)) |>
  cor(method = "spearman", use = "pairwise.complete.obs")

rank_consistency |>
  as_tibble(rownames = "metric") |>
  mutate(metric = unname(ladder_labels[metric])) |>
  rename_with(~ unname(ladder_labels[.x]), all_of(metric_cols)) |>
  gt() |>
  fmt_number(columns = -metric, decimals = 2) |>
  sub_missing(missing_text = "—") |>
  cols_label(metric = "")
```

#### Design analysis: what the metrics can resolve at this sample size

Before locking the metric ladder, we check — from design constants and literature priors only, touching no data — whether each rung has *dynamic range* at this design: whether a genuinely good submission can score visibly better than the floor. The inputs are the design's cell sizes (≈500 per intervention and ≈1,000 in control per human half), a typical 0–100 slider SD of ≈25 points, and plausible spreads of true megastudy ATEs (SDs of roughly 0.5–2 points; large message-based megastudies rarely move attitude scales by more). Three conclusions follow. **First**, the reference half-sample ATE SE is ≈1.4 points — the noise floor under every raw metric. **Second**, the expected human-replication correlation and the expected correlation of a *perfect* predictor can be computed for each plausible true-ATE spread; the perfect predictor's is always the square root of the replication row's, which is the quantitative form of "the replication reference is beatable" (see *Overview*), and when the true spread is small relative to 1.4 the r~adj~ NA-guard is expected to bind for many entries. **Third**, v1's relative TOST bound Δ = 0.5 × |ATE~h~| (≈0.25–1 point) is *smaller than the smallest difference the test could ever confirm* (1.645 × se~diff~ ≈ 2.3 points even for a noiseless submission), so that rung would read ~0% for every entry including the human replication row — a dead metric. The preregistered fixed bound of **3 scale points** (3% of range) gives a perfect predictor roughly 40% per-pair equivalence power while remaining unreachable for badly wrong predictions: dynamic range restored, at the cost of a laxer notion of equivalence, which we accept and state.

```{r}
#| label: tbl-design-analysis
#| tbl-cap: "Blinded design analysis (no data touched): expected values of key scoring quantities from design constants (cell sizes ≈500/1,000 per half, outcome SD ≈25) under three plausible spreads of true ATEs. r (replication) is the expected human-replication correlation; r (perfect) the expected raw correlation of an error-free predictor — its square root relationship to the replication value is why the reference row is beatable. TOST columns give the per-pair probability that a perfect predictor declares equivalence under the old relative bound and the new fixed bound Δ = 3."

sd_y     <- 25
n_arm_h  <- 500
n_ctrl_h <- 1000
se_ate_h <- sd_y * sqrt(1 / n_arm_h + 1 / n_ctrl_h)
z05      <- qnorm(0.95)

tost_power <- function(delta, se) {
  margin <- delta - z05 * se
  if (margin <= 0) 0 else pnorm(margin / se) - pnorm(-margin / se)
}

tibble(tau = c(0.5, 1, 2)) |>
  mutate(
    se_reference    = se_ate_h,
    r_replication   = tau^2 / (tau^2 + se_ate_h^2),
    r_perfect       = sqrt(r_replication),
    tost_relative   = map_dbl(0.5 * tau, tost_power, se = se_ate_h),
    tost_fixed_3    = tost_power(3, se_ate_h)
  ) |>
  gt() |>
  fmt_number(columns = -tau, decimals = 2) |>
  cols_label(
    tau           = "SD of true ATEs (pts)",
    se_reference  = "Reference SE (pts)",
    r_replication = "r (replication)",
    r_perfect     = "r (perfect predictor)",
    tost_relative = "TOST power, relative Δ",
    tost_fixed_3  = "TOST power, fixed Δ = 3"
  )
```

#### Robustness of the scoring pipeline: split-seed sensitivity

The scoring pipeline has researcher degrees of freedom of its own — chiefly the preregistered random seed that splits the human sample into Human 1 and Human 2, and secondarily the TOST bound. @cumminsThreatAnalyticFlexibility2025b show that such analytic choices can move fidelity conclusions substantially, and a skeptic could reasonably ask whether the reported reference value reflects the split we happened to draw. We therefore report a **resplit sensitivity check**: the half-split is redrawn many times with different seeds, the human replication reference is recomputed on each, and the spread of the reference's metrics across resplits is reported alongside the primary (preregistered-seed) values. A narrow spread demonstrates that the seed is not driving the results. When real data arrive, the same check is run for the full field: every submission is rescored against each resplit's Human 1, and the stability of the field summary (and of leaderboard ranks) across resplits is reported.

```{r}
#| label: tbl-resplit
#| tbl-cap: "Split-seed sensitivity of the human replication reference (placeholder data, 50 resplits): distribution of each pooled ATE-recovery metric for the Human 1 vs. Human 2 comparison across 50 alternative random half-splits, next to the value under the preregistered seed. On placeholder data the reference is itself noise-level, so the check demonstrates the code, not a substantive reference value. The 50 resplits are drawn from the same sample and so are mutually dependent; the spread quantifies split-conditional variability, not independent replications."

resplit_ceiling <- cached("amendment_resplit_ceiling", {
  all_ids <- unique(human_data$id)

  map_dfr(1:50, function(i) {
    set.seed(1000 + i)
    ids_i <- sample(all_ids, size = floor(length(all_ids) / 2))
    h1_i  <- human_data |> filter(id %in% ids_i)
    h2_i  <- human_data |> filter(!id %in% ids_i)

    side_of <- function(dat, suffix) {
      map_dfr(outcomes_continuous, function(out) {
        prep_ate(run_main_treatment_model(dat, outcome = out)) |>
          mutate(outcome = out)
      }) |>
        rename_with(~ paste0(.x, suffix), c(estimate, se, p_adj))
    }

    inner_join(side_of(h1_i, "_h"), side_of(h2_i, "_l"),
               by = c("condition", "outcome")) |>
      mutate(infer_l = classify_sig(estimate_l, p_adj_l)) |>
      pooled_metrics() |>
      mutate(resplit = i)
  })
})

resplit_ceiling |>
  pivot_longer(-resplit, names_to = "metric", values_to = "value") |>
  filter(metric %in% names(ladder_labels)) |>
  group_by(metric) |>
  summarise(mean = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE),
            min = min(value, na.rm = TRUE),  max = max(value, na.rm = TRUE),
            n_defined = sum(!is.na(value)), .groups = "drop") |>
  left_join(
    # Restricted to the continuous outcomes for comparability with the
    # resplits (whose side_of() maps outcomes_continuous only).
    pairs_list[["Human replication"]] |>
      filter(outcome %in% outcomes_continuous) |>
      pooled_metrics() |>
      pivot_longer(everything(), names_to = "metric",
                   values_to = "preregistered_seed"),
    by = "metric"
  ) |>
  mutate(metric = factor(unname(ladder_labels[metric]),
                         levels = unname(ladder_labels))) |>
  arrange(metric) |>
  gt() |>
  fmt_number(columns = c(mean, sd, min, max, preregistered_seed), decimals = 2) |>
  sub_missing(missing_text = "—") |>
  cols_label(metric = "Metric", mean = "Mean", sd = "SD", min = "Min",
             max = "Max", n_defined = "N defined",
             preregistered_seed = "Preregistered seed")
```

When real submissions arrive, both findings are recomputed by replacing the placeholder datasets with the submitted prediction files routed through their tier's builder (`build_pairs_t1()` / `build_pairs_t2()` / `build_pairs_t3()`); the subgroup and distribution figures are produced for the tiers eligible for them, alongside the demographic-baseline and stereotyping diagnostics of Section 3 — including the demographic parity gap (`demographic_parity_gap()`). The approach-characteristics table coded from the registration forms is then joined to the leaderboard for the exploratory analysis of which design choices accompany better predictions (publicly disclosed features only; Class C entries excluded from this analysis, not from the leaderboard).

