---
title: "The Silicon Sample Benchmark: A multi-team preregistration for predicting behavioral-experiment effects"

abstract: |
    This is the preregistration for the **Silicon Sample Benchmark**, a multi-team prediction benchmark. We invite independent research teams to predict the results of a novel behavioral megastudy (16 message interventions to increase trust in climate scientists, 13 preregistered outcomes, a representative U.S. sample of ~18,000) with any AI-based approach they choose, from behavioral clones to group-level simulation to direct effect forecasting. Submissions enter at one of three tiers defined by how fine-grained the prediction is (individual-level synthetic data, cell-level statistics, or effect-level estimates with uncertainty intervals), and every approach is scored on the same preregistered metric ladder against the same human replication reference. At the time of writing, human data collection is complete (~18,000 responses). None of these data have been shared, published, or made accessible to any prospective team, and no team will be granted access before the prediction lock. Before any human data are released, every team registers its approach using a standardized method-registration form (an extended version of the GUIDE-LLM reporting checklist) and deposits its predictions on Zenodo, either publicly or partly escrowed under a disclosure policy that keeps every entry locked and verifiable even where it cannot be fully public.

date: "2026-06-11"

format:
  html:
    code-fold: true
    toc: true
    toc-depth: 4
  # PDF disabled for now to speed up iteration — re-enable for the final render.
  # pdf:
  #   toc: true
  #   toc-depth: 5
  #   include-before-body:
  #     text: |
  #       \clearpage
  #   colorlinks: true
  #   tbl-colwidths: true
  #   header-includes:
  #     - \usepackage{float}
  #     - \floatplacement{figure}{H}
  #     - \floatplacement{table}{H}
  #     - \usepackage{listings}
  #     - \lstset{breaklines=true, basicstyle=\small\ttfamily}

execute:
  echo: true
  warning: false
  message: false
  fig-width: 8
  fig-height: 5
  fig-dpi: 300
---

```{r}
#| label: setup
#| output: false

# packages
library(tidyverse)
library(sandwich)
library(lmtest)
library(marginaleffects)
library(broom)
library(MetBrewer)
library(gghalves)
library(patchwork)
library(gt)
library(tinytable)
library(here)

# custom functions (reused unchanged from the original preregistration)
source(here("R/functions/statistics.R"))
source(here("R/functions/plots.R"))
source(here("R/functions/tables.R"))
source(here("R/functions/reporting.R"))

# simulated (random) placeholder data
human_data     <- readRDS(here("data/simulation_preregistration/human_data_preregistration.rds"))
llm_data_teams <- readRDS(here("data/simulation_preregistration/llm_data_preregistration_teams.rds"))

# Restrict the human data to the scored conditions (the 16 non-interactive
# interventions + control) BEFORE any model is fit. The real cleaned human
# data will additionally contain the three chatbot arms and the value-
# similarity arm, which no submission predicts; fitting the human models on
# those arms would put the human side's BH adjustment over a different test
# family than the submissions'. A no-op on the placeholder data (which has
# scored conditions only), load-bearing when the real loader is wired in.
scored_conditions <- levels(droplevels(factor(llm_data_teams$condition)))
human_data <- human_data |>
  filter(condition %in% scored_conditions) |>
  mutate(condition = droplevels(condition))

# Same preregistered random split of the human sample as in the original
# preregistration. Human 1 is the reference half all submissions are scored
# against; Human 2 is the human replication reference.
set.seed(42)
split_ids    <- sample(unique(human_data$id), size = floor(length(unique(human_data$id)) / 2))
human_data_1 <- human_data |> filter(id %in% split_ids)
human_data_2 <- human_data |> filter(!id %in% split_ids)
```


::: {.callout-warning}
**This is an ongoing draft, to share with collaborators. This preregistration is not yet registered and might change still. We will register right after the deadline for the intent to participate (August 1st). This allows us to be exact about the number of participating teams.**
:::

::: {.callout-important}
**This is the canonical preregistration for the benchmark.** It is self-contained. The full analysis plan is restated here (see *Preregistered analysis plan*), so a participating team does not need to open the project's earlier single-team preregistration ([v1](preregistration.qmd)). The section *About this preregistration* explains why this second preregistration exists, what it keeps from v1, and every way it changes.

The benchmark's public-facing design (motivation, rules, the three tiers, the full method-registration form, the disclosure policy, scoring, and the timeline) is also described in the [call for participation](index.html). To avoid duplication, this document summarizes that design only as far as the analysis needs it, and concentrates on the **preregistered analysis plan**. Where the two overlap, the call is the canonical description of the procedure and this document is the canonical description of the analysis.

All datasets used for demonstration here are simulated random noise with the correct structure.
:::

## Motivation

Whether large language models can stand in for human participants is a consequential open question in behavioral science. The evidence so far is mixed, and has methodological issues: Most existing evaluations target experiments that were already published, and plausibly in the models' training data. Almost none are preregistered, so reported successes are entangled with researcher degrees of freedom in model and prompt choice [@cumminsThreatAnalyticFlexibility2025b].

We therefore aim to provide a preregistered, multi-team prediction benchmark, the *Silicon Sample Benchmark*. Independent teams predict the same novel megastudy (N ≈ 18,000) with any AI-based approach, and every approach is scored on the same preregistered metric ladder. The design addresses both problems. The target is novel, since none of the interventions has been tested in this form, and every prediction is deposited and time-stamped before any human data are released. The full rationale, rules, and incentives for participating teams are laid out in the [call for participation](index.html).

The benchmark scores the 16 non-interactive text interventions of the megastudy^[The original experiment contains 20 conditions. The benchmark scores only the 16 non-interactive text interventions. The four left out are the three interactive LLM-chatbot arms and the "Value similarity" quiz, all of which depend on a live back-and-forth that would substantially complicate the simulation task.], across self-reported attitudes (e.g., trust in climate scientists, support for funding of climate science) and behavioral measures (a donation to a scientific organization, a newsletter sign-up).

@tbl-interventions gives an overview of the 16 interventions, grouped by their content. All are short text messages (e.g., an article, an interview excerpt, corrected statistics, a portrait) shown once before the outcome battery. Together with the shared control condition they make up the 17 conditions every submission needs to cover.

```{r}
#| label: tbl-interventions
#| tbl-cap: "The 16 text-based interventions scored in the benchmark, grouped by the broad angle each takes on trust in climate scientists. The **Code name** column gives the internal condition identifier used in the raw survey files (`survey.qsf` / `survey_definition.json`), for teams that parse those directly. Each intervention has exactly one code name; where intervention-author teams were combined, the code name is a semicolon-joined list of animal pairs, and the semicolons are part of the name, so match on the full string and never split it (see `survey/condition_codenames.csv` in the benchmark repository). The original megastudy also fields three interactive LLM-chatbot arms and a 'Value similarity' quiz; these four are interactive and excluded from the simulation benchmark, leaving these 16 plus a shared control condition (17 conditions in total)."

interventions <- read_csv(here("data/interventions.csv"), show_col_types = FALSE)

tag_order <- c(
  "Collaboration and peer-review",
  "Scientific methods and results",
  "Applications and impact",
  "Others' endorsement",
  "Values",
  "Other"
)

# Same table layout as v1: category headers inserted
# as bold full-width rows, with a running intervention number.
intervention_table_data <- interventions |>
  mutate(across(c(title, summary, tag, code_name), str_trim)) |>
  filter(tag != "LLM-chatbot", title != "Value similarity") |>
  mutate(tag = factor(tag, levels = tag_order)) |>
  arrange(tag, title) |>
  mutate(number = as.character(row_number())) |>
  # internal condition code names exactly as in the raw survey files
  # (survey.qsf / survey_definition.json); semicolon-joined names are a
  # single identifier — semicolons are part of the name, never split.
  mutate(codename = code_name) |>
  group_by(tag) |>
  group_modify(~ bind_rows(
    tibble(title = as.character(.y$tag), codename = "", summary = ""),
    .x
  )) |>
  ungroup() |>
  mutate(` ` = ifelse(title %in% levels(tag), "", number)) |>
  select(` `, title, codename, summary) |>
  rename(`Intervention Title` = title, `Code name` = codename, Summary = summary)

# identify header rows
header_rows <- which(intervention_table_data$` ` == "")

intervention_table_data |>
  tt(width = c(0.04, 0.19, 0.17, 0.60)) |>
  format_tt(escape = TRUE) |>
  style_tt(align = "l", alignv = "t") |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = header_rows, bold = TRUE, j = 2, colspan = 3) |>
  style_tt(j = 3, monospace = TRUE) |>
  style_tt(fontsize = 0.85)
```

The goal of the benchmark is *not* to crown a single champion team. Instead, we aim to map the current design space of synthetic sample approaches and to evaluate, as a field, how well they predict human behavior. This benchmark study will provide a **catalogue**---a structured, comparable description of how research teams currently approach AI prediction of behavioral experiments (open- vs. closed-weight models, persona sources, administration protocols, fine-tuning, in-context learning) and which of these choices appear to matter. We will evaluate approaches on a ladder of metrics, from directional agreement to formal equivalence, plus distributional fidelity and stereotyping diagnostics (all specified in full under *Preregistered analysis plan*, below). We expect different approaches to win on different rungs.

## About this preregistration

**Why a second preregistration.** The project began as a single-team simulation study, preregistered as *"Predicting the effects of behavioral interventions on humans with LLMs"* (the original preregistration, **v1**). That plan asked how well one approach, an individual-level simulation built by members of the core team from demographically representative U.S. profiles with one particular model and prompt, recovers a novel megastudy. But a single team's result is only one point in a large space of ways to build a synthetic sample. We therefore decided to turn the project into a **multi-team benchmark** in which many independent teams predict the same novel experiment with any AI-based approach, and every approach is scored identically. This new preregistration plan supersedes the original **v1** preregistration.

**What it keeps from v1.** The human target study and the estimands (average treatment effects per intervention × outcome, subgroup interactions, control-condition response distributions, and demographic baselines), the three-section analysis structure, the least-to-most-strict metric ladder, and the preregistered split-half **human replication reference** are all carried over from v1. All are restated below (see *Preregistered analysis plan*) to keep this new preregistation self-contained. The model and comparison functions are reused unchanged.

**What it changes and adds.** The changes fall into three groups. Group A redesigns the study, group B changes what it reports, and group C adds scoring tools from the recent silicon-sample literature. Each change is summarized here, with a pointer to the section that specifies it in full.

Because this preregistration is written for teams who are blinded to the human outcome data, revising the analysis plan does not carry the usual risk of data-contingent researcher degrees of freedom.

**A. From one approach to a field.** These change the *design*, leaving v1's analysis pipeline untouched.

1. **Single team → multi-team benchmark.** The project now evaluates a *field* of independent submissions rather than one core-team approach. It allows for submissions at different levels (Tier 1 individual-level synthetic data, Tier 2 cell-level statistics, Tier 3 effect-level estimates with uncertainty intervals). See *Benchmark design*.
2. **Method registration and disclosure.** Every team documents its pipeline on a standardized method-registration form (an extended GUIDE-LLM checklist) and deposits its predictions under a disclosure policy that keeps even proprietary entries verifiable. See *Method registration and disclosure*.
3. **Per-tier submission specifications.** Each tier has an exact file schema. We will compare predictions across the different tiers. For instance, for Tiers 2-3, we will reconstruct effect standard errors from the uncertainty intervals. See *Submission specification* and *Tier-specific estimand recovery*.

**B. Additional benchmark design choices.** Benchmark specific results are mere add-ons to the orignial v1 pipeline. No estimand, model, or metric inherited from v1 is redefined.

4. **The headline result is the field, not a single scorecard.** The primary result is no longer one approach's report card but the distribution of predictive quality across the whole field of approaches. Every approach is scored against a split-half **human replication reference** (how well one half of the human sample predicts the other). Teams may enter more than once, so the field summary counts one primary entry per team. A team may also choose to predict only part of the study rather than every intervention and outcome. The main leaderboard is limited to *full-coverage* entries, so its scores are all measured on the same basis, and partial entries are reported separately. See *Cross-team comparison*.
5. **One score per metric, pooled across outcomes.** A single data point is one predicted effect placed next to the matching human effect, for example how much a particular message shifted trust in climate scientists. We deliberately do *not* collapse the metrics into one overall "champion" score. An approach can get the direction of effects right while mis-scaling their size, or do well on the main outcome and badly on the rest, and a single number would hide exactly the structure the benchmark exists to reveal. Each metric is therefore read on its own. See *Cross-team comparison* (*Scoring principle*). That said, in our main figures, we will report correlations (and adjusted correlations).
6. **An uncertainty interval on every score.** The benchmark tests 16 interventions, but these are only a sample of all possible kinds of messages one could have tested. A different study could have used 16 other messages. An approach's score is therefore an estimate, and like any estimate it would move somewhat under a different draw of interventions. With bootstrap sampling, we attach an interval to every score showing how much it might move based on message selection. For the same reason, we never call one approach better than another from the ranking alone. See *Uncertainty intervals on the leaderboard*.

**C. New evaluation add-ons from the silicon-sample literature.** The remaining additions extend the orignally proposed evaluation metrics from v1. The new metrics are motivated by how recent studies have evaluated synthetic samples. They supplement or improve the inherited metrics from v1.

7. **A noise-corrected correlation.** The human effects an approach is scored against are estimated from a (here, half-split) sample which is not perfectly representing the population.
The resulting sampling noise can drag down the correlations between human and synthetic sample effects. Because the sampling noise is unrelated to predictive skill, we try to account for it. Following @ashokkumarLargeLanguageModels2026a, we report a noise-corrected correlation beside the raw one (the raw value stays the ranked metric). See *Comparison functions*. 
8. **Top interventions match** Megastudies with many interventions are often run to provide practically relevant advice on which interventions are worth acting on. We therefore score directly, whether the same advice emerges from the human vs. the synthetic samples[@ashokkumarLargeLanguageModels2026a]? See *Section 1*.
9. **A floor for the rungs without a natural zero.** Some rungs (inferential agreement, equivalence, and RMSE) have no natural zero to read them against. For example, on inferential agreement a lazy "predict no effect for everything" strategy could score deceptively well when most true effects are small and statistically insignificant. Adding a naive "do-nothing" approach as a floor reference makes these scores more interpretable. Each approach can then be interpreted by its distance to that floor, and to the human reference [@ashokkumarLargeLanguageModels2026a; @cumminsThreatAnalyticFlexibility2025b]. See *Comparison functions*.
10. **A fixed bound for the equivalence test.** One rung asks whether a predicted effect is *close enough* to the human effect to count as equivalent, which needs a bound for "close enough." v1 set that bound as a fraction of the human effect being tested. But that effect is itself measured with error, so the bound moved with the noise and the test was not valid. We now fix the bound in advance, at a set fraction of each outcome's scale, so it no longer depends on the numbers under test. See *Comparison functions*.
11. **Subgroup distributions.** Averaging outcome distributions over demographic groups can hide an approach that serves most groups well but consistently misses one. We report the gap between the worst- and best-served group in terms of distribution similarity [@parkLLMAgentsGrounded2026]. Because too-similar synthetic responses (under-dispersion) are the failure mode the literature flags most often, we make under-dispersion the headline distribution diagnostic. See *Section 3*, and *Section 1* (response distributions).
12. **Human sample split robustness.** The human sample is divided in half at random to build the reference. That split is fixed before any data are seen, so it is not a researcher degree of freedom. But the particular split we drew is still arbitrary, and another random split would serve just as well. To show the results do not hinge on the one we happened to draw, we recompute them across many alternative splits and report how stable they are [@cumminsThreatAnalyticFlexibility2025b]. See *Cross-team comparison*.
15. **Scoring the prediction intervals themselves.** Teams in Tiers 2–3 submit not only a predicted effect but an uncertainty interval around it, and we use that interval to decide whether the effect counts as statistically significant (does the interval exclude zero?). This opens a loophole. A team that submits very wide intervals might have every interval include zero, so every prediction reads "not significant". Since most effects in behavioral experiments are expected to be relatively small, potentially statistically insignificant, matching "not significant" everywhere might earn a high agreement score with no real skill, just by being vague. To close the loophole we also score the intervals directly, checking whether they are well-calibrated (containing the true effect about as often as they claim to) and appropriately narrow. Being vague then carries its own cost. See *Tier-specific estimand recovery*.
16. **One unit for all outcomes.** The outcome measures use different scales, with 0–100 attitude sliders, a $0–10 donation, and a 0–1 signup probability. To make effects on these scales comparable, we convert them to **percentage points of its outcome's scale range** before any cross-outcome scoring, following the practice of @ashokkumarLargeLanguageModels2026a. The binary newsletter outcome is added to the subgroup analysis on the same footing, via the linear probability model. See *Cross-team comparison* (*Scoring principle*) and *Section 2*.

## The human target sample

Predictions are scored against the human data from the megastudy itself. Participants are recruited from a non-probability opt-in panel, quota-matched (age, gender and race) to the population of U.S. residents, all aged 18 or older, who pass a series of attention and bot-detection checks before treatment assignment.

The benchmark targets **N ≈ 18,000 complete responses**, with 1,000 per intervention across the 16 text-based interventions plus 2,000 in the shared control condition. Recruitment uses **census-based cross quotas** on gender × age and gender × race/ethnicity to approximate the U.S. adult population. The targets below match the proportions of the parent megastudy, with counts rescaled to a total N ≈ 18,000. Participants selecting "Other" as their gender are not quota-constrained.

The targets below are drawn from the 2024 U.S. Census Bureau Population Estimates Program. *Total* is the share within each category. *Male* and *Female* give the within-category gender split. The table describes the key demographics of the human sample. Participating teams are not required to use this information for their own synthetic sample.

```{r}
#| label: tbl-quotas
#| tbl-cap: "Sampling quota targets (N ≈ 18,000), from the 2024 U.S. Census Bureau Population Estimates Program. Proportions match the parent megastudy, with counts rescaled to N ≈ 18,000, while the target of the megastudy is N≈ 22,000."

# Same layout as the interventions table: category headers inserted as bold
# full-width rows above their levels (tinytable, for a consistent look).
quotas <- read_csv(here("data/quotas_18000.csv"), show_col_types = FALSE)

quota_table_data <- quotas |>
  group_by(Variable) |>
  group_modify(~ bind_rows(
    tibble(Category = as.character(.y$Variable), Total = "", Male = "", Female = ""),
    .x
  )) |>
  ungroup() |>
  select(Category, Total, Male, Female) |>
  rename(` ` = Category)

header_rows_q <- which(quota_table_data$Total == "")

quota_table_data |>
  tt(width = c(0.40, 0.20, 0.20, 0.20)) |>
  format_tt(escape = TRUE) |>
  style_tt(align = "l", alignv = "t") |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = header_rows_q, bold = TRUE) |>
  style_tt(fontsize = 0.85)
```

Full sampling and exclusion detail is in the [human study preregistration](https://doi.org/10.5281/zenodo.20160212).

## Benchmark design

Participation is by invitation. Each team builds an AI-based prediction pipeline from the shipped materials, including intervention texts, the full survey instrument, the codebook, the registration template. 

Teams can use any models, personas, prompts, fine-tuning, retrieval, or prior literature they want. Before the prediction lock deadline, each team registers its approach by depositing its synthetic sample/predictions and materials on Zenodo, hashed and time-stamped. After the prediction lock deadline, every submission runs through the preregistered analysis pipeline unchanged and is reported on a public leaderboard alongside the human replication reference, and in a joint manuscript with all participating teams as co-authors. There are three ground rules for an entry to be valid: **No data access.** By design, no participating team has accesses to the human data. **Pipelines are AI-based and automated.** Predictions may be made based on previously published human data, but no new human data specific to this project may be collected. **Everything is registered** (see the disclosure policy below). The full procedure, ground rules, timeline, and incentives are in the [call for participation](index.html).

**Blinding and data access.** The collected human data (all ~18,000 responses; collection complete at the time of writing) have been accessed only by the research lead, Jan Pfänder, for data-quality monitoring. As a result, members or affiliates of the core team who have had no access to any human outcome data may enter the benchmark as regular teams under the same rules as everyone else (same registration form, same blinding attestation, same deadlines). 

For full disclosure, Jan Pfänder presented some very preliminary results (human megastudy sample at the time N ≈ 3000) from an initial simulation attempt (following preregistration **v1**) in a five-minute talk at a closed academic workshop at the Max Planck Institute for Human Development in Berlin (May 2026). This is the only presentation that has targeted an audience of potential collaborators on the benchmark study. It focused on the prediction ability of the simulation approach and revealed nothing about how humans behaved in the megastudy. Two other talks were explicitly about the human megastudy results. Jan Pfänder presented preliminary results from the human megastudy at an invited talk at the Max Planck Institute for Human Development in Berlin (May 2026, N at the time ≈ 3000 of 22000), and at a small conference at the University of Hohenheim (June 2026, N at the time ≈ 13000 of 22000). No one in the audience of these talks will be invited to participate in the benchmark study. In addition, since we ask for extensive disclosure of materials, a "cheating" prompt (e.g., "make sure intervention XX does best") would surface in the reporting framework that teams are required to fill out.

### Three ways to compete

Submissions enter at one of **three tiers**, defined by the granularity of the prediction. Teams choose the highest tier their approach supports. Lower tiers are simply eligible for fewer analyses.

- **Tier 1 · Individual-level synthetic data.** Eligible for everything, meaning ATE recovery, calibration regression, response distributions, subgroup heterogeneity, demographic baselines, and stereotyping diagnostics. *Grain: 1 row = 1 synthetic respondent.*
- **Tier 2 · Group-level treatment effect predictions.** Eligible for ATEs, calibration, subgroup heterogeneity, and demographic baselines. Not eligible for the distribution-shape metrics (OVL, KS) or the predictability regressions. *Grain: 1 row = 1 cell (condition × [moderator level ×] outcome).*
- **Tier 3 · Average treatment effect predictions.** Eligible for directional agreement, Spearman ρ, Pearson r, RMSE, and calibration regression. Inferential agreement and TOST are computed from the submitted intervals. *Grain: 1 row = 1 intervention × outcome estimate.*

### Analysis eligibility by tier

| Preregistered analysis | Tier 1 | Tier 2 | Tier 3 |
|---|:-:|:-:|:-:|
| ATE recovery, six-metric ladder (Section 1) | ✓ | ✓ | ✓ * |
| Calibration regression α / β (Section 1) | ✓ | ✓ | ✓ |
| Response distributions, variance ratio, OVL, KS (Section 1) | ✓ | — | — |
| Subgroup heterogeneity, condition × moderator (Section 2) | ✓ | ✓ | — |
| Within-subgroup distributions (Section 2) | ✓ | — | — |
| Demographic baseline calibration (Section 3) | ✓ | ✓ | — |
| Demographic predictability / stereotyping (Section 3) | ✓ | — | — |

\* For Tiers 2 and 3, inferential agreement and equivalence (TOST) are evaluated against the submitted 95% uncertainty intervals rather than model-based standard errors. For Tier 1, standard errors fall out of the fitted models, so no separate uncertainty input is required.

## Method registration and disclosure

Before the prediction lock, every team documents its pipeline on a structured **method-registration form** and deposits it alongside the predictions. The form builds on the GUIDE-LLM reporting checklist [@feuerriegelReportingChecklistLarge2026]. The form and the **disclosure policy** (which lets proprietary teams seal items under escrow while keeping the lock publicly verifiable) are published in the [call for participation](index.html#registration-form). We provide a template in the [benchmark submission template repository](https://github.com/janpfander/silicon-sample-submission). 

## Submission specification

The deposit logistics (Zenodo record structure, the `metadata.json` file, file naming, the self-check validator, and the public/escrowed split) are specified on the [materials page](materials.html) and in the [benchmark submission template repository](https://github.com/janpfander/silicon-sample-submission). Two requirements bear directly on the analysis. Condition and column labels must match the published codebook exactly. Submissions are validated before the lock, and cells that fail to parse are dropped. Because the share of usable output is itself an analytic choice that can vary widely across pipelines [@cumminsThreatAnalyticFlexibility2025b], the share of dropped cells is reported per submission, alongside its coverage, on the leaderboard.

### Tier 1 — individual-level dataset

One row per synthetic respondent, matching the structure of the cleaned analysis data (`cleaned_llm.rds` schema). Each team supplies its own synthetic respondents. `profile_id` is a team-assigned identifier, unique within the submission, and assignment of respondents to conditions (between-subjects, covering all 17 conditions) is the team's responsibility, registered under form items D.1 and D.3.

| Column(s) | Type | Specification |
|---|---|------|
| `profile_id` | string | A unique respondent identifier you assign (unique within the submission). |
| `condition` | factor | Control or one of the 16 intervention labels, exactly as in the codebook. |
| `gender`, `age_band`, `race`, `education`, `income`, `party` | factor | The six moderators, levels per codebook (from your own profiles). |
| `trust_multidimensional` + 12 trust items | numeric | Primary composite 0–100. The 12 underlying items per codebook scale. |
| `trust_post`, `distrust_post`, `funding_perceptions`, `policy_role_mean`, `inst_trust_mean` | numeric | Secondary outcomes, original scales per codebook. |
| `belief_post`, `concern_mean`, `policy_general`, `policy_specific_mean`, `behavior_mean` | numeric | Tertiary outcomes, original scales per codebook. |
| `donation_ams` | numeric | Donation to AMS in $, per codebook. |
| `newsletter_signup` | 0 / 1 | Binary behavioral outcome. |

Because a Tier 1 submission has the same shape as the human participant data, the same analysis code applies. 

**Precision requirement.** An approach's estimated effects are only as precise as the synthetic sample behind them. The fewer synthetic respondents, the noisier each estimated effect, and a very small sample would score poorly for a reason that has nothing to do with the method. A Tier 1 submission must therefore contain at least as many synthetic respondents as the human half it is scored against, meaning **500 per intervention and 1,000 in the control condition**. At that size the submission's effects carry about the same sampling noise as the human reference. We encourage teams to go well beyond this minimum. A larger synthetic sample gives more precise effect estimates that reflect the method itself rather than the luck of a particular draw, and synthetic respondents are relatively cheap to generate.

### Tier 2 — cell-level file(s)

Two long-format files. `*_cells_main.csv` has one row per condition × outcome, and `*_cells_moderator.csv` one row per condition × moderator × moderator-level × outcome.

| Column | Type | Specification |
|---|---|------|
| `condition` | factor | Per codebook. |
| `moderator`, `moderator_level` | factor | Moderator file only. Levels per codebook. |
| `outcome` | string | Outcome variable name per codebook. |
| `mean` | numeric | Predicted cell mean on the original scale (proportion for `newsletter_signup`). |
| `pi_lower`, `pi_upper` | numeric | 95% uncertainty interval bounds on the cell mean. This is the team's own uncertainty about the *human* cell mean, not a within-simulation SD. The ATE interval is derived by combining the treatment- and control-cell intervals under an independence assumption (see *Tier-specific estimand recovery*), so uncertainty about the overall response level — which is shared across cells and cancels in the ATE — should be left out of the cell intervals; otherwise the derived ATE intervals come out too wide. Used for inferential-agreement and equivalence scoring. |

### Tier 3 — effect-level file

One row per intervention × outcome (16 × 13 = 208 rows for full coverage). Estimates are ATEs versus control on the original outcome scale. The binary outcome is on the probability scale.

| Column | Type | Specification |
|---|---|------|
| `condition` | factor | Intervention label per codebook. |
| `outcome` | string | Outcome variable name per codebook. |
| `ate` | numeric | Predicted ATE, original units. |
| `pi_lower`, `pi_upper` | numeric | 95% uncertainty interval bounds. This is the team's own uncertainty about the *human* ATE, not a standard error. Used for inferential-agreement and equivalence scoring. |

## Preregistered analysis plan

### Overview and the human replication reference

Each submission is scored against the same human sample. The human sample is split into two halves by a preregistered random seed. **Human 1** is the reference half every submission is scored against. **Human 2** predicts Human 1 like any submission would, and its scores define the **human replication reference**, which shows how well a fresh human sample of this size reproduces the reference half. This human reference is evaluated in the same way a Tier-1 synthetic submission. A **floor baseline** ("no effect") anchors the other end for the metrics without a natural null (see *Cross-team comparison*). Because the split seed is an arbitrary choice, we will provide robustness checks across many alternative splits, so that results do not hinge on the seed we happened to draw (*Robustness of the scoring pipeline*, below). 

Which of the analyses below applies to a given submission follows from its tier (see *Analysis eligibility by tier*). Tier 1 is eligible for all of them, Tiers 2–3 for the estimate-based subset.

### Estimation functions

Every effect size compared below is produced by the same model functions, applied in the same way to the human reference and to each submission. They produce the average treatment effects per intervention for Section 1 and the condition × moderator interactions for Section 2. These functions are unchanged from v1, and their full statistical rationale is in the [human study preregistration](https://doi.org/10.5281/zenodo.20160212). We reproduce them here so this document stands alone.

`run_main_treatment_model()` estimates each intervention's ATE against the shared control by OLS with HC2 heteroskedasticity-robust standard errors and BH-adjusted p-values.

```{r, results='asis'}
#| label: fn-run-main-treatment-model
print_a_function_from_file("run_main_treatment_model")
```

`run_main_treatment_model_binary()` is the counterpart for the binary newsletter-signup outcome. It fits a logistic regression and returns marginal effects on the probability scale.

```{r, results='asis'}
#| label: fn-run-main-treatment-model-binary
print_a_function_from_file("run_main_treatment_model_binary")
```

`run_moderator_model()` estimates the condition × moderator interactions used for the subgroup analysis (Section 2). For the binary newsletter-signup outcome the benchmark's subgroup scoring runs this same OLS function as a **linear probability model**. 

```{r, results='asis'}
#| label: fn-run-moderator-model
print_a_function_from_file("run_moderator_model")
```

### Comparison functions

The comparison metrics are implemented by a small set of helper functions, applied symmetrically to the human reference and to each submission. They are unchanged from v1 and shown here in full.

#### Comparison metrics

`compare_estimates()` is used throughout, for ATEs (Section 1), moderator interaction estimates (Section 2), and secondary analyses. It computes six metrics, reported in increasing order of strictness.

- **Directional agreement (%)**. The share of estimates with the same sign as the human estimate, where chance is 50%. Does the approach at least get the direction of each effect right?
- **Spearman ρ**. The rank correlation of the point estimates. Does the approach rank the interventions in the same order as humans? It is insensitive to scale or offset, so an approach that inflates all effects equally still scores ρ = 1.
- **Pearson r**. The linear correlation of the point estimates. It captures whether predicted and human effects rise and fall together in proportion, but not whether their absolute size matches.
- **Inferential agreement (%)**. The share of estimates with the same BH-adjusted conclusion (significantly positive, significantly negative, or not significant at α = .05). Getting this right requires both magnitude and uncertainty. Would a researcher using the submission reach the same conclusion as one using the human data?
- **RMSE**. The root mean squared error, the average size of the prediction errors. In the cross-team analysis every estimate is expressed in **percentage points (pp) of its outcome's scale range** before scoring (see *Scoring principle*, below), so RMSE reads in pp and is reported both pooled across outcomes and per outcome, matching @ashokkumarLargeLanguageModels2026a, whose primary archive reports RMSE pooled across experiments in exactly these units. To judge whether an RMSE is good, read it against the **spread (SD) of the human ATEs**. That spread is the error of the best possible constant prediction, so a submission with real intervention-specific skill must come in below it. (In v1, which scores each outcome on its own raw scale, RMSE stays per outcome and is omitted from the pooled row.)
- **TOST (%)**. The share of estimate pairs passing an equivalence test (two one-sided tests, α = .05) with a **fixed, prespecified bound of 3% of each outcome's scale range**, which is a uniform Δ = 3 once effects are in pp (3 points on the 0–100 sliders, $0.30 on the donation, 3 percentage points on the signup proportion). This replaces v1's relative bound Δ_k = 0.5 × |ATE_human_k| for two reasons. First, a bound computed from the noisy estimate under test is not a valid equivalence procedure, since bounds must be set independently of the observed data. Second, the design analysis (see *Cross-team comparison*) shows the relative bound is smaller than the smallest difference the test could confirm at this design's reference SEs, so that rung would read about 0% for every entry. 

```{r, results='asis'}
#| label: fn-compare-estimates
print_a_function_from_file("compare_estimates")
```

#### Adjusted correlation and RMSE (noise-corrected)

The effects in the human sample are noisy. That noise mechanically drags the correlation between effects of the human and synthetic sample down, because it inflates the observed spread of the reference effects. The raw Pearson r divides by that inflated spread and is biased toward zero. The same noise inflates the RMSE.

Wherever Pearson r appears in the cross-team analysis we therefore report, beside the raw value, a noise-corrected correlation **r~adj~** and a noise-corrected RMSE (`adjusted_metrics()`). The correction works by subtraction. Because the sampling noise is independent (that's an assumption we make), the true spread of the reference effects is their observed spread minus the average known sampling variance. The adjusted correlation divides the covariance by that corrected spread (Spearman disattenuation with known error variances). This method is similar to the meta-analytic r~adj~ of @ashokkumarLargeLanguageModels2026a, but uses fewer assumptions. As @ashokkumarLargeLanguageModels2026a, we will report r and r~adj~ side by side in the manuscript. This allows us to directly compare our results to their larger megastudy archive, the closest analogue to this benchmark^[We take Ashokkumar et al.'s megastudy figures, not their headline r = .85, as the reference because the two numbers measure different tasks. The .85 is a correlation over 469 effects pooled from 70 unrelated survey experiments, spanning many topics, treatments, and outcomes. Their megastudy figures instead measure how well the model distinguishes the several interventions of a single megastudy against that study's shared control, which is exactly this benchmark's setting (16 interventions, one control). Ashokkumar et al. report the megastudies as a separate secondary archive (15 megastudies, 606 effects) and note that these correlations run lower than the primary archive's.], survey experiments reach r = .43 (r~adj~ = .52) and text-based treatments r = .45 (r~adj~ = .54).

The adjusted RMSE applies the same logic of adjustement to the RMSE rung. It subtracts the average reference sampling variance from the average squared prediction error and takes the square root.

```{r, results='asis'}
#| label: fn-adjusted-metrics
print_a_function_from_file("adjusted_metrics")
```

#### Distribution comparison metrics

`compare_distributions()` compares response distributions for a single condition and outcome using three metrics:

- **Variance ratio** (submission variance / human variance). It captures only relative spread. A ratio of 1 means equal spread and says nothing about shape. A ratio below 1 means the synthetic responses vary less than the human ones, the most widely reported failure mode of synthetic samples.
- **OVL (overlapping coefficient)**. The share of area the two kernel density estimates have in common (1 = identical distributions, 0 = no overlap). It summarizes overall shape similarity.
- **KS D-statistic**. The largest gap between the two empirical cumulative distributions (0 = indistinguishable everywhere). It is sensitive to local departures that OVL might average away.

```{r, results='asis'}
#| label: fn-compare-distributions
print_a_function_from_file("compute_ovl")
```

#### Calibration regression

Correlation tells us only whether the human and predicted effects covary. It does not capture baseline differences or scaling. An approach could inflate every effect by a factor of two and still score Pearson r = 1. The calibration regression captures this. 

`run_calibration()` regresses human ATEs on predicted ATEs, `ATE_human = α + β × ATE_predicted`. The intercept α captures the constant offset. A positive α means the approach systematically underestimates, a negative α that it overestimates. The slope β captures the scaling. β = 1 is perfect, β > 1 means predicted effects are compressed, β < 1 exaggerated. We report 95% CIs for both, and R² as the share of human-ATE variance linearly recoverable from the predictions. Calibration is then tested at two strictness levels. The **joint F-test of (α = 0, β = 1)** tests perfect calibration, meaning the predictions can be read directly as human ATEs with no correction. The weaker **slope-only F-test of (β = 1)** asks whether the predictions are proportionally calibrated, permitting a constant offset. In the benchmark the regression runs with **HC2 heteroskedasticity-robust inference** (`robust = TRUE`), because the outcome side carries known, condition-varying sampling error that classical F-tests ignore. A precision-weighted variant (weights 1/SE²~h~, `weight_by_precision = TRUE`) is reported as a sensitivity check with the real data. The regression has only 16 points, so the tests have little power and **non-rejection is not evidence of calibration**. Reference noise is not a problem in this regression, since the human ATEs sit on the outcome side, but for Tiers 2–3 the *predicted* ATEs carry the team's stated uncertainty as regressor noise, which pulls β̂ toward zero and biases the verdict toward "exaggerated". And the leaderboard's β and R² columns are computed on the primary outcome only, as stated in the table caption.

```{r, results='asis'}
#| label: fn-calibration
print_a_function_from_file("run_calibration")
```

#### Demographic baseline and predictability metrics

`compare_demographic_baselines()` computes, for each moderator, the Pearson correlation and RMSE between human and synthetic sample cell means in the control condition. `compare_demographic_predictability()` fits separate OLS regressions of the outcome on each moderator plus condition fixed effects, returning R² and dummy-coded coefficients for humans and for the submission. `demographic_parity_gap()` condenses the group specific baseline errors into one fairness number per moderator, the gap between the worst- and best-served group (see *Section 3*).

```{r, results='asis'}
#| label: fn-demographic-baselines
print_a_function_from_file("compare_demographic_baselines")
```

```{r, results='asis'}
#| label: fn-demographic-predictability
print_a_function_from_file("compare_demographic_predictability")
```

```{r, results='asis'}
#| label: fn-demographic-parity-gap
print_a_function_from_file("demographic_parity_gap")
```

### Section 1 — Average Treatment Effects

**ATE recovery.** For each intervention, the average treatment effect (ATE) is the difference in mean outcome between participants assigned to that intervention and those in the control condition. ATEs are estimated with `run_main_treatment_model()`, run on the human reference and on the submission separately for each preregistered outcome. For the binary `newsletter_signup`, `run_main_treatment_model_binary()` returns marginal effects on the probability scale. Before any cross-outcome scoring, every estimate and standard error is converted to percentage points of its outcome's scale range (see *Scoring principle*), so effects on the 0–100 sliders, the $0–10 donation, and the signup probability have comparable units. The full six-metric ladder (`compare_estimates()`) and the calibration regression (`run_calibration()`) are then applied. The main analysis pools across all outcomes to maximize power. 

**Top-intervention selection.** The ranking correlation already looks at how well the order of effect sizes aligns between human and synthetic sample effects. However, in the context of megastudies, people are often interested in only a subset of interventions that are worth acting upon [@ashokkumarLargeLanguageModels2026a]. We therefore use `topk_selection()`, which, per outcome, picks an approach's top **k = 3** interventions by predicted ATE (about the top 20% of the 16 arms) are compared with the truly best three by the human reference. The **hit rate** is the share of picks that are in the true best set, with chance at 3/16. The **regret ratio** is the mean true ATE of the picked set divided by the mean true ATE of the best set, so 1 is a perfect pick. It is scale-free within outcome and undefined when the best achievable mean ATE is not positive. 

**Response distributions.** Even an approach that matches every ATE may distribute responses differently. Two groups with the same mean can differ in spread, skewness, or tail shape. In the control condition, across all continuous outcomes, we compare the full response distribution (`compare_distributions()`, reporting the variance ratio, OVL, and KS D). The **variance ratio** is the headline diagnostic of the three, because under-dispersion (synthetic respondents answering too much alike; ratio < 1) is a widely reported issue in the recent evaluation literature [@ashokkumarLargeLanguageModels2026a; @cumminsThreatAnalyticFlexibility2025b; @parkLLMAgentsGrounded2026; @arrudaCollectiveCooperationIndividual2026], is **under-dispersion**. `newsletter_signup` is excluded, since kernel density is not meaningful for a binary outcome. This analysis requires respondent-level data and so is **Tier 1 only**.

### Section 2 — Subgroup Effects

**Heterogeneity in treatment effects.** Interventions may work differently across demographic groups. A consensus message might shift trust more among Republicans than Democrats, or more among younger than older respondents. For each of six categorical moderators (gender, age band, race, education, income, partisan identity), condition × moderator interactions are estimated with `run_moderator_model()` on the human reference and on the submission. For the binary outcome the same OLS fit is a linear probability model, so the interaction coefficient is the difference-in-differences of cell signup proportions on the probability scale (see *Estimation functions*). As in Section 1, all interaction estimates are converted to percentage points of each outcome's scale range before pooling. The interaction estimates are then compared with the same ladder ordering as Section 1. Because these pairs carry no inferential category per interaction cell, the pooled subgroup score uses the reduced signed-effect ladder of directional agreement, Spearman ρ, and Pearson r (`signed_metrics()`), plus the adjusted correlation r~adj~ computed from the human interaction estimates' standard errors. The subgroup section is where the noise correction matters most, and r~adj~ is the emphasized value here. Small demographic cells make the reference noise large and uneven, so raw correlations are unevenly deflated across groups (see *Adjusted correlation and RMSE*). 

One caveat is that if the human data turn out to contain little true moderation, subgroup recovery cannot separate approaches: a submission that predicts no heterogeneity at all then scores well, and a high correlation may reflect generalized predictions rather than subgroup-specific skill. This is a limitation @ashokkumarLargeLanguageModels2026a flag in their subgroup analysis, too. As context for reading the subgroup scores, we therefore report the **share of human interaction estimates that are individually significant** (normal approximation from the interaction standard errors, BH-adjusted within outcome × moderator) next to the subgroup leaderboard; when that share is near the 5% false-positive rate, the subgroup rungs are read as uninformative rather than as evidence for or against any approach. (The r~adj~ NA-guard flags the same condition from the noise side.) Eligible for Tiers 1–2.

**Response distributions by subgroup.** Mirroring Section 1, within each demographic subgroup in the control condition we compare response distributions (OVL, KS D, variance ratio), to catch an approach that reproduces the overall distribution while miscalibrating within-group spreads. Tier 1 only.

### Section 3 — Demographic Baseline Calibration

Section 2 tests whether an approach exaggerates *effect* heterogeneity. Section 3 asks a separate question. Does the approach start from the right **baseline**? It uses only the control condition. An approach could assign far too little trust to Republicans, or too much to Democrats, before any treatment, a case of baseline stereotyping.

**Demographic baseline calibration.** For each moderator, we compute control condition group means for humans and for the submission, then correlate and compare them (`compare_demographic_baselines()`, which reports Pearson r and RMSE of the cell means). High correlation and low RMSE mean the approach starts from the right response level for each demographic group. Eligible for Tiers 1–2.

**Demographic parity gap (DPD).** The baseline-calibration metrics average over groups, and an average can hide a simulation approach that may serve most groups well while consistently missing one. Following the demographic-parity-difference summary of @parkLLMAgentsGrounded2026^[Park et al. compute the gap in per-group predictive accuracy over paired agent–person data, while this benchmark has no person-level pairing, so the gap is taken over group-mean baseline errors instead.], `demographic_parity_gap()` reports, per moderator, the gap between the worst- and best-served demographic group (the difference between the largest and smallest absolute cell-mean error in the control condition), together with which groups those are. The worst-served group's absolute error is reported alongside it, because a zero gap can also mean every group is served equally *badly*, which the worst-group error would reaveal. Eligible for Tiers 1–2.

**Demographic predictability.** For each of the six moderators we fit a separate OLS regression of the outcome on that moderator plus condition fixed effects, on the human reference and on the submission independently (`compare_demographic_predictability()`). This answers questions such as how much does partisan identity alone explain outcome variance in humans versus in the synthetic samples? An R² substantially higher for the submission than for humans flags exaggeration of demographic variables (stereotyping). The dummy-coded coefficients identify which groups are over- or under-weighted. Tier 1 only.

### Tier-specific estimand recovery

The comparison-metric ladder requires, per intervention × outcome, an estimate, a standard error, and an inferential category. How these are obtained depends on the tier:

- **Tier 1.** The submission runs through `run_main_treatment_model()` exactly like the human data, which produces estimates and HC2 robust standard errors. Submission-side inference is then standardized to the human design's precision. Each SE is rescaled by the two-cell design factor $\sqrt{(1/n_{h,k} + 1/n_{h,ctrl})\,/\,(1/n_{l,k} + 1/n_{l,ctrl})}$ and p-values are recomputed (normal approximation, BH within outcome, mirroring the Tier-2 path). The reason is that a team may simulate any pool size, so unstandardized significance would measure that choice rather than the method, and the Tier-1 precision requirement would mechanically hurt precise teams wherever a human effect is not significant. Point estimates are untouched, so the correlation, directional, and RMSE rungs are unaffected. The standardization makes the inferential and TOST columns comparable across tiers.
- **Tier 2.** The ATE for each intervention × outcome is the submitted cell mean minus the submitted control mean. Because only cell summaries are available, the preregistered OLS cannot be refit on a Tier 2 submission. Instead the ATE's standard error comes from the submitted cell uncertainty intervals. Each cell interval implies a standard error on that cell mean, $\mathrm{SE}_{cell} = (\mathrm{pi}_{upper} - \mathrm{pi}_{lower}) / (2 \times z_{0.975})$, and the treatment and control cells combine in the usual way for a between-subjects contrast, $\mathrm{SE} = \sqrt{\mathrm{SE}_t^2 + \mathrm{SE}_c^2}$. The test statistic $z = \mathrm{ATE} / \mathrm{SE}$ is referred to the standard normal distribution for a two-sided p-value, $p = 2\,\Phi(-|z|)$. These p-values are BH-adjusted within outcome, mirroring the Tier 1 models. Subgroup ATEs and demographic baselines are computed analogously from the moderator cell file. The combination rule assumes the team's uncertainties about the two cells are **independent**. That is why the schema asks teams to leave shared level uncertainty out of the cell intervals (see *Submission specification*): uncertainty about the overall response level cancels in a treatment-minus-control contrast, and a team that folds it into every cell gets derived ATE intervals that are too wide — which reads as above-nominal PI coverage and a worse Winkler score (below), not as a scoring advantage.
- **Tier 3.** The submitted `ate` is the estimate. Inferential agreement uses the submitted 95% uncertainty interval. An effect counts as significantly positive (negative) if the interval lies entirely above (below) zero, the direct analog of a two-sided test at α = .05, so the significance criterion matches the one applied to Tiers 1–2. For the TOST equivalence test, the interval implies a standard error, $\mathrm{SE} = (\mathrm{pi}_{upper} - \mathrm{pi}_{lower}) / (2 \times z_{0.975})$.

**The submitted intervals are also scored directly (Tiers 2–3).** An uncertainty interval that only feeds the inferential rung would open a loophole. Arbitrarily wide intervals turn every prediction "not significant", which matches a mostly-null world for free. Two preregistered scores close it (`pi_metrics()`). **PI coverage** is the share of reference ATEs falling inside the submitted (Tier 3) or implied (Tier 2) 95% intervals. Honest intervals land near 95%. The **Winkler interval score** is the interval's width plus $2/\alpha$ times the distance by which the reference value misses it, averaged over pairs. It is a *proper* score, minimized in expectation by reporting one's true predictive interval, so neither padding intervals wide nor squeezing them tight pays. Like RMSE, it is unit-bound, and like RMSE it pools across **all outcomes** once every estimate is in percentage points of its outcome's scale range (see *Scoring principle*). Both appear as leaderboard columns for Tiers 2–3. They are undefined for Tier 1, whose SEs are sampling rather than stated uncertainty, and for the reference rows.

### Cross-team comparison

The primary goal of this project is to assess how well the research field of synthetic samples can predict the results of the behavioral experiment. It reports the distribution of predictions across submissions on each metric, its center and spread. The secondary goal is to rank individual approaches on a ladder of evaluation metrics. 

**Scoring principle.** Each approach is scored on every rung of the ladder, with the intervention × outcome estimates pooled across all 13 outcomes into a single value per metric (the 11 continuous sliders plus the two behavioral outcomes, the donation via the OLS model and the newsletter signup via logistic marginal effects on the probability scale, as in v1's pooled row). Before pooling, every estimate and standard error is converted to percentage points (pp) of its outcome's scale range. This is the unit of the predicted-vs-observed analysis of @ashokkumarLargeLanguageModels2026a, whose primary archive pools effects only after this conversion and who avoid SD standardization because synthetic under-dispersion corrupts an SD denominator. The same rationale applies here. The conversion is a modeling choice which equates 3 points on a 0–100 slider with $0.30 on the donation and 3 pp of signup probability, the same equation the TOST bound already makes. We deliberately do not collapse the rungs into one composite "champion" score. An approach can recover the direction of effects well while mis-scaling them, or do well on the primary outcome while failing on the rest, and indeed this inconsistency across outcome metrics has been observed in other studies [@cumminsThreatAnalyticFlexibility2025b]. Pooling across outcomes can be inflated by a submission that predicts *which outcomes* are generally movable while misranking the interventions within every outcome. The leaderboard therefore carries the **Fisher-z average of the per-outcome correlations** beside the pooled value (`fisher_z_avg()`). When the two diverge, that is the signature of exactly this artifact, and the per-outcome breakdown shows where. With all estimates in pp, RMSE and its noise-corrected companion RMSE~adj~ are comparable across outcomes and are reported **pooled as well as per outcome** (in v1's raw-unit, per-outcome framing they could not be pooled). The per-intervention breakdown keeps its restriction to `outcomes_continuous`. The pp conversion removes the *unit* mix, but a per-intervention correlation across all 13 outcomes would still mix attitude and behavioral constructs, so the conservative subset is retained.

Teams may enter up to two submissions per tier. For the main findings on the field (Finding 1) we will only une pre-designated primary entry per team. Every valid entry still appears on the leaderboard (Finding 2), labeled primary or secondary. Approaches that cover only a subset of study outcomes are in principle allowed. However, a self-chosen, potentially more easy-to-predict subset would be an unfair comparison to full-coverage entries. The primary leaderboard is therefore restricted to full-coverage entries. Partial-coverage entries are reported in a separate annex, scored pairwise on what they cover, with their coverage stated. Finally, the field distribution describes an invited, selected set of expert teams, a purposive sample of the current generation of serious approaches rather than a random sample of all possible ones. We state this wherever the field summary is interpreted.

**Presentation.** Every preregistered metric is computed and reported. However, for the key figures of the manuscript, we will pick one key metric per question. **Pearson *r*** is the key metric for ATE recovery (Section 1) and, on the same axis, for subgroup recovery (Section 2). The **calibration slope β and intercept α** answer whether those effects are the right *size*. The **variance ratio** is the key metric for response-shape recovery (Section 1b/2b). Every other rung of the preregistered ladder (directional agreement, Spearman ρ, the adjusted correlations, inferential and TOST agreement, RMSE, OVL, KS, W1) is still computed and scored for every submission, and reported in a companion table beside each figure.

The headline display (@fig-headline) puts the key measures in one figure with two panels. Panel A plots predicted against human ATEs across all approaches and outcomes, in percentage points of each outcome's scale range. This is the direct visual of the prediction task, in the units of the predicted-vs-observed figure of @ashokkumarLargeLanguageModels2026a. It is overlaid with the average field and human **calibration regressions** and a corner readout of their correlation and calibration coefficients (slope β, perfect = 1, and intercept α, perfect = 0, with 95% CIs). Panel B **ranks** the approaches on the pooled Pearson *r*, each with a cluster-bootstrap **uncertainty interval** over interventions (new in this preregistration, since v1 reported point estimates only, and detailed under *Uncertainty intervals* below). The human replication reference is pinned on top, and an **all-approaches summary row** at the bottom carries the field's density and mean with its CI, so this one panel shows both the field distribution (Finding 1) and the ranking (Finding 2). Three further figures break the key metrics down, by outcome and by intervention (@fig-field-by-group), by moderator for subgroup recovery (@fig-subgroup-headline), and as per-outcome response densities with the variance ratio and Wasserstein-1 distance stamped on each panel (@fig-distribution-headline). Every entry carries its disclosure-class badge (open / escrowed / sealed) in the leaderboard table, and the precise numbers accompany every figure as a companion table.

#### Uncertainty intervals on the leaderboard

Each point in the ranking panel is a single pooled score, one number per approach per metric. A bare point invites over-reading. If one approach scores Pearson *r* = 0.60 and another 0.55, is the first genuinely better, or is the gap an artifact of the particular interventions that happened to be tested? The interval answers that question. It guards against ranking approaches on differences that would not survive a different, equally reasonable test.

The pooled metric is computed once on the full data to give the point estimate. We then form a *cluster bootstrap over interventions*. The 16 interventions are resampled with replacement (e.g., some are drawn twice and some are dropped), and the pooled metric is recomputed on each resampled set, and the 2.5th to 97.5th percentiles of those recomputed scores give the 95% interval (`cluster_boot()` in `R/functions/statistics.R`). Resampling happens at the level of whole interventions, each with its full block of outcome rows, because the outcome rows belonging to one intervention are not independent. 

No submitted prediction is ever altered. Each approach's predicted effect for each intervention stays fixed at the value the team submitted. What is resampled is which interventions enter the score. The interval therefore measures the spread of the recomputed score, not the spread of the raw effects. Averaging or correlating cancels much of the effect-to-effect spread, which is why the bar is typically far narrower than the range of the underlying estimates.

**Assumptions.** The interval treats the 16 interventions as a *test set* standing in for the broader space of interventions the benchmark cares about, just as a held-out test set stands in for future cases in predictive modeling. Read this way, the bar answers a generalization question. Would this approach's score, and its rank, survive on a different set of interventions of the same kind? Resampling does **not** assume that a team that predicted a given intervention well could have gotten it wrong, since the predictions are fixed. It assumes only that a replication might have tested a different set of interventions.

The metrics are bounded agreement scores with a ceiling at perfect agreement. An approach that tracks the human effects closely sits near that ceiling and stays there under resampling, because dropping or duplicating interventions barely moves a score already pinned near perfect agreement. Accurate approaches therefore carry *narrow* intervals. An approach whose predictions do not track the human effects has a score floating in the interior, free to swing with the resampled set, and so carries a *wide* interval. Interval width is entangled with accuracy, but in the helpful direction. A wide bar signals that a score is flimsy and may reflect test-set luck rather than skill, exactly the warning one wants attached to a weak entry. (Spread in the size of the true effects does not by itself widen the bar. The correlation and agreement metrics are scale-free, so what matters is whether predictions hug the line, not how far apart the interventions are. Width also partly reflects differences in per-intervention difficulty, so it is a flimsiness *signal*, not a pure standard error.)

Overlapping bootstrap intervals do **not** mean two approaches are indistinguishable, and conversely, non-overlap is not required to claim a difference. That is the standard misreading. Because every submission is scored against the *same* reference on the *same* intervention clusters with the *same* preregistered bootstrap seed, the correct comparison is paired. In each replicate one set of clusters is drawn, both submissions are scored on it, and the difference is taken, so the shared reference noise cancels within each replicate (`cluster_boot_diff()`). **We preregister the rule that any claim that one approach outperforms another on a metric requires the paired-difference 95% interval to exclude zero.** Otherwise the two are reported as statistically indistinguishable on that metric, the most common outcome among top performers in comparable mass-collaboration benchmarks [@salganikMeasuringPredictabilityLife2020] and one we expect here. Three caveats apply to all bootstrap intervals. With 16 intervention clusters, percentile intervals are known to undercover somewhat. The 16 ATEs share the single control arm, which induces a dependence the intervention-level resampling does not capture. And the BH-adjusted significance labels are computed once on the full intervention family and held fixed under resampling, so the inferential and TOST rungs' intervals do not propagate how the adjustment itself would shift on a different draw. The intervals are read as honest approximations, not exact coverage statements.

**Relation to the per-effect confidence and uncertainty intervals.** The bootstrap interval is a different layer of uncertainty from the intervals attached to individual treatment effects, and the two should not be confused. A per-effect interval is *within-intervention*. It says how precisely a single effect is known, whether from sampling noise in a Tier 1 model fit or from the team's own stated uncertainty in a Tier 2 or 3 submission. The bootstrap interval is *across-intervention*. It says how much the pooled score depends on which interventions were tested. The two are **nested, not competing**. The per-effect intervals enter the *value* of each row. For the correlation and directional rungs only the point estimates are used, so the per-effect intervals do not affect those scores at all and the bootstrap is their sole source of an error bar. For the inferential-agreement and TOST rungs the per-effect intervals set each row's significance or equivalence label, which the bootstrap then pools and resamples. In all cases the bootstrap holds each intervention's estimate fixed at its point value. It propagates the which-interventions uncertainty into the bars but **not** the within-effect sampling noise.

#### Scoring functions

The cross-team analysis reuses the comparison functions specified under *Comparison functions*, above, and adds the scoring helpers below (in `R/functions/statistics.R`), each shown in full and explained before it.

`pooled_metrics()` computes the ATE-recovery ladder on intervention × outcome pairs **pooled across all outcomes**. It reports directional agreement, Spearman ρ, Pearson r, the noise-corrected companion r~adj~ (via `adjusted_metrics()`, see *Comparison functions*), the Fisher-z average of the per-outcome correlations (via `fisher_z_avg()`, the pooled-vs-averaged divergence diagnostic, see *Scoring principle*), inferential agreement, and TOST equivalence with the fixed bound (`tost_delta`, a uniform 3 pp in the cross-team scoring). It mirrors the original `compare_estimates()` but operates on the pooled set of pairs rather than a single outcome. The pairs it receives are already in pp of scale range, converted once at pair-building time, so `include_rmse = TRUE` adds RMSE and RMSE~adj~ in pp, used for the pooled leaderboard as well as the per-outcome breakdown. It expects pre-joined pairs with the reference (Human 1) columns `estimate_h`, `se_h`, `p_adj_h` and the submission columns `estimate_l`, `se_l`, `infer_l`. `pi_metrics()` scores the Tier-2/3 uncertainty intervals (coverage and the Winkler interval score). `cluster_boot_diff()` implements the paired-replicate decision rule for ranking claims.

```{r, results='asis'}
#| label: fn-pooled-metrics
print_a_function_from_file("pooled_metrics")
```

`signed_metrics()` is the reduced ladder for the subgroup leaderboard, where only the signed-effect rungs are pooled (directional agreement, Spearman ρ, Pearson r). This matches Section 2, above, which reports these three on estimate-only pairs (no standard error or inferential category).

```{r, results='asis'}
#| label: fn-signed-metrics
print_a_function_from_file("signed_metrics")
```

`cluster_boot()` attaches the uncertainty interval described above. It resamples the levels of a clustering variable (interventions for the ATE and subgroup panels, outcomes for the distributional panel) with replacement, recomputes the supplied metric function on each replicate, and returns the point estimate together with percentile confidence bounds in long form (one row per metric, with `value`, `lo`, `hi`). It is generic. The metric function is passed in, so the same bootstrap serves every panel.

```{r, results='asis'}
#| label: fn-cluster-boot
print_a_function_from_file("cluster_boot")
```

`metrics_by_group()` scores within each level of a grouping variable instead of pooling everything into one number. With `group = "outcome"` it returns one ladder per outcome (scored across the interventions), and with `group = "condition"` one ladder per intervention (scored across the outcomes). It drives the by-outcome and by-intervention distribution figures. Per intervention the score runs across `outcomes_continuous` only (see *Scoring principle*).

```{r, results='asis'}
#| label: fn-metrics-by-group
print_a_function_from_file("metrics_by_group")
```

`summarise_field()` reduces the per-approach scores to the field-level summary printed beside the headline figure, giving the mean, median, SD, min, and max of each metric across approaches (excluding the reference row).

```{r, results='asis'}
#| label: fn-summarise-field
print_a_function_from_file("summarise_field")
```

The remainder of this section demonstrates the code on placeholder data with **30 mock "teams"**, each a stratified bootstrap resample (within condition) of the single individual-level placeholder dataset, enough approaches to give the field distributions a meaningful shape. To exercise all three tier code paths, 28 teams enter at Tier 1 (individual-level), one is collapsed into a Tier 2 cell-level file, and one into a Tier 3 effect-level file. All of this placeholder data, including the resampled teams and the collapsed Tier 2/3 files, is built in the simulation script (`R/simulate_data_preregistration.R`), not here. This document only reads it. Because every mock team is just a resample of the same dataset, all approaches are random noise relative to the human reference. On placeholder data the field distributions therefore cluster near the null, and the figures only show that the code runs. Real submissions will spread out. The adjusted correlation r~adj~ behaves accordingly. With no true signal, the reference effects' observed spread is all noise, so the NA guard binds for many mock teams and the surviving values are unstable. That is the intended behavior, demonstrated rather than hidden. The code applies to real submissions without modification.

```{r}
#| label: variables

outcomes_primary   <- "trust_multidimensional"
outcomes_secondary <- c("trust_post", "distrust_post",
                         "funding_perceptions", "policy_role_mean", "inst_trust_mean")
outcomes_tertiary  <- c("belief_post", "concern_mean", "policy_general",
                        "policy_specific_mean", "behavior_mean")
outcomes_continuous <- c(outcomes_primary, outcomes_secondary, outcomes_tertiary)

# The two behavioral outcomes enter the pooled cross-team ladder alongside the
# 0-100 sliders (as in v1's pooled ATE row): donation in dollars via the OLS
# model, newsletter signup via logistic marginal effects on the probability
# scale. Scale-bound quantities (RMSE, the Winkler interval score, the
# by-intervention ladder) stay on `outcomes_continuous` only.
outcomes_behavioral <- c("donation_ams", "newsletter_signup")
outcomes_pooled     <- c(outcomes_continuous, outcomes_behavioral)

# Scale ranges: all cross-team scoring runs in percentage points (pp) of each
# outcome's scale range — effects and SEs are converted once, at pair-building
# time, and every downstream metric, figure, and table reads the same unit.
# This follows Ashokkumar et al.'s primary archive, which pools effects only
# after expressing them in pp of the original scales, and deliberately avoids
# SD standardization (an under-dispersed submission would corrupt an SD
# denominator). Pooling raw mixed units instead would let the 0-100 sliders
# drown out the two behavioral outcomes (dollar and probability effects are
# 10-100x smaller numerically) and hand every submission 32 free near-origin
# points. The 0-100 sliders are already in pp; only the donation (0-10
# dollars) and signup probability (0-1) rescale.
scale_range <- c(
  set_names(rep(100, length(outcomes_continuous)), outcomes_continuous),
  donation_ams = 10, newsletter_signup = 1
)

# Fixed, prespecified TOST equivalence bound: 3% of each outcome's scale
# range, which in pp units is a uniform 3 (3 points on the 0-100 sliders,
# $0.30 on the 0-10 donation, 3 percentage points on the signup proportion).
# Fixed rather than |ATE_h|-relative (v1's bound): a bound computed from the
# noisy estimate under test is not a valid alpha-level procedure, and the
# design analysis below shows the relative bound has no dynamic range at this
# design's reference SEs.
tost_delta_pp <- 3

# Bootstrap replicates: the PREREGISTERED value for the real-data run is
# 2,000. This placeholder render is a demonstration of the procedure, not of
# its precision, so it uses a small demo value to keep the render cheap —
# only this parameter changes when the submissions arrive, nothing else.
n_boot_real <- 2000
n_boot_demo <- 200

# Resplit sensitivity check (see "Robustness of the scoring pipeline"): same
# demo/real split as the bootstrap. The PREREGISTERED value for the real-data
# run is 200 resplits; the placeholder render uses 50 to keep the build cheap.
n_resplit_real <- 200
n_resplit_demo <- 50

# Illustrative outcome subset and categorical moderators — same definitions as
# Section 2 above. The subgroup-effects leaderboard scores the illustrative
# outcomes plus the binary newsletter signup (see Section 2: its interaction
# difference-in-differences come from the linear probability model, so they
# live on the probability scale and pp-rescale like everything else).
outcomes_illustrative <- c("trust_multidimensional", "donation_ams",
                           "funding_perceptions", "policy_general")
outcomes_subgroup     <- c(outcomes_illustrative, "newsletter_signup")
moderators_cat <- c("gender", "age_band", "race", "education", "income", "party")

outcome_label_map <- c(
  trust_multidimensional = "Multidimensional trust (primary)",
  trust_post             = "Single-item trust",
  distrust_post          = "Distrust",
  funding_perceptions    = "Funding perceptions",
  policy_role_mean       = "Policy role",
  inst_trust_mean        = "Institutional trust",
  belief_post            = "Climate belief",
  concern_mean           = "Climate concern",
  policy_general         = "General climate policy",
  policy_specific_mean   = "Specific climate policy",
  behavior_mean          = "Pro-climate behavior",
  donation_ams           = "Donation to AMS ($)",
  newsletter_signup      = "Newsletter signup"
)
```

```{r}
#| label: leaderboard-helpers

# Extract ATE pairs (estimate, SE, adjusted p) — same helper as the original
# preregistration's rq1-all chunk.
prep_ate <- function(df) {
  df |>
    select(condition, estimate) |>
    mutate(
      se    = if ("std.error"        %in% names(df)) df$std.error        else NA_real_,
      p_adj = if ("p.value_adjusted" %in% names(df)) df$p.value_adjusted else NA_real_
    )
}

# Inferential category from estimate sign + adjusted p-value (Tiers 1-2)
classify_sig <- function(estimate, p_adj) {
  case_when(
    !is.na(p_adj) & p_adj < .05 & estimate > 0 ~ "pos",
    !is.na(p_adj) & p_adj < .05 & estimate < 0 ~ "neg",
    TRUE                                       ~ "ns"
  )
}

control_lab <- levels(human_data$condition)[1]
z95 <- qnorm(0.975)

# Reference (Human 1) ATE models on the continuous outcomes, fit once and
# reused (the calibration regression reads the primary outcome's model).
human_models <- map(outcomes_continuous,
                    ~ run_main_treatment_model(human_data_1, outcome = .x)) |>
  set_names(outcomes_continuous)

# One ATE side (estimate, SE, BH-adjusted p) for an individual-level dataset,
# across all pooled outcomes: OLS with HC2 SEs for the continuous outcomes and
# the donation; logistic marginal effects on the probability scale for the
# binary newsletter signup (as in v1's pooled row). The binary SE is recovered
# from the marginal-effect CI.
ate_side <- function(dat, models = NULL) {
  cont <- map_dfr(c(outcomes_continuous, "donation_ams"), function(out) {
    res <- if (!is.null(models) && out %in% names(models)) models[[out]]
           else run_main_treatment_model(dat, outcome = out)
    prep_ate(res) |> mutate(outcome = out)
  })
  bin <- run_main_treatment_model_binary(dat,
                                         outcome = "newsletter_signup")$marginal_effects |>
    transmute(condition,
              estimate,
              se      = (conf.high - conf.low) / (2 * z95),
              p_adj   = p.value_adjusted,
              outcome = "newsletter_signup")
  bind_rows(cont, bin)
}

human_side <- ate_side(human_data_1, models = human_models) |>
  rename(estimate_h = estimate, se_h = se, p_adj_h = p_adj)

# Common unit for all scoring: estimates and SEs in pp of each outcome's scale
# range (see `scale_range` in the variables chunk). Applied once to every set
# of joined pairs — the three tier builders, the floor baseline, and the
# subgroup pairs — so all pooled metrics, figures, and tables read one unit.
# A positive multiplicative factor: signs, p-values, and inferential
# categories are untouched.
pp_scale_pairs <- function(pairs) {
  pairs |>
    mutate(across(any_of(c("estimate_h", "se_h", "estimate_l", "se_l")),
                  ~ .x * 100 / scale_range[outcome]))
}

# Human per-condition cell sizes (named vector), for standardizing Tier-1
# submission-side inference to the human design's precision.
n_cells_h <- human_data_1 |> count(condition) |> deframe()

# Tier 1: pool ATE pairs from an individual-level dataset by refitting the
# preregistered treatment models on the submission. Submission-side INFERENCE
# (the inferential-agreement category and the TOST SE) is evaluated at the
# HUMAN design's per-arm precision: a team may simulate any pool size, so
# model-based significance would otherwise measure that choice rather than the
# method — and the Tier-1 precision requirement would mechanically hurt
# precise teams wherever a human effect is n.s. Each submission SE is rescaled
# by the two-cell design factor sqrt((1/n_h,k + 1/n_h,ctrl) /
# (1/n_l,k + 1/n_l,ctrl)) and p-values recomputed (normal approximation, BH
# within outcome, mirroring the Tier-2 path). Point estimates are untouched,
# so the correlation, directional, and RMSE rungs are unaffected. This also
# makes the inferential and TOST columns comparable across tiers: every
# submission's "significance" is evaluated at the same design precision.
build_pairs_t1 <- function(pred_data) {
  n_cells_l <- pred_data |> count(condition) |> deframe()
  fac <- sqrt((1 / n_cells_h + 1 / n_cells_h[[control_lab]]) /
              (1 / n_cells_l[names(n_cells_h)] + 1 / n_cells_l[[control_lab]]))
  ate_side(pred_data) |>
    mutate(se_std = se * unname(fac[condition])) |>
    group_by(outcome) |>
    mutate(p_adj = p.adjust(2 * pnorm(-abs(estimate / se_std)), method = "BH")) |>
    ungroup() |>
    mutate(infer_l = classify_sig(estimate, p_adj)) |>
    select(condition, outcome, estimate_l = estimate, se_l = se_std, infer_l) |>
    inner_join(human_side, by = c("condition", "outcome")) |>
    pp_scale_pairs()
}

# Tier 2: ATEs and SEs from submitted cell means and their 95% prediction
# intervals (mean, pi_lower, pi_upper per condition x outcome). Each cell
# interval implies an SE on the cell mean; the treatment and control cells
# combine for the between-subjects contrast. Normal-approximation p-values,
# BH-adjusted within outcome to mirror the Tier 1 models.
build_pairs_t2 <- function(cells_main) {
  cells_main |>
    mutate(
      condition = as.character(condition),
      se_cell   = (pi_upper - pi_lower) / (2 * z95)
    ) |>
    group_by(outcome) |>
    group_modify(function(cells, key) {
      ctrl <- cells |> filter(condition == control_lab)
      cells |>
        filter(condition != control_lab) |>
        mutate(
          estimate_l = mean - ctrl$mean,
          se_l       = sqrt(se_cell^2 + ctrl$se_cell^2),
          p_adj      = p.adjust(2 * pnorm(-abs(estimate_l / se_l)), method = "BH"),
          infer_l    = classify_sig(estimate_l, p_adj)
        ) |>
        select(condition, estimate_l, se_l, infer_l)
    }) |>
    ungroup() |>
    inner_join(human_side, by = c("condition", "outcome")) |>
    pp_scale_pairs()
}

# Tier 3: submitted ATEs taken as-is; inference from the submitted 95%
# uncertainty interval (significant iff the interval excludes zero, the direct
# analog of a two-sided test at alpha = .05); implied SE from the interval
# width for the TOST computation.
build_pairs_t3 <- function(effects) {
  effects |>
    mutate(
      condition  = as.character(condition),
      estimate_l = ate,
      se_l       = (pi_upper - pi_lower) / (2 * z95),
      infer_l    = case_when(
        pi_lower > 0 ~ "pos",
        pi_upper < 0 ~ "neg",
        TRUE         ~ "ns"
      )
    ) |>
    inner_join(human_side, by = c("condition", "outcome")) |>
    select(condition, outcome, estimate_l, se_l, infer_l,
           estimate_h, se_h, p_adj_h) |>
    pp_scale_pairs()
}

# `pooled_metrics()`, `signed_metrics()`, and `cluster_boot()` are defined in
# R/functions/statistics.R and shown under "Scoring functions" above.

# One leaderboard row: pooled ladder + calibration slope and R^2 on the
# primary outcome (HC2-robust inference: the outcome side of the calibration
# regression carries known, condition-varying sampling error) + the two
# interval scores for Tiers 2-3 (pi_metrics(); the Winkler score is
# unit-bound, and the pairs arrive in pp of scale range, so like RMSE both
# scores pool across all outcomes). Interval scores are NA
# for Tier 1 (its SEs are sampling, not epistemic, uncertainty) and for the
# reference rows.
leaderboard_row <- function(pairs, submission, tier, disclosure) {
  calib <- run_calibration(
    human_models[[outcomes_primary]],
    pairs |> filter(outcome == outcomes_primary) |>
      select(condition, estimate = estimate_l),
    robust = TRUE
  )
  pi <- if (tier %in% c("2", "3"))
    pi_metrics(pairs)
  else tibble(pi_coverage_pct = NA_real_, interval_score = NA_real_)
  pooled_metrics(pairs, tost_delta = tost_delta_pp, include_rmse = TRUE) |>
    mutate(submission = submission, tier = tier, disclosure = disclosure,
           beta = calib$beta, calib_r2 = calib$r_squared,
           pi_coverage_pct = pi$pi_coverage_pct,
           interval_score  = pi$interval_score)
}
```

```{r}
#| label: mock-submissions

# Mock Tier 2 (cell-level) and Tier 3 (effect-level) submissions, generated
# outside this script alongside the other placeholder data (see
# R/simulate_data_preregistration.R) so that no mock-data simulation lives in
# the preregistration itself. Tier 2 collapses one placeholder team to the
# cell-level schema (*_cells_main.csv); Tier 3 gives ATEs vs. control with a
# widened 95% uncertainty interval standing in for a team's epistemic uncertainty.
cells_t2   <- readRDS(here("data/simulation_preregistration/mock_submission_tier2_cells.rds"))
effects_t3 <- readRDS(here("data/simulation_preregistration/mock_submission_tier3_effects.rds"))
```

```{r}
#| label: leaderboard-build

# 30 mock submissions plus the fixed human replication reference (Human 1 vs. Human 2).
# Teams 1-28 enter as individual-level Tier 1 data; team 29 as a Tier 2
# cell-level file and team 30 as a Tier 3 effect-level file, so all three tier
# code paths are exercised. Each submission's ATE pairs are built once here and
# reused by the headline and breakdown figures, the leaderboard table, and the
# companion tables.
tier1_ids <- 1:28

# Cached: the 29 Tier-1 builds each refit the treatment models incl. the
# logistic marginal effects for the newsletter outcome, the slow step of a
# cold render.
pairs_list <- cached("amendment_pairs_list", {
  c(
    set_names(
      map(tier1_ids,
          ~ build_pairs_t1(llm_data_teams |> filter(team == paste0("team_", .x)))),
      paste0("Team ", tier1_ids)
    ),
    list(
      "Team 29"           = build_pairs_t2(cells_t2),
      "Team 30"           = build_pairs_t3(effects_t3),
      "Human replication" = build_pairs_t1(human_data_2)
    )
  )
})

set.seed(2026)
submission_meta <- tibble(
  submission = c(paste0("Team ", 1:30), "Human replication"),
  tier       = c(rep("1", 28), "2", "3", "—"),
  disclosure = c(sample(c("A", "B", "C"), 30, replace = TRUE), "—")
)

# Floor baseline: a naive reference row for the rungs without a natural null
# (inferential %, TOST %, RMSE). "No effect" predicts a zero ATE everywhere —
# a fully naive, blinded-achievable strategy whose inferential agreement is
# whatever the share of truly null human effects hands it for free. It is a
# reference row like the reference, not a member of the field: excluded from the
# field violins and the field summary. (An oracle-mean floor — predict each
# outcome's mean human ATE everywhere — was considered and rejected: the mean
# is the in-sample least-squares constant fit to the very ATEs being scored,
# so it competes rather than floors; its interpretive content survives as the
# RMSE scale reference, the SD of the human ATEs.)
floor_pairs <- list(
  "Floor: no effect" = human_side |>
    mutate(estimate_l = 0, se_l = 0, infer_l = "ns") |>
    pp_scale_pairs()
)
floor_labels <- names(floor_pairs)

# Calibration is undefined for the floor (a constant prediction within the
# primary outcome -> zero-variance regressor), so their rows carry NA there.
# Its directional % reads 0, not 50: sign(0) matches neither sign, an artifact
# of the all-zero prediction, not below-chance skill (footnoted in the table).
floor_rows <- imap_dfr(floor_pairs, function(pr, sub) {
  pooled_metrics(pr, tost_delta = tost_delta_pp, include_rmse = TRUE) |>
    mutate(submission = sub, tier = "—", disclosure = "—",
           beta = NA_real_, calib_r2 = NA_real_,
           pi_coverage_pct = NA_real_, interval_score = NA_real_)
})

leaderboard <- bind_rows(
  imap_dfr(pairs_list, function(pr, sub) {
    m <- submission_meta |> filter(submission == sub)
    leaderboard_row(pr, sub, tier = m$tier, disclosure = m$disclosure)
  }),
  floor_rows
) |>
  arrange(desc(submission == "Human replication"), desc(pearson_r))
```

```{r}
#| label: ate-ladder-labels

# Display labels / panel order for the metric ladder, and the null reference
# line per metric (50 = chance for directional agreement, 0 for the
# correlations; omitted for the percentage-agreement metrics where no clean
# null exists).
ladder_labels <- c(
  directional_pct = "Directional %",
  spearman_rho    = "Spearman ρ",
  pearson_r       = "Pearson r",
  pearson_adj     = "Pearson r (adj.)",
  inferential_pct = "Inferential %",
  rmse            = "RMSE (pp)",
  rmse_adj        = "RMSE (adj., pp)",
  tost_pct        = "TOST %"
)
ladder_nulls <- c("Directional %" = 50, "Spearman ρ" = 0, "Pearson r" = 0,
                  "Pearson r (adj.)" = 0)

relabel_metrics <- function(df, labels) {
  df |> mutate(metric = factor(labels[metric], levels = labels))
}
```

#### Finding 1 — How well can the field predict the experiment?

The headline question is not *who wins* but how well the current generation of approaches can predict a novel experiment at all. To keep that readable, each question gets **one key metric in the figures**. Pearson *r* serves ATE recovery and, on the interaction estimates, subgroup recovery, and the variance ratio serves response shape. The correlation is the literature's common currency [@ashokkumarLargeLanguageModels2026a], and one shared axis across the ATE and subgroup sections makes them directly comparable. The variance ratio carries the field's one directional expectation (under-dispersion). The full preregistered ladders are computed for everything and reported in the companion tables.

@fig-headline shows the raw prediction task with the field's calibration lines and coefficients (A) and the ranking of the field on the pooled Pearson *r* (B). One rendering note on panel B. This placeholder demonstration runs the bootstrap with a small replicate count (`n_boot_demo = 200`) to keep the document cheap to build. The **preregistered value for the real-data run is `n_boot_real = 2,000`**. When the submissions arrive, only that parameter changes.

```{r}
#| label: fig-headline
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "Headline cross-team figure (placeholder data). (A) Predicted against human ATEs across all outcomes and approaches, in percentage points of each outcome's scale range (as in the predicted-vs-observed figure of Ashokkumar et al.), one point per approach × intervention × outcome. Axes are centered on zero, the dashed line is the identity, and red crosses mark the human replication reference (Human 2 predicting Human 1). The solid lines are the average calibration regression of human on predicted ATEs, pooled across all outcomes, dark blue for the field (per-approach fits averaged) and red for the human replication reference. The corner prints the mean field Pearson r (with its noise-corrected value) and the human replication's, the same pooled metric ranked in B, together with the field and human calibration coefficients (slope β, perfect = 1, and intercept α, perfect = 0) with 95% CIs. (B) Approaches ranked by Pearson r pooled across all outcomes against Human 1. Filled points carry 95% cluster-bootstrap intervals over interventions, open points show the noise-corrected r (adj.) where its guard allows a value, and the human replication reference is pinned on top with its value dropped as a dotted line. The bottom row summarizes the field, with the density of the approaches' scores, a diamond at their mean with its 95% CI, and an open point at the mean noise-corrected r. On placeholder data every approach is a noise resample of the same dataset, so the field clusters near r = 0 in panel B, as expected of a random resample."

# (A) the raw prediction task: all approaches, all outcomes. The pairs are
# already in pp of scale range (pp_scale_pairs() at pair-building), the same
# unit as every pooled metric.
headline_pairs <- imap_dfr(pairs_list, ~ mutate(.x, submission = .y))
p_scatter <- plot_pred_scatter(
  headline_pairs,
  x_lab       = "Predicted ATE (pp of scale range)",
  y_lab       = "Human ATE (pp of scale range)",
  tag         = "A",
  point_alpha = 0.15, point_size = 0.5,
  ref_alpha   = 0.6,  ref_size   = 1,
  symmetric   = TRUE
)

# Average field / human calibration lines + a corner readout of correlation
# (r, adj.) and coefficients (β, α with CI). Mean field Pearson r and its
# noise-corrected value come from the pooled leaderboard — the identical pp
# quantity ranked in panel B; the coefficients are fit from the same pp pairs
# by calibration_overlay().
field_ld <- leaderboard |> filter(!submission %in% floor_labels,
                                  submission != "Human replication")
human_ld <- leaderboard |> filter(submission == "Human replication")
p_scatter <- p_scatter +
  calibration_overlay(headline_pairs,
                      r_field    = mean(field_ld$pearson_r,   na.rm = TRUE),
                      radj_field = mean(field_ld$pearson_adj, na.rm = TRUE),
                      r_human    = human_ld$pearson_r,
                      radj_human = human_ld$pearson_adj)

# (B) ranked field on the key metric, with cluster-bootstrap intervals.
# n_boot_demo here; the real-data run uses n_boot_real = 2,000 (see variables).
pooled_f <- \(p) pooled_metrics(p, tost_delta = tost_delta_pp)
ate_forest <- cached("amendment_ate_forest", {
  imap_dfr(pairs_list, function(pr, sub) {
    cluster_boot(pr, pooled_f, cluster = "condition", n_boot = n_boot_demo) |>
      mutate(submission = sub)
  }) |>
    relabel_metrics(ladder_labels)
})
p_forest <- plot_field_forest_single(ate_forest, tag = "B")

(p_scatter | p_forest) + plot_layout(widths = c(1, 1.15))
```

The full pooled ladder stays scored for every approach. @tbl-field-summary gives the field's center and spread on each rung.

```{r}
#| label: field-pooled-compute

field_pooled <- leaderboard |>
  filter(!submission %in% floor_labels) |>   # floors are reference rows, not field members
  select(submission, all_of(names(ladder_labels))) |>
  pivot_longer(-submission, names_to = "metric", values_to = "value") |>
  relabel_metrics(ladder_labels)
```

```{r}
#| label: tbl-field-summary
#| tbl-cap: "Field summary (placeholder data). Center and spread of each pooled ATE-recovery metric across the 30 approaches (the human replication reference excluded)."

summarise_field(field_pooled) |>
  gt() |>
  fmt_number(columns = c(mean, median, sd, min, max), decimals = 2) |>
  cols_label(metric = "Metric", mean = "Mean", median = "Median",
             sd = "SD", min = "Min", max = "Max")
```

##### Selecting the best interventions

The correlation rungs summarize the whole effect pattern. A decision-maker's question is narrower. If I trust this approach and act on its top picks, do I get the interventions that truly work best? `topk_selection()` scores exactly that (see *Section 1* for the full specification). The table shows the hit rate (share of an approach's top-3 picks that are among the truly best 3, chance ≈ 0.19) and the regret ratio (mean true ATE of the picked set as a share of the best achievable, 1 = perfect pick), on the primary outcome and averaged across the continuous outcomes.

```{r}
#| label: tbl-topk
#| tbl-cap: "Top-3 intervention selection (placeholder data). Per approach, the hit rate (share of its top-3 picks that are among the truly best 3, chance ≈ 0.19) and the regret ratio (mean true ATE of the picked set / best achievable, 1 = perfect), on the primary outcome and averaged across the continuous outcomes. On placeholder data picks are random, so hit rates sit near chance."

topk_table <- imap_dfr(pairs_list, function(pr, sub) {
  scores <- topk_selection(pr, k = 3)
  bind_cols(
    scores |>
      filter(outcome == outcomes_primary) |>
      transmute(hit_primary = hit_rate, regret_primary = regret_ratio),
    scores |>
      summarise(hit_mean    = mean(hit_rate, na.rm = TRUE),
                regret_mean = mean(regret_ratio, na.rm = TRUE))
  ) |>
    mutate(submission = sub, .before = 1)
}) |>
  arrange(desc(submission == "Human replication"), desc(hit_mean))

topk_table |>
  gt() |>
  fmt_number(columns = c(hit_primary, regret_primary, hit_mean, regret_mean),
             decimals = 2) |>
  sub_missing(missing_text = "—") |>
  cols_label(
    submission     = "Submission",
    hit_primary    = "Hit rate (primary)",
    regret_primary = "Regret ratio (primary)",
    hit_mean       = "Hit rate (mean)",
    regret_mean    = "Regret ratio (mean)"
  ) |>
  tab_style(
    style     = cell_fill(color = "#eef3f7"),
    locations = cells_body(rows = submission == "Human replication")
  )
```

##### Where the field predicts well

Split **by outcome** (one distribution per outcome, scored across the interventions) and **by intervention** (one per intervention, scored across the continuous outcomes, see *Scoring principle*), the key metric shows *where* the field predicts well and what resists prediction (@fig-field-by-group). The full ladders per outcome and per intervention are in @tbl-field-by-outcome and @tbl-field-by-intervention.

```{r}
#| label: field-by-group-compute

# include_rmse: with all pairs in pp, RMSE and its noise-corrected companion
# RMSE_adj are defined per outcome and pooled alike; here they localise where
# the absolute error concentrates.
field_by_outcome <- imap_dfr(pairs_list, function(pr, sub) {
  metrics_by_group(pr, "outcome",
                   f = \(p) pooled_metrics(p, tost_delta = tost_delta_pp,
                                           include_rmse = TRUE)) |>
    mutate(submission = sub)
}) |>
  mutate(outcome = factor(unname(outcome_label_map[outcome]),
                          levels = unname(outcome_label_map))) |>
  pivot_longer(c(directional_pct, spearman_rho, pearson_r, pearson_adj,
                 inferential_pct, rmse, rmse_adj, tost_pct),
               names_to = "metric", values_to = "value") |>
  relabel_metrics(ladder_labels)

# Restricted to the 0-100 continuous outcomes: with pp units the scale
# confound is gone, but a per-intervention correlation across all 13 outcomes
# would still mix constructs (attitude sliders vs. behaviors), so the
# conservative restriction is kept.
field_by_intervention <- imap_dfr(pairs_list, function(pr, sub) {
  metrics_by_group(pr |> filter(outcome %in% outcomes_continuous), "condition") |>
    mutate(submission = sub)
}) |>
  pivot_longer(c(directional_pct, spearman_rho, pearson_r, pearson_adj,
                 inferential_pct, tost_pct),
               names_to = "metric", values_to = "value") |>
  relabel_metrics(ladder_labels)

# Companion-table helper: field median per metric within each group level,
# with the human replication reference value in parentheses.
field_group_table <- function(field_long, group,
                              pct_metrics = c("Directional %", "Inferential %",
                                              "TOST %")) {
  fmt <- function(v, pct) if_else(is.na(v), "—",
                                  if_else(pct, sprintf("%.0f", v),
                                          sprintf("%.2f", v)))
  med <- field_long |>
    filter(submission != "Human replication") |>
    group_by(across(all_of(group)), metric) |>
    summarise(med = median(value, na.rm = TRUE), .groups = "drop")
  ref <- field_long |>
    filter(submission == "Human replication") |>
    select(all_of(group), metric, ref = value)
  med |>
    left_join(ref, by = c(group, "metric")) |>
    mutate(pct  = metric %in% pct_metrics,
           cell = paste0(fmt(med, pct), " (", fmt(ref, pct), ")")) |>
    select(all_of(group), metric, cell) |>
    pivot_wider(names_from = metric, values_from = cell)
}
```

```{r}
#| label: fig-field-by-group
#| fig-width: 11
#| fig-height: 6.5
#| fig-cap: "Where the field predicts well (placeholder data). Distribution of each approach's Pearson r by outcome (A, scored across the interventions) and by intervention (B, scored across the continuous outcomes). One dot per approach over the field's half-violin, with the field mean and its 95% CI (white diamond) on each row. Red crosses mark the human replication reference, the dotted line the null."

by_outcome_r <- field_by_outcome |>
  filter(metric == "Pearson r") |>
  mutate(metric = factor("Outcomes"))
by_intervention_r <- field_by_intervention |>
  filter(metric == "Pearson r") |>
  mutate(metric = factor("Interventions"))

p_by_outcome <- plot_metric_distribution(
  by_outcome_r, x = "outcome", mean_ci = TRUE,
  ref_lines = c("Outcomes" = 0)
) + labs(tag = "A", y = "Pearson r")
p_by_intervention <- plot_metric_distribution(
  by_intervention_r, x = "condition", mean_ci = TRUE,
  ref_lines = c("Interventions" = 0)
) + labs(tag = "B", y = "Pearson r")

p_by_outcome | p_by_intervention
```

```{r}
#| label: tbl-field-by-outcome
#| tbl-cap: "Field summary by outcome (placeholder data). Median across approaches of each ladder metric, scored per outcome across the interventions, with the human replication reference in parentheses. RMSE and RMSE (adj.) are in pp of scale range, the same unit as the pooled row. Per outcome they localize where the absolute error concentrates."

field_group_table(field_by_outcome, "outcome") |>
  gt() |>
  sub_missing(missing_text = "—") |>
  cols_label(outcome = "Outcome")
```

```{r}
#| label: tbl-field-by-intervention
#| tbl-cap: "Field summary by intervention (placeholder data). Median across approaches of each ladder metric, scored per intervention across the continuous outcomes, with the human replication reference in parentheses."

field_group_table(field_by_intervention, "condition") |>
  gt() |>
  sub_missing(missing_text = "—") |>
  cols_label(condition = "Intervention")
```

##### Subgroup recovery

Whether an approach reproduces *who* an intervention moves, not just its average effect. For each Tier-1 submission we fit the preregistered condition × moderator interaction model (`run_moderator_model()`) across the Section 2 outcome set (the illustrative outcomes plus the binary newsletter signup, the latter as a linear probability model, see *Estimation functions*) over the six demographic moderators, then pool the interaction estimates against Human 1. All interaction estimates are placed in percentage points of each outcome's scale range before pooling, the same unit as Section 1. Because interaction estimates require respondent-level data, this view covers Tier 1 only. As specified in Section 2 above, the signed-effect rungs are scored (directional agreement, Spearman ρ, Pearson r) plus the adjusted correlation r~adj~, emphasized here because small subgroup cells make the reference noise uneven across cells (see *Adjusted correlation and RMSE*).

```{r}
#| label: subgroup-pairs

# Human-1 interaction estimates, computed once and reused for every submission.
# The cached() keys keep their historical "amendment_" prefix (the analysis is
# unchanged); they are internal identifiers for analysis/results/*.rds, not tied
# to this file's name, so renaming the document does not require a recompute.
# The binary newsletter outcome runs through the same OLS machinery as a
# linear probability model: in the saturated condition × moderator
# specification the interaction coefficient is exactly the
# difference-in-differences of cell signup proportions — the interaction on
# the probability scale — and the HC2 robust SEs are valid for a binary
# response. It then pp-rescales (× 100) like every other outcome.
human_mod_side <- cached("amendment_human_mod_side", {
  map_dfr(outcomes_subgroup, function(out) {
    map_dfr(moderators_cat, function(mod) {
      run_moderator_model(human_data_1, outcome = out, moderator = mod,
                          compute_predicted = FALSE)$interaction_effects |>
        # se_h feeds the adjusted correlation; only the human reference side
        # is corrected (see "Adjusted correlation and RMSE").
        select(condition, moderator_level,
               estimate_h = estimate, se_h = std.error) |>
        mutate(outcome = out, moderator = mod)
    })
  })
})

build_subgroup_pairs <- function(pred_data) {
  pred_side <- map_dfr(outcomes_subgroup, function(out) {
    map_dfr(moderators_cat, function(mod) {
      run_moderator_model(pred_data, outcome = out, moderator = mod,
                          compute_predicted = FALSE)$interaction_effects |>
        select(condition, moderator_level, estimate_l = estimate) |>
        mutate(outcome = out, moderator = mod)
    })
  })
  inner_join(human_mod_side, pred_side,
             by = c("condition", "moderator_level", "outcome", "moderator")) |>
    pp_scale_pairs()
}

# Tier-1 submissions plus the reference are eligible. Interaction pairs are built
# once per submission and reused by the subgroup figure and its companion table.
subgroup_inputs <- c(
  set_names(
    map(tier1_ids, ~ llm_data_teams |> filter(team == paste0("team_", .x))),
    paste0("Team ", tier1_ids)
  ),
  list("Human replication" = human_data_2)
)
subgroup_pairs_list <- cached("amendment_subgroup_pairs", {
  map(subgroup_inputs, build_subgroup_pairs)
})

# Context statistic for reading the subgroup scores (see Section 2): the share
# of human interaction estimates that are individually significant, from the
# normal approximation on the interaction SEs, BH-adjusted within outcome ×
# moderator. Near 5% (the false-positive rate) = the human data show little
# true moderation, and the subgroup rungs are read as uninformative.
human_int_sig_share <- human_mod_side |>
  group_by(outcome, moderator) |>
  mutate(p_adj = p.adjust(2 * pnorm(-abs(estimate_h / se_h)), method = "BH")) |>
  ungroup() |>
  summarise(share = mean(p_adj < .05, na.rm = TRUE)) |>
  pull(share)
```

@fig-subgroup-headline shows the moderation analogue of the headline figure. Panel A shows the raw prediction task on the interaction estimates, and panel B the key metric per moderator, showing which kinds of heterogeneity the field can predict at all. The full signed-effect ladder, pooled, is in @tbl-field-subgroup. As preregistered in Section 2, the share of human interaction estimates that are individually significant is reported here as context: on the placeholder data it is `r sprintf("%.0f%%", 100 * human_int_sig_share)`, and a value near the 5% false-positive rate on the real data will mean the human sample shows little true moderation, so the subgroup rungs cannot separate approaches and are read as uninformative rather than as evidence about any submission.

```{r}
#| label: fig-subgroup-headline
#| fig-width: 11
#| fig-height: 4.2
#| fig-cap: "Subgroup recovery (placeholder data, Tier 1). (A) Predicted against human condition × moderator interaction estimates pooled across all approaches and the Section 2 outcome set (the four illustrative outcomes plus the binary newsletter signup), in percentage points of each outcome's scale range. Axes are centered on zero, the dashed line is the identity, and red crosses mark the human replication reference. The solid lines are the average calibration regression of human on predicted estimates (dark blue the field, red the human replication), with a corner readout of the mean field and human Pearson r (with its noise-corrected value) and calibration coefficients (β and α with 95% CIs). (B) Distribution of each approach's Pearson r on the interaction estimates, per moderator (pooled across the Section 2 outcomes), with the field mean and its 95% CI (white diamond) per row. Red crosses mark the reference, the dotted line the null."

moderator_label_map <- c(gender = "Gender", age_band = "Age",
                         race = "Race/ethnicity", education = "Education",
                         income = "Income", party = "Party ID")

# Pairs are already in pp of scale range (pp_scale_pairs() in the builder).
subgroup_scatter_pairs <- imap_dfr(subgroup_pairs_list,
                                   ~ mutate(.x, submission = .y))
# ~37,000 pooled pairs: much lighter marks than the ATE scatter, or it reads
# as a solid blob.
p_sub_scatter <- plot_pred_scatter(
  subgroup_scatter_pairs,
  x_lab       = "Predicted interaction estimate (pp)",
  y_lab       = "Human interaction estimate (pp)",
  tag         = "A",
  point_alpha = 0.05, point_size = 0.3,
  ref_alpha   = 0.2,  ref_size   = 0.5,
  symmetric   = TRUE
)

# Mean field / human Pearson r on the interaction estimates (from the pooled
# signed-effect ladder, matching panel B), and the calibration lines + readout.
subgroup_ld <- imap_dfr(subgroup_pairs_list,
                        ~ signed_metrics(.x) |> mutate(submission = .y))
sub_field_ld <- subgroup_ld |> filter(submission != "Human replication")
sub_human_ld <- subgroup_ld |> filter(submission == "Human replication")
p_sub_scatter <- p_sub_scatter +
  calibration_overlay(subgroup_scatter_pairs,
                      r_field    = mean(sub_field_ld$pearson_r,   na.rm = TRUE),
                      radj_field = mean(sub_field_ld$pearson_adj, na.rm = TRUE),
                      r_human    = sub_human_ld$pearson_r,
                      radj_human = sub_human_ld$pearson_adj)

field_subgroup_by_mod <- imap_dfr(subgroup_pairs_list, function(pr, sub) {
  pr |>
    group_by(moderator) |>
    group_modify(~ signed_metrics(.x)) |>
    ungroup() |>
    mutate(submission = sub)
}) |>
  transmute(submission,
            value      = pearson_r,
            moderator  = factor(unname(moderator_label_map[moderator]),
                                levels = unname(moderator_label_map)),
            metric     = factor("Moderators"))
p_sub_field <- plot_metric_distribution(
  field_subgroup_by_mod, x = "moderator", mean_ci = TRUE,
  ref_lines = c("Moderators" = 0)
) + labs(tag = "B", y = "Pearson r")

(p_sub_scatter | p_sub_field) + plot_layout(widths = c(1, 1.3))
```

```{r}
#| label: tbl-field-subgroup
#| tbl-cap: "Field summary of subgroup recovery (placeholder data, Tier 1). Center and spread across approaches of each signed-effect metric on the condition × moderator interaction estimates (in pp of scale range), pooled across the Section 2 outcomes (illustrative subset plus newsletter signup) and six moderators (human replication reference excluded)."

field_subgroup <- imap_dfr(subgroup_pairs_list, function(pr, sub) {
  signed_metrics(pr) |>
    pivot_longer(everything(), names_to = "metric", values_to = "value") |>
    mutate(submission = sub)
}) |>
  relabel_metrics(ladder_labels[c("directional_pct", "spearman_rho",
                                  "pearson_r", "pearson_adj")])

summarise_field(field_subgroup) |>
  gt() |>
  fmt_number(columns = c(mean, median, sd, min, max), decimals = 2) |>
  cols_label(metric = "Metric", mean = "Mean", median = "Median",
             sd = "SD", min = "Min", max = "Max")
```

##### Response-shape recovery

How well an approach reproduces the **shape** of the human response distribution, not only its mean. Using the preregistered `compare_distributions()` in the shared control condition, we score the overlap coefficient (OVL, 1 = identical densities), the **Wasserstein-1 distance** (W1, 0 = identical, computed as the integral of the absolute ECDF difference, in scale points), the Kolmogorov–Smirnov distance (KS, 0 = indistinguishable), and the variance ratio (clone/human, 1 = matched spread) for each continuous outcome. W1 is the field's current standard distributional metric for bounded ordinal scales and is added here as the bandwidth-free companion to OVL. The KDE-based OVL depends on kernel bandwidths, which scale with the two samples' sizes (a large synthetic sample is smoothed differently from the human half) and leak density past the 0/100 scale bounds. W1 has no tuning parameter and reads directly in scale points. The **variance ratio stays the headline diagnostic**, with the directional expectation of under-dispersion (ratio < 1). These are inherently per-outcome, control-condition quantities. @fig-distribution-headline therefore shows the raw response densities for every continuous outcome, one facet per outcome. Each facet stamps the field's median variance ratio and Wasserstein-1 distance next to the human replication's, so spread (VR) and overall distributional distance (W1) can be read at a glance for both the synthetic field and the human reference. All four metrics per outcome are in @tbl-field-distribution.

```{r}
#| label: distribution-pairs

dist_per_outcome <- function(pred_data) {
  map_dfr(outcomes_continuous, function(out) {
    compare_distributions(human_data_1, pred_data, out, control_lab,
                          include_w1 = TRUE) |>
      mutate(outcome = out)
  })
}

dist_labels <- c(ovl = "Overlap (OVL)", w1 = "Wasserstein-1",
                 ks_d = "KS distance", variance_ratio = "Variance ratio")

# Per-outcome shape metrics built once per submission, reused by the
# distribution figure and its companion table (cached: 29 x 11 KDE/ECDF
# passes over full respondent-level data).
dist_per_outcome_list <- cached("amendment_dist_per_outcome", {
  map(subgroup_inputs, dist_per_outcome)
})
```

```{r}
#| label: fig-distribution-headline
#| fig-width: 11
#| fig-height: 7
#| fig-cap: "Response-shape recovery (placeholder data, Tier 1). Control-condition response densities for every continuous outcome. In each facet, every Tier-1 approach is a thin line, Human 1 (the reference half) the bold line, and Human 2 (the replication half) the dashed red line. The corner stamp reports two numbers for that outcome, the variance ratio (VR, approach ÷ human, 1 = matched spread, below 1 = under-dispersion) and the Wasserstein-1 distance (W1, in scale points, 0 = identical distributions, larger = more different), for both the synthetic field (median across approaches, 'Field') and the human replication ('Human', Human 2 vs Human 1). VR captures spread, W1 captures how far the whole distribution sits from the human one in both location and shape. The full shape metrics per outcome are in the companion table."

# Per facet, stamp the field's median variance ratio and Wasserstein-1 distance
# alongside the human replication's (Human 2 vs Human 1), so each panel shows
# both spread (VR) and overall distributional distance (W1) for the synthetic
# field and the human reference.
field_dist_ann <- imap_dfr(dist_per_outcome_list,
                           ~ mutate(.x, submission = .y)) |>
  filter(submission != "Human replication") |>
  group_by(outcome) |>
  summarise(vr = median(variance_ratio, na.rm = TRUE),
            w1 = median(w1, na.rm = TRUE), .groups = "drop")

human_dist_ann <- dist_per_outcome_list[["Human replication"]] |>
  select(outcome, w1, vr = variance_ratio)

dist_ann <- left_join(field_dist_ann, human_dist_ann,
                      by = "outcome", suffix = c("_field", "_human")) |>
  mutate(label = sprintf("Field   VR %.2f   W1 %.2f\nHuman  VR %.2f   W1 %.2f",
                         vr_field, w1_field, vr_human, w1_human)) |>
  select(outcome, label)

# wrapped primary-outcome label: the full one clips its facet strip
density_labels <- outcome_label_map
density_labels["trust_multidimensional"] <- "Multidimensional trust\n(primary)"

plot_density_overlay(
  human_data_1, human_data_2, llm_data_teams,
  outcomes       = outcomes_continuous,
  condition_val  = control_lab,
  outcome_labels = density_labels,
  annotations    = dist_ann,
  x_lab          = "Response (0–100, control condition)"
)
```

```{r}
#| label: tbl-field-distribution
#| tbl-cap: "Field summary of response-shape recovery by outcome (placeholder data, Tier 1). Median across approaches of each shape metric in the control condition, per continuous outcome, with the human replication reference in parentheses. Ideal values are OVL = 1, W1 = 0, KS = 0, and variance ratio = 1."

field_distribution <- imap_dfr(dist_per_outcome_list, function(d, sub) {
  d |>
    pivot_longer(c(ovl, w1, ks_d, variance_ratio),
                 names_to = "metric", values_to = "value") |>
    mutate(submission = sub)
}) |>
  mutate(outcome = factor(unname(outcome_label_map[outcome]),
                          levels = unname(outcome_label_map))) |>
  relabel_metrics(dist_labels)

field_group_table(field_distribution, "outcome", pct_metrics = character()) |>
  gt() |>
  sub_missing(missing_text = "—") |>
  cols_label(outcome = "Outcome")
```

#### Finding 2 — Ranking approaches (secondary)

Having characterized the field, we rank individual approaches. This is the secondary read. Each metric is its own ranking, and the interesting result is *which* approaches lead on *which* rung. The leaderboard **table** is the main summary. The ranking *visual* for the key metric, with its cluster-bootstrap uncertainty intervals, is panel B of @fig-headline.

```{r}
#| label: tbl-leaderboard
#| tbl-cap: "Cross-team leaderboard (placeholder data). Pooled ATE-recovery metrics across all outcomes (every estimate in percentage points of its outcome's scale range), plus calibration slope and R² on the primary outcome and interval scores for Tiers 2–3, scored against the Human 1 reference half. The Human replication row (Human 1 vs. Human 2) shows how well two human halves of this size agree. Metrics run least to most strict, left to right. Class is the disclosure-class badge (A open, B escrowed, C sealed)."

leaderboard |>
  select(submission, tier, disclosure, directional_pct, spearman_rho, pearson_r,
         pearson_adj, pearson_fz, inferential_pct, rmse, rmse_adj, tost_pct,
         beta, calib_r2, pi_coverage_pct, interval_score) |>
  gt() |>
  fmt_number(columns = c(spearman_rho, pearson_r, pearson_adj, pearson_fz,
                         rmse, rmse_adj, beta, calib_r2, interval_score),
             decimals = 2) |>
  fmt_number(columns = c(directional_pct, inferential_pct, tost_pct,
                         pi_coverage_pct), decimals = 0) |>
  sub_missing(missing_text = "—") |>
  cols_label(
    submission      = "Submission",
    tier            = "Tier",
    disclosure      = "Class",
    directional_pct = "Directional %",
    spearman_rho    = "Spearman ρ",
    pearson_r       = "Pearson r",
    pearson_adj     = "Pearson r (adj.)",
    pearson_fz      = "Pearson r (per-outcome)",
    inferential_pct = "Inferential %",
    rmse            = "RMSE (pp)",
    rmse_adj        = "RMSE (adj., pp)",
    tost_pct        = "TOST %",
    beta            = "Calib. β",
    calib_r2        = "Calib. R²",
    pi_coverage_pct = "PI coverage %",
    interval_score  = "Interval score"
  ) |>
  tab_footnote(
    footnote = "Fisher-z average of the per-outcome correlations. Divergence from the pooled Pearson r flags a submission that predicts which outcomes move without ranking the interventions within them.",
    locations = cells_column_labels(columns = pearson_fz)
  ) |>
  tab_footnote(
    footnote = "Tiers 2–3 only. Coverage of the submitted 95% uncertainty intervals (honest intervals land near 95) and the mean Winkler interval score in pp of scale range, pooled across all outcomes (lower is better, and it is a proper score, so neither padding intervals wide nor squeezing them tight pays).",
    locations = cells_column_labels(columns = pi_coverage_pct)
  ) |>
  tab_footnote(
    footnote = "The floor's directional % reads 0, not 50, because sign(0) matches neither sign. This is an artifact of the all-zero prediction, not below-chance skill.",
    locations = cells_body(columns = directional_pct,
                           rows = submission %in% floor_labels)
  ) |>
  tab_style(
    style     = cell_fill(color = "#eef3f7"),
    locations = cells_body(rows = submission == "Human replication")
  ) |>
  tab_style(
    style     = cell_fill(color = "#f4f4f4"),
    locations = cells_body(rows = submission %in% floor_labels)
  )
```

##### Cross-metric consistency

A single-metric verdict can mislead. An approach that leads on one rung may trail on another, and @cumminsThreatAnalyticFlexibility2025b show empirically that performance on one data feature routinely fails to transfer to others. Rather than leave that to a reader's cross-panel comparison, we report it directly as the Spearman correlation between the approaches' rankings on each pair of ladder rungs (reference and floor rows excluded). High values mean the rungs sort the field similarly, so a summary rank would be defensible. Low or negative values mean each rung tells its own story and no single number should be trusted alone.

```{r}
#| label: tbl-metric-consistency
#| tbl-cap: "Cross-metric rank consistency (placeholder data). Spearman correlations between the approaches' rankings on each pair of pooled ladder metrics (reference and floor rows excluded). On placeholder data all scores are noise, so consistency is expected to be near zero."

metric_cols <- names(ladder_labels)

rank_consistency <- leaderboard |>
  filter(submission != "Human replication", !submission %in% floor_labels) |>
  select(all_of(metric_cols)) |>
  cor(method = "spearman", use = "pairwise.complete.obs")

rank_consistency |>
  as_tibble(rownames = "metric") |>
  mutate(metric = unname(ladder_labels[metric])) |>
  rename_with(~ unname(ladder_labels[.x]), all_of(metric_cols)) |>
  gt() |>
  fmt_number(columns = -metric, decimals = 2) |>
  sub_missing(missing_text = "—") |>
  cols_label(metric = "")
```

#### Design analysis of what the metrics can resolve at this sample size

Before locking the metric ladder, we check whether each rung has *dynamic range* at this design, meaning whether a genuinely good submission can score visibly better than the floor. The check uses design constants and literature priors only and touches no data. The inputs are the design's cell sizes (≈500 per intervention and ≈1,000 in control per human half), a typical 0–100 slider SD of ≈25 points, and plausible spreads of true megastudy ATEs (SDs of roughly 0.5–2 points, since large message-based megastudies rarely move attitude scales by more). Three conclusions follow. **First**, the reference half-sample ATE SE is ≈1.4 points, the noise floor under every raw metric. **Second**, we can compute the expected human-replication correlation and the expected correlation of a *perfect*, noiseless predictor for each plausible true-ATE spread. The perfect predictor's is the square root of the replication row's. When the true spread is small relative to 1.4, the r~adj~ NA-guard is expected to bind for many entries. **Third**, v1's relative TOST bound Δ = 0.5 × |ATE~h~| (≈0.25–1 point) is *smaller than the smallest difference the test could ever confirm* (1.645 × se~diff~ ≈ 2.3 points even for a noiseless submission), so that rung would read about 0% for every entry including the human replication row. That is a dead metric. The preregistered fixed bound of **3 scale points** (3% of range) gives a perfect predictor roughly 40% per-pair equivalence power while remaining unreachable for badly wrong predictions. Dynamic range is restored, at the cost of a laxer notion of equivalence, which we accept and state.

```{r}
#| label: tbl-design-analysis
#| tbl-cap: "Blinded design analysis (no data touched). Expected values of key scoring quantities from design constants (cell sizes ≈500/1,000 per half, outcome SD ≈25) under three plausible spreads of true ATEs. r (replication) is the expected human-replication correlation. r (perfect) is the expected raw correlation of an error-free predictor, the square root of the replication value. The TOST columns give the per-pair probability that a perfect predictor declares equivalence under the old relative bound and the new fixed bound Δ = 3."

sd_y     <- 25
n_arm_h  <- 500
n_ctrl_h <- 1000
se_ate_h <- sd_y * sqrt(1 / n_arm_h + 1 / n_ctrl_h)
z05      <- qnorm(0.95)

tost_power <- function(delta, se) {
  margin <- delta - z05 * se
  if (margin <= 0) 0 else pnorm(margin / se) - pnorm(-margin / se)
}

tibble(tau = c(0.5, 1, 2)) |>
  mutate(
    se_reference    = se_ate_h,
    r_replication   = tau^2 / (tau^2 + se_ate_h^2),
    r_perfect       = sqrt(r_replication),
    tost_relative   = map_dbl(0.5 * tau, tost_power, se = se_ate_h),
    tost_fixed_3    = tost_power(3, se_ate_h)
  ) |>
  gt() |>
  fmt_number(columns = -tau, decimals = 2) |>
  cols_label(
    tau           = "SD of true ATEs (pts)",
    se_reference  = "Reference SE (pts)",
    r_replication = "r (replication)",
    r_perfect     = "r (perfect predictor)",
    tost_relative = "TOST power, relative Δ",
    tost_fixed_3  = "TOST power, fixed Δ = 3"
  )
```

#### Robustness of the scoring pipeline

The scoring pipeline has researcher degrees of freedom of its own, chiefly the preregistered random seed that splits the human sample into Human 1 and Human 2, and secondarily the TOST bound. @cumminsThreatAnalyticFlexibility2025b show that such analytic choices can move fidelity conclusions substantially, and a skeptic could reasonably ask whether the reported reference value reflects the split we happened to draw. We therefore run a **resplit sensitivity check**. The half-split is redrawn many times with different seeds (`n_resplit_real = 200` for the real-data run; this placeholder render uses `n_resplit_demo = 50` to keep the build cheap, and as with the bootstrap only that parameter changes when the submissions arrive), the human replication reference is recomputed on each, and the spread of the reference's metrics across resplits is reported alongside the primary (preregistered-seed) values. A narrow spread shows that the seed is not driving the results. When real data arrive, the same check is run for the full field. Every submission is rescored against each resplit's Human 1, and the stability of the field summary (and of leaderboard ranks) across resplits is reported.

```{r}
#| label: tbl-resplit
#| tbl-cap: "Split-seed sensitivity of the human replication reference (placeholder data, 50 resplits; the real-data run uses the preregistered 200). Distribution of each pooled ATE-recovery metric for the Human 1 vs. Human 2 comparison across the alternative random half-splits, next to the value under the preregistered seed. On placeholder data the reference is itself noise-level, so the check demonstrates the code, not a substantive reference value. The resplits are drawn from the same sample and so are mutually dependent. The spread quantifies split-conditional variability, not independent replications."

resplit_ceiling <- cached("amendment_resplit_ceiling", {
  all_ids <- unique(human_data$id)

  map_dfr(seq_len(n_resplit_demo), function(i) {
    set.seed(1000 + i)
    ids_i <- sample(all_ids, size = floor(length(all_ids) / 2))
    h1_i  <- human_data |> filter(id %in% ids_i)
    h2_i  <- human_data |> filter(!id %in% ids_i)

    side_of <- function(dat, suffix) {
      map_dfr(outcomes_continuous, function(out) {
        prep_ate(run_main_treatment_model(dat, outcome = out)) |>
          mutate(outcome = out)
      }) |>
        rename_with(~ paste0(.x, suffix), c(estimate, se, p_adj))
    }

    inner_join(side_of(h1_i, "_h"), side_of(h2_i, "_l"),
               by = c("condition", "outcome")) |>
      mutate(infer_l = classify_sig(estimate_l, p_adj_l)) |>
      # continuous outcomes only, so raw units are already pp of scale range
      pooled_metrics(include_rmse = TRUE) |>
      mutate(resplit = i)
  })
})

resplit_ceiling |>
  pivot_longer(-resplit, names_to = "metric", values_to = "value") |>
  filter(metric %in% names(ladder_labels)) |>
  group_by(metric) |>
  summarise(mean = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE),
            min = min(value, na.rm = TRUE),  max = max(value, na.rm = TRUE),
            n_defined = sum(!is.na(value)), .groups = "drop") |>
  left_join(
    # Restricted to the continuous outcomes for comparability with the
    # resplits (whose side_of() maps outcomes_continuous only).
    pairs_list[["Human replication"]] |>
      filter(outcome %in% outcomes_continuous) |>
      pooled_metrics(include_rmse = TRUE) |>
      pivot_longer(everything(), names_to = "metric",
                   values_to = "preregistered_seed"),
    by = "metric"
  ) |>
  mutate(metric = factor(unname(ladder_labels[metric]),
                         levels = unname(ladder_labels))) |>
  arrange(metric) |>
  gt() |>
  fmt_number(columns = c(mean, sd, min, max, preregistered_seed), decimals = 2) |>
  sub_missing(missing_text = "—") |>
  cols_label(metric = "Metric", mean = "Mean", sd = "SD", min = "Min",
             max = "Max", n_defined = "N defined",
             preregistered_seed = "Preregistered seed")
```

When real submissions arrive, both findings are recomputed by replacing the placeholder datasets with the submitted prediction files routed through their tier's builder (`build_pairs_t1()` / `build_pairs_t2()` / `build_pairs_t3()`). The subgroup and distribution figures are produced for the tiers eligible for them, alongside the demographic-baseline and stereotyping diagnostics of Section 3, including the demographic parity gap (`demographic_parity_gap()`). The approach-characteristics table coded from the registration forms is then joined to the leaderboard for the exploratory analysis of which design choices accompany better predictions (publicly disclosed features only, with Class C entries excluded from this analysis, not from the leaderboard).

