---
title: "Building the behavioral clone profile bank from real human profiles"
subtitle: "GSS-as-spine sampling for the climate-trust megastudy"
author: "[Your name]"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
  pdf:
    toc: true
    number-sections: true
execute:
  echo: true
  eval: false
  warning: false
  message: false
---

```{r setup}
#| eval: false

# Required packages
# install.packages(c("tidyverse", "haven", "labelled", "fs", "glue", "here"))
# remotes::install_github("kjhealy/gssr")
# remotes::install_github("kjhealy/gssrdoc")

library(tidyverse)
library(gssr)        # GSS data
library(gssrdoc)     # GSS variable documentation
library(haven)       # labelled data handling
library(labelled)    # label utilities
library(glue)        # string interpolation
library(here)        # project paths
library(fs)          # file paths
library(tinytable)

# Reproducibility for any tie-breaking in deterministic replication
set.seed(2026)
```

# Overview

This document builds the profile bank used to prompt the LLM clones for our climate-trust megastudy. The main idea here is to use the demographics and attitudes of *real human personas*. This avoids creating fictional personas based on quotas (e.g., the Black, college-educated, evangelical Republican rancher in rural Maine). A major advantage is that, for the real personas, we do not only have demographic data, but also potentially relevant data on attitudes and beliefs (e.g., trust in scientists in general). The structure of this script is the following:

1. Loads the most recent (2024) General Social Survey (GSS) cross-section.
2. Selects demographic and opinion variables that match (or most closely correspond to) the items in our human survey.
3. Drops respondents with missing values on any selected variable (no imputation).
4. Draws **N = 500 unique profiles** by **weighted sampling without replacement**, with selection probability proportional to each respondent's post-stratification weight.
5. Generates two parallel sets of prompt profiles --- a **basic** one (Hewitt-style minimal demographics) and a **detailed** one (full demographic + opinion profile) --- so we can compare prompt richness as a robustness analysis.

Each of the 500 profiles will be queried in **every** experimental condition, in independent API sessions. This way, we have no sampling noise, and the different conditions represent actual counterfactuals for the 500 different profiles. 

# Design rationale

We use the most recent wave of the GSS (2024) rather than Census data as the source for the profiles because GSS contains both the demographic variables we need *and* the most directly relevant opinion variable for our outcome (`confsci`, confidence in the scientific community). ACS PUMS would buy us finer demographic granularity but no attitudes.

We will use **Weighted sampling *without* replacement.** Draw N=500 unique rows, probability proportional to weight. Every donor has a nonzero chance of being in the bank, no donor appears twice, and the realized marginal distributions converge to the weighted GSS marginals.

Per-condition N is therefore 500 across all 21 conditions, yielding 500 × 21 = 10,500 clone responses per outcome. This corresponds to roughly half of the human sample size *per condition*.

We generate two parallel prompt sets:

- **Basic profile:** age, sex, race, education, party identification.
- **Detailed profile:** all basic profile variables, as well as income, social class, urban/rural, religion, religiosity, political ideology, and confidence in the scientific community.

Whether richer profiles improve clone accuracy is empirically open. Argyle et al. found rich profiles helped; Wang et al. and Bisbee et al. found that more identity information can produce flattened, stereotyped responses.

# Variable mapping

The table below shows how each demographic and opinion variable in the human survey maps to a GSS variable. Variables not in GSS are skipped (and noted as such).

```{r}
#| label: tbl-variable-mapping
#| tbl-cap: "Mapping of human-survey items to GSS variables."


variable_mapping <- tribble(
  ~`Human-survey item`,                  ~`GSS variable`,        ~Notes,

  "Gender",                              "`sex`",                "GSS is binary (1 = male, 2 = female); the human 'Other' response has no GSS analogue.",
  "Age",                                 "`age`",                "GSS records age directly (top-coded at 89). Recovered from year of birth in the human survey.",
  "Race",                                "`race` + `hispanic`",  "Combined post-hoc; matching to the human multi-select schema is downstream.",
  "Education",                           "`degree`",             "0 = LT high school, 1 = HS, 2 = junior college, 3 = bachelor, 4 = graduate.",
  "Education climate science",           "--",                   "Not in GSS. Omitted from the profile.",
  "Income",                              "`income16`",           "26-bracket variable in dollars; mapping to the human 5-bracket scheme is downstream.",
  "Household size",                      "`hompop`",             "Direct count.",
  "Social class",                        "`class`",              "1 = lower, 2 = working, 3 = middle, 4 = upper. Direct match.",
  "Urban rural",                         "`srcbelt`",            "7-category place-size variable.",
  "Geographical location (postcode)",    "`region`",             "GSS gives Census region only. We retain `region` as the geography proxy.",
  "Partisan identity -- Party",          "`partyid`",            "0–7 scale; can be collapsed downstream.",
  "Partisan identity -- Importance",     "--",                   "Not in GSS. Strength is partly captured by `partyid`'s endpoints (strong Dem / strong Rep).",
  "Religion -- Affiliation",             "`relig`",              "Categorical.",
  "Religion -- Born-again",              "`reborn`",             "Y/N.",
  "Religion -- Religiosity",             "`reliten`",            "Strength of religious affiliation.",
  "Religious attendance (supplement)",   "`attend`",             "Frequency of attendance; not in the human survey, retained for richness.",
  "Need for epistemic autonomy",         "--",                   "Not in GSS.",
  "Belief in climate change",            "--",                   "Not in GSS regularly.",
  "Trust in climate scientists (DV)",    "--",                   "Outcome variable; not used as a profile attribute.",
  "(Opinion) Confidence in science",     "`confsci`",            "Closest GSS analogue of the trust outcome; included in the *detailed* profile only.",
  "(Opinion) Political views",           "`polviews`",           "1–7 liberal–conservative; included in the *detailed* profile only."
)

variable_mapping |>
  tt(width = c(0.28, 0.18, 0.54)) |>
  format_tt(j = c("GSS variable", "Notes"), markdown = TRUE)
```

# Step 1. Load the most recent GSS cross-section

```{r load-gss}
#| eval: false

# Load the most recent available wave. Update `target_year` if a newer
# release is available.
target_year <- 2024

gss_recent <- gss_get_yr(2024)

glue("GSS {target_year}: {nrow(gss_recent)} respondents loaded.")
```

# Step 2. Select and clean the relevant variables

We select only the variables we will use, attach human-readable factor labels, and convert weights to numeric.

```{r select-vars}
#| eval: false

selected_vars <- c(
  # weighting and identifiers
  "year", "id", "wtssps",
  # core demographics
  "age", "sex", "race", "hispanic",
  "degree", "income16", "hompop", "class",
  "region",
  # politics
  "partyid", "polviews",
  # religion
  "relig", "reborn", "reliten", "attend",
  # opinion
  "consci"
)

numeric_vars <- c("age", "hompop", "wtssps")
factor_vars  <- setdiff(selected_vars, c("year", "id", numeric_vars))

gss_sel <- gss_recent |>
  select(any_of(selected_vars)) |>
  mutate(across(all_of(numeric_vars), ~ as.numeric(haven::zap_labels(.x)))) |>
  mutate(across(all_of(factor_vars),  ~ haven::as_factor(.x)))
```

# Step 3. Drop missing data

Respondents must be non-missing on the post-stratification weight and on the core demographic variables, but `NA` is allowed on opinion / intensity variables and is handled later in the prompt builder.

**Weight NAs are structural, not random.** `wtssps` is missing for ~17% of GSS 2024 respondents because they could not be placed in a post-stratification cell (typically due to refusing one of the post-strat variables themselves). We cannot weight these respondents, so they have to come out.

## Step 3a. Recode GSS missing-value labels to proper NA

In GSS labelled data, missingness appears as named factor levels rather than real `NA`s. We recode the four common labels (`"don't know"`, `"no answer"`, `"skipped on web"`, `"iap"`) before applying the missing-data rule.

```{r recode-na}
#| eval: false

gss_missing_labels <- c("don't know", "no answer", "skipped on web", "iap")

gss_sel <- gss_sel |>
  mutate(across(where(is.factor), ~ {
    chr <- as.character(.x)
    chr[tolower(chr) %in% gss_missing_labels] <- NA
    factor(chr, levels = setdiff(levels(.x), gss_missing_labels))
  }))
```

## Step 3b. Tiered missing-data rule

```{r drop-missing}
#| eval: false

# Required: respondents missing any of these are dropped.
# Includes the post-stratification weight itself (~17% NA in GSS 2024)
# plus the core demographic variables we need for prompt construction.
required_vars <- c(
  "wtssps",                           # post-strat weight; drop NA
  "age", "sex", "race", "hispanic",
  "degree", "income16", "class",
  "region",
  "partyid", "polviews",
  "relig"
)

gss_complete <- gss_sel |>
  drop_na(all_of(required_vars))
```

## Step 3c. Audit

```{r audit}
#| eval: false

n_before <- nrow(gss_sel)
n_after  <- nrow(gss_complete)

cat(sprintf(
  "Required-vars complete cases: %d of %d (%.1f%%).\n",
  n_after, n_before, 100 * n_after / n_before
))

optional_vars <- c("reborn", "reliten", "attend", "consci", "hompop")

# Per-variable NA rates among the kept respondents (optional variables only)
gss_complete |>
  summarise(across(all_of(optional_vars), ~ mean(is.na(.x)))) |>
  pivot_longer(everything(),
               names_to  = "variable",
               values_to = "pct_na") |>
  mutate(pct_na = round(100 * pct_na, 1))
```

# Step 4. Weighted sampling without replacement {#sec-replicate-step}

We draw 500 unique GSS respondents, with selection probability proportional to their `wtssps` post-stratification weight.

```{r sample}
#| eval: false

target_n <- 500

profiles <- gss_complete |>
  slice_sample(n = target_n, weight_by = wtssps, replace = FALSE) |>
  mutate(profile_id = sprintf("p%04d", row_number())) |>
  select(profile_id, everything())

glue("Profile bank built: N = {nrow(profiles)} unique clones from ",
     "{nrow(gss_complete)} eligible GSS respondents.")
```

Because sampling is random, the script is reproducible only with the seed set in the setup (`set.seed(2026)`). Re-running with a different seed will produce a different 500-profile bank.

# Step 5. Validate the profile bank against weighted GSS

A sanity check: the marginal distribution of each variable in `profiles` should approximate the *weighted* marginal distribution in the GSS source. With N = 500 there will be sampling variation, so deviations of a few percentage points on small categories are expected. We compute side-by-side summaries.

```{r validate}
#| eval: false

validate_marginal <- function(var) {
  src <- gss_complete |>
    count(.data[[var]], wt = wtssps) |>
    mutate(prop_gss = n / sum(n)) |>
    select(-n)
  prof <- profiles |>
    count(.data[[var]]) |>
    mutate(prop_profiles = n / sum(n)) |>
    select(-n)
  full_join(src, prof, by = var) |>
    mutate(diff = prop_profiles - prop_gss) |>
    mutate(across(c(prop_gss, prop_profiles, diff), ~ round(.x, 4)))
}

vars_to_check <- selected_vars

walk(vars_to_check, \(v) {
  cat("\n----", v, "----\n")
  print(validate_marginal(v))
})
```

With N = 500 drawn from a donor pool of ~2,000+, the largest absolute deviation on a multi-category variable will typically be a few percentage points. Systematically large deviations (e.g., > 5 pp on a major category) indicate either a coding error or that the seed produced an unrepresentative draw.

# Step 6. Construct prompt-profile strings

We build two parallel profile strings per clone: a **basic** one and a **detailed** one. Each is a self-contained natural-language description of the simulated respondent. The downstream simulation script will concatenate this string with (a) a role/framing message, (b) the experimental stimulus, and (c) the outcome question.

The detailed version retains GSS labels verbatim. We write small helpers that translate factor levels into natural prose, but we do **not** recode GSS values into the human-survey response categories --- that mapping is a separate step.

```{r prompt-helpers}
#| eval: false

library(purrr)   # for compact()

lbl <- function(x) {
  v <- if (is.factor(x)) as.character(x) else as.character(x)
  if (length(v) == 0 || is.na(v)) NA_character_ else v
}

# Returns NULL when value is NA, so compact() can drop the line
opt_line <- function(label, value, transform = tolower) {
  v <- lbl(value)
  if (is.na(v)) return(NULL)
  glue("- {label}: {transform(v)}.")
}

format_relig_line <- function(relig, reliten, attend, reborn) {
  r <- lbl(relig)
  if (is.na(r)) return(NULL)
  if (tolower(r) == "none") return("- Religion: not religious.")
  parts <- glue("- Religion: {tolower(r)}.")
  rl <- lbl(reliten); if (!is.na(rl)) parts <- c(parts, glue("- Strength of religious affiliation: {tolower(rl)}."))
  at <- lbl(attend);  if (!is.na(at)) parts <- c(parts, glue("- Religious service attendance: {tolower(at)}."))
  rb <- lbl(reborn);  if (!is.na(rb) && tolower(rb) == "yes") {
    parts <- c(parts, "- Self-identifies as a born-again or evangelical Christian.")
  }
  paste(parts, collapse = "\n")
}

build_basic_profile <- function(row) {
  glue(
    "You are simulating the response of a U.S. resident with the following ",
    "characteristics: ",
    "age {row$age}; ",
    "sex {tolower(lbl(row$sex))}; ",
    "race {tolower(lbl(row$race))}; ",
    "highest educational degree: {tolower(lbl(row$degree))}; ",
    "party identification: {tolower(lbl(row$partyid))}."
  )
}

build_detailed_profile <- function(row) {
  lines <- list(
    "You are simulating the response of a U.S. resident with the following characteristics.",
    glue("- Age: {row$age} years old."),
    glue("- Sex: {tolower(lbl(row$sex))}."),
    glue("- Race: {tolower(lbl(row$race))}."),
    opt_line("Hispanic origin",                            row$hispanic),
    glue("- Highest educational degree: {tolower(lbl(row$degree))}."),
    glue("- Household income (GSS bracket): {tolower(lbl(row$income16))}."),
    opt_line("Household size",                             row$hompop, transform = identity),
    glue("- Self-described social class: {tolower(lbl(row$class))}."),
    glue("- Census region: {tolower(lbl(row$region))}."),
    glue("- Party identification: {tolower(lbl(row$partyid))}."),
    glue("- Political views (liberal-conservative): {tolower(lbl(row$polviews))}."),
    format_relig_line(row$relig, row$reliten, row$attend, row$reborn),
    opt_line("Confidence in the scientific community",     row$consci)
  )
  paste(compact(lines), collapse = "\n")
}
```

```{r build-profile-strings}
#| eval: false

profiles <- profiles |>
  rowwise() |>
  mutate(
    profile_basic    = build_basic_profile(pick(everything())),
    profile_detailed = build_detailed_profile(pick(everything()))
  ) |>
  ungroup()
```

# Step 7. Save outputs

```{r save}
#| eval: false

# (a) Full profile table with raw GSS values + both prompt strings
write_csv(profiles, "clone_profiles/profiles.csv")

# (b) Compact prompt-only file: one row per clone, two prompt columns
profiles |>
  select(profile_id, profile_basic, profile_detailed) |>
  write_csv("clone_profiles/prompts.csv")
```

