---
title: "Marginal Quotas for Clone Sampling: Census vs. GSS"
subtitle: "Non-crossed quotas for AGE, GENDER, PARTY, RACE, EDUCATION, POLITICAL_ORIENTATION"
author: "[Your name]"
date: today
format:
  html:
    toc: true
    number-sections: true
execute:
  echo: true
  eval: false
  warning: false
  message: false
---

# Purpose

This script produces **non-crossed (marginal) quotas** for the six profile variables we need for an alternative prompting strategy: `AGE`, `GENDER`, `PARTY`, `RACE`, `EDUCATION`, `POLITICAL_ORIENTATION`.

We compute each marginal twice:

- **Census-based** quotas, using `tidycensus` against the most recent Census Population Estimates (vintage 2024) for `gender`, `age`, `race`, and the most recent ACS 1-year release for `education`. Census has no measure of party or political orientation, so those rows are `NA` on the Census side.
- **GSS-based** quotas, computed from the most recent GSS wave with post-stratification weights (`wtssps`). GSS covers all six variables.

The script binds the two sets and saves a comparison table.

# Setup

```{r setup}
#| eval: false

# install.packages(c("tidycensus", "tidyverse", "haven", "labelled", "fs", "here"))
# remotes::install_github("kjhealy/gssr")

library(tidycensus)
library(tidyverse)
library(gssr)
library(haven)
library(labelled)
library(fs)
library(here)

tidycensus::census_api_key(Sys.getenv("my_census_api"))

out_dir <- here("outputs", "quotas_clones")
dir_create(out_dir)
```

# Census quotas

We mirror the approach used in the main megastudy's `quotas.R` for the variables Census measures, and add education from ACS 1-year estimates.

## Gender (18+)

```{r census-gender}
#| eval: false

census_gender <- get_estimates(
  geography        = "state",
  product          = "characteristics",
  breakdown        = c("SEX", "AGE"),
  breakdown_labels = TRUE,
  vintage          = 2024
) |>
  filter(SEX %in% c("Male", "Female"), AGE >= 18) |>
  group_by(SEX) |>
  summarise(value = sum(value), .groups = "drop") |>
  mutate(
    variable    = "Gender",
    category    = SEX,
    pct_census  = value / sum(value)
  ) |>
  select(variable, category, pct_census)
```

## Age (18+)

```{r census-age}
#| eval: false

census_age <- get_estimates(
  geography        = "state",
  product          = "characteristics",
  breakdown        = "AGE",
  breakdown_labels = TRUE,
  vintage          = 2024
) |>
  filter(AGE >= 18) |>
  mutate(
    category = case_when(
      AGE <= 29 ~ "18-29",
      AGE <= 44 ~ "30-44",
      AGE <= 59 ~ "45-59",
      TRUE      ~ "60+"
    )
  ) |>
  group_by(category) |>
  summarise(value = sum(value), .groups = "drop") |>
  mutate(
    variable   = "Age",
    pct_census = value / sum(value)
  ) |>
  select(variable, category, pct_census)
```

## Race / Ethnicity (18+)

```{r census-race}
#| eval: false

census_race <- get_estimates(
  geography        = "state",
  product          = "characteristics",
  breakdown        = c("RACE", "HISP", "AGE"),
  breakdown_labels = TRUE,
  vintage          = 2024
) |>
  filter(HISP != "Both Hispanic Origins", AGE >= 18) |>
  mutate(
    category = case_when(
      HISP == "Hispanic"      ~ "Hispanic / Latino",
      RACE == "White alone"   ~ "White (non-Hispanic)",
      RACE == "Black alone"   ~ "Black / African American",
      RACE == "Asian alone"   ~ "Asian / Asian American",
      TRUE                    ~ "Other"
    )
  ) |>
  group_by(category) |>
  summarise(value = sum(value), .groups = "drop") |>
  mutate(
    variable   = "Race / Ethnicity",
    pct_census = value / sum(value)
  ) |>
  select(variable, category, pct_census)
```

## Education (25+)

::: {.callout-note}
ACS table `B15003` reports educational attainment for the population aged 25 and older — the standard table for educational distributions. The 18–24 group is excluded because most of them are still in school. This means the resulting education marginal is for the *25+* adult population, slightly older than the survey target (18+). For quota purposes the deviation is small (the 18–24 group is ~12% of adults and skews lower in completed education), and using B15003 keeps the script simple. If you want strict 18+ education, swap in table B15001 and aggregate by sex × age × education.
:::

We collapse the 25 disaggregated education codes in B15003 into the same 5 brackets used in the human survey (with Master's, Professional, and Doctorate lumped, since GSS lumps them into a single "Graduate" code).

```{r census-education}
#| eval: false

# Most recent 1-year ACS available; update if a newer release exists
acs_year <- 2023

census_edu <- get_acs(
  geography = "us",
  table     = "B15003",
  year      = acs_year,
  survey    = "acs1"
) |>
  separate(variable, into = c("table", "code"), sep = "_") |>
  mutate(
    code_num = as.integer(code),
    category = case_when(
      code_num %in% 2:16            ~ "Less than high school",
      code_num %in% c(17, 18)       ~ "High school / GED",
      code_num %in% 19:21           ~ "Some college / Associate's",
      code_num == 22                ~ "Bachelor's",
      code_num %in% 23:25           ~ "Graduate (Master's / Prof / Doctorate)",
      TRUE                          ~ NA_character_
    )
  ) |>
  filter(!is.na(category)) |>
  group_by(category) |>
  summarise(value = sum(estimate), .groups = "drop") |>
  mutate(
    variable   = "Education",
    pct_census = value / sum(value)
  ) |>
  select(variable, category, pct_census)
```

## Party and political orientation (not in Census)

```{r census-politics-na}
#| eval: false

census_party <- tibble(
  variable   = "Party",
  category   = c("Democrat", "Republican", "Independent", "Other"),
  pct_census = NA_real_
)

census_polorient <- tibble(
  variable   = "Political orientation",
  category   = c("Liberal", "Moderate", "Conservative"),
  pct_census = NA_real_
)
```

## Bind census quotas

```{r census-bind}
#| eval: false

census_quotas <- bind_rows(
  census_gender, census_age, census_race,
  census_edu, census_party, census_polorient
)
```

# GSS quotas

We use the most recent GSS wave with post-stratification weights. All six variables come from the same source, so categories are internally consistent.

```{r gss-load}
#| eval: false

target_year <- 2022

data(gss_all)

gss_recent <- gss_all |>
  filter(year == target_year) |>
  filter(!is.na(wtssps)) |>
  select(year, age, sex, race, hispanic, degree, partyid, polviews, wtssps) |>
  mutate(across(c(sex, race, hispanic, degree, partyid, polviews),
                ~ as_factor(.x))) |>
  mutate(age = as.numeric(age),
         wtssps = as.numeric(wtssps))
```

## Helper: weighted marginal proportion

```{r gss-helper}
#| eval: false

weighted_pct <- function(df, var, weight = "wtssps") {
  df |>
    drop_na({{ var }}) |>
    count({{ var }}, wt = .data[[weight]], name = "value") |>
    mutate(pct_gss = value / sum(value)) |>
    rename(category = {{ var }}) |>
    mutate(category = as.character(category))
}
```

## Gender

```{r gss-gender}
#| eval: false

gss_gender <- gss_recent |>
  weighted_pct(sex) |>
  mutate(variable = "Gender",
         category = recode(category,
                           "MALE"   = "Male",
                           "FEMALE" = "Female")) |>
  select(variable, category, pct_gss)
```

## Age

```{r gss-age}
#| eval: false

gss_age <- gss_recent |>
  filter(!is.na(age)) |>
  mutate(category = case_when(
    age <= 29 ~ "18-29",
    age <= 44 ~ "30-44",
    age <= 59 ~ "45-59",
    TRUE      ~ "60+"
  )) |>
  count(category, wt = wtssps, name = "value") |>
  mutate(pct_gss   = value / sum(value),
         variable  = "Age") |>
  select(variable, category, pct_gss)
```

## Race / Ethnicity

```{r gss-race}
#| eval: false

# GSS race + hispanic combined into the same 5-category schema used for Census
gss_race <- gss_recent |>
  drop_na(race, hispanic) |>
  mutate(
    is_hisp = !grepl("NOT HISPANIC", toupper(as.character(hispanic))),
    category = case_when(
      is_hisp                                 ~ "Hispanic / Latino",
      toupper(as.character(race)) == "WHITE"  ~ "White (non-Hispanic)",
      toupper(as.character(race)) == "BLACK"  ~ "Black / African American",
      TRUE                                    ~ "Other"  # GSS lumps Asian into "Other"
    )
  ) |>
  count(category, wt = wtssps, name = "value") |>
  mutate(pct_gss   = value / sum(value),
         variable  = "Race / Ethnicity") |>
  select(variable, category, pct_gss)
```

::: {.callout-note}
GSS's `race` variable has only three levels (White, Black, Other), so Asian respondents are folded into Other. This is a known limitation of GSS race coding. The Census categories are more granular but we collapse to GSS's bins to make the comparison meaningful. If you want to keep Asian separate, you would need to use `racecen1` (the more detailed Census-recoded race variable, available in some years).
:::

## Education

```{r gss-education}
#| eval: false

gss_edu <- gss_recent |>
  drop_na(degree) |>
  mutate(category = case_when(
    grepl("LT HIGH SCHOOL|LESS THAN", toupper(as.character(degree))) ~ "Less than high school",
    grepl("HIGH SCHOOL", toupper(as.character(degree)))              ~ "High school / GED",
    grepl("JUNIOR|ASSOCIATE", toupper(as.character(degree)))         ~ "Some college / Associate's",
    grepl("BACHELOR", toupper(as.character(degree)))                  ~ "Bachelor's",
    grepl("GRADUATE", toupper(as.character(degree)))                  ~ "Graduate (Master's / Prof / Doctorate)",
    TRUE                                                              ~ NA_character_
  )) |>
  drop_na(category) |>
  count(category, wt = wtssps, name = "value") |>
  mutate(pct_gss   = value / sum(value),
         variable  = "Education") |>
  select(variable, category, pct_gss)
```

## Party

```{r gss-party}
#| eval: false

# partyid: 0 = strong Dem ... 6 = strong Rep, 7 = other
gss_party <- gss_recent |>
  drop_na(partyid) |>
  mutate(
    pid_chr = toupper(as.character(partyid)),
    category = case_when(
      grepl("STRONG DEMOCRAT|NOT STR DEMOCRAT|IND,NEAR DEM", pid_chr) ~ "Democrat",
      grepl("STRONG REPUBLICAN|NOT STR REPUBLICAN|IND,NEAR REP", pid_chr) ~ "Republican",
      grepl("INDEPENDENT", pid_chr)                                       ~ "Independent",
      grepl("OTHER", pid_chr)                                             ~ "Other",
      TRUE                                                                ~ NA_character_
    )
  ) |>
  drop_na(category) |>
  count(category, wt = wtssps, name = "value") |>
  mutate(pct_gss   = value / sum(value),
         variable  = "Party") |>
  select(variable, category, pct_gss)
```

::: {.callout-note}
GSS `partyid` distinguishes strong/weak partisans and "leaners" (independents who lean toward a party). We collapse leaners *with* their party (the standard "branched" recoding); pure independents stay as "Independent". Adjust the `case_when` if you prefer the strict-categorical recoding (where leaners count as Independent).
:::

## Political orientation

```{r gss-polorient}
#| eval: false

# polviews: 1 = extremely liberal ... 7 = extremely conservative
gss_polorient <- gss_recent |>
  drop_na(polviews) |>
  mutate(
    pv_chr = toupper(as.character(polviews)),
    category = case_when(
      grepl("LIBERAL", pv_chr)      ~ "Liberal",
      grepl("MODERATE", pv_chr)     ~ "Moderate",
      grepl("CONSERVATIVE", pv_chr) ~ "Conservative",
      TRUE                          ~ NA_character_
    )
  ) |>
  drop_na(category) |>
  count(category, wt = wtssps, name = "value") |>
  mutate(pct_gss   = value / sum(value),
         variable  = "Political orientation") |>
  select(variable, category, pct_gss)
```

## Bind GSS quotas

```{r gss-bind}
#| eval: false

gss_quotas <- bind_rows(
  gss_gender, gss_age, gss_race,
  gss_edu, gss_party, gss_polorient
)
```

# Comparison data frame

We join the two sources on `variable` and `category`, compute the difference, and order the rows for readability.

```{r compare}
#| eval: false

quota_compare <- full_join(
  census_quotas, gss_quotas,
  by = c("variable", "category")
) |>
  mutate(
    diff_gss_minus_census = pct_gss - pct_census,
    pct_census            = round(pct_census, 4),
    pct_gss               = round(pct_gss, 4),
    diff_gss_minus_census = round(diff_gss_minus_census, 4)
  ) |>
  arrange(
    factor(variable, levels = c("Gender", "Age", "Race / Ethnicity",
                                "Education", "Party", "Political orientation")),
    category
  )

quota_compare
```

# Save

```{r save}
#| eval: false

write_csv(quota_compare, path(out_dir, "quotas_comparison.csv"))
write_csv(census_quotas, path(out_dir, "quotas_census.csv"))
write_csv(gss_quotas,    path(out_dir, "quotas_gss.csv"))

cat(sprintf(
  "Saved 3 files to %s:\n  - quotas_comparison.csv (Census vs GSS, side-by-side)\n  - quotas_census.csv\n  - quotas_gss.csv\n",
  out_dir
))
```

# Notes

- **Census party / political orientation** are intentionally `NA` — the Census Bureau does not measure these. Any clone-sampling strategy that requires party or ideology quotas must use a survey-based source (GSS, ANES, Pew).
- **Education age range** is 25+ for Census (table B15003) and 18+ for GSS (the standard adult population). The two are not perfectly comparable; expect Census to show modestly higher educational attainment because younger adults still in school are excluded. Swap in B15001 if strict 18+ is needed.
- **Race coding** is collapsed to the GSS schema (which lacks a separate Asian category). To preserve Asian as a distinct quota, use `racecen1` from GSS (where available) and re-run the race section.
- **GSS party recoding** treats leaners as partisans (the standard branched recoding). Adjust the `case_when` in the `gss_party` block if you prefer pure-independent treatment.
- **Output schema** of `quotas_comparison.csv` is one row per (variable × category) with columns `pct_census`, `pct_gss`, `diff_gss_minus_census`. Plug this into your simulation framework as the quota target.
