---
title: "Building the 9,000-profile pool by pooling multiple GSS cross-sections"
subtitle: "Alternative to single-year weight duplication"
author: "[Your name]"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
  pdf:
    toc: true
    number-sections: true
execute:
  echo: true
  eval: false
  warning: false
  message: false
---

```{r setup}
#| eval: false

# Required packages
# install.packages(c("tidyverse", "haven", "labelled", "fs", "glue", "here"))
# remotes::install_github("kjhealy/gssr")

library(tidyverse)
library(gssr)       # GSS data (cumulative gss_all + gss_get_yr)
library(haven)      # labelled data handling
library(labelled)   # label utilities
library(glue)       # string interpolation
library(here)       # project paths
library(fs)         # file paths

set.seed(2026)
```

# Overview

This script builds a 9,000-profile demographic pool for LLM simulation by pooling
multiple recent GSS cross-sectional waves rather than duplicating respondents by
weight from a single year.

**Key design decisions:**

- **Years used**: The 4 most recent distinct cross-sectional GSS waves (2016, 2018,
  2022, 2024). GSS was not fielded in 2020 (COVID); 2021 was a panel recontact of
  prior respondents, not a fresh cross-section.
- **Weight reconciliation**: Each year's `wtssps` is normalized within year (divides
  by that year's weight sum), so each wave contributes equal total weight to the
  sampling lottery while preserving within-year relative weights.
- **Sampling**: Weighted without replacement (N = 9,000). If the 4-year pool falls
  short of 9,000 complete cases, the script extends to 2014 automatically and warns.
- **Output**: Demographic dataset only. Prompt strings are not generated here —
  that step is handled by a collaborator in a separate project.

This approach avoids the artificial duplication inherent in weight-cloning from a
single year (where high-weight respondents would appear 3–4×), replacing it with
genuinely different individuals surveyed in different years.

# Variable selection

Same 19 GSS variables as `build_realistic_clone_profiles.qmd`, minus prompt output.

```{r variable-setup}
#| eval: false

selected_vars <- c(
  # weighting and identifiers
  "year", "id", "wtssps",
  # core demographics
  "age", "sex", "race", "hispanic",
  "degree", "income16", "hompop", "class",
  "region", "srcbelt",
  # politics
  "partyid", "polviews",
  # religion
  "relig", "reborn", "reliten", 
  # opinion
  "consci"
)

# Variables where NA means "drop respondent"
required_vars <- c(
  "wtssps",
  "age", "sex", "race", "hispanic",
  "degree", "income16", "class",
  "region",
  "partyid", "polviews",
  "relig", "srcbelt"
)

# Variables where NA is acceptable (handled downstream)
optional_vars <- c("hompop", "reborn", "reliten", "consci")

numeric_vars <- c("age", "hompop", "wtssps")
factor_vars  <- setdiff(selected_vars, c("year", "id", numeric_vars))
```

# Step 1. Load the cumulative GSS and select target years

We use `gss_all` (the full cumulative GSS file shipped with the `gssr` package) and
filter to our target years. This is more efficient than four separate `gss_get_yr()`
calls.

```{r load-gss}
#| eval: false

# The initial target: 4 most recent true cross-sectional waves
target_years <- c(2018, 2021, 2022, 2024)

data(gss_all)

gss_multi <- gss_all |>
  filter(year %in% target_years) |>
  select(any_of(selected_vars))

# Year audit — verify all target years loaded and check raw N per year
cat("Raw respondent counts by year:\n")
print(table(gss_multi$year))
cat(glue("\nTotal raw respondents across {length(target_years)} years: {nrow(gss_multi)}\n"))
```

# Step 2. Select, type-convert, and recode GSS missing labels

```{r clean-vars}
#| eval: false

# Convert weight and numeric vars; attach factor labels to all others
gss_sel <- gss_multi |>
  mutate(across(all_of(numeric_vars), ~ as.numeric(haven::zap_labels(.x)))) |>
  mutate(across(all_of(factor_vars),  ~ haven::as_factor(.x)))

# GSS encodes missingness as named factor levels rather than real NAs.
# Recode the four common "missing" labels to proper NA.
gss_missing_labels <- c("don't know", "no answer", "skipped on web", "iap")

gss_sel <- gss_sel |>
  mutate(across(where(is.factor), ~ {
    chr <- as.character(.x)
    chr[tolower(chr) %in% gss_missing_labels] <- NA
    factor(chr, levels = setdiff(levels(.x), gss_missing_labels))
  }))

# check
gss_sel |>
  summarise(across(everything(),
                   ~ mean(is.na(.x), na.rm = FALSE),
                   .names = "{.col}_na_share")) |>
  tidyr::pivot_longer(everything(),
                      names_to = "variable",
                      values_to = "na_share") |>
  mutate(variable = stringr::str_remove(variable, "_na_share"),
         na_share = scales::percent(na_share, accuracy = 0.1))
```

# Step 3. Remove missing data for required variables

```{r drop-missing}
#| eval: false

gss_complete <- gss_sel |>
  drop_na(all_of(required_vars))

# Per-year complete-case audit
cat("Complete cases by year (after required-var filter):\n")
gss_complete |>
  count(year) |>
  mutate(pct_of_raw = round(100 * n / table(gss_multi$year)[as.character(year)], 1)) |>
  print()

cat(glue(
  "\nTotal eligible pool: {nrow(gss_complete)} respondents ",
  "({round(100 * nrow(gss_complete) / nrow(gss_sel), 1)}% of {nrow(gss_sel)} multi-year rows)\n"
))

# Safety check: if pool is still under 9000, extend to 2014
if (nrow(gss_complete) < 9000) {
  warning(
    "Pool of ", nrow(gss_complete), " complete cases is below the target N = 9000. ",
    "Extending to 2014. Re-run from Step 1 with target_years <- c(2014, 2016, 2018, 2022, 2024)."
  )
}

# Per-variable NA rates among kept respondents (optional variables only)
cat("\nNA rates on optional variables (kept respondents):\n")
gss_complete |>
  summarise(across(all_of(optional_vars), ~ round(100 * mean(is.na(.x)), 1))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "pct_na") |>
  print()
```

# Step 4. Within-year weight normalization

Each year's `wtssps` values sum to that year's effective sample size, which differs
across waves. Normalizing within year ensures each wave contributes equal total weight
to the sampling lottery, while preserving the relative ordering of respondents within
each year.

```{r normalize-weights}
#| eval: false

gss_complete <- gss_complete |>
  group_by(year) |>
  mutate(w_pooled = wtssps / sum(wtssps, na.rm = TRUE)) |>
  ungroup()

# Sanity check: each year's normalized weights should sum to 1.0
cat("Sum of normalized weights by year (should be 1.0 each):\n")
gss_complete |>
  group_by(year) |>
  summarise(w_sum = round(sum(w_pooled, na.rm = TRUE), 6)) |>
  print()
```

# Step 5. Weighted sampling without replacement (N = 9,000)

```{r sample}
#| eval: false

target_n <- 9000

# Guard: pool must be large enough for without-replacement sampling
stopifnot(
  "Pool too small for without-replacement sampling — extend target_years to include 2014" =
    nrow(gss_complete) >= target_n
)

profiles <- gss_complete |>
  slice_sample(n = target_n, weight_by = w_pooled, replace = FALSE) |>
  select(-w_pooled) |>
  mutate(profile_id = sprintf("p%05d", row_number())) |>
  select(profile_id, everything())

glue("Profile pool built: N = {nrow(profiles)} unique individuals ",
     "drawn from {nrow(gss_complete)} eligible respondents across ",
     "{n_distinct(profiles$year)} GSS waves.")
```

# Step 6. Validate the profile pool

## Year balance

Normalized within-year weights means each year contributes equal total sampling
probability. In expectation, the sample should be split roughly equally across years
(~2,250 per year for a 4-year pool).

```{r validate-years}
#| eval: false

cat("Sample composition by GSS year:\n")
profiles |>
  count(year) |>
  mutate(pct = round(100 * n / sum(n), 1)) |>
  print()
```

## Uniqueness check

```{r validate-unique}
#| eval: false

n_unique_pairs <- nrow(distinct(profiles, year, id))
cat(glue(
  "Unique (year, id) pairs: {n_unique_pairs} of {nrow(profiles)} ",
  "({if (n_unique_pairs == nrow(profiles)) 'PASS — no duplicates' else 'FAIL — duplicates present'})\n"
))
```

## Marginal distributions vs. weighted 2024 GSS

We use the 2024 wave (most recent) as the reference population. The pooled sample
should approximate 2024-weighted GSS marginals on key variables; deviations are
expected because older waves represent slightly different demographic compositions.

```{r validate-marginals}
#| eval: false

gss_2024 <- gss_complete |> filter(year == 2024)

validate_marginal <- function(var) {
  src <- gss_2024 |>
    count(.data[[var]], wt = wtssps) |>
    mutate(prop_gss2024 = n / sum(n)) |>
    select(-n)
  prof <- profiles |>
    count(.data[[var]]) |>
    mutate(prop_pool = n / sum(n)) |>
    select(-n)
  full_join(src, prof, by = var) |>
    mutate(diff = round(prop_pool - prop_gss2024, 4),
           prop_gss2024 = round(prop_gss2024, 4),
           prop_pool    = round(prop_pool, 4))
}

vars_to_check <- c("sex", "race", "degree", "partyid", "polviews", "region")

walk(vars_to_check, \(v) {
  cat("\n----", v, "----\n")
  print(validate_marginal(v))
})
```

# Step 7. Save output

```{r save}
#| eval: false

out_path <- here("clone_profiles", "profiles.csv")
write_csv(profiles, out_path)

cat(glue(
  "Saved: {out_path}\n",
  "  Rows    : {nrow(profiles)}\n",
  "  Columns : {ncol(profiles)}\n",
  "  Years   : {paste(sort(unique(profiles$year)), collapse = ', ')}\n"
))
```
